### 2016-12-06 qualification.tex
initial commit

\documentclass{article}
\usepackage{graphics}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage{helvet}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[margin=1.25in]{geometry}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{cite}
\usepackage{microtype}

\DisableLigatures{encoding = *, family = * }

\newcommand{\name}{Renan Greca}
\newcommand{\website}{www.renangreca.com}
\newcommand{\email}{renangreca@gmail.com}

\usepackage[font=small,labelfont=bf,labelsep=space]{caption}
\captionsetup{%
  figurename=fig.,
  tablename=tab.
}

\bibliographystyle{alpha}

\begin{document}

%\fontfamily{phv}\selectfont
\pagestyle{fancy}
\lhead{\name}
\chead{}
\rhead{\email}
\lfoot{}
\cfoot{\thepage}
\rfoot{}

\inputencoding{latin1}

\begin{center}
    \Large
    \textbf{} \\
    \textit{Renan Greca} \\
    %\coursetitle - Research Paper\\
    \small
\end{center}

\large
\doublespacing

\section{Introduction}

\section{Trust in VANETs}

[description of VANETs]

With the increasing feasibility of autonomous vehicles, the study of Vehicular Ad-hoc Networks (VANETs) is ever more important.
Before vehicles are completely autonomous, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
By allowing vehicles to share their locations, speeds, and destinations with each other, traffic can become more stable and predictable.

[description of trust in context of VANETs]

However, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) of the network.
If one node is malicious or faulty, it can spread incorrect data that may compromise the utility of the VANET.
Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.

[malicious and faulty nodes]

When it comes to trust in VANETs, there is the important distinction between a malicious node and a faulty one; both of them may be sharing false data, but for different reasons and with different consequences.
For example, a malicious node may lie about its location in order to make routing protocols use it as a hop \cite{leinmuller2005influence}, while a faulty GPS module may cause an accident because its position data was incorrect.
However, when using a binary trust model (one node can either trust or not another), that distinction is less relevant, so malicious and faulty nodes can be treated as the same in such a model.

[details of security and safety issues in VANETs that can be diminished by trust]

\textbf{[existing trust models applied in VANETs]}

Systems to handle trust in vehicular networks have been studied since, at least, 2005.
\cite{dotzer2005vars} is one of the earliest examples, establishing a system called VARS, based on the reputation of nodes and messages throughout the network. The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node will append its opinion of the message's contents. That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message. This trust model works as a way of adding credibility of a message through validation by its sender's peers in a non-centralized fashion. However, opinion piggybacking has its own share of problems. First, it means that forwarding nodes must be able to access (at least some of) the contents of a message so it can form an opinion on it, diminishing privacy; a malicious forwarding node could even attempt to alter those contents. Second, using previously appended opinions from other nodes to form a new opinion means that the first nodes to forward the message will have a substantially greater impact over the final opinion than the later ones. Finally, there is an issue with scalability, since appending new information to a message on each hop may add a significant overhead to the transmission. Additionally, the authors provide little to no experimentation or proof that their approach would be sound in a real-world network.

\section{Trust in Complex Networks}
[description of complex networks]

[examples of complex networks]

[details of how trust can be established in complex networks]

[existing trust models applied in complex networks]

[small world phenomenon]

\textbf{\cite{vernize2015malicious}}

\section{Complex network trust model applied in VANETs}

\section{Proposal}

[how we plan on adapting the previous works]

[what do we expect to change]

[what do we expect to be the challenges]

[scenario context (speed, density, etc)]

\bibliography{qualification}

\end{document}

### 2017-03-23 ufpr/1-intro/chapter1.tex
new version of Qualification text

\chapter{Introduction}

%=====================================================

% A introdução geral do documento pode ser apresentada através das seguintes seções: Desafio, Motivação, Proposta, Contribuição e Organização do documento (especificando o que será tratado em cada um dos capítulos). O Capítulo 1 não contém subseções\footnote{Ver o Capítulo \ref{cap-exemplos} para comentários e exemplos de subseções.}.

%=====================================================

### 2017-03-23 ufpr/2-complexnetworks/chapter2.tex
new version of Qualification text

\chapter{Complex Networks}
\label{chap:complexnetworks}

Complex networks can describe many systems which are observed in nature and society \cite{newmannetworks}, such as computer networks, food chains, protein structures, etc. 
They are generally divided into four categories:

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs), networks of mobile devices which communicate with each other without the need of a centralized agent.
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		In \autoref{chap:vanets}, a special type of MANET will be introduced, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as is contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		The brain contains a neural network formed by neurons, cells which enable information processing; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.

\end{enumerate}

As this work relates to trust issues in complex networks, it is interesting to further examine social networks and the importance of trust within them.
While the end result of the study will be applied to technological networks, certain aspects of social networks provide useful analogies to trust management in other kinds of networks.

\section{Trust in Social Networks}
\label{section:trustsocial}
In a traditional social network, it is simple to perceive how trust is relevant and how it works, since trust relationships between people (friends, relatives, colleagues, etc.) are used on a daily basis to make decisions.
When adapted to a digital environment, these social relationships can be used to automatically increase the relevance of certain information.
For instance, upon reading an online review for a certain product, a user will be more likely to accept the review's conclusion if it was written by a close friend than if it were written by a stranger.
Furthermore, if the review was written by a person who the user knows to be malicious or uninformed, its contents will be even less relevant.
In short, trust is a way of estimating how much a certain recommendation will lead to a positive outcome. \cite{golbeck2006inferring}

Social networks also have the property of carrying trust from one relationship to another:
information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust.
Friends of friends \cite{boissevain1974friends} might not have very high trust values, but could still be considered more trustworthy than the average stranger.
This property is similar, but not identical, to transitivity, since trust is diminished for each extra step an origin node needs to reach a destination, and there is also the possibility that one node distrusts another even if they share a mutual friend.

Social networks also have the trait of being mostly static.
Although friendships are formed and ended frequently, those connections do not disrupt the general shape of the network.
In fact, the ending of a friendship might indicate the presence of a \textit{negative} trust relationship between two people.
Such negative information can be useful in digital social networks since, frequently, only positive relationships are available.
% \textemdash for example, on Facebook, people are usually only connected if there is a positive relationship between them.

\section{Existing trust models for Social Networks}

Many works exist regarding trust relationships in complex networks, and in social networks in particular.
Here, only a few examples will be described, in order to show the general idea of trust research for these types of networks.

The most important previous work on complex network trust for the trust model proposed in this paper will be detailed in \autoref{chap:proposal}.

%=====================================================

### 2017-03-23 ufpr/3-vanets/chapter3.tex
new version of Qualification text

\chapter{Vehicular Ad-hoc Networks}
\label{chap:vanets}

The prospect of autonomous vehicles makes the study of Vehicular Ad-hoc Networks (VANETs) important.
However, it will still take many years before completely autonomous vehicles are the norm and, until then, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
Thousands of people die each year due to traffic accidents, many of which could have been avoided if vehicles were smarter and able to provide important alerts both to their drivers and to neighboring vehicles.
The vehicles' on-board computers can aid drivers to increase safety and reduce traffic, but data needs to be shared among the various vehicles in the network.
Such data includes location, speed and destination, which neighboring vehicles can use to adjust their trajectory and increase their own efficiency.
% By allowing vehicles to share their locations, speeds, and destinations with each other, traffic can become more stable and predictable.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/vanet.png}
    \caption{Basic elements of a VANET: OBUs and RSUs. \cite{saini2015close}}
    \label{fig:vanet}
\end{figure}

Current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
Amongst other aspects of the wireless technology, the WAVE standard describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
On-board units are computers placed within each vehicle which monitor the vehicle's data and are able to communicate with other nodes using wireless signals.
Road-side units are placed in static locations near roads; they may also have wired interfaces with other RSUs and the Internet, so it is possible to use them as anchor points for Internet access for passing vehicles.

%[description of trust in context of VANETs]

\section{Trust in VANETs}

Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) of the network.
If one node is malicious or faulty, it can spread incorrect data that may compromise the utility of the VANET.
Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
Examples include triggering false alarms about inexistent accidents, lying about the average speed in one road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms. Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.

There is an important distinction between a malicious node and a faulty one; both of them may be sharing false data, but for different reasons and with different consequences.
For example, a malicious node may lie about its location in order to make routing protocols use it \cite{leinmuller2005influence}, in order to try to store or alter messages, while a faulty GPS module may cause an accident because its position data was incorrect.
However, that distinction can be hard to make, because a close inspection is necessary to determine whether the incorrect data is erratic or deliberate.
Since both types of nodes are problematic to the proper functioning of a network, malicious and faulty nodes can be treated as the same in a trust model.

%\textbf{/*********}
%Trust can be a related, yet distinct, issue to cryptography and authentication in vehicular networks.
%Many studies propose the use of a PKI (Public Key Infrastructure) to guarantee legitimate identities for nodes, although this solution requires, by definition, a certifying authority to manage and distribute public keys.
%VANETs must be able to function in a completely independent fashion, regardless of the amount of active nodes in a given region, and without relying on a centralized service.
%Both trust and authentication must work together to provide an adequate solution for VANETs, since it is difficult to establish trust without means to verify an identity (a malicious node may provide false data about its identity to bypass trust solutions), while a cryptography system requires a pre-existing infrastructure and may add unacceptable latency to high-priority messages.
%\textbf{[moved to chapter 4]}
%\textbf{*********/}

In general, trust management solutions for VANETs use \textit{data-oriented trust}, \textit{entity-based trust}, or a combination of the two.
The solutions that use data-oriented trust (or \textit{data-centric trust}) \cite{raya2008data} focus on validating messages instead of entities.
This is important when vehicles share messages about a specific event, such as a collision, which must be quickly validated by neighbors and distributed to other nodes within a relevant area.
In this scenario, vehicles sharing the same road might be complete strangers to each other, and therefore will not have any trust relationship, so neighboring nodes must decide if a message is true by its contents and by other nodes' observations of the event.
On the other hand, when dealing with frequent messages which contain basic information such as geolocation and speed (used for traffic-diminishing solutions), it is too costly to judge each individual message.
Therefore, \textit{entity-based trust} becomes more appealing, since benign nodes can quickly identify a malicious node and isolate it from the network.
These models take advantage of the possibility of nodes meeting more than once and, therefore, being able to form a long-term trust relationship with each other.

%\textbf{/*********}
%In \cite{gerlach2007trust}, the author shows a general framework of how security and trust can be organized in a vehicular network.
%It proposes three layers: a service plane, which contains the core functionalities of the network, such as communication technologies, routing algorithms and the security measures directly related to them (such as cryptography); a middleware plane which handles trust and privacy issues; and the application layer, which contains the applications that will use the other layers' data to make decisions.
%The model shows neatly how security and trust can interact with each other to provide best results.
%\textbf{*********/}

%[malicious and faulty nodes]

%[details of security and safety issues in VANETs that can be diminished by trust]

\section{Existing trust models for VANETs}

Several models have been proposed to solve the problem of trust in vehicular networks. 
In this section, some of the most relevant ones will be described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
In this section, some of the most important ones will be detailed.
There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.

\cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node will append its opinion of the message's contents.
That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
This process adds credibility to a message through validation by nodes in a non-centralized fashion.
It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
One interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
However, opinion piggybacking has its own share of problems.
First, it allows forwarding nodes to access (at least some of) the contents of a message so it can form an opinion on it, diminishing privacy; a malicious forwarding node could even attempt to alter those contents.
Second, using previously appended opinions from other nodes to form a new opinion causes the first nodes to forward the message to have a substantially greater impact over the final opinion than the later ones.
Finally, there is an issue with scalability, since appending new information to a message on each hop may add a significant overhead to the transmission. Additionally, the authors provide little to no experimentation or proof that their approach would be sound in a real-world network.

The model proposed in \cite{minhas2010towards} uses several criteria to judge whether or not a received message is trustworthy.
First, nodes are classified by their roles and previous experience with them.
Roles are used for vehicles which should be more trustworthy than the average: government official cars, traffic report vans, buses, cabs, etc.
Nodes also store their experience each time an event message is received (if one neighboring node reported an event which did not turn out to be true, its trust value is reduced).
Additionally, messages have higher reliability when their senders are closer in time and space to the reported event.
When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \autoref{chap:proposal}.

In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
When a node sends a message, one node in the cluster (the leader) must aggregate the other node's opinions on that message.
Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; furthermore, nodes that receive the message will only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
Furthermore, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy. 
The model considers role-based trust (i.e. static trust of pre-authenticated vehicles such as police units) and experience-based trust, which is calculated using knowledge of the outcomes related to previous messages.

The trust model in \cite{park2011long} takes advantage of daily commutes.
In this article, the focus is on the early stages of VANETs, in which a very small percentage of vehicles will be equipped with OBUs.
To make trust viable in such a scenario, the authors rely on RSUs to store reputation information from passing vehicles.
Each vehicle must have an ``Agent RSU'', which will be tasked with sharing that vehicle's trust data to other vehicles and RSUs as well as keeping the data updated when the vehicle is near it once again.
To make this viable, the properties of daily commutes are used: it is assumed the vehicle will be near that RSU with reasonable frequency because it is located within the driver's home-to-work route.
The main problem with this model is that it relies on the presence of frequent RSUs, which might not always be viable.
It also does not make it clear what should happen when a vehicle stops using a route or does not have a daily predictable path (it does, however, handle occasions in which a vehicle chooses an alternate route or is absent for some days such as weekends and holidays).

The authors of \cite{huang2014social} take special note of two characteristics from social networks that can also be found in many VANET trust models: \textit{information cascading} and \textit{oversampling}.
That is, information reported by a number of original nodes (i.e. the ones that witnessed an event) may be diluted as nodes that forward it append their own opinions on the matter.
An algorithm is proposed to diminish that effect by assigning higher weights to the opinions of origin nodes and lower weights  to others.
However, the authors conclude that the optimal scenario is to assign no weight at all to forwarding nodes, therefore allowing each node to form an opinion based only on the original nodes' reports.
Furthermore, the authors are quick to dismiss the validity of entity-based trust, instead opting for a pure data-oriented approach.
Since their model is based only on events, it also does not consider the usefulness of trust in more mundane cases such as the frequent sharing of location and speed among vehicles.

\begin{table}[h!]
\centering
\begin{tabular}{ |p{3cm}||p{1cm}|p{1cm}|p{1cm}|p{1cm}| }
 \hline
 & Dotzer & Chen & Huang & Park\\
 \hline
 Decentralized   & AF    &AFG&   004 &\\
 \hline
\end{tabular}
\caption{Features of VANET trust models}
\label{table:1}
\end{table}

%=====================================================

### 2017-03-23 ufpr/4-socialvanets/chapter4.tex
new version of Qualification text

\chapter{Special properties of VANETs}
\label{chap:socialvanets}

%=====================================================

As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
However, VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
Some of these properties include:

\begin{enumerate}

	\item 	Rapidly changing topology.
			Since the nodes are vehicles, they move frequently and at relatively high speeds.
			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.

	\item	Node mobility is constrained to a pre-existing grid of roads.
			Within those roads, nodes will usually travel in predictable directions according to local laws and historical data.
			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.

	\item	VANETs are prone to fragmentation, since a gap in the network topology can make two parts of it unable to communicate with each other.
			Combined with the property above, this fragmentation can appear and disappear frequently, depending on the node density.

	\item	Due to the changing topology and possible disconnection, connection with distant nodes is not reliable.
			Therefore, the effective diameter of the network is relatively small for important applications.

	\item 	Compared to devices like smartphones, vehicles have no notable power constraints.

	\item 	In certain locations and/or moments, large vehicle density will create a large-scale network, since there will be many nodes concentrated in a relatively small space.

	\item 	The topology is susceptible to driver behavior.
			First, this means the topology can occasionally change in unpredictable ways.
			Second, contents of a message sent through the network can alter the driver's behavior and therefore change the topology.

\end{enumerate}

\section{Implications for trust management}

Despite the technological similarities between VANETs and MANETs, the differences in node behavior in VANETs creates the necessity of distinguished trust management solutions for those types of network.
Trust or reputation management systems in MANETs are usually developed around properties of MANETs such as: devices with constrained battery power, low-speed or no node mobility, and long contact times during each encounter.
As is described above, such properties are not valid for VANETs.

A few examples of trust management models for MANETs are shown below, and the reasons why they could not be applied to VANETs without significant modifications are explained.

### 2017-03-23 ufpr/5-proposal/chapter5.tex
new version of Qualification text

\chapter{Proposal}
\label{chap:proposal}

As described above, properties derived from social networks will be used to adapt a trust model developed for static complex networks to dynamic ones.
These properties allow for a long-term trust relationships to be established amongst nodes of a network and, therefore, previously existing trust models can be adapted to the vehicular environment.

In this chapter, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment.
Afterwards, a list of changes that will be made to the model, in order for it to handle dynamic networks, is presented.

\section{Social Networks and VANETs}
Several of the aforementioned articles state that vehicular networks are, by definition, ephemeral.
That is, the likelihood of two nodes meeting each other twice is too low to be relevant.
In this work, it is argued that such claims are not entirely true.
It stands to reason that, throughout the course of several days, many drivers will take similar routes at similar times of day (e.g. to commute to work) and, therefore, their vehicles will be in similar locations each day.
Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
Since that is true for a notable percentage of a city's fleet, we can also assume that those vehicles may frequently encounter each other during their commute.
While two vehicles that share a commute route may not be direct neighbors every day, they will likely be relatively close to each other, meaning few hops separate them in the ad-hoc network.

Additionally, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
Examples of these include vehicles whose owners are drivers or coworkers.
Such vehicles' trust relationship should become steady over time and, in the case of positive trust, they can use each other's information to learn more about other nodes in the network.

%\textbf{[WORKING DAY MOBILITY MODEL]}

In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
%They do not, however, present data regarding specific pairs of nodes.
%[Should we propose to show this information in our work?]

%Regardless of these properties of commuting vehicles, the probability of two specific vehicles meeting each other more than once, in given time, tends to 1.
%So it is also reasonable to assume that two vehicles, whose drivers reside in the same city, over a long period of time (e.g. one year) may meet each other more than once.
%However, if they only meet rarely, their trust opinions about each other may not be very useful.

Most cities also have one or more types of mass transit systems (buses or trains).
Those vehicles can also be part of a VANET and communicate with private cars.
Buses share the same roads as cars, but instead of having specific destinations, they travel a predefined route during the whole day, usually tied to a tight schedule.
Trains travel on rails, so their contact with cars is less frequent, but it can also happen on railroad crossings; they travel long distances in relatively short amount of time, which helps the dissemination of data in a VANET.
In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they will meet the same buses and/or trains frequently.

%Another way of applying trust models from social networks to VANETs would be to take into account the relationships between drivers.
%If a neighboring vehicle happens to be driven by a close friend or relative, there is reason to believe that it is not a malicious node in the network.
%That might not always be the case, since there may be malicious hackers controlling the node and its driver is oblivious to it.
%In any case, this approach is not covered in this work.

While certification of identity is an issue in vehicular network, this work will not address it.
There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing} 
\cite{kumar2015intelligent}.
One study even proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time is a prerequisite for the model proposed here.

\section{Working Day Movement Model}

To make use of the above properties, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
Therefore, the Working Day Movement Model \cite{ekman2008working} will be used.
The model, developed for used in Delay-Tolerant Network (DTN) simulations, includes many of the features that will be necessary to simulate the daily movement of a vehicular network.

As the name implies, the Working Day model abstracts people's movement from their homes to their offices and back.
Each node has a home and a workplace and they need to travel back and forth between those locations on a daily basis.
Occasionally, nodes can also go to other locations for leisure.
Most VANET trust models use the Random Waypoint mobility model, i.e. each node has an origin point, chooses a random location, gets to that location, then chooses another random location and goes there, and so forth.
While this model is efficient for testing trust protocols, it doesn't truly represent vehicle mobility in the real world.
As mentioned above, many drivers have routes they travel on daily, so the Working Day model is a more accurate representation, although it represents the mobility of humans instead of vehicle and requires some adjustments, which will be listed later in this section.

The model proposed by the authors makes use of several other models for specific tasks.
The main mobility model defines nodes and gives them their destinations.
Within it, five other models are used:
\begin{enumerate}
\item 
The \textbf{home activity submodel} describes what nodes do at night, within their homes.
No movement is modeled.
Nodes can be relatives or neighbors, and therefore share the same home.
\item 
The \textbf{office activity submodel} describes the nodes' routines within their offices.
Nodes can go to other locations within the office (such as meeting rooms) and such movement is modeled.
Nodes with share the same office are coworkers.
\item 
The \textbf{evening activity submodel} is responsible for mobility outside the nodes' standard routine. 
They can meet at certain locations (such as restaurants) and spend a few hours gathered with friends.
\item
The \textbf{transport submodel} shows how nodes move around the city.
It includes another tier of submodels, responsible for modeling three different types of transportation: walking, driving, and riding a bus.
Nodes which own a car always use it, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
The walking and driving submodels represent similar types of movement, although at different speeds, while the bus submodel follows cyclical routes and can take or deliver passengers at bus stops.
\item
The \textbf{map} represents the city in which the simulation runs.
Its streets constrain the movement of nodes and all homes and offices must be within the map boundaries.
The map can be divided into districts, which increases what the authors define as \textit{locality}.
In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes will rarely leave the district.
Nodes which reside and work in different districts serve to connect the network with their commutes.
\end{enumerate}

By thinking of these submodels for vehicles instead of people, it can become apparent how the frequency and length of encounters between nodes are similar in both instances.
If two vehicles belong to family members or neighbors, they will likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
Cars can also meet each other frequently if their drivers go out with friends.
In the vehicular case, there is the added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, although the drivers are likely complete strangers.

In the original article, nodes are devices (such as smartphones) being carried by humans.
Therefore, the Working Day model represents not only people's movements inside their cars, but also within their offices, walking on foot, or riding a bus.
To adapt it to a VANET environment, changes need to be made to their definition of node, since they now represent vehicles instead of people. Some of those changes are as follows:
\begin{enumerate}
\item
The office activity submodel no longer needs to model movement within the office and can be identical to the home submodel.
In both, a node can move a small amount once after reaching the office or home, to simulate parking.
This can be done using the Working Day model's parameters.
\item
The walking submodel needs to be disabled, since all nodes will be either cars or buses.
\item
The bus submodel needs to be changed so that each bus is one node in the network, which follows a predefined route with bus stops.
In the original model, each bus could carry several nodes, but this is no longer necessary.
\end{enumerate}
Other necessary changes might become apparent during the development of the study.

One important topic raised in the Working Day movement model article is the use of two metrics: \textit{inter-contact times} and \textit{contact duration}.
The choice of this movement model for tests is more strongly related to inter-contact times, i.e. how much time it takes for two nodes to meet again.
On the other hand, the contact duration is how long each meeting lasts.
For the reasons explained earlier in this section, relatively short inter-contact times is important for the proposed trust model.
Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although taking it into consideration might add excessive complexity to the model.

\section{Original trust model}

The basis of the proposed trust model is \cite{vernize2015malicious}.
This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.

The input graph $G = (V,E)$ is a static, connected, and directed graph containing all trust relationships in the network.
Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.

The process for identifying malicious nodes within $G$ is as follows:

First, $G$ is separated into strongly components using Tarjan's algorithm \cite{tarjan1972depth}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another completely; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $T = (V', E')$.

The creation of the graph $T$ simplifies the remaining computation.
Since each node of $T$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $G$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
After the formation of $T$, one or two heuristics can be used to classify the nodes as benign or malicious.

The \textit{coloring heuristic} uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
After running either algorithms with graph $T$, the color with the most nodes in $T$ is classified as correct, and all others are classified as malicious.
Once this information in $T$ is brought back to graph $G$, it is trivial to label the nodes in $G$ as either benign or malicious based on their components' classifications.

The other alternative is the \textit{articulation heuristic}.
It uses a graph articulation algorithm, which does a depth-first search and classifies nodes based on their roles in the spanning tree.
A node $n \in V$ is an articulation if its removal from $G$ makes $G$ a disconnected graph or, in other words, increases the number of connected components of $G$.
Once articulation nodes are found, the heuristic classifies each node as correct if it does not have neighbors that are also articulations.

In the experiments shown in the article, the coloring heuristic shows promising results, identifying a high ratio of the malicious nodes in the network.
The coloring and articulation heuristics can be combined, although the experiments using this combination display a high number of false positives.
Therefore, for the purposes of this paper, only the coloring heuristic will be considered.

\section{Changes}
While the algorithm described above works well for static graphs, some important modifications will have to be made to accommodate dynamic graphs.
The main changes planned are described below.

\subsection{Trust graph}
The original work's scope does not include the generation of the trust graph, as it is a parameter to the algorithm.
In dynamic graphs, trust relationships change over time, so how those changes affect our trust model have to be considered.
This includes not only the full figure of the graph, but also each node's views on its neighbors.

\subsection{Strongly connected components}
Strongly connected components, or components in which all nodes have positive trust relationships with each other, are somewhat different in dynamic graphs.
As nodes enter, leave, or change locations in the network, the components will change accordingly.
Therefore, our algorithm will have to work using ``snapshots'' of the graph, meaning each measurement will be valid for a specific timestamp.
In a later step, it might be useful to consider how specific nodes may change over several snapshots.

\subsection{Graph knowledge}
In the original paper, the malicious nodes are identified by a central entity with access to all nodes' trust values.
Not only does our work plan to decentralize this procedure, having complete knowledge of the graph is unfeasible in large-scale dynamic networks such as VANETs.
Therefore, our algorithm needs to work using only a partial understanding of the network.
For this to be viable, each node will likely have to gather information not only from its views on its neighbors, but also from its neighbor's views and so forth.

The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).

\subsection{Binary trust}
A model using binary trust is one in which trust values can only be 0 or 1.
If node $A$ positively trusts node $B$, the value of the edge $A\rightarrow B$ is 1.
If it does not trust $B$, or is unsure, the value is 0.
Such a scheme makes sense in a static graph but, in a dynamic one, it is interesting to have a range of possible trust values.

Since nodes will interact with each other occasionally and might learn information from other nodes, a range such as $[0,1]$ is useful to express how one trust value can change over time.
As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.

\subsection{Datasets}
The datasets used for experimentation must be changed accordingly.
Instead of using data gathered from social networks, which are static, dynamic network datasets will need to be used, such as traffic information for cities.

An example of such dataset is the Zurich Trace \cite{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
This dataset doesn't necessarily need to be for VANETs per se, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.
The mobility of the nodes is the most important aspect of the dataset.

%=====================================================

### 2017-03-28 qualification.tex
initial commit
new version

@@
 \usepackage{microtype}
+\graphicspath{ {images1/} }
 
@@
 
-\section{Trust in VANETs}
+\section{Complex Networks}
+Complex networks can describe many systems which are observed in nature and society, such as computer networks, food chains, protein structures, etc. \cite{newmannetworks}
+They are frequently divided into four categories: technological networks, social networks, information networks, and biological networks.
 
-[description of VANETs]
+\textbf{/*********}
+Many networks have some concept of trust.
+In peer-to-peer information networks, nodes have complete control over their own data \cite{vernize2015malicious}, so other nodes must trust that the data they receive is the one they requested.
+Trust in the context of social networks will be elaborated in this section, while technological networks will be further examined in the next section.
+\textbf{*********/}
 
-With the increasing feasibility of autonomous vehicles, the study of Vehicular Ad-hoc Networks (VANETs) is ever more important.
-Before vehicles are completely autonomous, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
-By allowing vehicles to share their locations, speeds, and destinations with each other, traffic can become more stable and predictable.
+\subsection{Trust in Social Networks}
+In a traditional social network, it is simple to perceive how trust is relevant and how it works, since trust relationships between people (friends, relatives, colleagues, etc.) are used on a daily basis to make decisions.
+When adapted to a digital environment, these social relationships can be used to automatically increase the relevance of certain information.
+For instance, upon reading an online review for a certain product, a user will be more likely to accept the review's conclusion if it was written by a close friend than if it were written by a stranger.
+Furthermore, if the review was written by a person who the user knows to be malicious or uninformed, its contents will be even less relevant.
+In short, trust is a way of estimating how much a certain recommendation will lead to a positive outcome. \cite{golbeck2006inferring}
 
-[description of trust in context of VANETs]
+Social networks also have the property of carrying trust from one relationship to another:
+information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
+Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust.
+Friends of friends \cite{boissevain1974friends} might not have very high trust values, but could still be considered more trustworthy than the average stranger.
+This property is similar, but not identical, to transitivity, since trust is diminished for each extra step an origin node needs to take to reach it (a friend of a friend is more trustworthy than a stranger, but is less than a direct friend).
 
-However, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) of the network.
+Social networks also have the trait of being mostly static.
+Although friendships are formed and ended frequently, those connections do not disrupt the general shape of the network.
+In fact, the ending of a friendship might indicate the presence of a \textit{negative} trust relationship between two people.
+Such negative information can be useful in digital social networks since that data might not be available.
+% \textemdash for example, on Facebook, people are usually only connected if there is a positive relationship between them.
+
+\subsection{Existing trust models for Complex Networks}
+
+%[description of complex networks]
+
+%[examples of complex networks]
+
+%[details of how trust can be established in complex networks]
+
+%[existing trust models applied in complex networks]
+
+%[small world phenomenon]
+
+\section{VANETs}
+
+%[description of VANETs]
+
+%With the increasing feasibility of smart and autonomous vehicles, the study of Vehicular Ad-hoc Networks (VANETs) is ever more important \textbf{[CITE]}.
+The prospect of autonomous vehicles makes the study of Vehicular Ad-hoc Networks (VANETs) important.
+However, it will still take many years before completely autonomous vehicles are the norm and, until then, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
+The vehicles' on-board computers can aid drivers to increase safety and reduce traffic, but data needs to be shared among the various vehicles in the network.
+Such data includes location, speed and destination, which neighboring vehicles can use to adjust their trajectory and increase their own efficiency.
+% By allowing vehicles to share their locations, speeds, and destinations with each other, traffic can become more stable and predictable.
+Thousands of people die each year due to traffic accidents, many of which could have been avoided if vehicles were smarter and able to provide important alerts both to their drivers and to neighboring vehicles.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/vanet.png}
+    \caption{Basic elements of a VANET: OBUs and RSUs. \cite{saini2015close}}
+    \label{fig:vanet}
+\end{figure}
+
+Current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
+Amongst other aspects of the wireless technology, the WAVE standard describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
+On-board units are computers placed within each vehicle which monitor the vehicle's data and are able to communicate with other nodes using wireless signals.
+Road-side units are placed in static locations near roads; they may also have wired interfaces with other RSUs and the Internet, so it is possible to use them as anchor points for Internet access for passing vehicles.
+
+%[description of trust in context of VANETs]
+
+\subsection{Trust in VANETs}
+
+Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) of the network.
 If one node is malicious or faulty, it can spread incorrect data that may compromise the utility of the VANET.
+Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
+Examples include triggering false alarms about inexistent accidents, lying about the average speed in one road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms.
 Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.
 
-[malicious and faulty nodes]
+There is an important distinction between a malicious node and a faulty one; both of them may be sharing false data, but for different reasons and with different consequences.
+For example, a malicious node may lie about its location in order to make routing protocols use it \cite{leinmuller2005influence}, while a faulty GPS module may cause an accident because its position data was incorrect.
+However, that distinction can be hard to make, because a close inspection is necessary to determine whether the incorrect data is erratic or deliberate.
+Since both types of nodes are problematic to the proper functioning of a network, malicious and faulty nodes can be treated as the same in a trust model.
 
-When it comes to trust in VANETs, there is the important distinction between a malicious node and a faulty one; both of them may be sharing false data, but for different reasons and with different consequences.
-For example, a malicious node may lie about its location in order to make routing protocols use it as a hop \cite{leinmuller2005influence}, while a faulty GPS module may cause an accident because its position data was incorrect.
-However, when using a binary trust model (one node can either trust or not another), that distinction is less relevant, so malicious and faulty nodes can be treated as the same in such a model.
+%\textbf{/*********}
+%Trust can be a related, yet distinct, issue to cryptography and authentication in vehicular networks.
+%Many studies propose the use of a PKI (Public Key Infrastructure) to guarantee legitimate identities for nodes, although this solution requires, by definition, a certifying authority to manage and distribute public keys.
+%VANETs must be able to function in a completely independent fashion, regardless of the amount of active nodes in a given region, and without relying on a centralized service.
+%Both trust and authentication must work together to provide an adequate solution for VANETs, since it is difficult to establish trust without means to verify an identity (a malicious node may provide false data about its identity to bypass trust solutions), while a cryptography system requires a pre-existing infrastructure and may add unacceptable latency to high-priority messages.
+%\textbf{[moved to chapter 4]}
+%\textbf{*********/}
 
-[details of security and safety issues in VANETs that can be diminished by trust]
+Some solutions use \textit{data-oriented trust} (or \textit{data-centric trust}) \cite{raya2008data}.
+Those solutions focus on validating messages instead of entities, and are especially important when dealing with event-based trust.
+This is important when vehicles share messages about a specific event, such as a collision, which must be quickly validated by neighbors and distributed to other nodes within a relevant area.
+In this scenario, vehicles sharing the same road might be complete strangers to each other, and therefore will not have any trust relationship, so neighboring nodes must decide if a message is true purely by its contents.
 
-\textbf{[existing trust models applied in VANETs]}
+On the other hand, when dealing with frequent messages which contain basic information such as geolocation and speed (used for traffic-diminishing solutions), it is too costly to judge each individual message.
+Therefore, \textit{entity-based trust} becomes more appealing, since benign nodes can quickly identify a malicious node and isolate it from the network.
+These models take advantage of the possibility of nodes meeting more than once and, therefore, being able to form a long-term trust relationship with each other.
 
-Systems to handle trust in vehicular networks have been studied since, at least, 2005.
-\cite{dotzer2005vars} is one of the earliest examples, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
+%\textbf{/*********}
+%In \cite{gerlach2007trust}, the author shows a general framework of how security and trust can be organized in a vehicular network.
+%It proposes three layers: a service plane, which contains the core functionalities of the network, such as communication technologies, routing algorithms and the security measures directly related to them (such as cryptography); a middleware plane which handles trust and privacy issues; and the application layer, which contains the applications that will use the other layers' data to make decisions.
+%The model shows neatly how security and trust can interact with each other to provide best results.
+%\textbf{*********/}
+
+%[malicious and faulty nodes]
+
+%[details of security and safety issues in VANETs that can be diminished by trust]
+
+\subsection{Existing trust models for VANETs}
+
+Several models have been proposed to solve the problem of trust in vehicular networks. 
+In this section, some of the most relevant ones will be described, taking into consideration the time in which they were proposed, the advantages they bring and their contributions to later study. 
+None of them provide a complete solution, but serve as pieces of a puzzle that is, at the time of writing, still incomplete. 
+Some review and/or survey articles have been written on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{mejri2014survey}, \cite{sengar2016survey}, and \cite{dwivedi2016review}.
+
+\cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
 The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node will append its opinion of the message's contents.
@@
 This trust model works as a way of adding credibility of a message through validation by its sender's peers in a non-centralized fashion.
+It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
+One interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
 However, opinion piggybacking has its own share of problems.
@@
 
-\section{Trust in Complex Networks}
-[description of complex networks]
+In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
+By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
+When a node sends a message, one node in the cluster (the leader) must aggregate the other node's opinions on that message.
+Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; furthermore, nodes that receive the message will only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
+However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
+Furthermore, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy. 
+The model takes into consideration role-based trust (i.e.
+static trust of pre-authenticated vehicles such as police units) and experience-based trust, which is calculated using knowledge of the outcomes related to previous messages.
 
-[examples of complex networks]
+The authors of \cite{huang2014social} take special note of two characteristics from social networks that can also be found in many VANET trust models: \textit{information cascading} and \textit{oversampling}.
+That is, information reported by a number of original nodes (i.e. the ones that witnessed an event) may be diluted as nodes that forward it append their own opinions on the matter.
+An algorithm is proposed to diminish that effect by assigning higher weights to the opinions of origin nodes and lower weights  to others.
+However, the authors conclude that the optimal scenario is to assign no weight at all to forwarding nodes, therefore allowing each node to form an opinion based only on the original nodes' reports.
+Furthermore, the authors are quick to dismiss the validity of [node-based reputation, check proper name], instead opting for a pure message-based approach.
+Since their model is based only on events, it also does not consider the usefulness of trust in more mundane cases such as the frequent sharing of location and speed among vehicles.
 
-[details of how trust can be established in complex networks]
+The trust model in \cite{park2011long} takes advantage of daily commutes.
+In this article, the focus is on the early stages of VANETs, in which a very small percentage of vehicles will be equipped with OBUs.
+To make trust viable in such a scenario, the authors rely on RSUs to store reputation information from passing vehicles.
+Each vehicle must have an ``Agent RSU'', which will be tasked with sharing that vehicle's trust data to other vehicles and RSUs as well as keeping the data updated when the vehicle is near it once again.
+To make this viable, the properties of daily commutes are used: it is assumed the vehicle will be near that RSU with reasonable frequency because it is located within the driver's home-to-work route.
+The main problem with this model is that it relies on the presence of frequent RSUs, which might not always be viable.
+It also does not make it clear what should happen when a vehicle stops using a route or does not have a daily predictable path (it does, however, handle occasions in which a vehicle chooses an alternate route or is absent for some days such as weekends and holidays).
 
-[existing trust models applied in complex networks]
-
-[small world phenomenon]
-
-\textbf{\cite{vernize2015malicious}}
+%[data-oriented vs entity-oriented trust]
 
@@
 
+Several of the aforementioned articles state that vehicular networks are, by definition, ephemeral.
+That is, the likelihood of two nodes meeting each other twice is too low to be relevant.
+In this work, it is argued that such claims are not entirely true.
+It stands to reason that, throughout the course of several days, many drivers will take similar routes at similar times of day (e.g.
+to commute to work) and, therefore, their vehicles will be in similar locations each day.
+Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
+Since that is true for a notable percentage of a city's fleet, we can also assume that those vehicles may frequently encounter each other during their commute.
+While two vehicles that share a commute route may not be direct neighbors every day, they will likely be relatively close to each other, meaning few hops separate them in the ad-hoc network.
+
+%\textbf{[WORKING DAY MOBILITY MODEL]}
+
+In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
+By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
+They do not, however, present data regarding specific pairs of nodes.
+%[Should we propose to show this information in our work?]
+
+Additionally, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
+Examples of these include vehicles whose owners are drivers or coworkers.
+Such vehicles' trust relationship should become steady over time and, in the case of positive trust, they can use each other's information to learn more about other nodes in the network.
+
+Regardless of these properties of commuting vehicles, the probability of two specific vehicles meeting each other more than once, in given time, tends to 1.
+So it is also reasonable to assume that two vehicles, whose drivers reside in the same city, over a long period of time (e.g. one year) may meet each other more than once.
+However, if they only meet rarely, their trust opinions about each other may not be very useful.
+
+Most cities also have one or more types of mass transit systems (buses or trains).
+Those vehicles can also be part of a VANET and communicate with private cars.
+Buses share the same roads as cars, but instead of having specific destinations, they travel a predefined route during the whole day, usually tied to a tight schedule.
+Trains travel on rails, so their contact with cars is less frequent, but it can also happen on railroad crossings; they travel long distances in relatively short amount of time, which helps the dissemination of data in a VANET.
+In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they will meet the same buses and/or trains frequently.
+
+Another way of applying trust models from social networks to VANETs would be to take into account the relationships between drivers.
+If a neighboring vehicle happens to be driven by a close friend or relative, there is reason to believe that it is not a malicious node in the network.
+That might not always be the case, since there may be malicious hackers controlling the node and its driver is oblivious to it.
+In any case, this approach is not covered in this work.
+
+While certification of identity is an issue in vehicular network, this work will not address it.
+There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}
+\cite{kumar2015intelligent}.
+One study even proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time is a prerequisite for the model proposed here.
+
+\subsection{Working Day Movement Model}
+
+To make use of the above properties, it is important to choose a mobility model that properly represents the way vehicles move on a daily basis in the real world.
+Therefore, the Working Day Movement Model \cite{ekman2008working} will be used.
+The model, developed for used in Delay-Tolerant Network (DTN) simulations, includes many of the features that will be necessary to simulate the daily movement of a vehicular network.
+
+As the name implies, the Working Day model abstracts people's movement from their homes to their offices and back.
+Each node has a home and a workplace and they need to travel back and forth between those locations on a daily basis.
+Occasionally, nodes can also go to other locations for leisure.
+Most VANET trust models use the Random Waypoint mobility model, i.e., each node has an origin point, chooses a random location, gets to that location, then chooses another random location and goes there, and so forth.
+While this model is efficient for testing trust protocols, it doesn't truly represent vehicle mobility in the real world.
+As mentioned above, many drivers have routes they travel on daily, so the Working Day model is a more accurate representation, although it represents the mobility of humans instead of vehicle and requires some adjustments, which will be listed later in this section.
+
+The model proposed by the authors makes use of several other models for specific tasks.
+The main mobility model defines nodes and gives them their destinations.
+Within it, five other models are used:
+\begin{enumerate}
+\item 
+The \textbf{home activity submodel} describes what nodes do at night, within their homes.
+No movement is modeled.
+Nodes can be relatives or neighbors, and therefore share the same home.
+\item 
+The \textbf{office activity submodel} describes the nodes' routines within their offices.
+Nodes can go to other locations within the office (such as meeting rooms) and such movement is modeled.
+Nodes with share the same office are coworkers.
+\item 
+The \textbf{evening activity submodel} is responsible for mobility outside the nodes' standard routine. 
+They can meet at certain locations (such as restaurants) and spend a few hours gathered with friends.
+\item
+The \textbf{transport submodel} shows how nodes move around the city.
+It includes another tier of submodels, responsible for modeling three different types of transportation: walking, driving, and riding a bus.
+Nodes which own a car always use it, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
+The walking and driving submodels represent similar types of movement, although at different speeds, while the bus submodel follows cyclical routes and can take or deliver passengers at bus stops.
+\item
+The \textbf{map} represents the city in which the simulation runs.
+Its streets constrain the movement of nodes and all homes and offices must be within the map boundaries.
+The map can be divided into districts, which increases what the authors define as \textit{locality}.
+In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes will rarely leave the district.
+Nodes which reside and work in different districts serve to connect the network with their commutes.
+\end{enumerate}
+
+By thinking of these submodels for vehicles instead of people, it can become apparent how the frequency and length of encounters between nodes are similar in both instances.
+If two vehicles belong to family members or neighbors, they will likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
+Cars can also meet each other frequently if their drivers go out with friends.
+In the vehicular case, there is the added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, although the drivers are likely complete strangers.
+
+In the original article, nodes are devices (such as smartphones) being carried by humans.
+Therefore, the Working Day model represents not only people's movements inside their cars, but also within their offices, walking on foot, or riding a bus.
+To adapt it to a VANET environment, changes need to be made to their definition of node, since they now represent vehicles instead of people. Some of those changes are as follows:
+\begin{enumerate}
+\item
+The office activity submodel no longer needs to model movement within the office and can be identical to the home submodel.
+In both these submodels, a node can move a small amount once after reaching the office or home, to simulate parking.
+This can be done using the Working Day model's parameters.
+\item
+The walking submodel needs to be disabled, since all nodes will be either cars or buses.
+\item
+The bus submodel needs to be changed so that each bus is one node in the network, which follows a predefined route with bus stops.
+In the original model, each bus could carry several nodes, but this is no longer necessary.
+\end{enumerate}
+Other necessary changes might become apparent during the development of the study.
+
+One important topic raised in the Working Day movement model article is the use of two metrics: \textit{inter-contact times} and \textit{contact duration}.
+The choice of this movement model for tests is more strongly related to inter-contact times, i.e. how much time it takes for two nodes to meet again.
+On the other hand, the contact duration is how long each meeting lasts.
+For the reasons explained earlier in this section, relatively short inter-contact times is important for the proposed trust model.
+Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although taking it into consideration might add excessive complexity to the model.
+
 \section{Proposal}
 
-[how we plan on adapting the previous works]
+As described above, properties derived from social networks will be used to adapt a trust model developed for static complex networks to dynamic ones.
+The basis of the proposed trust model is \cite{vernize2015malicious}.
+This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
 
-[what do we expect to change]
+First, a graph $G$ is separated into components so that each component contains only nodes that trust each other (in other words, any distrust in a relationship between two nodes is enough to separate them into different components).
+Then, on the supergraph $T$ whose nodes are $G$'s components, a graph coloring algorithm is used.
+The color with the most nodes in $T$ is deemed correct, while the others are deemed incorrect.
+This means the correct color contains only trustworthy nodes, and the others contain nodes of dubious trustworthiness.
 
-[what do we expect to be the challenges]
+\subsection{Changes}
+While the algorithm described above works well for static graphs, some important modifications will have to be made to accommodate dynamic graphs.
+The main changes planned are described below.
 
-[scenario context (speed, density, etc)]
+\subsubsection{Trust graph}
+The original work's scope does not include the generation of the trust graph, as it is a parameter to the algorithm.
+In dynamic graphs, trust relationships change over time, so how those changes affect our trust model have to be taken into consideration.
+This includes not only the full figure of the graph, but also each node's views on its neighbors.
+
+\subsubsection{Strongly connected components}
+Strongly connected components, or components in which all nodes have positive trust relationships with each other, are somewhat different in dynamic graphs.
+As nodes enter, leave, or change locations in the network, the components will change accordingly.
+Therefore, our algorithm will have to work using ``snapshots'' of the graph, meaning each measurement will be valid for a specific timestamp.
+In a later step, it might be useful to take into consideration how specific nodes may change over several snapshots.
+
+\subsubsection{Graph knowledge}
+In the original paper, the malicious nodes are identified by a central entity with access to all nodes' trust values.
+Not only does our work plan to decentralize this procedure, having complete knowledge of the graph is unfeasible in large-scale dynamic networks such as VANETs.
+Therefore, our algorithm needs to work using only a partial understanding of the network.
+For this to be viable, each node will likely have to gather information not only from its views on its neighbors, but also from its neighbor's views and so forth.
+
+The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
+Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
+That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
+They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
+The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).
+
+\subsubsection{Binary trust}
+A model using binary trust is one in which trust values can only be 0 or 1.
+If node $A$ positively trusts node $B$, the value of the edge $A\rightarrow B$ is 1.
+If it does not trust $B$, or is unsure, the value is 0.
+Such a scheme makes sense in a static graph but, in a dynamic one, it is interesting to have a range of possible trust values.
+
+Since nodes will interact with each other occasionally and might learn information from other nodes, a range such as $[0,1]$ is useful to express how one trust value can change over time.
+As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.
+
+\subsubsection{Datasets}
+The datasets used for experimentation must be changed accordingly.
+Instead of using data gathered from social networks, we will need to use dynamic network datasets, such as traffic information for cities.
+
+%[how we plan on adapting the previous works]
+
+%[what do we expect to change]
+
+%[what do we expect to be the challenges]
+
+%[scenario context (speed, density, etc)]
 

### 2017-03-28 ufpr/1-intro/chapter1.tex
new version

@@
 %=====================================================
+
+As computers grow in power and shrink in size, more aspects of everyday life can be enhanced by adding processing units to common devices.
+While many of these applications focus on conveniences, such as smart purchasing suggestions or automatic household control, computers can also be important to save time and save lives.
+One way of achieving this is by adding computers and wireless transmitters to vehicles — such as cars, buses, and trains — so they can share data which may increase traffic efficiency or reduce the chance of accidents.
+
+In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \cite{whotraffic}.
+While this number has greatly reduced over the past decades \cite{johnson2010traffic} fthanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \cite{whofactsheet}, so further action is necessary.
+Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \cite{inrixtraffic}.
+
+Smart vehicles and vehicular networks are ways that technology can aid both of the aforementioned problems.
+Through the use of sensors and wireless communications, these vehicles will be able to avoid accidents by alerting distracted drivers, or by knowing in advance another vehicle's position and speed.
+By communicating, they can also collaborate to distribute traffic over alternative routes and, therefore, reduce the possibility of traffic jams.
+These features are possible with the development of a vehicular ad-hoc network (VANET), in which vehicles can quickly share data amongst themselves without the need of a server or an Internet connection.
+
+However, as is the case for many technological networks, VANETs must be reliable in order to be functional.
+In the case of ad-hoc networks, the nodes themselves must be reliable, since data can only propagate safely through benign cooperation.
+Thus, the problem of trust arises in the context of vehicular networks.
+Without trust management\footnote{There is an important distinction between \textit{trust} and \textit{trust management}.
+The former relates to the actual trust relationship between a pair of nodes, while the latter concerns the management of previously generated trust data, such as previous tests or information received from neighboring nodes, and how this data can be used to make decisions.
+While this work addresses trust management rather than trust itself, both terms will be used when referring to trust management for the sake of brevity.} features, a vehicle cannot judge whether or not received messages are benign or malicious in order to make an informed decision and, therefore, incorrect data might be presented to the driver.
+
+\section{Document organization}
+In this study, a method to identify malicious nodes in a vehicular network is proposed, taking advantage of a previously existing study, which was limited to static networks.
+\Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
+Then, \cref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
+Finally, \cref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2017-03-28 ufpr/3-vanets/chapter3.tex
new version

@@
 \label{chap:vanets}
+
+Vehicular ad-hoc networks are a special instance of MANETs, in which the nodes of the network are related to vehicles and traffic.
 
@@
 %[description of trust in context of VANETs]
+
+\section{Special properties of VANETs}
+As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
+However, VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
+Some of these properties include:
+
+\begin{enumerate}
+
+	\item 	Rapidly changing topology.
+			Since the nodes are vehicles, they move frequently and at relatively high speeds.
+			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
+			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
+			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
+
+	\item	Node mobility is constrained to a pre-existing grid of roads.
+			Within those roads, nodes will usually travel in predictable directions according to local laws and historical data.
+			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.
+
+	\item	VANETs are prone to fragmentation, since a gap in the network topology can make two parts of it unable to communicate with each other.
+			Combined with the property above, this fragmentation can appear and disappear frequently, depending on the node density.
+
+	\item	Due to the changing topology and possible disconnection, connection with distant nodes is not reliable.
+			Therefore, the effective diameter of the network is relatively small for important applications.
+
+	\item 	Compared to devices like smartphones, vehicles have no notable power constraints.
+
+	\item 	In certain locations and/or moments, large vehicle density will create a large-scale network, since there will be many nodes concentrated in a relatively small space.
+
+	\item 	The topology is susceptible to driver behavior.
+			First, this means the topology can occasionally change in unpredictable ways.
+			Second, contents of a message sent through the network can alter the driver's behavior and therefore change the topology.
+
+\end{enumerate}
 
@@
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{ |p{3cm}||p{1cm}|p{1cm}|p{1cm}|p{1cm}| }
- \hline
- & Dotzer & Chen & Huang & Park\\
- \hline
- Decentralized   & AF    &AFG&   004 &\\
- \hline
-\end{tabular}
-\caption{Features of VANET trust models}
-\label{table:1}
-\end{table}
+%\begin{table}[h!]
+%\centering
+%\begin{tabular}{ |p{3cm}||p{1cm}|p{1cm}|p{1cm}|p{1cm}| }
+% \hline
+% & Dotzer & Chen & Huang & Park\\
+% \hline
+% Decentralized   & AF    &AFG&   004 &\\
+% \hline
+%\end{tabular}
+%\caption{Features of VANET trust models}
+%\label{table:1}
+%\end{table}
 

### 2017-03-28 ufpr/4-socialvanets/chapter4.tex
new version

@@
 %=====================================================
-
-As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
-However, VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
-Some of these properties include:
-
-\begin{enumerate}
-
-	\item 	Rapidly changing topology.
-			Since the nodes are vehicles, they move frequently and at relatively high speeds.
-			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
-			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
-			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
-
-	\item	Node mobility is constrained to a pre-existing grid of roads.
-			Within those roads, nodes will usually travel in predictable directions according to local laws and historical data.
-			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.
-
-	\item	VANETs are prone to fragmentation, since a gap in the network topology can make two parts of it unable to communicate with each other.
-			Combined with the property above, this fragmentation can appear and disappear frequently, depending on the node density.
-
-	\item	Due to the changing topology and possible disconnection, connection with distant nodes is not reliable.
-			Therefore, the effective diameter of the network is relatively small for important applications.
-
-	\item 	Compared to devices like smartphones, vehicles have no notable power constraints.
-
-	\item 	In certain locations and/or moments, large vehicle density will create a large-scale network, since there will be many nodes concentrated in a relatively small space.
-
-	\item 	The topology is susceptible to driver behavior.
-			First, this means the topology can occasionally change in unpredictable ways.
-			Second, contents of a message sent through the network can alter the driver's behavior and therefore change the topology.
-
-\end{enumerate}
 

### 2017-03-28 ufpr/5-proposal/chapter5.tex
new version

@@
 
-As described above, properties derived from social networks will be used to adapt a trust model developed for static complex networks to dynamic ones.
-These properties allow for a long-term trust relationships to be established amongst nodes of a network and, therefore, previously existing trust models can be adapted to the vehicular environment.
+This work proposes that it is possible to adapt a malicious node identification scheme originally designed for static networks to a dynamic environment, such as a VANET.
+In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
+The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
+Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
 
-In this chapter, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment.
-Afterwards, a list of changes that will be made to the model, in order for it to handle dynamic networks, is presented.
+By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
+This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as input for a malicious node detection algorithm.
+
+%The previous chapters show why vehicular networks require distinguished trust models compared to social networks and traditional MANETs.
+%In this chapter, it is proposed that certain aspects from social networks can be found in vehicular networks and, through those aspects, it is possible to adapt a trust model designed for traditional complex networks to a vehicular environment.
+%These social properties allow for long-term trust relationships to be established amongst nodes of a network and, therefore, previously existing trust models can be adapted to the vehicular environment.
+
+In this chapter, the details for the process above are presented.
+First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
+Then, the Working Day Movement Model is presented, along with the reasons why it is best fitted for this work.
+Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
+Finally, a schedule for the next steps of the project is shown.
 
 \section{Social Networks and VANETs}
+\label{section:socialvanets}
 Several of the aforementioned articles state that vehicular networks are, by definition, ephemeral.
@@
 Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
-Since that is true for a notable percentage of a city's fleet, we can also assume that those vehicles may frequently encounter each other during their commute.
+Since that is true for a notable percentage of a city's fleet, it can also be assumed that those vehicles may frequently encounter each other during their commute.
 While two vehicles that share a commute route may not be direct neighbors every day, they will likely be relatively close to each other, meaning few hops separate them in the ad-hoc network.
-
-Additionally, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
+Furthermore, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
 Examples of these include vehicles whose owners are drivers or coworkers.
@@
 %\textbf{[WORKING DAY MOBILITY MODEL]}
-
-In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
-By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
 %They do not, however, present data regarding specific pairs of nodes.
@@
 
+In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
+By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
+
 %Another way of applying trust models from social networks to VANETs would be to take into account the relationships between drivers.
@@
 \cite{kumar2015intelligent}.
-One study even proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
-Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time is a prerequisite for the model proposed here.
+One study proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time, so that nodes can identify each other over a long period, is a prerequisite for the model proposed here.
 
 \section{Working Day Movement Model}
+\label{section:workingday}
 
@@
 
-\section{Changes}
+\section{Necessary changes to the malicious node identification scheme}
+\label{section:changes}
 While the algorithm described above works well for static graphs, some important modifications will have to be made to accommodate dynamic graphs.
@@
 
+Since the graph changes with time, for every timestamp $t$, there is a corresponding graph $G_t=(V_t, E_t)$, which includes the available vertices and edges in that moment in time.
+
 \subsection{Graph knowledge}
@@
 
+Thus, each node will have its own model of the network, and will generate a trust graph $G_i=(V_i, E_i)$, in which $i$ is the node's identifier.
+Combined with the necessity of timestamps explained above, this will result in a graph $G_{it}=(V_{it}, E_{it})$ for each node and each timestamp, which will be the input of the new algorithm.
+Accordingly, there are also component graphs $T_{it}=(V'_{it}, E'_{it})$, generated on each instance of the algorithm.
+
+During the early stages of development, it will be considered that vertices cannot enter or leave the network and, therefore, only the layout of the edges changes.
+This is sufficient to experiment with the algorithm's performance with the added complexity of a dynamic network.
+
 \subsection{Binary trust}
@@
 As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.
+The rate in which trust increases or decreases must be experimented with; it could be useful to have trust decrease faster than it increases, but that might also cause false positives in the case of occasional mistakes.
 
@@
 
+At first, the objective is to test using the Working Day Movement Model, setting parameters that contribute to the validation of the algorithm.
+When the simulation begins, a set number of nodes will be malicious and, therefore, other nodes will have low trust in them after interacting.
+In time, benign and malicious nodes can change their behavior and, accordingly, their trustworthiness will be updated.
+
+It might still be useful to simulate using real-world movement traces.
 An example of such dataset is the Zurich Trace \cite{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
-This dataset doesn't necessarily need to be for VANETs per se, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.
+The chosen dataset doesn't necessarily need to be created for VANETs, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.
 The mobility of the nodes is the most important aspect of the dataset.
 
+\section{Tasks and schedule}
+
+Given the proposed trust management scheme, the following list of tasks has been developed:
+
+\begin{enumerate}
+	\item \textbf{Changes to the malicious node identification algorithm}. Develop and adapt the algorithm using the changes listed in \autoref{section:changes}.
+	\item \textbf{Changes to the movement model}. Develop and adapt the Working Day Movement Model according to \autoref{section:workingday}.
+	\item \textbf{Study and selection of a simulator}.
+Several simulators for vehicular networks exist, and one must be chosen in order to run simulations using the newly developed model.
+The Working Day Movement Model was tested with the ONE Simulator \cite{keranen2009one}, however it was made for DTNs rather than VANETs.
+Many VANET simulators, such as TraNS \cite{piorkowski2008trans}, were developed by combining a network simulator (such as ns-2) and a vehicular mobility model (such as SUMO \cite{behrisch2011sumo}).
+One possible solution is to adapt the Working Day model to an existing network simulator.
+	\item \textbf{Simulations and adjustments}.
+Once the core of the project has been developed and a simulation has been chosen, simulations begin.
+The models may be tweaked in order to achieve better results, and all data from the process will be collected.
+	\item \textbf{Analysis of data and results}.
+With the data collected from simulations, it will be possible to observe the outcome of the proposed trust model.
+If tweaks were made to the model during simulations, those changes can be listed, showing which ones provided better results.
+	\item \textbf{Writing and submission of article}.
+Having concluded development, an article will be written explaining the motivations, challenges and results of the study.
+This article will be submitted to a congress.
+	\item \textbf{Writing of dissertation}. Finally, the whole process will be detailed within the dissertation.
+\end{enumerate}
+
+An approximate schedule is shown in \autoref{table:schedule}.
+The first column displays the task, using the numbers from the list above, while the remaining ones are marked when work on that task is expected for the given month.
+The months of April through December are in 2017 and January through March are in 2018.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{ |p{0.7cm}||p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}| }
+ \hline
+ \textbf{Task} & Apr & May & Jun & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
+ \hline
+ \hline
+ \textbf{1} & \checkmark & \checkmark & \checkmark & & & & & & & & &\\
+ \hline
+ \textbf{2} & & & \checkmark & \checkmark & \checkmark & & & & & & &\\
+ \hline
+ \textbf{3} & \checkmark & \checkmark & \checkmark & \checkmark & & & & & & & &\\
+ \hline
+ \textbf{4} & & & & & \checkmark & \checkmark & & & & & &\\
+ \hline
+ \textbf{5} & & & & & \checkmark & \checkmark & & & & & &\\
+ \hline
+ \textbf{6} &  &  & & & & \checkmark & \checkmark & \checkmark & & & &\\
+ \hline
+ \textbf{7} &  &  & & & & & & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
+ \hline
+\end{tabular}
+\caption{Expected schedule}
+\label{table:schedule}
+\end{table}
+
+%Cronograma
+%
+%1. Adaptação do modelo de nós maliciosos
+%2. Adaptação do modelo de mobilidade
+%3. Escolha e estudo do simulador
+%4. Simulações
+%5. Análise de dados e resultados
+%6. Escrita e submissão de artigo (procurar congresso)
+%7. Escrita da dissertação
+
 %=====================================================

### 2017-03-30 ufpr/1-intro/chapter1.tex
newest revision

@@
 
-However, as is the case for many technological networks, VANETs must be reliable in order to be functional.
+When dealing with safety or traffic-efficiency applications, it is crucial that network communications occur with low latency, which is why an ad-hoc solution is preferred.
+Current technology could easily be used by connecting vehicles to the Internet through the cellular network, but the milliseconds added by the transmission could be enough to make certain applications unfeasible.
+Cellular connections also have other problems: the connection would require an active subscription with a carrier; the connection depends on available infrastructure; the wireless frequency would be shared with phones and other mobile devices, increasing the possibility of interference and congestion; server-side issues could impact the vehicles' communications.
+
+All these issues are solved by VANETs, since vehicles share data amongst themselves, using their own wireless transmitters, without having to rely on external devices.
+Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, autonomous vehicles to plan a platooning approach to traffic \cite{amoozadeh2015platoon}.
+In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes will pick up and forward.
+That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{images/collision.png}
+    \caption{Propagation of a collision alert in a VANET}
+    \label{fig:collision}
+\end{figure}
+
+However, as is the case for many types of networks, VANETs must be reliable in order to be functional.
 In the case of ad-hoc networks, the nodes themselves must be reliable, since data can only propagate safely through benign cooperation.
@@
 
+Often, when people hear the term ``network'', they think only of technological examples, such as computer networks.
+However, the study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}; technological networks, and VANETs by extension, are only a subset of that much larger field of study.
+
+Trust issues and their solutions are an important aspect of the study of complex networks, and of social or technological networks in particular.
+Each different type of network requires a specific approach to trust, according to the features that distinguish it from others, so it is interesting to review the solutions proposed for social networks and technological networks in general, in order to then narrow down on how dynamic networks, and VANETs in particular, are handled.
+Despite the differences, there are some concepts that overlap across most types of networks.
+
+This work proposes a method for identifying malicious nodes in a dynamic and decentralized network, using VANETs as a use-case scenario.
+The solution is based on a previously existing algorithm \cite{vernize2015malicious}, which was limited to static networks.
+In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
+These features are found by tracing analogies to social networks, i.e.
+identifying how and why two nodes (in the case of VANETs, vehicles) would meet more than once and at a somewhat predictable interval.
+
+%Further explain VANETs
+
+%VANETs can be complex networks
+
+%why mention social networks
+
 \section{Document organization}
-In this study, a method to identify malicious nodes in a vehicular network is proposed, taking advantage of a previously existing study, which was limited to static networks.
+The remainder of this document is organized as follows:
+
 \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.

### 2017-03-30 ufpr/2-complexnetworks/chapter2.tex
new version
newest revision

@@
 		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
-		In \autoref{chap:vanets}, a special type of MANET will be introduced, along with several details regarding trust in those types of networks.
+		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
+		VANETs, which are special types of MANETs, will be introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
 	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
@@
 
-As this work relates to trust issues in complex networks, it is interesting to further examine social networks and the importance of trust within them.
-While the end result of the study will be applied to technological networks, certain aspects of social networks provide useful analogies to trust management in other kinds of networks.
+While issues of trust have been studied on most types of networks, this work will emphasize on social and technological networks, for two reasons: first, these are the most studied types of networks in the field of computer science; and second, previous study on both can provide useful insights towards trust in vehicular networks.
+It is clear that VANETs are part of technological networks; in this work, however, it is also argued that they have social network features, which is why they will also be described in this chapter.
+
+%As this work relates to trust issues in complex networks, it is interesting to further examine social networks and the importance of trust within them.
+%While the end result of the study will be applied to technological networks, certain aspects of social networks provide useful analogies to trust management in other kinds of networks.
 
@@
 
-Social networks also have the trait of being mostly static.
-Although friendships are formed and ended frequently, those connections do not disrupt the general shape of the network.
-In fact, the ending of a friendship might indicate the presence of a \textit{negative} trust relationship between two people.
-Such negative information can be useful in digital social networks since, frequently, only positive relationships are available.
-% \textemdash for example, on Facebook, people are usually only connected if there is a positive relationship between them.
+In general, social networks are mostly static.
+Although friendships are formed and ended frequently, those connections do not disrupt the general shape of the network, because members of the network will usually have other friends to rely on.
+Even if a certain person's trust integrity is compromised due to a specific incident, that person's friends are not necessarily also deemed untrustworthy, preserving part of the trust graph.
 
-\section{Existing trust models for Social Networks}
+%Many works exist regarding trust relationships in complex networks, and in social networks in particular.
+%Here, only a few examples will be described, in order to show the general idea of trust research for these types of networks.
 
-Many works exist regarding trust relationships in complex networks, and in social networks in particular.
-Here, only a few examples will be described, in order to show the general idea of trust research for these types of networks.
+\section{Trust in Technological Networks}
+\label{section:trusttechnological}
+In conventional technological networks, such as the Internet, trust is defined and applied quite differently from social networks.
+Generally, trust is very centralized through services that offer security to users.
+Examples of this can be a IP filtering to avoid distributed denial of service (DDoS) attacks or web browser extensions that block requests to domains in a blacklist.
+However, all these centralized solutions would be too slow to be viable in a dynamic and ad-hoc network.
 
-The most important previous work on complex network trust for the trust model proposed in this paper will be detailed in \autoref{chap:proposal}.
+MANETs require a decentralized approach to trust management and, therefore, are more similar to social networks.
+In both cases, each member of a network has its own opinions about other members, and these opinions can change over time.
+However, while social networks are mostly static, the topology of a MANET changes frequently as nodes move around.
+[Perhaps this is a good place to mention trust vs. trust management]
+Nodes must be able to handle trust with its immediate neighbors for any given moment, so it can properly filter out potentially malicious data.
+In networks in which nodes can meet more than once, it may be valuable to store information from previous encounters to use in the future, although this process can be too slow or resource-consuming to be viable in certain devices.
+
+The most common instance of MANETs is using mobile devices, such as smartphones, being carried by humans.
+Although these networks are dynamic, their mobility is relatively low in relation to the wireless range of the devices — if two people are walking in opposite directions, their phones may communicate for several seconds before they leave each other's range.
+Trust solutions for MANETs use this property to their advantage, since it allows one node to test another and check several of the messages sent between them.
+
+Another important aspect of trust in ad-hoc networks is that information is, generally, uncertain and incomplete \cite{baras2005cooperation}.
+That is, since nodes form their own model and opinions of the surrounding network, it is unlikely that this data will be certain and accurate with reality.
+For this reason, it is also possible to use data gathered by neighboring nodes to complement the model.
+This data itself is subject to the trust evaluation of the neighbors, but it is crucial to better approximate the trust values of other nodes.
+Incompleteness is an inherit trace of MANETs, since it is entirely possible for nodes to be too distant to communicate, and only occasionally come into contact.
+
+Finally, MANETs of mobile devices must take into consideration the processing and battery limitations of those devices.
+Nodes may disable wireless communications to save power and therefore become uncooperative, or it may be too slow to be a reliable source of information.
 

### 2017-03-30 ufpr/5-proposal/chapter5.tex
newest revision

@@
 To make use of the above properties, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
-Therefore, the Working Day Movement Model \cite{ekman2008working} will be used.
+Therefore, the Working Day Movement Model \cite{ekman2008working} (WDM) will be used.
 The model, developed for used in Delay-Tolerant Network (DTN) simulations, includes many of the features that will be necessary to simulate the daily movement of a vehicular network.
@@
 Several simulators for vehicular networks exist, and one must be chosen in order to run simulations using the newly developed model.
-The Working Day Movement Model was tested with the ONE Simulator \cite{keranen2009one}, however it was made for DTNs rather than VANETs.
-Many VANET simulators, such as TraNS \cite{piorkowski2008trans}, were developed by combining a network simulator (such as ns-2) and a vehicular mobility model (such as SUMO \cite{behrisch2011sumo}).
-One possible solution is to adapt the Working Day model to an existing network simulator.
+The Working Day Movement Model was tested with the ONE Simulator \cite{keranen2009one}.
+The source code for both the simulator and the WDM generation can be found online \cite{onerepo}, and will likely be used for the necessary simulations.
 	\item \textbf{Simulations and adjustments}.

### 2017-06-01 qualification/0-preambulo/abstract.tex
final version of qualification text

\begin{otherlanguage}{english}

\begin{abstract}

% em inglês, o primeiro parágrafo não deve ser indentado
\noindent
As computers become small and powerful, the possibility of integrating them into everyday objects is ever more appealing.
By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular ad-hoc network (VANET), in which cars share data amongst themselves in order to cooperate and make roads safer and more efficient.
A decentralized ad-hoc solution, which doesn't rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
To generate these trust values, nodes use information acquired from past interactions; nodes which frequently share false or irrelevant data will have lower trust values than the ones which appear to be reliable.
This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
This movement model allows the comparison of VANETs to traditional social networks, because it can be observed that pairs of vehicles are likely to meet more than once in several scenarios: for example, they can belong to neighbors or work colleagues, or simply take similar routes every day.
Through these repeated encounters, a trust relationship can be developed between a pair of nodes.
The resulting trust value can also be used to aid other nodes which might not have a developed relationship with each other.
The proposed algorithm is based on a previously existing one, which was developed for centralized networks and focused on static ad-hoc models; the previous algorithm will be adapted to serve a decentralized and dynamic network, which is the case of VANETs.
Using existing trust values, a directed graph is modeled in which edges represent the trust relationship between pairs of nodes.
Then, graph components are formed in which no pair of nodes within a single component has a negative trust relationship.
A graph coloring algorithm is used on the resulting components graph and, using the coloring results, it is possible to infer which nodes are considered malicious by the consensus of the network.
The complete algorithm is expected to be fast, so it can be executed frequently, and will allow nodes to maintain a model of the surrounding networks indicating which neighboring nodes can be trusted or not.

%With the possibility of integrating smart features into vehicles,

%The abstract should be the English translation of the ``resumo'', no more, no less.

\end{abstract}

\end{otherlanguage}

### 2017-06-01 qualification/1-intro/chapter1.tex
final version of qualification text

\chapter{Introduction}
\label{chap:introduction}
%=====================================================

% A introdução geral do documento pode ser apresentada através das seguintes seções: Desafio, Motivação, Proposta, Contribuição e Organização do documento (especificando o que será tratado em cada um dos capítulos). O Capítulo 1 não contém subseções\footnote{Ver o Capítulo \ref{cap-exemplos} para comentários e exemplos de subseções.}.

%=====================================================

As computers grow in power and shrink in size, more aspects of everyday life can be enhanced by adding processing units to common devices.
While many of these applications focus on conveniences, such as home automation \cite{mccole2016} (the collection of connected and smart devices is dubbed the Internet of Things or IoT \cite{morgan2014}), the integration of computers with other objects and devices can also be important to save time and save lives \cite{rti2014}.
One way of achieving this is by adding computers and wireless transmitters to vehicles — such as cars, buses, and trains — so they can share data which may increase traffic efficiency or reduce the chance of accidents \cite{saini2015close}.

In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \cite{whotraffic}.
While this number has greatly reduced over the past decades \cite{johnson2010traffic} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \cite{whofactsheet}, so further actions are necessary.
Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \cite{inrixtraffic}.

Smart vehicles and vehicular networks are ways that technology can aid both of the aforementioned problems.
Through the use of sensors and wireless communications, these vehicles are able to avoid accidents by alerting distracted drivers \cite{lee2004collision}, or by knowing in advance another vehicle's position and speed \cite{hafner2011automated}.
By communicating, they can also collaborate to offer driving and route suggestions, therefore reducing the possibility of traffic jams \cite{knorr2012reducing}.
%These features are possible with the development of a vehicular ad-hoc network (VANET), in which vehicles can quickly share data amongst themselves without the need of a server or an Internet connection.

When dealing with safety or traffic-efficiency applications, it is crucial that network communications occur with low latency (approximately 100 milliseconds \cite{camp2005vehicle}).
Current cellular technology, such as LTE, could be used to connect vehicles to the Internet, but the delay added by the transmission would make safety applications unfeasible or unreliable \cite{mangel2010comparison}.
Cellular connections also have other problems: the connection would require an active subscription with a carrier; the connection depends on available infrastructure; the wireless frequency would be shared with phones and other mobile devices, increasing the possibility of interference and congestion; server-side issues could impact the vehicles' communications.

For these reasons, ad-hoc solutions are preferred over centralized ones.
An ad-hoc network is one that has no reliance on pre-existing infrastructure (such as routers or access points) \cite{wu2004ad}.
Instead, each node is able to communicate directly with others and a routing protocol allows for messages to be forwarded until they reach their destinations.
Every time a node wants to send a message and the recipient is not a direct neighbor, it must choose which nearby node is the most likely one to get the message to its destination.
Routing techniques can use either the network's topology or geographical coordinates \cite{saini2015close} to choose which node should be the next hop.

These issues — the additional safety and efficiency as well as the low-latency communications — can be tackled through the use of a vehicular ad-hoc network (VANET), in which vehicles share data amongst themselves without relying on external devices, an Internet connection or server availability.
Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \cite{amoozadeh2015platoon}.
In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \cite{li2007routing}.
That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/collision.png}
    \caption{Propagation of a collision alert in a VANET}
    \label{fig:collision}
\end{figure}

As is the case with most new technologies, VANETs are expected to be a notable target of attacks for a diversity of reasons \cite{isaac2010security}.
A local malicious user might alter the data his or her vehicle broadcasts in order to manipulate traffic conditions, while remote attackers could invade vehicles' computers and obtain partial control of the network  \cite{garip2015congestion}.
These attacks can vary from time-consuming annoyances to life-threatening, so it is important that real-world implementations of vehicular networks are prepared to handle them.

%However, as is the case for many types of networks, VANETs must be reliable in order to be functional.

%In the case of ad-hoc networks, the nodes themselves must be reliable, since data can only propagate safely through benign cooperation.

%In ad-hoc networks, the secure propagation of data depends on the reliability of each node, since messages often travel

%While in a conventional technological network (such as the Internet), the 
In ad-hoc networks, one way to mitigate a number of attacks is through each node using data collected from previous experiences to filter out incoming messages that seem to be malicious, incorrect, or irrelevant.
A node's degree of confidence that some data is correct and useful is called \textit{trust}.
%The ability of nodes to verify one another and judge messages based on past experiences is called \textit{trust}.
% The product of a series of past experiences with one other node is called trust.
For instance, in the example of a single malicious user broadcasting false data, nodes receiving these messages can use their own sensors to verify whether or not the data was correct, and update the \textit{trust value} of the sender vehicle.
In case the trust value of a sender is too low, a receiver node can choose to ignore the data contained in a message, as it concluded that the sender is not trustworthy.
Trust allows for better cooperation of nodes in a network, since incorrect messages might be detected and discarded.

Furthermore, once nodes form their own opinions about others, they can propagate pre-existing trust values when necessary.
For example, if two nodes are not direct neighbors and do not have any pre-existing trust information about each other, they can ask intermediary nodes for their opinions on the other node \cite{wang2009trust}.
The management of trust values (i.e. how one node acquires and updates trust values) and the use of these values to derive further information (such as designating nodes as malicious or not) is called \textit{trust management} \cite{ma2011survey}. 
%The combination of one node's trust of another with the opinions of other nearby nodes form what is called \textit{trust management}.
The effective use of trust management allows for the detection of malicious, misbehaving or faulty nodes in the network, and for such information to be shared amongst the benign participants. 
Throughout this document, trust and trust management may be used interchangeably.

%There is an important distinction between \textit{trust} and \textit{trust management}.
%The former relates to the actual trust relationship between a pair of nodes, while the latter concerns the management of previously generated trust data, such as previous tests or information received from neighboring nodes, and how this data can be used to make decisions.
%While this work addresses trust management rather than trust itself, both terms will be used when referring to trust management for the sake of brevity.
%Without trust management\footnote{} features, a vehicle cannot judge whether or not received messages are benign or malicious in order to make an informed decision and, therefore, incorrect data might be presented to the driver.

While the detection of incorrect nodes and/or messages is an important aspect of security and safety in vehicular networks, it does not address all of the problems.
Trust solutions are not viable without a solid identity verification scheme, for instance, since nodes would not be able to form trust opinions without being sure of the others' identities (these schemes often use a Public Key Infrastructure \cite{wasef2010complementing}).
They also do not address issues such as driver and passenger privacy when using the facilities of a VANET.
Furthermore, cryptography must be used in order to guarantee that the secrecy of messages is not violated.
Therefore, trust and trust management must be viewed as an important aspect of vehicular ad-hoc networks, but not as a definitive solutions to all of the related concerns.

In order to study the implications of trust in vehicular networks, it is interesting to first take a closer look at trust in other kinds of networks.
VANETs are a subset of technological networks, therefore it is useful to consider how the Internet or other ad-hoc networks handle trust.
Furthermore, VANETs contain several features often found in social networks, which can be directly tied to how nodes form and develop trust relationships with each other.

%While this work emphasizes on the characteristics of VANETs, they are only a subset of the much broader field of complex networks.
%The study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}.
%Trust issues are present in many of these types of networks, although each one requires a specific approach that addresses the peculiarities involved.

%Often, when people hear the term ``network'', they think only of technological examples, such as computer networks.
%Such networks are only a subset of a much larger field of study: that of complex networks.
%The study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}.
%Technological networks, and VANETs by extension, are only a subset of that much larger field of study.
%
%Trust issues and their solutions are an important aspect of the study of complex networks, and of social or technological networks in particular.
%Each different type of network requires a specific approach to trust, according to the features that distinguish it from others.
%It is interesting to review the solutions proposed for social networks and technological networks in general, and then narrow down on how dynamic networks, and VANETs in particular, are handled.
%Despite the differences, there are some concepts that overlap across most types of networks.

This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.
The solution is based on a previously existing algorithm \cite{vernize2015malicious}, which was limited to centralized networks.
In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
These features are found by tracing analogies to social networks, i.e. identifying how and why two nodes (in the case of VANETs, vehicles) would meet more than once and at a somewhat predictable interval.

% Seems like there's a paragraph missing to explain why we mention social and technological networks at all.

%Further explain VANETs

%VANETs can be complex networks

%why mention social networks

\section{Document organization}
The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2017-06-01 qualification/3-vanets/chapter3.tex
final version of qualification text

--- 2017-03-28 ufpr/3-vanets/chapter3.tex
@@
 
-Vehicular ad-hoc networks are a special instance of MANETs, in which the nodes of the network are related to vehicles and traffic.
+Today, most premium vehicles come equipped with hardware that allow for connectivity features; it is expected that, by 2022, many standard vehicles will also come with such features built-in, accounting for a substantial share of the automotive industry's revenue \cite{connectedcar2016}.
+Although these features can be useful tools to aid drivers, reducing traffic and risk of accidents, they are merely a gateway to the long-term goal of truly autonomous vehicles, which might become a reality within the next decade; many automakers and technology companies have laid out their plans for the upcoming years \cite{tesla2016}.
+However, the proper functioning and utility of both connected and fully autonomous vehicles rely on technologies, protocols and applications that allow for the fast communication between vehicle's on-board computers.
 
-The prospect of autonomous vehicles makes the study of Vehicular Ad-hoc Networks (VANETs) important.
-However, it will still take many years before completely autonomous vehicles are the norm and, until then, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
-Thousands of people die each year due to traffic accidents, many of which could have been avoided if vehicles were smarter and able to provide important alerts both to their drivers and to neighboring vehicles.
-The vehicles' on-board computers can aid drivers to increase safety and reduce traffic, but data needs to be shared among the various vehicles in the network.
-Such data includes location, speed and destination, which neighboring vehicles can use to adjust their trajectory and increase their own efficiency.
+Vehicular ad-hoc networks, which are a special instance of MANETs, are a much-studied solution to the problems in the way of smart and autonomous vehicles.
+In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
+By quickly sharing data with neighboring vehicles, without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
+
+%The advantages of an ad-hoc approach, as opposed to using cellular infrastructure or other server-based solutions, are explained in \autoref{chap:introduction}.
+%For messages to travel to nodes out of the sender's wireless range, they must use other nodes as \textit{hops}.
+%This approach requires appropriate routing protocols that consider the dynamic topology of VANETs.
+%While traditional MANETs often use protocols which rely on the network topology, the ones proposed for VANETs generally use geographic position instead \cite{saini2015close}; these protocols use GPS coordinates in order to send messages to the available neighbor physically closest to the destination node.
+
+%Certain cars available for purchase today are able to achieve some of those features (including autonomy) through a variety of sensors (cameras, accelerometers, microphones, etc.) which capture information about the world around them.
+%However, these sensors don't provide a complete solution, since they are subject to a variety of interferences.
+%The combination of built-in sensors with connectivity may provide the
+
+%The prospect of autonomous vehicles makes the study of Vehicular Ad-hoc Networks (VANETs) important.
+%However, it will still take many years before completely autonomous vehicles are the norm and, until then, VANETs can be a valuable tool to help drivers reduce travel times and diminish the risk of collisions.
+%Thousands of people die each year due to traffic accidents, many of which could have been avoided if vehicles were smarter and able to provide important alerts both to their drivers and to neighboring vehicles.
+%The vehicles' on-board computers can aid drivers to increase safety and reduce traffic, but data needs to be shared among the various vehicles in the network.
+%Such data includes location, speed and destination, which neighboring vehicles can use to adjust their trajectory and increase their own efficiency.
 % By allowing vehicles to share their locations, speeds, and destinations with each other, traffic can become more stable and predictable.
@@
 
-Current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
-Amongst other aspects of the wireless technology, the WAVE standard describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
+Several current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
+Among other aspects of the wireless technology, the WAVE standard describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
 On-board units are computers placed within each vehicle which monitor the vehicle's data and are able to communicate with other nodes using wireless signals.
 Road-side units are placed in static locations near roads; they may also have wired interfaces with other RSUs and the Internet, so it is possible to use them as anchor points for Internet access for passing vehicles.
+When referring to the communication between two OBUs, the term vehicle-to-vehicle (V2V) communication is used \cite{yang2004vehicle}; when both an OBU and an RSU are involved, it is called vehicle-to-infrastructure (V2I) communication \cite{chou2009feasibility}.
+Although this nomenclature is important to understand other studies on the subject of VANETs, this study does not consider RSUs and focuses only on vehicles themselves as nodes, so references to VANETs and vehicular networks are exclusively tied to V2V communications.
+
+%When referring to VANETs, the terms vehicle-to-vehicle (V2V) and vehicle-to-infrastructure (V2I) are used.
+%Both are types of communications that may occur within vehicular networks, depending on which nodes are actively participating in the networks.
+%Aside from the vehicles themselves, VANETs can use road-side infrastructure to provide additional features, such as Internet access, to its users.
+%In the model proposed in this work, however, such infrastructure is not used, so references to VANETs and vehicular networks are exclusively tied to V2V communications.
 
@@
 
+In traditional networks (ad-hoc or not), routing protocols usually use the topology to choose where to forward packets; in other words, the primary metric used is the number of hops required to reach the destination.
+This metric is not as useful in vehicular networks, since the high mobility causes the topology to change frequently.
+Instead, most VANET routing protocols use geographical coordinates to forward packets \cite{saini2015close}, that is, the physical distance between two nodes is used as the primary metric.
+The implication is that, even if a packet requires more hops to reach its destination, it will always be traveling the generally correct direction.
+
+%[Explain basic security features]
+
+As expected for a new technology being introduced, vehicular communications can become an appealing target for malicious users and attackers.
+These are some examples of possible issues in a VANET:
+
+\begin{enumerate}
+	\item 	Vehicles with faulty GPS modules, speedometers or other sensors.
+			If a vehicle is broadcasting incorrect data (perhaps unknowingly) because of a hardware or software fault, it can be a serious hinderance to efficiency and safety applications.
+			It might behaving appropriately according to protocol, but the data it sends is not reliable \cite{isaac2010security}.
+	\item	Vehicles might be deliberately broadcasting false data.
+			In this case, there might be a specific purpose (by either the vehicle's driver or a remote attacker), like altering traffic or even cause an accident \cite{golle2004detecting}.
+	\item   Attackers with control of several vehicles can propagate junk data in an attempt to flood the network, causing a distributed denial of service (DDoS) attack.
+			Alternatively, the data propagated might have some reasoning behind it, like lying about road conditions in order to divert traffic \cite{garip2015congestion}.
+	\item	Instead of sending data, some vehicles might try to eavesdrop on others' communications.
+			The hop-based routing protocols used in VANETs facilitate this, since any node can be asked to be a hop.
+			If the intermediary node is malicious, it may attempt to extract data contained in messages or refuse to forward them.
+			Related to this, there is the Sybil attack \cite{isaac2010security}, in which a node lies about its position in order to seemingly alter the physical topology of the network and be chosen as a hop \cite{leinmuller2005influence}.
+	\item	Malicious vehicles may use signal jammers or other devices in order to affect other vehicles' sensors and communications \cite{isaac2010security}.
+			That can cause other vehicles to broadcast incorrect data, therefore obfuscating the origin of the attack.
+	\item	A malicious user or remote attacker can monitor messages shared across the network in an attempt to stalk one specific vehicle \cite{isaac2010security}.
+
+\end{enumerate}
+
+Each of these possible attacks requires a unique approach, though there are some broader ways to help the security and safety of VANET users.
+Trust, as is described in \autoref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
+It does not, however, avoid all possible attacks, such as a signal jammer or stalking.
+Rather, different mechanisms must be explored in order to avoid most problems.
+
+%Related to trust is the issue of identity in vehicular networks.
+%Solutions to this issue can be important elements of security and privacy, and there are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}.
+%In fact, at least one study proposes that these identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+%However, refreshing identities can make maintaining a trust relationship between two nodes difficult or even impossible, since there is no way to know which nodes changed identities.
+%Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, persistent identities is a prerequisite for the model proposed here.
+
+% While certification of identity is an issue in vehicular networks, and , this work will not address it.
+%There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}
+%\cite{kumar2015intelligent}.
+%One study proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+%Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time, so that nodes can identify each other over a long period, is a prerequisite for the model proposed here.
+
 \section{Special properties of VANETs}
-As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
-However, VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
+%As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
+%However, 
+VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
 Some of these properties include:
@@
 			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
-			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
-			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
+%			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
+%			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
 
 	\item	Node mobility is constrained to a pre-existing grid of roads.
-			Within those roads, nodes will usually travel in predictable directions according to local laws and historical data.
+			Within those roads, nodes usually travel in predictable directions according to local laws and historical data.
 			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.
@@
 
-	\item 	In certain locations and/or moments, large vehicle density will create a large-scale network, since there will be many nodes concentrated in a relatively small space.
+	\item 	In certain locations and/or moments, large vehicle density results in a large-scale network, since there are many nodes concentrated in a relatively small space.
 
@@
 
+Some of these properties provide advantages or disadvantages when developing trust models for vehicular networks, although all of them must be considered.
+
 \section{Trust in VANETs}
-
-Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) of the network.
-If one node is malicious or faulty, it can spread incorrect data that may compromise the utility of the VANET.
+\label{section:trustvanet}
+Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) which participate in it.
+If one node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
 Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
@@
 This is important when vehicles share messages about a specific event, such as a collision, which must be quickly validated by neighbors and distributed to other nodes within a relevant area.
-In this scenario, vehicles sharing the same road might be complete strangers to each other, and therefore will not have any trust relationship, so neighboring nodes must decide if a message is true by its contents and by other nodes' observations of the event.
+In this scenario, vehicles sharing the same road might be complete strangers to each other, and therefore would not have any trust relationship, so neighboring nodes must decide if a message is true by its contents and by other nodes' observations of the event.
 On the other hand, when dealing with frequent messages which contain basic information such as geolocation and speed (used for traffic-diminishing solutions), it is too costly to judge each individual message.
 Therefore, \textit{entity-based trust} becomes more appealing, since benign nodes can quickly identify a malicious node and isolate it from the network.
-These models take advantage of the possibility of nodes meeting more than once and, therefore, being able to form a long-term trust relationship with each other.
+% These models take advantage of the possibility of nodes meeting more than once and, therefore, being able to form a long-term trust relationship with each other.
+Within entity-based trust, there are also two often-used methods of establishing trust: first, there is \textit{role-based trust}, which is the static trust of pre-authenticated vehicles such as police units; second, there is \textit{experience-based trust}, which is built through previous encounters shared between pairs of nodes.
+The model proposed in this work utilizes entity-based and experience-based trust, as it is based on the possibility of nodes meeting more than once and, therefore, being able to form a long-term trust relationship with each other.
+%It is not, however, mutually exclusive to other solutions
 
@@
 Several models have been proposed to solve the problem of trust in vehicular networks. 
-In this section, some of the most relevant ones will be described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
+In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
 None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
 Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
-In this section, some of the most important ones will be detailed.
 There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
@@
 \cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
-The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node will append its opinion of the message's contents.
+The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents.
 That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
@@
 When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
-Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \autoref{chap:proposal}.
+The model considers both role-based trust and experience-based trust; although the work proposed here does not use role-based trust, the authors provide a useful method of calculating and updating an experience-based trust value, which might be used or adapted.
+However, their model relies only on direct interaction between pairs of nodes, so no form of indirect trust (that is, trust values received from other nodes) is considered.
+
+%Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \autoref{chap:proposal}.
 
@@
 By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
-When a node sends a message, one node in the cluster (the leader) must aggregate the other node's opinions on that message.
-Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; furthermore, nodes that receive the message will only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
+When a node sends a message, one node in the cluster (the leader) must aggregate the other nodes' opinions on that message.
+Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; furthermore, nodes that receive the message only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
 However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
-Furthermore, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy. 
-The model considers role-based trust (i.e.
+Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy.
+%The whole premise of organizing a VANET in clusters is  
+%The model considers role-based trust (i.e.
 static trust of pre-authenticated vehicles such as police units) and experience-based trust, which is calculated using knowledge of the outcomes related to previous messages.
+%This model considers both role-based and experience-based trust.
 
 The trust model in \cite{park2011long} takes advantage of daily commutes.
-In this article, the focus is on the early stages of VANETs, in which a very small percentage of vehicles will be equipped with OBUs.
+In this article, the focus is on the early stages of VANETs, in which a very small percentage of vehicles are equipped with OBUs.
 To make trust viable in such a scenario, the authors rely on RSUs to store reputation information from passing vehicles.
-Each vehicle must have an ``Agent RSU'', which will be tasked with sharing that vehicle's trust data to other vehicles and RSUs as well as keeping the data updated when the vehicle is near it once again.
-To make this viable, the properties of daily commutes are used: it is assumed the vehicle will be near that RSU with reasonable frequency because it is located within the driver's home-to-work route.
+Each vehicle must have an ``Agent RSU'', which is be tasked with sharing that vehicle's trust data to other vehicles and RSUs as well as keeping the data updated when the vehicle is near it once again.
+To make this viable, the properties of daily commutes are used: it is assumed the vehicle would be near that RSU with reasonable frequency because it is located within the driver's home-to-work route.
 The main problem with this model is that it relies on the presence of frequent RSUs, which might not always be viable.
@@
 Furthermore, the authors are quick to dismiss the validity of entity-based trust, instead opting for a pure data-oriented approach.
-Since their model is based only on events, it also does not consider the usefulness of trust in more mundane cases such as the frequent sharing of location and speed among vehicles.
+%Since their model is based only on events, it also does not consider the usefulness of trust in more mundane cases such as the frequent sharing of location and speed among vehicles.
+Although it is true that data-oriented trust is efficient for events, which is what their model is based on, it is not ideal for sharing data quickly and frequently.
+When a collision or other major incident occurs, it is useful to judge each message on its own, since not all members of the network will have existing trust relationships with each other.
+However, when sharing location and velocity data several times per second, it is not reasonable to expect that each message will be analyzed so carefully; rather, it makes sense to form an opinion about the sender of the message and use the resulting trust value to choose which messages are relevant or not.
 

### 2017-06-01 qualification/5-proposal/chapter5.tex
final version of qualification text

--- 2017-03-30 ufpr/5-proposal/chapter5.tex
@@
 
-This work proposes that it is possible to adapt a malicious node identification scheme originally designed for static networks to a dynamic environment, such as a VANET.
+This work proposes that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
 In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
-The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
+%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
 Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
-
 By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
@@
 
-In this chapter, the details for the process above are presented.
+In this chapter, the details for the above process are presented.
 First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
-Then, the Working Day Movement Model is presented, along with the reasons why it is best fitted for this work.
+Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
 Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
@@
 \label{section:socialvanets}
-Several of the aforementioned articles state that vehicular networks are, by definition, ephemeral.
-That is, the likelihood of two nodes meeting each other twice is too low to be relevant.
-In this work, it is argued that such claims are not entirely true.
-It stands to reason that, throughout the course of several days, many drivers will take similar routes at similar times of day (e.g.
-to commute to work) and, therefore, their vehicles will be in similar locations each day.
+Some proposed trust models for vehicular networks, such as \cite{huang2014social}, state that the likelihood of two nodes meeting each other twice is too low to be relevant.
+However, it stands to reason that, throughout the course of several days, many drivers take similar routes at similar times of day (e.g.
+to commute to work) and, therefore, their vehicles are in similar locations each day.
 Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
 Since that is true for a notable percentage of a city's fleet, it can also be assumed that those vehicles may frequently encounter each other during their commute.
-While two vehicles that share a commute route may not be direct neighbors every day, they will likely be relatively close to each other, meaning few hops separate them in the ad-hoc network.
+While two vehicles that share a commute route may not be direct neighbors every day, they are likely to be relatively close to each other most days, meaning few hops separate them in the ad-hoc network.
 Furthermore, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
-Examples of these include vehicles whose owners are drivers or coworkers.
+Examples of these include vehicles whose owners are neighbors or coworkers.
 Such vehicles' trust relationship should become steady over time and, in the case of positive trust, they can use each other's information to learn more about other nodes in the network.
@@
 Trains travel on rails, so their contact with cars is less frequent, but it can also happen on railroad crossings; they travel long distances in relatively short amount of time, which helps the dissemination of data in a VANET.
-In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they will meet the same buses and/or trains frequently.
+In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they meet the same buses and/or trains frequently.
 
@@
 By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
+They note that, during rush hours, the diameter of the graph decreases to around 6 hops; additionally, the frequency of total encounters between pairs of nodes in the network increases during those hours.
+Although the authors do not quantify the encounters between specific pairs of nodes, these numbers support the idea that daily commutes do indeed cause vehicular networks to exhibit social network features.
 
@@
 
-While certification of identity is an issue in vehicular network, this work will not address it.
-There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}
-\cite{kumar2015intelligent}.
-One study proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
-Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time, so that nodes can identify each other over a long period, is a prerequisite for the model proposed here.
-
 \section{Working Day Movement Model}
@@
 
-To make use of the above properties, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
-Therefore, the Working Day Movement Model \cite{ekman2008working} (WDM) will be used.
-The model, developed for used in Delay-Tolerant Network (DTN) simulations, includes many of the features that will be necessary to simulate the daily movement of a vehicular network.
+Most VANET trust models consider the Random Waypoint mobility model, i.e.
+each node has an origin point, chooses a random location, gets to that location, then chooses another random location and goes there, and so forth.
+While this model is efficient for testing trust protocols, it doesn't truly represent vehicle mobility in the real world.
 
-As the name implies, the Working Day model abstracts people's movement from their homes to their offices and back.
+To make use of the properties described in \autoref{section:socialvanets}, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
+Therefore, the Working Day Movement Model \cite{ekman2008working} (WDM) is useful.
+The model, developed for use in Delay-Tolerant Network (DTN) simulations, includes many of the features that are necessary to simulate the daily movement of a vehicular network.
+
+As the name implies, the Working Day Model abstracts people's movement from their homes to their offices and back.
 Each node has a home and a workplace and they need to travel back and forth between those locations on a daily basis.
 Occasionally, nodes can also go to other locations for leisure.
-Most VANET trust models use the Random Waypoint mobility model, i.e.
-each node has an origin point, chooses a random location, gets to that location, then chooses another random location and goes there, and so forth.
-While this model is efficient for testing trust protocols, it doesn't truly represent vehicle mobility in the real world.
-As mentioned above, many drivers have routes they travel on daily, so the Working Day model is a more accurate representation, although it represents the mobility of humans instead of vehicle and requires some adjustments, which will be listed later in this section.
+As mentioned above, many drivers have routes they travel on daily, so the Working Day model is a more accurate representation, although it represents the mobility of humans instead of vehicle and requires some adjustments, which are listed later in this section.
 
@@
 The map can be divided into districts, which increases what the authors define as \textit{locality}.
-In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes will rarely leave the district.
+In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes rarely leave the district.
 Nodes which reside and work in different districts serve to connect the network with their commutes.
@@
 By thinking of these submodels for vehicles instead of people, it can become apparent how the frequency and length of encounters between nodes are similar in both instances.
-If two vehicles belong to family members or neighbors, they will likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
+If two vehicles belong to family members or neighbors, they likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
 Cars can also meet each other frequently if their drivers go out with friends.
@@
 \item
-The walking submodel needs to be disabled, since all nodes will be either cars or buses.
+The walking submodel needs to be disabled, since all nodes are either cars or buses.
 \item
@@
 
-The input graph $G = (V,E)$ is a static, connected, and directed graph containing all trust relationships in the network.
+The input graph $T = (V,O)$ is a static, connected, and directed graph containing all trust relationships in the network.
 Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
@@
 
-The process for identifying malicious nodes within $G$ is as follows:
+The process for identifying malicious nodes within $T$ is as follows:
 
-First, $G$ is separated into strongly components using Tarjan's algorithm \cite{tarjan1972depth}.
+First, $T$ is separated into strongly components using Tarjan's algorithm \cite{tarjan1972depth}.
 In each of these components, all nodes are connected by edges of value $1$.
 In other words, within a single component, all nodes trust one another completely; nodes connected by edges of value $0$ are separated into different components.
-Each of these components becomes a node of a component graph $T = (V', E')$.
+Each of these components becomes a node of a component graph $T' = (V', O')$.
 
-The creation of the graph $T$ simplifies the remaining computation.
-Since each node of $T$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $G$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
+The creation of the graph $T'$ simplifies the remaining computation.
+Since each node of $T'$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
 They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
-After the formation of $T$, one or two heuristics can be used to classify the nodes as benign or malicious.
+After the formation of $T'$, one or more heuristics can be used to classify the nodes as benign or malicious.
 
-The \textit{coloring heuristic} uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
+The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
 Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
-After running either algorithms with graph $T$, the color with the most nodes in $T$ is classified as correct, and all others are classified as malicious.
-Once this information in $T$ is brought back to graph $G$, it is trivial to label the nodes in $G$ as either benign or malicious based on their components' classifications.
-
-The other alternative is the \textit{articulation heuristic}.
-It uses a graph articulation algorithm, which does a depth-first search and classifies nodes based on their roles in the spanning tree.
-A node $n \in V$ is an articulation if its removal from $G$ makes $G$ a disconnected graph or, in other words, increases the number of connected components of $G$.
-Once articulation nodes are found, the heuristic classifies each node as correct if it does not have neighbors that are also articulations.
+After running either algorithms with graph $T'$, the color with the most nodes in $T'$ is classified as correct, and all others are classified as malicious.
+Once this information in $T'$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
 
 In the experiments shown in the article, the coloring heuristic shows promising results, identifying a high ratio of the malicious nodes in the network.
-The coloring and articulation heuristics can be combined, although the experiments using this combination display a high number of false positives.
-Therefore, for the purposes of this paper, only the coloring heuristic will be considered.
+Other heuristics are experimented with, but were less effective in detecting malicious nodes, or provided too many false positives.
+Therefore, for the purposes of this paper, only the coloring heuristic is considered.
 
-\section{Necessary changes to the malicious node identification scheme}
+Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
+In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
+Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
+
+The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
+By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
+
+%$G = (V,E)$: undirected topology graph, in which $V$ are the vertices and $E$ are the edges.
+%
+%$T = (V,O)$: directed trust graph, in which $O$ are the nodes' opinions about each other.
+%
+%$G_t,T_t$: snapshots of $G$ and $T$ in a timestamp $t$.
+%
+%$G_i,T_i$: local knowledge of $G$ and $T$ of a node $i$.
+%
+%$T' = (V',O')$: component graph of $T$.
+
+\section{Necessary changes to the trust management scheme}
 \label{section:changes}
@@
 The original work's scope does not include the generation of the trust graph, as it is a parameter to the algorithm.
-In dynamic graphs, trust relationships change over time, so how those changes affect our trust model have to be considered.
+In dynamic graphs, trust relationships change over time, so how those changes affect the proposed trust model have to be considered.
 This includes not only the full figure of the graph, but also each node's views on its neighbors.
@@
 Strongly connected components, or components in which all nodes have positive trust relationships with each other, are somewhat different in dynamic graphs.
-As nodes enter, leave, or change locations in the network, the components will change accordingly.
+As nodes enter, leave, or change locations in the network, the components change accordingly.
 Therefore, our algorithm will have to work using ``snapshots'' of the graph, meaning each measurement will be valid for a specific timestamp.
+Since the graph changes with time, for every timestamp $t$, there is a corresponding topology graph $G_t=(V_t, E_t)$, which includes the available vertices and edges in that moment in time, and a trust graph $T_t=(V_t, O_t)$, with the nodes' opinions at that moment.
+
 In a later step, it might be useful to consider how specific nodes may change over several snapshots.
-
-Since the graph changes with time, for every timestamp $t$, there is a corresponding graph $G_t=(V_t, E_t)$, which includes the available vertices and edges in that moment in time.
+Specifically, the suggestion of running the algorithm in iterations, by removing the malicious nodes in each step, may be applied with the peculiarities of a dynamic network.
 
@@
 Therefore, our algorithm needs to work using only a partial understanding of the network.
-For this to be viable, each node will likely have to gather information not only from its views on its neighbors, but also from its neighbor's views and so forth.
+For this to be viable, each node might have to gather information not only from its views on its neighbors, but also from its neighbors' views.
 
@@
 
-Thus, each node will have its own model of the network, and will generate a trust graph $G_i=(V_i, E_i)$, in which $i$ is the node's identifier.
-Combined with the necessity of timestamps explained above, this will result in a graph $G_{it}=(V_{it}, E_{it})$ for each node and each timestamp, which will be the input of the new algorithm.
-Accordingly, there are also component graphs $T_{it}=(V'_{it}, E'_{it})$, generated on each instance of the algorithm.
+Thus, each node has its own model of the network topology, $G_i$, and generates a trust graph $T_i=(V_i, O_i)$, in which $i$ is the node's identifier.
+Combined with the necessity of timestamps explained above, this results in a graph $T_{it}=(V_{it}, O_{it})$ for each node and each timestamp, which is the input of the new algorithm.
+Accordingly, there are also component graphs $T'_{it}=(V'_{it}, O'_{it})$, generated on each instance of the algorithm.
 
@@
 
-Since nodes will interact with each other occasionally and might learn information from other nodes, a range such as $[0,1]$ is useful to express how one trust value can change over time.
+Since nodes interact with each other occasionally and might learn information from other nodes, a range of $[0,1]$ is useful to express how one trust value can change over time.
 As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.
@@
 The datasets used for experimentation must be changed accordingly.
-Instead of using data gathered from social networks, which are static, dynamic network datasets will need to be used, such as traffic information for cities.
+Instead of using data gathered from social networks, which are static, dynamic network datasets must be used, such as traffic information for cities.
 
 At first, the objective is to test using the Working Day Movement Model, setting parameters that contribute to the validation of the algorithm.
-When the simulation begins, a set number of nodes will be malicious and, therefore, other nodes will have low trust in them after interacting.
+When the simulation begins, a set number of nodes are malicious and, therefore, other nodes will have low trust in them after interacting.
 In time, benign and malicious nodes can change their behavior and, accordingly, their trustworthiness will be updated.
@@
 	\item \textbf{Changes to the movement model}. Develop and adapt the Working Day Movement Model according to \autoref{section:workingday}.
-	\item \textbf{Study and selection of a simulator}.
+	\item \textbf{Study of the simulator}.
 Several simulators for vehicular networks exist, and one must be chosen in order to run simulations using the newly developed model.
 The Working Day Movement Model was tested with the ONE Simulator \cite{keranen2009one}.
-The source code for both the simulator and the WDM generation can be found online \cite{onerepo}, and will likely be used for the necessary simulations.
+The source code for both the simulator and the WDM generation can be found online \cite{onerepo}, and will be used for the necessary simulations.
 	\item \textbf{Simulations and adjustments}.
@@
 
-An approximate schedule is shown in \autoref{table:schedule}.
-The first column displays the task, using the numbers from the list above, while the remaining ones are marked when work on that task is expected for the given month.
-The months of April through December are in 2017 and January through March are in 2018.
-
 \begin{table}[h!]
 \centering
-\begin{tabular}{ |p{0.7cm}||p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}| }
+\begin{tabular}{ |p{0.7cm}||p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}| }
  \hline
- \textbf{Task} & Apr & May & Jun & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
+ \textbf{Task} & May & Jun & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
  \hline
  \hline
- \textbf{1} & \checkmark & \checkmark & \checkmark & & & & & & & & &\\
+ \textbf{1} & \checkmark & \checkmark & \checkmark & & & & & & & &\\
  \hline
- \textbf{2} & & & \checkmark & \checkmark & \checkmark & & & & & & &\\
+ \textbf{2} & & \checkmark & \checkmark & \checkmark & & & & & & &\\
  \hline
- \textbf{3} & \checkmark & \checkmark & \checkmark & \checkmark & & & & & & & &\\
+ \textbf{3} & \checkmark & \checkmark & \checkmark & \checkmark & & & & & & &\\
  \hline
- \textbf{4} & & & & & \checkmark & \checkmark & & & & & &\\
+ \textbf{4} & & & & \checkmark & \checkmark & \checkmark & \checkmark & & & &\\
  \hline
- \textbf{5} & & & & & \checkmark & \checkmark & & & & & &\\
+ \textbf{5} & & & & & & & \checkmark & \checkmark & \checkmark & &\\
  \hline
- \textbf{6} &  &  & & & & \checkmark & \checkmark & \checkmark & & & &\\
+ \textbf{6} &  &  & & & & & & \checkmark & \checkmark & &\\
  \hline
- \textbf{7} &  &  & & & & & & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
+ \textbf{7} &  &  & & & & & & \checkmark & \checkmark & \checkmark & \checkmark \\
  \hline
@@
 \end{table}
+
+An approximate schedule is shown in \autoref{table:schedule}.
+The first column displays the task, using the numbers from the list above, while the remaining ones are marked when work on that task is expected for the given month.
+The months of April through December are in 2017 and January through March are in 2018.
 

### 2017-06-01 google/proposal.tex
google award proposal

\documentclass[12pt]{article}

\usepackage{sbc-template}

\usepackage{graphicx,url}

\usepackage[english]{babel}   
%\usepackage[latin1]{inputenc}  
\usepackage[utf8]{inputenc}  
% UTF-8 encoding is recommended by ShareLaTex

\title{Fully Distributed Trust Management Scheme for Vehicular Networks}

\author{Renan Domingos M. Greca, Luiz Carlos P. Albini}

\address{Informatics Department -- Federal University of Paraná\\
  PoBox: 19081 -- 81.531-980 -- Curitiba -- PR -- Brazil 
   \email{ rdmgreca@inf.ufpr.br, albini@ufpr.br }\\ \\
   Google Contact: Lucas Radaelli
}

\begin{document}

%\maketitle

%\begin{abstract}
%  This meta-paper describes the style to be used in articles and short papers
%  for SBC conferences. For papers in English, you should add just an abstract
%  while for the papers in Portuguese, we also ask for an abstract in
%  Portuguese (``resumo''). In both cases, abstracts should not have more than
%  10 lines and must be in the first page of the paper.
%\end{abstract}
%     
%\begin{resumo} 
%  Este meta-artigo descreve o estilo a ser usado na confecção de artigos e
%  resumos de artigos para publicação nos anais das conferências organizadas
%  pela SBC. É solicitada a escrita de resumo e abstract apenas para os artigos
%  escritos em português. Artigos em inglês deverão apresentar apenas abstract.
%  Nos dois casos, o autor deve tomar cuidado para que o resumo (e o abstract)
%  não ultrapassem 10 linhas cada, sendo que ambos devem estar na primeira
%  página do artigo.
%\end{resumo}

%\section{Overview}
%
%\vspace{0.2cm}
%
%\textbf{Proposal Title:}
%Trust management for vehicular networks.\\
%\textbf{Professor:}
%Luiz Carlos Pessoa Albini \\ 
%\hspace*{1cm} Universidade Federal do Paraná, Departamento de Informática  \\ 
%\hspace*{1cm} Centro Politécnico, Jardim das Américas, PoBox: 19081,  Postal Code 81531-980 \\ 
%\hspace*{1cm} Curitiba, Paraná, Brazil \\
%\textbf{Google Contacts:} Lucas Radaelli
%
%\vspace{0.2cm}
\maketitle

\begin{abstract}
By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular communication network, in which cars share data, such as speed and position, amongst themselves in order to cooperate and make roads safer and more efficient. Vehicular communication networks are the first step to implement Smart Cities technological solutions for road safety.
However, as is the case with most new technologies, these networks might be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
In order to avoid such attacks, one important feature for vehicular networks is trust management. It allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
This work proposes a trust management model in the context of daily commutes, taking advantage of social features that can be found in vehicular networks.
\end{abstract}

%\subsection{Abstract}

\section{Goals and problem statement}
Considering the high speed and potentially life-threatening situations of vehicular networks, it is essential for messages to be delivered with low latency, which is why a fully distributed approach is preferred. Due to the highly dynamic scenario, it is also important that the message delivery does not depend on pre-defined infrastructures.
Therefore, ad-hoc networks might correctly fit this paradigm, creating a Vehicular Ad-Hoc Network (VANET). It does not rely on an Internet connection or existing infrastructure, meaning that all network services depend on the participating vehicles themselves.
However, it also means that any vehicle can join the network by being within the communication range.
While this is expected, as VANETs benefit from having more participants, it can be a problem when one or more members of the network turn out to be malicious.
Malicious nodes are ones that share falsified or otherwise incorrect information with other members, which can in turn cause traffic jams or even accidents.
One way of avoiding the dissemination of false data is to have nodes assign trust values to other nodes.
As they interact and observe each other's behaviors, nodes can decide whether or not data sent from others is reliable.

To create and maintain long-term trust relationships with one another, nodes in a vehicular network must have a reasonable chance of interacting somewhat frequently.
Although it cannot be expected that all pairs of nodes satisfy this requirement, there are social aspects observed in vehicular networks which show that, at least for some pairs of nodes, encounters can happen relatively frequently \cite{cunha2014possible}.
For example, vehicles belonging to family members, neighbors or workmates are very likely parked close to each other almost every day.
Furthermore, vehicles that perform daily commutes, as well as public transit vehicles like buses, are expected to travel along the same roads at around the same time every day.
%In order to simulate the daily mobility of a vehicular network, the Working Day Movement Model \cite{ekman2008working} has been chosen.
%The model is ideal because it replicates the daily mobility of a real city, including parking at home, at work, and groups of friends getting together.
%However, it uses people as nodes instead of vehicles, so some changes will have to be made.

Using these long term relationships, nodes can generate trust values and models of their surroundings. 
This information can be represented as a graph in which vertices are vehicles and edges are trust relationships between pairs of vehicles.
Information collected from previous encounters with known nodes can be used to estimate some of the edges in the model; meanwhile, further information can be extracted from neighboring nodes and ongoing encounters, filling out the remainder of the graph.

The goal of this work is to develop a trust management algorithm, which can be used to efficiently detect malicious nodes on the network. 
This work is an extension of \cite{vernize2015malicious}, which introduces the fastest and most precise malicious node identification mechanism based on trust management for centralized static social networks. This work will extend such trust management to dynamic and distributed vehicular networks.

%which takes as input a model of a node's surrounding network and  based on the network consensus (that is, the opinions of other nodes).
%It is based on a previously existing algorithm \cite{vernize2015malicious}, which identifies malicious nodes in networks by separating the model graph into components, in which all nodes within a component trust each other, and using a graph coloring function on the resulting component graph.
%The algorithm is fast and precise, but it was developed primarily for static networks which can have a centralized observer.
%This work will adapt the algorithm for it to work in the highly dynamic and decentralized environment of a vehicular network.

\section{Expected work and outcomes}
There are three main components of the proposed work.

First, \cite{vernize2015malicious} will be extended to serve a dynamic and decentralized network.
The main differences between the two approaches are:
(1) the graph used as input is incomplete, as nodes only have partial knowledge of the network;
(2) instead of a single graph, there will be multiple variations of it, for each node and for each moment in time;
and (3) trust will be measured in a range of $[0,1]$, instead of being binary.

Second, the evaluation will be made through simulations using the Working Day Movement Model \cite{ekman2008working}.
The model was developed for mobile devices (such as smartphones) being carried by humans, so it must be adapted to suit a vehicular network.
%These are reasonable changes, such as removing the pedestrian movement model, and having buses serve as a single node instead of a collection of them.
Finally, results will be analyzed to confirm the algorithm's efficiency and effectiveness.
%A certain amount, which may or not vary over time, of the nodes in the simulations will be malicious and will send incorrect data.
%At a certain frequency, nodes will run the algorithm in order to identify nearby malicious nodes.
%If successful, the algorithm will locate as many malicious nodes (and as few false positives) as possible.

\section{Previous work}

Several trust management models for vehicular networks have been proposed, with different degrees of success.
Some of them emphasize \textit{entity-based trust}, in which trust values are stored for each node and that value is used to judge messages originating from it.
This has the advantage of shortening evaluation time for each message delivered, since most trust information is previously obtained.
Others use \textit{data-oriented trust}, in which messages are individually evaluated regardless of the sender.
These solutions are useful when nodes are mostly strangers to one another and must quickly decide whether or not to trust an incoming alert sent by an unknown source.

VARS \cite{dotzer2005vars} is an early example of a VANET trust model, using a process the authors call opinion piggybacking.
That is, when an event occurs and a node broadcasts an alert message about it, other nodes append their opinion on both the message contents and the message sender onto the alert.
Therefore, nodes receive a combined opinion of several of its neighbors, helping it to form its own decision.
However, this process has privacy implications in case the message is not of public service, and the opinion piggybacking itself might add unnecessary latency and overhead to the broadcast.

The model in \cite{chen2010trust} proposes the formation of clusters of nodes, in which one leader aggregates the other nodes' opinions about each message sent.
It is exceedingly costly to maintain clusters in a highly dynamic network, and might be unfeasible if the network is too sparse to generate relevant clusters.

\cite{park2011long} uses daily commutes and road-side units (stationary equipment which may act as part of a VANET).
Each vehicle is assigned an ``Agent RSU'', which is responsible for maintaining the vehicle's trust value and sharing it with others.
The trust value maintenance is comparable to the one proposed here, although it relies heavily on existing infrastructure, which is not always reasonable.

In \cite{huang2014social}, authors identify features generally associated with social networks and use them to generate a VANET trust model.
The algorithm places heavy emphasis on the opinion of the nodes in which messages originate, which may be a problem if those nodes are malicious. Nevertheless, there is no trust management scheme for VANETs which take advantage of the users`  long term relationship as the one proposed here.

\section{Data Policy}
Once the research is concluded, all related code and datasets will be available online in an open-source fashion.
Along with the dissertation required for the Master of Science degree, articles will be published explaining the core methodology, algorithms and experiments performed.

\bibliographystyle{sbc}
\bibliography{sbc-template}

\end{document}

### 2017-12-01 dissertation/2-complexnetworks/chapter2.tex
dissertation v 2017-11-30

\chapter{Complex Networks}
\label{chap:complexnetworks}

%Often, the term ``network'' is associated only with computer

%Often, when people hear the term ``network'', they think only of technological examples, such as computer networks.
%Such networks are only a subset of a much larger field of study: that of complex networks.
%The study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}.
%Technological networks, and VANETs by extension, are only a subset of that much larger field of study.
%
%Trust issues and their solutions are an important aspect of the study of complex networks, and of social or technological networks in particular.
%Each different type of network requires a specific approach to trust, according to the features that distinguish it from others.
%It is interesting to review the solutions proposed for social networks and technological networks in general, and then narrow down on how dynamic networks, and VANETs in particular, are handled.
%Despite the differences, there are some concepts that overlap across most types of networks.

Complex networks can describe many systems which are observed in nature and society through a collection of \textit{vertices} (or \textit{nodes}) and \textit{edges} \cite{newmannetworks}.
They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).
The study of networks is tied to graph theory, since graphs are a useful way to generate models of networks, and therefore many concepts of the two fields overlap.
Mathematical, computational and statistical concepts developed for graphs can be translated to be useful for a variety of different types of networks, like, for example, finding articulation vertices in a graph to locate bandwidth bottlenecks in a computer network.

Complex networks are generally divided into four categories \cite{newmannetworks}:

\begin{enumerate}
	\item \textbf{Technological Networks} are grids purposefully engineered to provide services to consumers and/or citizens.
	 	The primary examples of these networks are the Internet, the telephone network, power grids, transportation and delivery networks.
	 	A commonly studied type of technological network are Mobile Ad-hoc Networks (MANETs).
		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
		VANETs, which are special types of MANETs, are introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
		These relationships can be familiar, friendships, acquaintance, etc.
		For the purposes of this work, the most relevant type of relationship is that of trust.
		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
		Often, information networks are directly associated with technological or social networks.
		For instance, while the World Wide Web is an information network (in which the nodes are webpages and the edges are the links that users click on to navigate), it relies on the Internet, as it contains the physical infrastructure that makes the web possible.
		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
		Trust in information networks can be observed in some instances, like peer-to-peer networks [cite], although its usages are not relevant for this work.
	\item \textbf{Biological Networks} are the networks found in nature.
		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
		An example is the brain, which contains a neural network formed by neurons, cells which enable information processing; connections in the network represent signals that are sent from one neuron to another.
		Another instance of biological networks are food chains, categorized as ecological webs.
		Species of animals are the nodes, while the predation of other species form the edges.
		In biological networks, it is difficult to clearly define trust, since nodes may not have any sort of awareness or intelligence (such as cells or proteins).
		Regardless, the study of trust in biological networks is not relevant for this proposal.

\end{enumerate}

%While issues of trust have been studied on most types of networks, this work will emphasize on social and technological networks, because previous study on both can provide useful insights towards trust in vehicular networks.
%It is clear that VANETs are part of technological networks; in this work, however, it is also argued that vehicular networks present features usually attributed to social networks.
%Therefore, further details, as well as trust issues and proposed solutions, for both social and technological networks are presented in this chapter.

In most networks, trust can be a useful tool to aid the security and safety of its members.
Therefore, the study of the concept and applications of trust is an important part of the study of networks.

Trust is a concept studied in fields such as psychology and economics, with specific definitions.
In complex networks, under the perspective of computer science, it is a measure of one entity's confidence that another will behave properly and provide valid and/or meaningful data \cite{sherchan2013survey}.
What follows is the basis of how a network can be modelled using a graph and how a trust model can be applied to it.

Consider an undirected graph $G = (V,E)$, which models one complex network of any kind.
The vertices are the members of the network (computers, humans, etc.) and each edge represents a pair of vertices' ability to exchange data freely.
This graph represents the network's \textit{topology}, that is, the basic structure of the network.
Then, there is also a directed graph $T = (V, O)$, called a \textit{trust graph}.
$T$ contains the same nodes as $G$, although its edges represent the degree of trust (or opinion) each node has towards another.

There are two main ways to describe the edges in $O$: they can be binary, either existing when there is trust or not otherwise, or they can hold a specific trust value within a certain range.
This means the shape of $T$ is not necessarily similar to that of $G$.
For instance, two people can have contact with each other but not maintain a trust relationship, thus altering the layout of the trust graph compared to the network topology.

In the following two sections of this chapter, trust in the contexts of social and technological networks is further explained.
Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which is expanded upon in \autoref{chap:vanets}.

%As this work relates to trust issues in complex networks, it is interesting to further examine social networks and the importance of trust within them.
%While the end result of the study will be applied to technological networks, certain aspects of social networks provide useful analogies to trust management in other kinds of networks.

\section{Trust in Social Networks}
\label{section:trustsocial}
There are two types of social networks: real-world ones formed by relationships between people, and online ones that attempt to abstract the former into a digital environment.
Examples of the first one are all around, present in any family, workplace, school or group of friends \cite{newmannetworks}.
Online social networks started by connecting people who already knew each other and giving them an additional form of interaction (analogous to what telephones and email did before), but, today, it is not unusual for people to form relationships with others whom they have only met online.

In a traditional social network, it is simple to perceive how trust is relevant and how it works, since trust relationships between people are used on a daily basis to make decisions.
When adapted to a digital environment, these social relationships can be used to automatically increase the relevance of certain information.
For instance, upon reading an online review for a certain product, a user will be more likely to accept the review's conclusion if it was written by a close friend than if it were written by a stranger.
Social trust is a way of estimating how much a certain recommendation will lead to a positive outcome \cite{golbeck2006inferring}.

The absence of trust or the presence of distrust have consequences as well.
Both in the real world and online, information which comes from a stranger is received with uncertainty; there is no reason to trust the sender, so the data itself must be analyzed and compared to other sources in order to judge whether or not it may be trusted.
When one person actively distrusts another (that is, the person believes the other is malicious or uninformed), receiving data from the untrustworthy source will be actively avoided.
In online social networks, for example, one user can ``block'' another in order to avoid seeing anything from the other.

%Furthermore, if the review was written by a person who the user knows to be malicious or uninformed, its contents will be even less relevant.

Social networks also have the property of carrying trust from one relationship to another:
information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust \cite{newmannetworks}.
Expanding on that property, there is the concept friends of friends \cite{boissevain1974friends}.
If, for example, nodes $A$ and $B$ have mutual trust and are considered friends, then it is reasonable to assume that some of $A$'s trust for $B$ carries over to other nodes that enjoy mutual trust with $B$.
In other words, a friend of a friend can be considered more trustworthy than the average stranger.
This property is similar, although not identical, to transitivity, since trust is diminished for each extra step an origin node needs to reach a destination, and there is also the possibility that one node distrusts another even if they share a mutual friend.
Naturally, social trust is not commutative ($A$ trusting $B$ does not imply that $B$ trusts $A$).

In general, social networks' topology and trust graphs are mostly static.
Although friendships are formed and ended frequently (i.e. the topology is dynamic), those connections do not disrupt the general shape of the network, because members of the network will usually have other friends whose relationships remain stable.
Even if a certain person's trust integrity is compromised due to a specific incident, that person's friends are not necessarily deemed untrustworthy, preserving part of the trust graph.
While positive trust is often tied to the social topology, it is not always the case: one example is two work colleagues who may have a professional relationship, but wouldn't trust each other on other matters; another is the trust people place in authority figures without necessarily having met. 
\autoref{fig:social_network} shows an example of a small social network containing a family and an office.

In \autoref{section:socialvanets}, the argument is made that VANETs can be considered social networks in several occasions and how this can be used to develop trust in vehicular networks.

%\begin{figure}[h]
%    \centering
%    \includegraphics[width=1.0\textwidth]{images/socialnetwork.png}
%    \caption{Example of a topology graph and a trust graph in a social network.}
%    \label{fig:social_network}
%\end{figure}

\begin{figure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/socialnetwork_topology.png}
\caption{Topology graph} \label{fig:2_1a}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/socialnetwork_trust.png}
\caption{Trust graph} \label{fig:2_1b}
\end{subfigure}
\caption{Example of a topology graph and a trust graph in a social network.} \label{fig:social_network}
\end{figure}

%\begin{figure}[htbp]
%\subfloat[Topology graph]{\includegraphics[0.5\textwidth]{images/socialnetwork.png}}
%\subfloat[Trust graph]{\includegraphics[0.5\textwidth]{images/socialnetwork.png}}
%\caption{Example of a topology graph and a trust graph in a social network.}
%\end{figure}

%Many works exist regarding trust relationships in complex networks, and in social networks in particular.
%Here, only a few examples will be described, in order to show the general idea of trust research for these types of networks.

\section{Trust in Technological Networks}
\label{section:trusttechnological}
In conventional technological networks, such as the Internet, trust is defined and applied quite differently from social networks since, generally, it is very centralized through services that offer security to users.
Examples of this can be an IP filtering scheme to avoid distributed denial of service (DDoS) attacks or web browser extensions that block requests to domains in a blacklist; a central agent, be it a hosting provider or the extension's publisher, must maintain and update a list of untrustworthy IP addresses or domains.
This means that, in the context of the Internet, trust is often derived from a secondary source: end users and their computers can't be expected to maintain their own blacklists, so they rely on external parties which may provide these lists along with other security services.
Similarly, when a user visits an e-commerce website, they must have some degree of trust on the website or the vendor; in this case as well, third-party services are used to certify the legitimacy of the transaction, based on feedback from other customers.

% In ad-hoc networks, such as vehicular networks, which are detailed in \autoref{chap:vanets}, trust can be handled in very different ways.
While the centralized trust solutions above serve their purpose on the security and privacy of Internet users, they would be too slow to be viable in a dynamic ad-hoc network, which cannot rely on a back-end infrastructure to distribute those lists.
The most common instance of mobile ad-hoc networks, or MANETs, is using mobile devices, such as smartphones, being carried by humans.
Although these networks are dynamic, their mobility is relatively low in relation to the wireless range of the devices — if two people are walking in opposite directions, their phones may communicate for several seconds before they leave each other's range.
Trust solutions for MANETs can use this property to their advantage, since it allows one node to test another and check several of the messages sent between them.

% As is the case with social networks, the topology and trust graphs of MANETs are distinct, but

Ad-hoc networks require a decentralized approach to trust management; each member of a network has its own opinions about other members, and these opinions can change over time.
For these opinions to be generated and updated, two nodes must have had previous contact with each other, or derive trust from a third, intermediary, node.
Hence, there is the correlation between the trust graph and the topology graph of the network.
Since MANETs are dynamic, the graph that represents one network's topology is frequently changing and, with that, the opportunities to create and update trust relationships also changes.
%For two nodes to have a trust relationship at all, the topology must changed in a way so that they are neighbors to each other, or close enough that asking for a third node's
In networks in which nodes can meet more than once, it might be valuable to store information from previous encounters to use in the future, although this process can be too slow or resource-consuming to be viable in certain devices; by doing this, the trust graph maintains edges between nodes that are no longer connected in the topology.

% Ad-hoc networks, such as MANETs and VANETs, are the ob

%For this work, it is necessary to examine closely the case of mobile ad-hoc networks, since it is in that category that VANETs

%Depending on how the network is structured and its use-cases, it may or not be useful to store these trust values for later use.

%With social networks, the graph that represents relationships

%With MANETs, however, the graph that represents trust relationships amongst network members is

%However, while social networks are mostly static, the topology of a MANET changes frequently as nodes move around.
%[Perhaps this is a good place to mention trust vs. trust management]
%Nodes must be able to handle trust with its immediate neighbors for any given moment, so it can properly filter out potentially malicious data.
%In networks in which nodes can meet more than once, it may be valuable to store information from previous encounters to use in the future, although this process can be too slow or resource-consuming to be viable in certain devices.

Another important aspect of trust in ad-hoc networks is that information is, generally, uncertain and incomplete \cite{baras2005cooperation}.
That is, since nodes form their own model and opinions of the surrounding network, it is unlikely that this data will be certain and accurate with reality.
For this reason, it is also possible to use data gathered by neighboring nodes to complement the model.
This data itself is subject to the trust evaluation of the neighbors, but it is crucial to better approximate the trust values of other nodes.
Incompleteness is an inherit trace of MANETs, since it is entirely possible for nodes to be too distant to communicate, and only occasionally come into contact.

Finally, MANETs must consider the processing and battery limitations of the devices that integrate it.
Nodes may disable wireless communications to save power and therefore become uncooperative, or it may be too slow to be a reliable source of information.

There are few examples of MANETs implemented for consumer devices.
Two examples are networks created to quickly share data between devices using Wi-Fi or Bluetooth \cite{krochmal2014discovery}, or ones that allow for multiplayer gaming sessions amongst multiple nearby devices \cite{sasaki2011wireless}.
In both cases, there is no need for a complicated system-level trust model, since those activities involve active participation from the users wielding the device (that is, the user chooses whether or not to communicate with other devices); rather, the trust relationship occurs socially amongst the users themselves.

Naturally, VANETs are an instance of MANETs and therefore share some of the same features.
However, the topology of vehicular networks is very different from standard ad-hoc networks, and possible trust solutions are accordingly also distinct.

\begin{figure}
\centering

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_1.png}
\caption{Initial topology and trust graph} \label{fig:2_2a}
\end{subfigure}

\vspace*{\fill} % separation between the subfigures

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_topology_2.png}
\caption{Final topology graph} \label{fig:2_2b}
\end{subfigure}

\vspace*{\fill} % separation between the subfigures

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\linewidth]{images/manet_trust.png}
\caption{Final trust graph} \label{fig:2_3b}
\end{subfigure}

\caption{Example of the changes node mobility causes to the topology and trust graphs.} \label{fig:manet}
\end{figure}

%=====================================================

### 2017-12-01 dissertation/4-project/chapter4.tex
dissertation v 2017-11-30

\chapter{Project}
\label{chap:project}

This work proposes that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as input for a malicious node detection algorithm.

%The previous chapters show why vehicular networks require distinguished trust models compared to social networks and traditional MANETs.
%In this chapter, it is proposed that certain aspects from social networks can be found in vehicular networks and, through those aspects, it is possible to adapt a trust model designed for traditional complex networks to a vehicular environment.
%These social properties allow for long-term trust relationships to be established amongst nodes of a network and, therefore, previously existing trust models can be adapted to the vehicular environment.

In this chapter, the details for the above process are presented.
First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
Finally, a schedule for the next steps of the project is shown.

\section{Goals}
\label{section:goals}

\section{Social Networks and VANETs}
\label{section:socialvanets}
Some proposed trust models for vehicular networks, such as \cite{huang2014social}, state that the likelihood of two nodes meeting each other twice is too low to be relevant.
However, it stands to reason that, throughout the course of several days, many drivers take similar routes at similar times of day (e.g. to commute to work) and, therefore, their vehicles are in similar locations each day.
Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
Since that is true for a notable percentage of a city's fleet, it can also be assumed that those vehicles may frequently encounter each other during their commute.
While two vehicles that share a commute route may not be direct neighbors every day, they are likely to be relatively close to each other most days, meaning few hops separate them in the ad-hoc network.
Furthermore, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
Examples of these include vehicles whose owners are neighbors or coworkers.
Such vehicles' trust relationship should become steady over time and, in the case of positive trust, they can use each other's information to learn more about other nodes in the network.

%\textbf{[WORKING DAY MOBILITY MODEL]}
%They do not, however, present data regarding specific pairs of nodes.
%[Should we propose to show this information in our work?]

%Regardless of these properties of commuting vehicles, the probability of two specific vehicles meeting each other more than once, in given time, tends to 1.
%So it is also reasonable to assume that two vehicles, whose drivers reside in the same city, over a long period of time (e.g. one year) may meet each other more than once.
%However, if they only meet rarely, their trust opinions about each other may not be very useful.

Most cities also have one or more types of mass transit systems (buses or trains).
Those vehicles can also be part of a VANET and communicate with private cars.
Buses share the same roads as cars, but instead of having specific destinations, they travel a predefined route during the whole day, usually tied to a tight schedule.
Trains travel on rails, so their contact with cars is less frequent, but it can also happen on railroad crossings; they travel long distances in relatively short amount of time, which helps the dissemination of data in a VANET.
In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they meet the same buses and/or trains frequently.

In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
They note that, during rush hours, the diameter of the graph decreases to around 6 hops; additionally, the frequency of total encounters between pairs of nodes in the network increases during those hours.
Although the authors do not quantify the encounters between specific pairs of nodes, these numbers support the idea that daily commutes do indeed cause vehicular networks to exhibit social network features.

%Another way of applying trust models from social networks to VANETs would be to take into account the relationships between drivers.
%If a neighboring vehicle happens to be driven by a close friend or relative, there is reason to believe that it is not a malicious node in the network.
%That might not always be the case, since there may be malicious hackers controlling the node and its driver is oblivious to it.
%In any case, this approach is not covered in this work.

\section{Malicious Node Identification Algorithm}

The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.

The input graph $T = (V,O)$ is a static, connected, and directed graph containing all trust relationships in the network.
Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.

The process for identifying malicious nodes within $T$ is as follows:

First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
In each of these components, all nodes are connected by edges of value $1$.
In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
Each of these components becomes a node of a component graph $C = (V', O')$.

The creation of the graph $C$ simplifies the remaining computation.
Since each node of $C$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
After the formation of $C$, one or more heuristics can be used to classify the nodes as benign or malicious.

The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
After running either algorithms with graph $C$, the color whose nodes in $C$ represent the most nodes in $T$ is classified as correct, and all others are classified as malicious.
Once this information in $C$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.

In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
Therefore, for the purposes of this research, only the coloring heuristic is considered.

Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.

The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.

%\begin{algorithm}
%\caption{MaNI algorithm}\label{algorithm:mani}
%\begin{algorithmic}[1]
%
%\Function{MaNI}{graph $T$}
%
%\State $C \gets$ \textbf{Tarjan}$(T[0])$
%\State $colors \gets$ \textbf{Coloring}$(C)$
%\State \textbf{color} nodes of $T$ \textbf{according to} $colors$
%\State \textbf{count} 
%
%\EndFunction
%\end{algorithmic}
%\end{algorithm}

\begin{figure}
\centering

\begin{subfigure}{0.25\textwidth}
\includegraphics[width=\linewidth]{images/mani/0-trust.png}
\caption{Initial graph $T$.} \label{fig:mani0}
\end{subfigure}
\hspace*{2cm} % separation between the subfigures
\begin{subfigure}{0.25\textwidth}
\includegraphics[width=\linewidth]{images/mani/1-tarjan.png}
\caption{Tarjan's algorithm is used to identify the strongly connected components.} \label{fig:mani1}
\end{subfigure}

\vspace{1cm} % separation between the subfigures

\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/mani/2-components.png}
\caption{The graph $C$ is formed from the components of $T$.} \label{fig:mani2}
\end{subfigure}
\hspace*{2cm} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/mani/3-colored-components.png}
\caption{The coloring algorithm is used to label the nodes of $C$.} \label{fig:mani3}
\end{subfigure}

\vspace{1cm} % separation between the subfigures

\begin{subfigure}{0.25\textwidth}
\includegraphics[width=\linewidth]{images/mani/4-result.png}
\caption{The colors from $C$ are used to label nodes in $T$ as correct or malicious.} \label{fig:mani4}
\end{subfigure}

\caption{Example of an execution of the MaNI algorithm.}\label{fig:mani}
\end{figure}

\autoref{fig:mani} illustrates the execution of the MaNI algorithm.
With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
\autoref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
The strongly connected components form a graph $C$ according to edges present in $T$.
In \autoref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \autoref{fig:mani3}.
Finally, each node in $T$ is colored according to which color its component received in $C$.
The color with the most nodes is deemed benign, while the others are considered malicious.

%$G = (V,E)$: undirected topology graph, in which $V$ are the vertices and $E$ are the edges.
%
%$T = (V,O)$: directed trust graph, in which $O$ are the nodes' opinions about each other.
%
%$G_t,T_t$: snapshots of $G$ and $T$ in a timestamp $t$.
%
%$G_i,T_i$: local knowledge of $G$ and $T$ of a node $i$.
%
%$C = (V',O')$: component graph of $T$.

\section{Tarjan's strongly connected components algorithm}
\label{section:tarjan}
An important aspect of \cite{vernize2015malicious} is the use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth}.
This allows a large graph to be abstracted into a smaller graph, which therefore reduces the input for further algorithms.
Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any $u, v \in O$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
Every node of the input graph $T$ must belong to a component.

The algorithm works by performing a depth-first search, adding nodes to $stack$ as they are visited.
If two nodes are present on $stack$, then there is a path from the first node to the second one (in the order they were added to the stack).
Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.

In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessed from every call of the function.
$index$ and $lowlink$ are arrays indexed by node IDs, $count$ is an integer and $stack$ is a last-in-first-out data structure.

In the call that visits a node $u$, the algorithm must loop through each node $v$ for which $u \rightarrow v$ exists.
If node $v$ has not yet been visited, the algorithm is called for  $v$.
The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
Therefore, nodes must be popped from the $stack$ until $u$ is found.
Each node popped, including $u$, is a member of a strongly connected component.

The number of components is, at most, $|v|$.
In a worst-case scenario, each node is put into its own component; however, this would not be the case for most useful input graphs.

Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth} \cite{vernize2013dissertation}.
The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.

\begin{algorithm}
\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
\begin{algorithmic}[1]

\Function{Tarjan}{vertex $u$}

\State $index[u] = count$
\State $lowlink[u] = count$
\State $count \gets count + 1$
\State \textbf{push} $u$ \textbf{to} $stack$

\For{$v$ \textbf{in} neighbors of $u$ }

	\If{$index[v] = -1$}
		\Comment{\emph{v has not been visited yet}}
		\State \textbf{Tarjan}($v$)
	\EndIf

	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
\EndFor

\If{$lowlink[u] = index[u]$}
	\Repeat
		\Comment{\emph{unstack nodes until u is found}}
		\State \textbf{pop} $w$ \textbf{from} $stack$
		\State \textbf{add} $w$ \textbf{to} $component$
	\Until {$w = u$}
\EndIf

\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{figure}
\centering

\begin{subfigure}{0.30\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/0.png}
\caption{Initial graph.} \label{fig:tarjan0}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/1.png}
\caption{The search follows the path $0 \rightarrow 1 \rightarrow 2 \rightarrow 3 
\rightarrow 4 \rightarrow 2$.} \label{fig:tarjan1}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/2.png}
\caption{Since $lowlink[2] = 2$, nodes 2, 3 and 4 are added into a component.} \label{fig:tarjan2}
\end{subfigure}

\vspace{1cm}

\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/3.png}
\caption{The search follows the path $1 \rightarrow 4$, but node $4$ has already been visited.} \label{fig:tarjan3}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/4.png}
\caption{The search follows the path $1 \rightarrow 5 \rightarrow 0$.} \label{fig:tarjan6}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/5.png}
\caption{The search follows the path $5 \rightarrow 6$.} \label{fig:tarjan5}
\end{subfigure}

\vspace{1cm}

\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/6.png}
\caption{Node $6$ does not have outgoing edges and is added to a component.} \label{fig:tarjan6}
\end{subfigure}
\hspace*{2cm} % separation between the subfigures
\begin{subfigure}{0.3\textwidth}
\includegraphics[width=\linewidth]{images/tarjan/7.png}
\caption{The algorithm returns to 0 and, since $lowlink[0] = 0$, adds 0, 1 and 5 to a component.} \label{fig:tarjan7}
\end{subfigure}

\caption{Example of an execution of Tarjan's strongly connected components algorithm.}\label{fig:tarjan}
\end{figure}

\autoref{fig:tarjan} illustrates the execution of Tarjan's algorithm.
The algorithm starts from node 0, with $index[0] = 0$ and $lowlink[0] = 0$.
With a depth-first search, the algorithm traces the path $0 \rightarrow 1 \rightarrow 2 \rightarrow 3 
\rightarrow 4 \rightarrow 2$.
Since node $2$ has already been visited and nodes 4 and 3 have no further outgoing edges, $lowlink[4]$ and $lowlink[3]$ receive the value 2 and the function calls return back to the first visit of node $2$.
At this point, nodes 2, 3 and 4 all have 2 as the smallest reachable index and, therefore, they form a strongly connected component.

Continuing from node 1, the algorithm traces $1 \rightarrow 5 \rightarrow 0$, but stops there since node 0 has already been visited.
Continuing from node 5, the algorithm traces $5 \rightarrow 7$.
Node 7 has no outgoing edges, so it forms a strongly connected component by itself.
Once the function calls return to node 0, a strongly connected component is formed with nodes 0, 1, and 5, since they all have 0 as their $lowlink$ value.

\section{Graph coloring with minimum colors}
\label{section:coloring}
The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
Graph coloring is one of the heuristics used in \cite{vernize2013dissertation} to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
Since it was the heuristic that presented the best results, it has been chosen as the heuristic for this current study as well.

The process of graph coloring consists of giving each node a label (represented by a color) so that no two neighboring nodes share the same color.
This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.

In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
The complexity of the algorithm is $O(|E|)$ for an undirected graph $G = (V,E)$.
As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.

Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph} \cite{vernize2013dissertation}.

\begin{algorithm}
\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
\begin{algorithmic}[1]

\Function{Coloring}{graph $G$}

\State \textbf{color} all nodes of $G$ \textbf{with} 0
\State $d \rightarrow 0$
\For{$e=(u,v)$ \textbf{in} edges of $G$}
	\If{$u$ and $v$ have the same color}
		\If{$color[v] = d$}
			\State $d \rightarrow d+1$
		\EndIf
		\State $color[v] \rightarrow d$
	\EndIf
\EndFor

\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{figure}
\centering

\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/0.png}
\caption{Initial graph with all nodes labeled 1. $d=1$ (green).\\} \label{fig:coloring0}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/1.png}
\caption{Edge $(0,1)$ is checked and node $1$ gets a new color. $d=2$ (magenta).} \label{fig:coloring1}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/2.png}
\caption{Edge $(0,5)$ is checked and node $5$ gets the current value of $d$.} \label{fig:coloring2}
\end{subfigure}
\hspace*{\fill} % separation between the subfigures
\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/3.png}
\caption{Edge $(1,5)$ is checked and node $5$ gets a new color. $d=3$ (cyan).} \label{fig:coloring3}
\end{subfigure}

\vspace{2cm}

\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/4.png}
\caption{Edge $(2,3)$ is checked and node $3$ gets the current value of $d$.} \label{fig:coloring5}
\end{subfigure}
\hspace*{2cm} % separation between the subfigures
\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/5.png}
\caption{Edge $(2,4)$ is checked and node $4$ gets the current value of $d$.} \label{fig:coloring6}
\end{subfigure}
\hspace*{2cm} % separation between the subfigures
\begin{subfigure}{0.20\textwidth}
\includegraphics[width=\linewidth]{images/coloring/6.png}
\caption{Edge $(3,4)$ is checked and node $4$ gets a new color. $d=4$ (yellow).} \label{fig:coloring6}
\end{subfigure}

\caption{Example of an execution of the graph coloring with minimum colors algorithm.}\label{fig:coloring}
\end{figure}

A limitation of this algorithm is that the edges must be sorted according to node indexes.
It doesn't matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
This is demonstrated in \cite{vernize2013dissertation}.

Figure \autoref{fig:coloring} illustrates the execution of the graph coloring algorithm.
It is notable how few iterations the algorithm takes to fully color the graph.
However, the fact that the result does not use the minimum amount of colors.
By coloring node 4 as cyan and node 3 as magenta, the sample graph could have been colored with only three colors.
However, as described above, this is not a substantial problem for the purposes of this study.

\section{SNAP library}
\label{section:snap}
In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
The Stanford Network Analysis Platform (SNAP) library \cite{snap} was chosen primarily because it is memory efficient.
The simulations require multiple graphs that share the same set of nodes (because each node in the network has its own knowledge of the surrounding network), and the SNAP library uses pointers to nodes and edges, saving memory by not having to duplicate the entire data structure.
It is written in C++, but, for these simulations, the Snap.py Python library was used.

The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.

\section{Trust management algorithm}
\label{section:algorithm}

The MaNI algorithm works well and is efficient for static graphs.
However, some important modifications had to be made to accommodate dynamic graphs.
The main changes made are described below.

\subsection{Graph knowledge and graph building}
First of all, there is an important change to the graph knowledge used to run the algorithm.
The complete trust graph is a parameter to the MaNI algorithm, which executes with complete knowledge of the graph.

Given the complete topology graph $G = (V, E)$, each node $u \in V$ has its own vision of the network and $u$ has to build a graph, $G_u$, to represent its surrounding network.
In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
For every neighbor $v \in V$, $G_u$ must me merged with $G_v$.
In sufficient iterations, $G_u$ forms a complete or nearly complete vision of the network.

Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.

%The original work's scope does not include the generation of the trust graph, as it is a parameter to the algorithm.
%In dynamic graphs, trust relationships change over time, so how those changes affect the proposed trust model have to be considered.
%This includes not only the full figure of the graph, but also each node's views on its neighbors.
%
%\subsection{Strongly connected components}
%Strongly connected components, or components in which all nodes have positive trust relationships with each other, are somewhat different in dynamic graphs.
%As nodes enter, leave, or change locations in the network, the components change accordingly.
%Therefore, our algorithm will have to work using ``snapshots'' of the graph, meaning each measurement will be valid for a specific timestamp.
%Since the graph changes with time, for every timestamp $t$, there is a corresponding topology graph $G_t=(V_t, E_t)$, which includes the available vertices and edges in that moment in time, and a trust graph $T_t=(V_t, O_t)$, with the nodes' opinions at that moment.
%
%In a later step, it might be useful to consider how specific nodes may change over several snapshots.
%Specifically, the suggestion of running the algorithm in iterations, by removing the malicious nodes in each step, may be applied with the peculiarities of a dynamic network.
%
%\subsection{Graph knowledge}
%In the original paper, the malicious nodes are identified by a central entity with access to all nodes' trust values.
%Not only does our work plan to decentralize this procedure, having complete knowledge of the graph is unfeasible in large-scale dynamic networks such as VANETs.
%Therefore, our algorithm needs to work using only a partial understanding of the network.
%For this to be viable, each node might have to gather information not only from its views on its neighbors, but also from its neighbors' views.
%
%The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
%Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
%That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
%They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
%The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).
%
%Thus, each node has its own model of the network topology, $G_i$, and generates a trust graph $T_i=(V_i, O_i)$, in which $i$ is the node's identifier.
%Combined with the necessity of timestamps explained above, this results in a graph $T_{it}=(V_{it}, O_{it})$ for each node and each timestamp, which is the input of the new algorithm.
%Accordingly, there are also component graphs $C_{it}=(V'_{it}, O'_{it})$, generated on each instance of the algorithm.
%
%During the early stages of development, it will be considered that vertices cannot enter or leave the network and, therefore, only the layout of the edges changes.
%This is sufficient to experiment with the algorithm's performance with the added complexity of a dynamic network.
%
%
%\subsection{Binary trust}
%A model using binary trust is one in which trust values can only be 0 or 1.
%If node $A$ positively trusts node $B$, the value of the edge $A\rightarrow B$ is 1.
%If it does not trust $B$, or is unsure, the value is 0.
%Such a scheme makes sense in a static graph but, in a dynamic one, it is interesting to have a range of possible trust values.
%
%Since nodes interact with each other occasionally and might learn information from other nodes, a range of $[0,1]$ is useful to express how one trust value can change over time.
%As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.
%The rate in which trust increases or decreases must be experimented with; it could be useful to have trust decrease faster than it increases, but that might also cause false positives in the case of occasional mistakes.
%
%\subsection{Datasets}
%The datasets used for experimentation must be changed accordingly.
%Instead of using data gathered from social networks, which are static, dynamic network datasets must be used, such as traffic information for cities.
%
%At first, the objective is to test using the Working Day Movement Model, setting parameters that contribute to the validation of the algorithm.
%When the simulation begins, a set number of nodes are malicious and, therefore, other nodes will have low trust in them after interacting.
%In time, benign and malicious nodes can change their behavior and, accordingly, their trustworthiness will be updated.
%
%It might still be useful to simulate using real-world movement traces.
%An example of such dataset is the Zurich Trace \cite{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
%The chosen dataset doesn't necessarily need to be created for VANETs, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.
%The mobility of the nodes is the most important aspect of the dataset.

### 2017-12-01 dissertation/5-simulations/chapter5.tex
dissertation v 2017-11-30

--- 2017-06-01 qualification/5-proposal/chapter5.tex
@@
-\chapter{Proposal}
-\label{chap:proposal}
+\chapter{Simulations}
+\label{chap:simulations}
 
-This work proposes that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
-In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
-%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
-Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
-By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
-This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as input for a malicious node detection algorithm.
-
-%The previous chapters show why vehicular networks require distinguished trust models compared to social networks and traditional MANETs.
-%In this chapter, it is proposed that certain aspects from social networks can be found in vehicular networks and, through those aspects, it is possible to adapt a trust model designed for traditional complex networks to a vehicular environment.
-%These social properties allow for long-term trust relationships to be established amongst nodes of a network and, therefore, previously existing trust models can be adapted to the vehicular environment.
-
-In this chapter, the details for the above process are presented.
-First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
-Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
-Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
-Finally, a schedule for the next steps of the project is shown.
-
-\section{Social Networks and VANETs}
-\label{section:socialvanets}
-Some proposed trust models for vehicular networks, such as \cite{huang2014social}, state that the likelihood of two nodes meeting each other twice is too low to be relevant.
-However, it stands to reason that, throughout the course of several days, many drivers take similar routes at similar times of day (e.g.
-to commute to work) and, therefore, their vehicles are in similar locations each day.
-Additionally, many cities rely on main roads to serve as backbones to their traffic, meaning there is a high density of vehicles on those roads during rush hours.
-Since that is true for a notable percentage of a city's fleet, it can also be assumed that those vehicles may frequently encounter each other during their commute.
-While two vehicles that share a commute route may not be direct neighbors every day, they are likely to be relatively close to each other most days, meaning few hops separate them in the ad-hoc network.
-Furthermore, certain pairs of vehicles are bound to be within communication range of each other nearly every day.
-Examples of these include vehicles whose owners are neighbors or coworkers.
-Such vehicles' trust relationship should become steady over time and, in the case of positive trust, they can use each other's information to learn more about other nodes in the network.
-
-%\textbf{[WORKING DAY MOBILITY MODEL]}
-%They do not, however, present data regarding specific pairs of nodes.
-%[Should we propose to show this information in our work?]
-
-%Regardless of these properties of commuting vehicles, the probability of two specific vehicles meeting each other more than once, in given time, tends to 1.
-%So it is also reasonable to assume that two vehicles, whose drivers reside in the same city, over a long period of time (e.g. one year) may meet each other more than once.
-%However, if they only meet rarely, their trust opinions about each other may not be very useful.
-
-Most cities also have one or more types of mass transit systems (buses or trains).
-Those vehicles can also be part of a VANET and communicate with private cars.
-Buses share the same roads as cars, but instead of having specific destinations, they travel a predefined route during the whole day, usually tied to a tight schedule.
-Trains travel on rails, so their contact with cars is less frequent, but it can also happen on railroad crossings; they travel long distances in relatively short amount of time, which helps the dissemination of data in a VANET.
-In the same way that cars have a high probability of meeting more than once during their commutes, it is also very likely that they meet the same buses and/or trains frequently.
-
-In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
-By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
-They note that, during rush hours, the diameter of the graph decreases to around 6 hops; additionally, the frequency of total encounters between pairs of nodes in the network increases during those hours.
-Although the authors do not quantify the encounters between specific pairs of nodes, these numbers support the idea that daily commutes do indeed cause vehicular networks to exhibit social network features.
-
-%Another way of applying trust models from social networks to VANETs would be to take into account the relationships between drivers.
-%If a neighboring vehicle happens to be driven by a close friend or relative, there is reason to believe that it is not a malicious node in the network.
-%That might not always be the case, since there may be malicious hackers controlling the node and its driver is oblivious to it.
-%In any case, this approach is not covered in this work.
+\section{The ONE Simulator}
+\label{section:theone}
 
@@
 
-\section{Original trust model}
+\section{Simulation parameters and methodology}
+\label{section:parameters}
 
-The basis of the proposed trust model is \cite{vernize2015malicious}.
-This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-Furthermore, the algorithm is executed by a global observer which has information about the complete network.
+\subsection{Validation}
 
-The input graph $T = (V,O)$ is a static, connected, and directed graph containing all trust relationships in the network.
-Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
-The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.
+\section{Restrictions}
+\label{section:restrictions}
 
-The process for identifying malicious nodes within $T$ is as follows:
-
-First, $T$ is separated into strongly components using Tarjan's algorithm \cite{tarjan1972depth}.
-In each of these components, all nodes are connected by edges of value $1$.
-In other words, within a single component, all nodes trust one another completely; nodes connected by edges of value $0$ are separated into different components.
-Each of these components becomes a node of a component graph $T' = (V', O')$.
-
-The creation of the graph $T'$ simplifies the remaining computation.
-Since each node of $T'$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
-They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
-After the formation of $T'$, one or more heuristics can be used to classify the nodes as benign or malicious.
-
-The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
-Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
-After running either algorithms with graph $T'$, the color with the most nodes in $T'$ is classified as correct, and all others are classified as malicious.
-Once this information in $T'$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
-
-In the experiments shown in the article, the coloring heuristic shows promising results, identifying a high ratio of the malicious nodes in the network.
-Other heuristics are experimented with, but were less effective in detecting malicious nodes, or provided too many false positives.
-Therefore, for the purposes of this paper, only the coloring heuristic is considered.
-
-Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
-In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
-Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
-
-The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
-By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
-
-%$G = (V,E)$: undirected topology graph, in which $V$ are the vertices and $E$ are the edges.
-%
-%$T = (V,O)$: directed trust graph, in which $O$ are the nodes' opinions about each other.
-%
-%$G_t,T_t$: snapshots of $G$ and $T$ in a timestamp $t$.
-%
-%$G_i,T_i$: local knowledge of $G$ and $T$ of a node $i$.
-%
-%$T' = (V',O')$: component graph of $T$.
-
-\section{Necessary changes to the trust management scheme}
-\label{section:changes}
-While the algorithm described above works well for static graphs, some important modifications will have to be made to accommodate dynamic graphs.
-The main changes planned are described below.
-
-\subsection{Trust graph}
-The original work's scope does not include the generation of the trust graph, as it is a parameter to the algorithm.
-In dynamic graphs, trust relationships change over time, so how those changes affect the proposed trust model have to be considered.
-This includes not only the full figure of the graph, but also each node's views on its neighbors.
-
-\subsection{Strongly connected components}
-Strongly connected components, or components in which all nodes have positive trust relationships with each other, are somewhat different in dynamic graphs.
-As nodes enter, leave, or change locations in the network, the components change accordingly.
-Therefore, our algorithm will have to work using ``snapshots'' of the graph, meaning each measurement will be valid for a specific timestamp.
-Since the graph changes with time, for every timestamp $t$, there is a corresponding topology graph $G_t=(V_t, E_t)$, which includes the available vertices and edges in that moment in time, and a trust graph $T_t=(V_t, O_t)$, with the nodes' opinions at that moment.
-
-In a later step, it might be useful to consider how specific nodes may change over several snapshots.
-Specifically, the suggestion of running the algorithm in iterations, by removing the malicious nodes in each step, may be applied with the peculiarities of a dynamic network.
-
-\subsection{Graph knowledge}
-In the original paper, the malicious nodes are identified by a central entity with access to all nodes' trust values.
-Not only does our work plan to decentralize this procedure, having complete knowledge of the graph is unfeasible in large-scale dynamic networks such as VANETs.
-Therefore, our algorithm needs to work using only a partial understanding of the network.
-For this to be viable, each node might have to gather information not only from its views on its neighbors, but also from its neighbors' views.
-
-The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
-Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
-That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
-They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
-The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).
-
-Thus, each node has its own model of the network topology, $G_i$, and generates a trust graph $T_i=(V_i, O_i)$, in which $i$ is the node's identifier.
-Combined with the necessity of timestamps explained above, this results in a graph $T_{it}=(V_{it}, O_{it})$ for each node and each timestamp, which is the input of the new algorithm.
-Accordingly, there are also component graphs $T'_{it}=(V'_{it}, O'_{it})$, generated on each instance of the algorithm.
-
-During the early stages of development, it will be considered that vertices cannot enter or leave the network and, therefore, only the layout of the edges changes.
-This is sufficient to experiment with the algorithm's performance with the added complexity of a dynamic network.
-
-\subsection{Binary trust}
-A model using binary trust is one in which trust values can only be 0 or 1.
-If node $A$ positively trusts node $B$, the value of the edge $A\rightarrow B$ is 1.
-If it does not trust $B$, or is unsure, the value is 0.
-Such a scheme makes sense in a static graph but, in a dynamic one, it is interesting to have a range of possible trust values.
-
-Since nodes interact with each other occasionally and might learn information from other nodes, a range of $[0,1]$ is useful to express how one trust value can change over time.
-As one node learns more about another, the value can grow or diminish and, once it goes above or below a certain threshold, the neighboring node can be deemed trustworthy or not.
-The rate in which trust increases or decreases must be experimented with; it could be useful to have trust decrease faster than it increases, but that might also cause false positives in the case of occasional mistakes.
-
-\subsection{Datasets}
-The datasets used for experimentation must be changed accordingly.
-Instead of using data gathered from social networks, which are static, dynamic network datasets must be used, such as traffic information for cities.
-
-At first, the objective is to test using the Working Day Movement Model, setting parameters that contribute to the validation of the algorithm.
-When the simulation begins, a set number of nodes are malicious and, therefore, other nodes will have low trust in them after interacting.
-In time, benign and malicious nodes can change their behavior and, accordingly, their trustworthiness will be updated.
-
-It might still be useful to simulate using real-world movement traces.
-An example of such dataset is the Zurich Trace \cite{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
-The chosen dataset doesn't necessarily need to be created for VANETs, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.
-The mobility of the nodes is the most important aspect of the dataset.
-
-\section{Tasks and schedule}
-
-Given the proposed trust management scheme, the following list of tasks has been developed:
-
-\begin{enumerate}
-	\item \textbf{Changes to the malicious node identification algorithm}. Develop and adapt the algorithm using the changes listed in \autoref{section:changes}.
-	\item \textbf{Changes to the movement model}. Develop and adapt the Working Day Movement Model according to \autoref{section:workingday}.
-	\item \textbf{Study of the simulator}.
-Several simulators for vehicular networks exist, and one must be chosen in order to run simulations using the newly developed model.
-The Working Day Movement Model was tested with the ONE Simulator \cite{keranen2009one}.
-The source code for both the simulator and the WDM generation can be found online \cite{onerepo}, and will be used for the necessary simulations.
-	\item \textbf{Simulations and adjustments}.
-Once the core of the project has been developed and a simulation has been chosen, simulations begin.
-The models may be tweaked in order to achieve better results, and all data from the process will be collected.
-	\item \textbf{Analysis of data and results}.
-With the data collected from simulations, it will be possible to observe the outcome of the proposed trust model.
-If tweaks were made to the model during simulations, those changes can be listed, showing which ones provided better results.
-	\item \textbf{Writing and submission of article}.
-Having concluded development, an article will be written explaining the motivations, challenges and results of the study.
-This article will be submitted to a congress.
-	\item \textbf{Writing of dissertation}. Finally, the whole process will be detailed within the dissertation.
-\end{enumerate}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{ |p{0.7cm}||p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}|p{0.7cm}| }
- \hline
- \textbf{Task} & May & Jun & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
- \hline
- \hline
- \textbf{1} & \checkmark & \checkmark & \checkmark & & & & & & & &\\
- \hline
- \textbf{2} & & \checkmark & \checkmark & \checkmark & & & & & & &\\
- \hline
- \textbf{3} & \checkmark & \checkmark & \checkmark & \checkmark & & & & & & &\\
- \hline
- \textbf{4} & & & & \checkmark & \checkmark & \checkmark & \checkmark & & & &\\
- \hline
- \textbf{5} & & & & & & & \checkmark & \checkmark & \checkmark & &\\
- \hline
- \textbf{6} &  &  & & & & & & \checkmark & \checkmark & &\\
- \hline
- \textbf{7} &  &  & & & & & & \checkmark & \checkmark & \checkmark & \checkmark \\
- \hline
-\end{tabular}
-\caption{Expected schedule}
-\label{table:schedule}
-\end{table}
-
-An approximate schedule is shown in \autoref{table:schedule}.
-The first column displays the task, using the numbers from the list above, while the remaining ones are marked when work on that task is expected for the given month.
-The months of April through December are in 2017 and January through March are in 2018.
-
-%Cronograma
-%
-%1. Adaptação do modelo de nós maliciosos
-%2. Adaptação do modelo de mobilidade
-%3. Escolha e estudo do simulador
-%4. Simulações
-%5. Análise de dados e resultados
-%6. Escrita e submissão de artigo (procurar congresso)
-%7. Escrita da dissertação
-
-%=====================================================
+\section{Results}
+\label{section:results}

### 2018-01-22 article/article.tex
article and dissertation updates

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

\usepackage[utf8]{inputenc}	% arquivos LaTeX em Unicode (UTF8)
\RequirePackage[plainpages,pdfpagelabels]{hyperref}	% PDF com links, metadados

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{../pdf/}{../jpeg/}{./plots/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

\usepackage{cleveref}

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\floatname{algorithm}{Algorithm}
\renewcommand{\algorithmiccomment}[1]{~~~// #1}
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Trust management for vehicular networks}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Renan Greca}
\IEEEauthorblockA{Departamento de Informática\\
Universidade Federal do Paraná\\
Curitiba, Brazil\\
Email: rdmgreca@inf.ufpr.br}
\and
\IEEEauthorblockN{Luiz Carlos Albini}
\IEEEauthorblockA{Departamento de Informática\\
Universidade Federal do Paraná\\
Curitiba, Brazil\\
Email: albini@inf.ufpr.br}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular ad-hoc network (VANET), in which cars share data amongst themselves in order to cooperate and make roads safer and more efficient.
A decentralized ad-hoc solution, which doesn't rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
To generate these trust values, nodes use information acquired from past interactions; nodes which frequently share false or irrelevant data will have lower trust values than the ones which appear to be reliable.
This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
The results prove to be accurate and efficient, thanks to the low complexity of the algorithms constituting the trust model.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\label{section:introduction}

Within the next few years, a substantial share of new vehicles will come equipped with networking features \cite{connectedcar2016}.
These features will allow vehicles to quickly share data with other nearby devices and can be useful tools to reduce traffic and the risk of accidents.
Over 1 million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
By quickly sharing data with neighboring vehicles, without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.

The accepted for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
It describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
Communications between two OBUs is called vehicle-to-vehicle (V2V) communication, while communication between an OBU and an RSU is called vehicle-to-infrastructure (V2I) communication.
This study focuses only on V2V scenarios, and therefore, any references to VANETs and their nodes refer exclusively to vehicles and their on-board units.

As is expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
Some issues that could be exploited in a VANET include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming and stalking \cite{isaac2010security}.

Each of these problems require specific solutions, although there are ways of making the network safer in general.
One way is taking advantage of the concept of trust between members of the network.
By having nodes remember previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
Trust solutions for VANETs are generally classified into data-oriented trust, which emphasizes the contents of a message, and entity-based trust, which emphasizes the sender of a message.
The solution described in this article is one of entity-based trust.

The remainder of this paper is organized as follows. \autoref{section:previouswork} shows some examples of existing trust models for vehicular networks and the trust model for static networks that served as basis for this one; \autoref{section:algorithm} explains the algorithms used to develop the trust model; \autoref{section:results} shows the simulations which validate the usefulness of the trust model; and \autoref{section:conclusion} presents closing thoughts on the project.

\section{Previous work}
\label{section:previouswork}

Several models have been proposed to solve the problem of trust in vehicular networks. 
In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.

For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.
It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.

\section{Algorithm}
\label{section:algorithm}

The objective of the trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
The algorithm that dictates the trust model runs continuously, with iterations happening in a preset interval.
In every iteration, a node checks its neighbors to see if there were changes to the network and runs a combination of algorithms that help it detect malicious nodes in the known network.
Then, it separates graph $T$ into components using Tarjan's strongly connected components algorithm.
Finally, it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.

%In order to do this, a node must maintain a graph $T=(V,O)$ that represents its knowledge of the network, in which $V$ is the set of known nodes and $O$ is the set of opinions nodes have about others.
%Opinions are directed edges, each one with a weight between 0 and 1 that represents the degree of trust from one node towards another.
%A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.

The detailed descriptions of both algorithms are below, followed by the complete process of each iteration of the trust model.

% ---------------------------------------------------------------------------------------------------
\subsection{Tarjan's strongly connected components algorithm}
\label{section:tarjan}
An important aspect of the trust model is the use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth}.
This allows a large graph to be abstracted into a smaller graph, which therefore reduces the input for further algorithms.
Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
For the purposes of trust management, this definition is extended to accept only paths of edges with weight 1.
Every node of the input graph $T$ must belong to a component.

In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessible from every call of the function.
$index$ and $lowlink$ are arrays indexed by node IDs (predefined unique identifiers), $count$ is an integer and $stack$ is a last-in-first-out data structure.

The algorithm works by performing a depth-first search, adding nodes to $stack$ as they are visited.
If two nodes are present on $stack$, then there is a path from the first node to the second one (in the order they were added to $stack$).
Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.

In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value 1).
If node $v$ has not yet been visited, the algorithm is called for  $v$.
The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
Therefore, nodes must be popped from the $stack$ until $u$ is found.
Each node popped, including $u$, is a member of a strongly connected component.

The number of components is, at most, $|v|$.
In a worst-case scenario, each node is put into its own component; however, this would not be the case for most useful input graphs.

Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
The complexity of the algorithm is $O(|V|+|O|)$ for a graph $T = (V,O)$.

With the results of Tarjan's algorithm, an undirected component graph $C = (V',O')$ is formed.
Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $o' \in O'$ are edges from $T$ between nodes that do not belong in the same component.

\begin{algorithm}
\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
\begin{algorithmic}[1]

\Function{Tarjan}{vertex $u$}

\State $index[u] = count$
\State $lowlink[u] = count$
\State $count \gets count + 1$
\State \textbf{push} $u$ \textbf{to} $stack$

\For{$v$ \textbf{in} neighbors of $u$ }
	\If{weight of $u \rightarrow v$ is 0}
		\State \textbf{continue}
	\EndIf

	\If{$index[v] = -1$}
		\Comment{\emph{v has not been visited yet}}
		\State \textbf{Tarjan}($v$)
	\EndIf

	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
\EndFor

\If{$lowlink[u] = index[u]$}
	\Repeat
		\Comment{\emph{unstack nodes until u is found}}
		\State \textbf{pop} $w$ \textbf{from} $stack$
		\State \textbf{add} $w$ \textbf{to} $component$
	\Until {$w = u$}
\EndIf

\EndFunction
\end{algorithmic}
\end{algorithm}

% ---------------------------------------------------------------------------------------------------
\subsection{Graph coloring with minimum colors}
\label{section:coloring}
The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
Graph coloring is one of the possible heuristics used to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
Out of the tested heuristics, it presented the best results, so it has been chosen as the heuristic for the trust model.

The process of graph coloring consists of labeling each node with a color so that no two neighboring nodes share the same color.
This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.

In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
The complexity of the algorithm is $O(|E|)$ for an undirected graph $G = (V,E)$.
As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.

Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph} \cite{vernize2013dissertation}.

A limitation of this algorithm is that the edges must be sorted according to node indexes beforehand.
It does not matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
This is demonstrated in \cite{vernize2013dissertation}.

\begin{algorithm}
\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
\begin{algorithmic}[1]

\Function{Coloring}{graph $G$}

\State \textbf{color} all nodes of $G$ \textbf{with} 0
\State $d \rightarrow 0$
\For{$e=(u,v)$ \textbf{in} edges of $G$}
	\If{$u$ and $v$ have the same color}
		\If{$color[v] = d$}
			\State $d \rightarrow d+1$
		\EndIf
		\State $color[v] \rightarrow d$
	\EndIf
\EndFor

\EndFunction
\end{algorithmic}
\end{algorithm}

% ---------------------------------------------------------------------------------------------------
\subsection{Trust management for a vehicular network}
\label{section:trustmanagement}
%\subsection{Malicious Node Identification Algorithm}
%\label{section:mani}

%The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
%The authors present a malicious node identification scheme based on strongly connected components and graph coloring.
%The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
%Furthermore, the algorithm is executed by a global observer which has information about the complete network.

In order to work with dynamic networks, an algorithm must consider snapshots as inputs.
The algorithm runs continuously, running new iterations at a predetermined interval.
Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
A snapshot $G_i = (V_i, E_i)$ represents the complete topology of the network at the iteration $i$; $V_i$ is the set of nodes participating in the network at that time and $E_i$ is the set of edges connecting pairs of nodes within communication range of each other.
As the network is dynamic, it is expected that each $G_i$ is different from $G_{i-1}$, but also that the changes follow a determined mobility model.

Furthermore, it the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
This is necessary because it is not reasonable to expect a supervisor of the network to have complete knowledge of the nodes and relationships in the network.
Each node starts knowing only about itself and maintains its own abstraction of the network surrounding it.
Every node $u$ has a static, connected and directed trust graph $T^u = (V^u, O^u)$, in which $V^u$ is the set of nodes $u$ is aware of and $O^u$ is the set of trust relationships (opinions) $u$ knows about between members of $V^u$.
Since $T^u$ changes over time, there is a $T^u_i$ for every iteration $i$.

In each iteration, every node $u$ runs the following steps to detect malicious nodes in the network:

\begin{enumerate}
	\item Node $u$ checks who are its neighbors (nodes within communication range).
		  Newly discovered nodes and newly formed edges are added to $T^u_i$.
		  Edges are created with weight 0.5.
	\item Node $u$ tests all of its neighbors to discover which ones can be directly trusted or not.
		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$, adding nodes and edges that were present on $T^v_{i-1}$ but not on $T^u_{i-1}$.
	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are identified as benign or malicious according to the same rules as the MaNI algorithm.
\end{enumerate}

During steps 1, 2 and 3, the node is collecting and organizing information.
The storage of an abstraction of the network in the form of a trust graph $T$ is crucial for the remainder of the algorithm.
A prerequisite of step 2 is a test that correctly classifies a neighboring node as benign or malicious.
After these steps, $T^u_i$ is formed, which is then used for the following steps.

Each edge $w \rightarrow v$ in $T^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
The closer the value is to 1, the more $w$ trusts $v$.
These values are not mutual, so the value of $w \rightarrow v$ can be different from the value of $v \rightarrow w$.

After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail on \autoref{section:tarjan}.
For each node in a component, there is a path formed by edges of weight higher than a threshold $0 < t < 1$ to each other node in the same component.
In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, O'^u_i)$.

The creation of $C^u_i$ simplifies the remaining computation.
Since each vertex $v' \in V'^u_i$ is a component of $T^u_i$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as the same.
They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
After the formation of $C^u_i$, one or more heuristics can be used to classify the nodes as benign or malicious.

The authors of \cite{vernize2015malicious} describe the coloring heuristic, which can use a graph coloring algorithm such as DSATUR \cite{brelaz1979new} or the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}.
After running either algorithms with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
Once this information in $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on their components' classifications.

In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
Therefore, for the purposes of this research, only the coloring heuristic is considered.

%Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
%In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
%Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
%
%The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
%By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.

% ---------------------------------------------------------------------------------------------------

%The MaNI algorithm works well and is efficient for static graphs.
%However, some important modifications had to be made to accommodate dynamic graphs in a decentralized fashion.
%The main changes made are described below.

%First of all,

%Another important change is on the trust values themselves.
%MaNI uses binary trust (either 0 or 1).
%This has been replaced by a floating-point number between 0 and 1; a predefined threshold defines which values are deemed trustworthy or not.
%This was not the case in the first versions of the trust model; it was added in order to avoid abrupt changes to opinions.
%Since the network is different on each iteration of the algorithm, there were instances in which a change to the topology caused a radical change to the output of the graph coloring algorithm.
%Using a trust value that gradually changes over time, output variations between two iterations became small and gradual.
%
%At the start of an execution, a node knows only about itself.
%From that point, iterations of the algorithm will run in a set interval of $l$ seconds.
%In each iteration (labeled $t$ after the timestamp it occurs on) and for every node $u$, the following steps are executed:

%To mention:
%- Graph building;
%- Change from binary trust to a weighted model;
%- Evolution of trust over time.

%First of all, there is an important change to the graph knowledge used to run the algorithm.
%The complete trust graph is a parameter to the MaNI algorithm, which is executed by a supervisor with global knowledge of the graph.
%
%In this case, given a complete trust graph $T = (V, O)$, each node $u \in V$ has its own vision of the network and has to build a graph, called $T_u$, to represent its surrounding network.
%In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
%For every neighbor $v \in V$, $T_u$ must me merged with $T_v$.
%An edge $u \rightarrow v$ exists if $u$ is aware of $v$.
%Given sufficient iterations, $T_u$ forms a complete or nearly complete vision of the network.

%Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
%Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
%When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
%If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.

%Furthermore, this implies that the malicious node detection algorithm runs using incomplete graph knowledge.
%This is expected and generally unavoidable in vehicular networks, which have high mobility and can be extremely large.

\section{Results}
\label{section:results}

In order to test the trust model, simulations were made using an implementation of the algorithm in Python.
To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a reasonable imitation of vehicle movement in real life.
Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.

Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model.
The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
%The number of nodes and their communication range vary for different simulations.
There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
The parameters regarding offices, meeting spots and shopping were kept intact.

\subsection{Network Density}
The communication range of nodes vary from 10m to 50m, to illustrate the impact of different network densities.
The network density is a value that abstracts the volume and frequency of connections in a vehicular network.
It is calculated using the communication range of the nodes, the amount of nodes, and the total area of the simulation.
For this trust model, higher densities yield better results, since nodes can construct and update their models of the network more quickly (this is demonstrated in \autoref{subsection:simulations}).

The simulations shown here have densities that vary between 0.001 (10m range) and 0.04 (50m range).
As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for the algorithm.

\subsection{Simulations}
\label{subsection:simulations}
To improve readability, all figures in this section follow the same format.
The $X$ axis shows the results of sequential iterations, ranging from 0 to 8639, while the $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100.
The blue line represents the percentage of nodes detected out of the complete network.
Magenta is the percentage of malicious nodes in the network (ground truth).
Finally, green represents the nodes correctly identified as malicious (true positives), cyan represents the undetected malicious nodes (false negatives) and red represents the benign nodes incorrectly identified as malicious (false positives).

\autoref{fig:random10}, \autoref{fig:random30} and \autoref{fig:random50} show the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
It is possible to see how the increase in communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.

\autoref{fig:random1} to \autoref{fig:random6} show the variation of results for different amounts of malicious nodes in the network.
By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
At 40\%, a small part of malicious nodes are yet to be detected.
At 50\%, as is expected, the results are inconsistent as control of the network is completely divided between benign and malicious nodes; at this point, the network is completely compromised.
The amount of malicious nodes also affects network discovery, since nodes do not trust information from malicious neighbors.

\autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
Most malicious nodes are identified by the end of the first day; in the following iterations, the algorithm finishes building the network model and sorts out remaining false negative or false positive results.
After iteration 20000, the results are completely consistent.

% ----
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\lin0.5ewidth]{Network_A_10.0/new_plots/10.png}
%%\includegraphics[width=0.45\linewidth]{Network_A_10.0/new_plots/30.png}
%\caption{Simulation without random nodes and range of 10m.} \label{fig:nonrandom1}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/30.png}
%\caption{Simulation without random nodes and range of 30m.} \label{fig:nonrandom2}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/50.png}
%\caption{Simulation without random nodes and range of 50m.} \label{fig:nonrandom3}
%\end{figure}

% ----

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/10.png}
\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/30.png}
\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/50.png}
\caption{Simulation with random nodes and range of 50m.} \label{fig:random50}
\end{figure}

% ----

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_1}
\caption{Simulation with random nodes, range of 10m and 1\% malicious.} \label{fig:random0}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_5}
\caption{Simulation with random nodes, range of 10m and 5\% malicious.} \label{fig:random1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_10}
\caption{Simulation with random nodes, range of 10m and 10\% malicious.} \label{fig:random2}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_20}
\caption{Simulation with random nodes, range of 10m and 20\% malicious.} \label{fig:random3}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_30}
\caption{Simulation with random nodes, range of 10m and 30\% malicious.} \label{fig:random4}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_40}
\caption{Simulation with random nodes, range of 10m and 40\% malicious.} \label{fig:random5}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA/10_50}
\caption{Simulation with random nodes, range of 10m and 50\% malicious.} \label{fig:random6}
\end{figure}

% ----

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{Network_rA7/10_10}
\caption{Simulation with random nodes, range of 10m and 10\% malicious during 7 days.} \label{fig:random7}
\end{figure}

%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/4.png}
%\caption{Edge $(2,3)$ is checked and node $3$ gets the current value of $d$.} \label{fig:coloring5}
%\end{subfigure}
%\hspace*{2cm} % separation between the subfigures
%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/5.png}
%\caption{Edge $(2,4)$ is checked and node $4$ gets the current value of $d$.} \label{fig:coloring6}
%\end{subfigure}
%\hspace*{2cm} % separation between the subfigures
%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/6.png}
%\caption{Edge $(3,4)$ is checked and node $4$ gets a new color. $d=4$ (yellow).} \label{fig:coloring6}
%\end{subfigure}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Conclusion}
\label{section:conclusion}

While malicious nodes can diminish the usefulness of vehicular networks, the concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.

As nodes travel across the network and collect more data from neighbor, they are able to form an abstraction of the network which can be used to detect malicious nodes.
By placing nodes into strongly connected component with Tarjan's algorithm, a network containing a large amount of node can be simplified into a much smaller one.
And, using a simple graph coloring algorithm, most malicious nodes stand out by having colors different than the majority of nodes.

The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a small amount of false positives.
Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.

% conference papers do not normally have an appendix

% use section* for acknowledgment
%\section*{Acknowledgment}
%
%
%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv, article}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2018-01-22 dissertation/4-project/chapter4.tex
article and dissertation updates

@@
 
-\section{SNAP library}
-\label{section:snap}
-In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
-The Stanford Network Analysis Platform (SNAP) library \cite{snap} was chosen primarily because it is memory efficient.
-The simulations require multiple graphs that share the same set of nodes (because each node in the network has its own knowledge of the surrounding network), and the SNAP library uses pointers to nodes and edges, saving memory by not having to duplicate the entire data structure.
-It is written in C++, but, for these simulations, the Snap.py Python library was used.
-
-The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
-
 \section{Trust management algorithm}
@@
 If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.
+
+[Illustrate graph building here]
+
+Furthermore, this implies that the malicious node detection algorithm runs using incomplete graph knowledge.
+This is expected and generally unavoidable in vehicular networks, which have high mobility and can be extremely large.
+
+-\\
+----------\\
+This paragraph is probably pertinent to delimiting the size of graphs:\\
+The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
+Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
+That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
+They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
+The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).\\
+----------\\
 

### 2018-01-22 dissertation/5-simulations/chapter5.tex
article and dissertation updates

@@
 \label{chap:simulations}
+
+\section{SNAP library}
+\label{section:snap}
+In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
+The Stanford Network Analysis Platform (SNAP) library \cite{snap} was chosen primarily because it is memory efficient.
+The simulations require multiple graphs that share the same set of nodes (because each node in the network has its own knowledge of the surrounding network), and the SNAP library uses pointers to nodes and edges, saving memory by not having to duplicate the entire data structure.
+It is written in C++, but, for these simulations, the Snap.py Python library was used.
+
+The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
 

### 2018-02-05 article/article.tex

@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
 
@@
 \RequirePackage[plainpages,pdfpagelabels]{hyperref}	% PDF com links, metadados
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
 
@@
   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
   \graphicspath{{../pdf/}{../jpeg/}{./plots/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
   \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
 
 \usepackage{cleveref}
+
+\usepackage{tikz}
+\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
+\def\scalecheck{\resizebox{\widthof{\checkmark}*\ratio{\widthof{x}}{\widthof{\normalsize x}}}{!}{\checkmark}}
 
@@
 \renewcommand{\algorithmiccomment}[1]{~~~// #1}
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
 
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{Trust management for vehicular networks}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
+\title{TruMan: Trust Management for Vehicular Networks}
+
 \author{\IEEEauthorblockN{Renan Greca}
@@
 \and
-\IEEEauthorblockN{Luiz Carlos Albini}
+\IEEEauthorblockN{Luiz Carlos P. Albini}
 \IEEEauthorblockA{Departamento de Informática\\
@@
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
 By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular ad-hoc network (VANET), in which cars share data amongst themselves in order to cooperate and make roads safer and more efficient.
-A decentralized ad-hoc solution, which doesn't rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
+A decentralized ad-hoc solution, which does not rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
 However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
@@
 
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 Over 1 million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
-Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
-In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
-By quickly sharing data with neighboring vehicles, without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
+%Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
+%In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
+By quickly sharing data with neighboring vehicles without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
 
-The accepted for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
+The communication stardard for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
 It describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
-Communications between two OBUs is called vehicle-to-vehicle (V2V) communication, while communication between an OBU and an RSU is called vehicle-to-infrastructure (V2I) communication.
-This study focuses only on V2V scenarios, and therefore, any references to VANETs and their nodes refer exclusively to vehicles and their on-board units.
+Communication between two OBUs is called vehicle-to-vehicle (V2V) communication, while communication between an OBU and an RSU is called vehicle-to-infrastructure (V2I) communication.
+This study focuses only on V2V scenarios, and therefore, any references to Vehicular ad-hoc networks (VANETs) and their nodes refer exclusively to vehicles with on-board units.
 
 As is expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
-Some issues that could be exploited in a VANET include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming and stalking \cite{isaac2010security}.
+Some issues that could be exploited in such network include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming or stalking \cite{isaac2010security}.
 
 Each of these problems require specific solutions, although there are ways of making the network safer in general.
-One way is taking advantage of the concept of trust between members of the network.
+One way is taking advantage of the concept of trust between network members.
 By having nodes remember previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
@@
 
+Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes might be malicious.
+Since the network is dynamic, nodes acquire more knowledge as time passes and the algorithm's results become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between nodes that connect frequently.
+After running a series of simulations, the results were very promising; after one simulated days, nodes were able to correctly identify over 95\% of malicious nodes in the network.
+
+%==> O QUE SE PRETENDE ATINGIR? O QUE TEM DE DIFERENTE? QUAIS RESULTADOS FORAM OBTIDOS? FALTA UM PARAGRAFO AQUI FALANDO DISSO.
+
 The remainder of this paper is organized as follows.
-\autoref{section:previouswork} shows some examples of existing trust models for vehicular networks and the trust model for static networks that served as basis for this one; \autoref{section:algorithm} explains the algorithms used to develop the trust model; \autoref{section:results} shows the simulations which validate the usefulness of the trust model; and \autoref{section:conclusion} presents closing thoughts on the project.
+Section \ref{section:previouswork} shows some examples of existing trust models for vehicular networks and the trust model for static networks that served as basis for this one;
+\autoref{section:algorithm} explains the algorithms used to develop the trust model;
+\autoref{section:results} shows the simulations which validate the usefulness of the trust model;
+and \autoref{section:conclusion} presents closing thoughts on the project.
 
-\section{Previous work}
-\label{section:previouswork}
-
-Several models have been proposed to solve the problem of trust in vehicular networks. 
-In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
-None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
-Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
-There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
-
-For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-Furthermore, the algorithm is executed by a global observer which has information about the complete network.
-It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.
-
-\section{Algorithm}
+\section{Understanding the algorithm}
 \label{section:algorithm}
 
-The objective of the trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
+%==> MUDAR O TITULO, PRACISA DIZER ALGO DO TIPO UNDERNEATH ALGORITHM
+
+The objective of the Truman trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
 The algorithm that dictates the trust model runs continuously, with iterations happening in a preset interval.
-In every iteration, a node checks its neighbors to see if there were changes to the network and runs a combination of algorithms that help it detect malicious nodes in the known network.
-Then, it separates graph $T$ into components using Tarjan's strongly connected components algorithm.
-Finally, it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.
+In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
+This information is maintained in a directed graph $T=(V,O)$, in which $V$ is the set of known vertices and $O$ is the trust relationship between pairs $u,v \in V$.
+Graph $T$ is known as the \textit{trust graph}.
+
+The edges $(u, v) \in O$ store a value between 0 and 1 that represents the degree of trust $u$ has for $v$.
+This value is called the \textit{opinion} of $u$ about $v$.
+At the start of the execution, edges have value $0.5$; this value increases when nodes have positive interactions and decreases otherwise.
+A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
+By default, the threshold is $0.5$.
+
+After collecting information from other nodes, Truman performs two steps: ($i$) it divides the network graph into strongly connected components using Tarjan's strongly connected components algorithm; ($ii$) it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.
 
@@
 
-The detailed descriptions of both algorithms are below, followed by the complete process of each iteration of the trust model.
+The detailed descriptions of both algorithms are below, followed by the complete process of each iteration of the Truman trust model.
+
+%==> VOCE AINDA NAO FALOU DOS PESOS DAS ARESTAS E COMO ELES FORAM CONSEGUIDOS
+%==> VOCE NAO FALOU DOS GRAFOS
 
@@
 \label{section:tarjan}
-An important aspect of the trust model is the use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth}.
-This allows a large graph to be abstracted into a smaller graph, which therefore reduces the input for further algorithms.
+The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of Truman's efficiency.
+This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further algorithms.
 Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
-For the purposes of trust management, this definition is extended to accept only paths of edges with weight 1.
+For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
 Every node of the input graph $T$ must belong to a component.
@@
 
-In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value 1).
+In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
 If node $v$ has not yet been visited, the algorithm is called for  $v$.
@@
 \For{$v$ \textbf{in} neighbors of $u$ }
-	\If{weight of $u \rightarrow v$ is 0}
+	\If{weight of $u \rightarrow v$ $<$ $h$}
 		\State \textbf{continue}
@@
 \label{section:coloring}
-The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
+The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring.
 Graph coloring is one of the possible heuristics used to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
@@
 
-The process of graph coloring consists of labeling each node with a color so that no two neighboring nodes share the same color.
-This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
+%The process of graph coloring consists of labeling each node with a color so that no two neighboring nodes share the same color.
+%This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
 It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
-
 In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
@@
 As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
-For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
+%For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
 
@@
 % ---------------------------------------------------------------------------------------------------
-\subsection{Trust management for a vehicular network}
+\subsection{The Truman algorithm}
 \label{section:trustmanagement}
@@
 
-In order to work with dynamic networks, an algorithm must consider snapshots as inputs.
+Truman is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm used here.
+However, MaNI was developed for static networks such as social networks, and is executed by a supervising agent, making it unsuitable for a vehicular network.
+The work describes here takes advantage of the efficiency of the two graph theory algorithms into a model that is viable for highly dynamic networks.
+
+In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
 The algorithm runs continuously, running new iterations at a predetermined interval.
@@
 
-Furthermore, it the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
-This is necessary because it is not reasonable to expect a supervisor of the network to have complete knowledge of the nodes and relationships in the network.
-Each node starts knowing only about itself and maintains its own abstraction of the network surrounding it.
-Every node $u$ has a static, connected and directed trust graph $T^u = (V^u, O^u)$, in which $V^u$ is the set of nodes $u$ is aware of and $O^u$ is the set of trust relationships (opinions) $u$ knows about between members of $V^u$.
-Since $T^u$ changes over time, there is a $T^u_i$ for every iteration $i$.
+Furthermore, the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
+%This is necessary because it is not reasonable to expect a supervisor of the network to have complete knowledge of the nodes and relationships in the network.
+Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
+Every node $u$ has a static, connected and directed trust graph $T^u = (V^u, O^u)$, in which $V^u$ is the set of nodes $u$ is aware of and $O^u$ is the set of trust relationships (opinions) $u$ knows between members of $V^u$.
+Since $T^u$ changes over time, there is a $T^u_i = (V^u_i, O^u_i)$ for every iteration $i$.
 
-In each iteration, every node $u$ runs the following steps to detect malicious nodes in the network:
+%==> VOCE NAO FALOU DOS PESOS DAS ARESTAS AINDA
 
-\begin{enumerate}
-	\item Node $u$ checks who are its neighbors (nodes within communication range).
-		  Newly discovered nodes and newly formed edges are added to $T^u_i$.
-		  Edges are created with weight 0.5.
-	\item Node $u$ tests all of its neighbors to discover which ones can be directly trusted or not.
-		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
-	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$, adding nodes and edges that were present on $T^v_{i-1}$ but not on $T^u_{i-1}$.
-	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
-	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are identified as benign or malicious according to the same rules as the MaNI algorithm.
-\end{enumerate}
+%==> ACHO QUE PRECISA FALAR AQUI DA PRINCIPAL DIFERENCA PARA O MaNI, QUE AQUI ESTAMOS FALANDO DE REDE MOVEL E ALGORITMO DISTRIBUIDO
 
-During steps 1, 2 and 3, the node is collecting and organizing information.
+At first, the node is collecting and organizing information.
 The storage of an abstraction of the network in the form of a trust graph $T$ is crucial for the remainder of the algorithm.
@@
 
-Each edge $w \rightarrow v$ in $T^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
+Each edge $w \rightarrow v \in O^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
 The closer the value is to 1, the more $w$ trusts $v$.
@@
 Therefore, for the purposes of this research, only the coloring heuristic is considered.
+
+In summary, every node $u$ runs the following steps in each iteration to detect malicious nodes in the network:
+
+\begin{enumerate}
+	\item Node $u$ checks who are its neighbors (nodes within its communication range).
+		  Newly discovered nodes and newly formed edges are added to $T^u_i$.
+		  Edges are created with weight 0.5.
+	\item Node $u$ tests all of its neighbors to discover which ones can be directly trusted or not.
+		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
+	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$.
+	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
+	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are classified as benign or malicious. %according to the same rules as the MaNI algorithm.
+\end{enumerate}
 
@@
 
-\autoref{fig:random10}, \autoref{fig:random30} and \autoref{fig:random50} show the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
+\autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
 It is possible to see how the increase in communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
 
-\autoref{fig:random1} to \autoref{fig:random6} show the variation of results for different amounts of malicious nodes in the network.
+\autoref{fig:random0} and \autoref{fig:random1} show the variation of results for different amounts of malicious nodes in the network.
 By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
@@
 After iteration 20000, the results are completely consistent.
+
+%==> ACHO QUE PODIA USAR O MULTIFIGURE PARA AGRUPAR UM POUCO AS FIGURAS. COLOCA 3 LADO A LADO. ASSIM SOBRE ESPAÇO NO FINAL PARA PODERMOS COMPARAR COM TRABALHOS RELACIONADOS.
 
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/10.png}
-\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
-\end{figure}
+\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
+%\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
+\hspace{1pt}
+\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
+%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
+\hspace{1pt}
+\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/50.png}
+\caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
+\label{fig:random103050}
+\end{figure*}
 
-\begin{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/30.png}
+%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/50.png}
+%\caption{Simulation with random nodes and range of 50m.} \label{fig:random50}
+%\end{figure}
+
+% ----
+
+\begin{figure*}
 \centering
-\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/30.png}
-\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
-\end{figure}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_1}
+\hspace{1pt}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_5}
+\hspace{1pt}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_10}
+\caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
+\label{fig:random0}
+\end{figure*}
 
-\begin{figure}
+%\begin{figure}
+%\centering0.5
+%\includegraphics[width=0.5\linewidth]{Network_rA/10_5}
+%\caption{Simulation with random nodes, range of 10m and 5\% malicious.} \label{fig:random1}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA/10_10}
+%\caption{Simulation with random nodes, range of 10m and 10\% malicious.} \label{fig:random2}
+%\end{figure}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA/10_20}
+%\caption{Simulation with random nodes, range of 10m and 20\% malicious.} \label{fig:random3}
+%\end{figure}
+
+\begin{figure*}
 \centering
-\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/50.png}
-\caption{Simulation with random nodes and range of 50m.} \label{fig:random50}
-\end{figure}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_30}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_40}
+\includegraphics[width=0.3\linewidth]{Network_rA/10_50}
+\caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
+\label{fig:random1}
+\end{figure*}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA/10_40}
+%\caption{Simulation with random nodes, range of 10m and 40\% malicious.} \label{fig:random5}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{Network_rA/10_50}
+%\caption{Simulation with random nodes, range of 10m and 50\% malicious.} \label{fig:random6}
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_1}
-\caption{Simulation with random nodes, range of 10m and 1\% malicious.} \label{fig:random0}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_5}
-\caption{Simulation with random nodes, range of 10m and 5\% malicious.} \label{fig:random1}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_10}
-\caption{Simulation with random nodes, range of 10m and 10\% malicious.} \label{fig:random2}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_20}
-\caption{Simulation with random nodes, range of 10m and 20\% malicious.} \label{fig:random3}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_30}
-\caption{Simulation with random nodes, range of 10m and 30\% malicious.} \label{fig:random4}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_40}
-\caption{Simulation with random nodes, range of 10m and 40\% malicious.} \label{fig:random5}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA/10_50}
-\caption{Simulation with random nodes, range of 10m and 50\% malicious.} \label{fig:random6}
-\end{figure}
-
-% ----
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{Network_rA7/10_10}
+\includegraphics[width=0.7\linewidth]{Network_rA7/10_10}
 \caption{Simulation with random nodes, range of 10m and 10\% malicious during 7 days.} \label{fig:random7}
@@
 
+\section{Previous work}
+\label{section:previouswork}
+
+==> FALTA DETALHAR ELES.
+
+Several models have been proposed to solve the problem of trust in vehicular networks. 
+In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
+None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
+Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
+There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
+
+For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
+The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
+Furthermore, the algorithm is executed by a global observer which has information about the complete network.
+It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
+The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.
+
+\begin{table}[h!]
+\caption{Properties of Truman and related work}
+\label{table:schedule}
+\centering
+\begin{tabular}{|p{1.5cm}||p{1.5cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
+ \hline
+ \textbf{Task} & Truman & \cite{vernize2013dissertation} & \cite{golle2004detecting} & \cite{patwardhan2006data} & \cite{minhas2010towards} \\
+ \hline
+ Decentralized & - & \checkmark & - & - & \\
+ \hline
+ Sparsity & - & \checkmark & - & - & \\
+ \hline
+ Dynamics & - & \checkmark & - & - & \\
+ \hline
+ Scalability & - & \checkmark & - & - & \\
+ \hline
+ Confidence & - & \checkmark & - & - & \\
+ \hline
+ Security & - & \checkmark & - & - & \\
+ \hline
+ Privacy & - & \checkmark & - & - & \\
+ \hline
+ Robustness & - & \checkmark & - & - & \\
+ \hline
+ Cost & $O(|V|* |O|)$ & n/a & n/a & n/a & n/a \\
+ \hline
+\end{tabular}
+\end{table}
+
 \section{Conclusion}
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv, article}
+\bibliography{article}
 %

### 2018-02-05 article/article.tex
minor corrections

@@
 Trust solutions for VANETs are generally classified into data-oriented trust, which emphasizes the contents of a message, and entity-based trust, which emphasizes the sender of a message.
-The solution described in this article is one of entity-based trust.
 
+This work proposes a new and improved entity-based trust model to compute the trust between any pair of nodes in a vehicular network.
 Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes might be malicious.
-Since the network is dynamic, nodes acquire more knowledge as time passes and the algorithm's results become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between nodes that connect frequently.
-After running a series of simulations, the results were very promising; after one simulated days, nodes were able to correctly identify over 95\% of malicious nodes in the network.
+Since the network is dynamic, nodes acquire more knowledge as time passes and the algorithm's results become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between frequently connected nodes.
+Simulations demonstrate that nodes are able to correctly identify more than 95\% of nodes in the network (malicious or not).
+Obviously, the algorithm correctness depends upon the velocity of the nodes, the frequency of the information exchange between them and the running time, i.e.
+the longevity, of the algorithm.
+%After running a series of simulations, the results were very promising; after one simulated day,
 
@@
 The objective of the Truman trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
-The algorithm that dictates the trust model runs continuously, with iterations happening in a preset interval.
+The algorithm that dictates the trust model runs continuously, with iterations happening in preset intervals.
 In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
@@
 
-The edges $(u, v) \in O$ store a value between 0 and 1 that represents the degree of trust $u$ has for $v$.
+The edges $(u, v) \in O$ store a value between 0 and 1 which represents the degree of trust $u$ has for $v$.
 This value is called the \textit{opinion} of $u$ about $v$.
@@
 A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
-By default, the threshold is $0.5$.
+%By default, the threshold is $0.5$.
 
@@
 The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of Truman's efficiency.
-This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further algorithms.
+This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
 Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
@@
 Every node of the input graph $T$ must belong to a component.
+
+-----ACHO QUE ESTA DEFINIÇAO ESTA COM MUITOS DETALHES DE IMPLEMENTACAO ---
+-----PODERIA DEIXAR ELA MAIS ALTO NIVEL, EM FORMA DE CONJUNTOS POR EXEMPLO ---
 
@@
 
-The number of components is, at most, $|v|$.
-In a worst-case scenario, each node is put into its own component; however, this would not be the case for most useful input graphs.
+---------------
+
+The number of components is, at most, $|v|$: in a worst-case scenario, each node is placed into its own component.
+%; however, this would not be the case for most useful input graphs.   ---> NÃO ENTENDI <---
 
@@
 Graph coloring is one of the possible heuristics used to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
-Out of the tested heuristics, it presented the best results, so it has been chosen as the heuristic for the trust model.
+Out of the tested heuristics, it presents the best results, so it has been chosen as the heuristic for the trust model.
 
@@
 %This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
-It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
+It has been mathematically proven that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
 In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
@@
 
+==> ALGUMAS PARTES ESTÃO MUITO REPETITIVAS COM A SEÇAO DE CIMA. PRECISA REVER PARA NÃO FICAR MUITO REPETITIVO. <==
+
 Truman is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm used here.
-However, MaNI was developed for static networks such as social networks, and is executed by a supervising agent, making it unsuitable for a vehicular network.
-The work describes here takes advantage of the efficiency of the two graph theory algorithms into a model that is viable for highly dynamic networks.
+However, MaNI was developed for static networks such as social networks, and is executed by an external supervising agent (i.e. outside of the network), making it unsuitable for a vehicular network.
+%The work describes here takes advantage of the efficiency of the two graph theory algorithms into a model that is viable for highly dynamic networks.
 
 In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
-The algorithm runs continuously, running new iterations at a predetermined interval.
+The algorithm runs continuously, running new iterations at predetermined intervals.
 Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
@@
 A prerequisite of step 2 is a test that correctly classifies a neighboring node as benign or malicious.
+There are several ways to perform such a test, but a study between these methods is out of the scope of this paper.
 After these steps, $T^u_i$ is formed, which is then used for the following steps.
@@
 They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
-After the formation of $C^u_i$, one or more heuristics can be used to classify the nodes as benign or malicious.
+After the formation of $C^u_i$, a heuristic is used to classify the nodes as benign or malicious.
 
-The authors of \cite{vernize2015malicious} describe the coloring heuristic, which can use a graph coloring algorithm such as DSATUR \cite{brelaz1979new} or the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}.
+==> PRIMEIRA FRASE ESTA ESTRANHA <==
+In this study, the coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
 After running either algorithms with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
@@
 
-In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
-Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
-Therefore, for the purposes of this research, only the coloring heuristic is considered.
+%In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
+%Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
+%Therefore, for the purposes of this research, only the coloring heuristic is considered.
 

### 2018-02-07 article/article.tex
further changes to sections II and III

@@
 In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
-This information is maintained in a directed graph $T=(V,O)$, in which $V$ is the set of known vertices and $O$ is the trust relationship between pairs $u,v \in V$.
+This information is maintained in a directed graph $T=(V,E)$, in which $V$ is the set of known vertices and $E$ is the set of trust relationships between pairs $u,v \in V$.
 Graph $T$ is known as the \textit{trust graph}.
 
-The edges $(u, v) \in O$ store a value between 0 and 1 which represents the degree of trust $u$ has for $v$.
+Each edge $(u, v) \in E$ stores a value between 0 and 1 which represents the degree of trust $u$ has for $v$.
 This value is called the \textit{opinion} of $u$ about $v$.
 At the start of the execution, edges have value $0.5$; this value increases when nodes have positive interactions and decreases otherwise.
-A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
+A threshold $0 < h < 1$ is used to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
 %By default, the threshold is $0.5$.
@@
 This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
-Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
+Given a directed graph $T = (V,E)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
 For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
@@
 
------ACHO QUE ESTA DEFINIÇAO ESTA COM MUITOS DETALHES DE IMPLEMENTACAO ---
------PODERIA DEIXAR ELA MAIS ALTO NIVEL, EM FORMA DE CONJUNTOS POR EXEMPLO ---
+%-----ACHO QUE ESTA DEFINIÇAO ESTA COM MUITOS DETALHES DE IMPLEMENTACAO ---
+%-----PODERIA DEIXAR ELA MAIS ALTO NIVEL, EM FORMA DE CONJUNTOS POR EXEMPLO ---
 
-In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessible from every call of the function.
-$index$ and $lowlink$ are arrays indexed by node IDs (predefined unique identifiers), $count$ is an integer and $stack$ is a last-in-first-out data structure.
+%Three data structures are used by the algorithm:
+%$index$ and $lowlink$ are indexed by node IDs (predefined unique identifiers) and $stack$ is a last-in-first-out stack.
+%The $count$ variable is incremented at each visited node.
+%In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessible from every call of the function.
+%$index$ and $lowlink$ are arrays indexed by node IDs (predefined unique identifiers), $count$ is an integer and $stack$ is a last-in-first-out data structure.
 
-The algorithm works by performing a depth-first search, adding nodes to $stack$ as they are visited.
-If two nodes are present on $stack$, then there is a path from the first node to the second one (in the order they were added to $stack$).
+The algorithm works by performing a depth-first search, adding nodes to a stack as they are visited.
+If two nodes are present on the stack, then there is a path from the first node to the second one (in the order they were added to the stack).
 Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.
@@
 After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
-Therefore, nodes must be popped from the $stack$ until $u$ is found.
+Therefore, nodes must be popped from the stack until $u$ is found.
 Each node popped, including $u$, is a member of a strongly connected component.
 
----------------
+%---------------
 
@@
 Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
-The complexity of the algorithm is $O(|V|+|O|)$ for a graph $T = (V,O)$.
+The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
 
-With the results of Tarjan's algorithm, an undirected component graph $C = (V',O')$ is formed.
-Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $o' \in O'$ are edges from $T$ between nodes that do not belong in the same component.
+With the results of Tarjan's algorithm, an undirected component graph $C = (V',E')$ is formed.
+Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $e' \in E'$ are edges from $T$ between nodes that do not belong in the same component.
 
@@
 Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
-The complexity of the algorithm is $O(|E|)$ for an undirected graph $G = (V,E)$.
-As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
+The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
+As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V'|^2)$ \cite{brelaz1979new}.
 %For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
@@
 
-==> ALGUMAS PARTES ESTÃO MUITO REPETITIVAS COM A SEÇAO DE CIMA. PRECISA REVER PARA NÃO FICAR MUITO REPETITIVO. <==
+%==> ALGUMAS PARTES ESTÃO MUITO REPETITIVAS COM A SEÇAO DE CIMA. PRECISA REVER PARA NÃO FICAR MUITO REPETITIVO. <==
 
@@
 
-In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
-The algorithm runs continuously, running new iterations at predetermined intervals.
-Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
-A snapshot $G_i = (V_i, E_i)$ represents the complete topology of the network at the iteration $i$; $V_i$ is the set of nodes participating in the network at that time and $E_i$ is the set of edges connecting pairs of nodes within communication range of each other.
-As the network is dynamic, it is expected that each $G_i$ is different from $G_{i-1}$, but also that the changes follow a determined mobility model.
+%In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
+%The algorithm runs continuously, running new iterations at predetermined intervals.
+%Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
+%A snapshot $S_i$ represents the complete topology of the network at the iteration $i$.
+%%; $V_i$ is the set of nodes participating in the network at that time and $E_i$ is the set of edges connecting pairs of nodes within communication range of each other.
+%As the network is dynamic, it is expected that each $S_i$ is different from $S_{i-1}$, but also that the changes follow a determined mobility model.
 
+In order to work with dynamic networks, the Truman algorithm runs iterations at predetermined intervals.
+%As nodes acquire more information, their model of the network will change.
 Furthermore, the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
-%This is necessary because it is not reasonable to expect a supervisor of the network to have complete knowledge of the nodes and relationships in the network.
 Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
-Every node $u$ has a static, connected and directed trust graph $T^u = (V^u, O^u)$, in which $V^u$ is the set of nodes $u$ is aware of and $O^u$ is the set of trust relationships (opinions) $u$ knows between members of $V^u$.
-Since $T^u$ changes over time, there is a $T^u_i = (V^u_i, O^u_i)$ for every iteration $i$.
+Every node $u$ stores a model of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows about between members of $V$.
+Since each node has its own model and the model changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
 
@@
 At first, the node is collecting and organizing information.
-The storage of an abstraction of the network in the form of a trust graph $T$ is crucial for the remainder of the algorithm.
-A prerequisite of step 2 is a test that correctly classifies a neighboring node as benign or malicious.
+A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
 There are several ways to perform such a test, but a study between these methods is out of the scope of this paper.
-After these steps, $T^u_i$ is formed, which is then used for the following steps.
+Every time a neighboring node $v$ is tested as benign, the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
+After this, a new $T^u_i$ is formed, which is then used for the following steps.
 
-Each edge $w \rightarrow v \in O^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
-The closer the value is to 1, the more $w$ trusts $v$.
-These values are not mutual, so the value of $w \rightarrow v$ can be different from the value of $v \rightarrow w$.
+%Each edge $w \rightarrow v \in E^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
+%The closer the value is to 1, the more $w$ trusts $v$.
+%These values are not mutual, so the value of $w \rightarrow v$ can be different from the value of $v \rightarrow w$.
 
-After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail on \autoref{section:tarjan}.
-For each node in a component, there is a path formed by edges of weight higher than a threshold $0 < t < 1$ to each other node in the same component.
+After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}.
+For each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
 In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
-Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, O'^u_i)$.
+Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, E'^u_i)$.
 
@@
 
-==> PRIMEIRA FRASE ESTA ESTRANHA <==
+%==> PRIMEIRA FRASE ESTA ESTRANHA <==
 In this study, the coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
-After running either algorithms with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
-Once this information in $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on their components' classifications.
+After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
+Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on their components' classifications.
 
@@
 %Therefore, for the purposes of this research, only the coloring heuristic is considered.
+
+The complexity of the whole algorithm can be calculated by adding the most costly operations involved.
+As discussed above, Tarjan's algorithm has a complexity of $O(|V|+|E|)$ (for the trust graph $T$), while the graph coloring algorithm has a complexity of $O(|E'|)$ (for the component graph $C$).
+The most costly part of the algorithm is the graph merge operation that happens between trusted nodes.
+The complexity of the graph merge algorithm is $O(|E|)$ for each neighbor a node has in an iteration; this number is at most $|V|$.
+The total complexity of Truman is therefore $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
+Since $|O'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
 
@@
 
-In order to test the trust model, simulations were made using an implementation of the algorithm in Python.
-To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a reasonable imitation of vehicle movement in real life.
+In order to test the Truman trust model, simulations were made using an implementation of the algorithm in Python.
+To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides strong similarity with vehicle movement in real life.
 Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
 
-Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model.
+Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}.
 The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
@@
 \subsection{Network Density}
+
+==> PRECISA MOSTRAR A EQUAÇÃO DE CALCULO DA DENSIDADE, SENÃO FICA DIFICIL DE ENTENDER O CALCULO.
+
 The communication range of nodes vary from 10m to 50m, to illustrate the impact of different network densities.
-The network density is a value that abstracts the volume and frequency of connections in a vehicular network.
-It is calculated using the communication range of the nodes, the amount of nodes, and the total area of the simulation.
-For this trust model, higher densities yield better results, since nodes can construct and update their models of the network more quickly (this is demonstrated in \autoref{subsection:simulations}).
+The network density is a value which abstracts the volume and frequency of connections in a vehicular network.
+It is calculated using the transmission range of the nodes, the amount of nodes, and the total area of the simulation.
+For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
 
-The simulations shown here have densities that vary between 0.001 (10m range) and 0.04 (50m range).
+Simulations shown here have densities that varying between 0.001 (10m range) and 0.04 (50m range).
 As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for the algorithm.
@@
 \label{subsection:simulations}
-To improve readability, all figures in this section follow the same format.
-The $X$ axis shows the results of sequential iterations, ranging from 0 to 8639, while the $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100.
-The blue line represents the percentage of nodes detected out of the complete network.
-Magenta is the percentage of malicious nodes in the network (ground truth).
-Finally, green represents the nodes correctly identified as malicious (true positives), cyan represents the undetected malicious nodes (false negatives) and red represents the benign nodes incorrectly identified as malicious (false positives).
+To improve readability, all figures in this section follow the same format:
+
+\begin{itemize}
+\item $X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
+\item $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
+\item Blue line represents the percentage of nodes detected out of the complete network;
+\item Magenta is the percentage of malicious nodes in the network (ground truth);
+\item Green represents the nodes correctly identified as malicious (true positives);
+\item Cyan represents the undetected malicious nodes (false negatives); 
+\item Red represents the benign nodes incorrectly identified as malicious (false positives).
+\end{itemize}
 
 \autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
-It is possible to see how the increase in communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
+It is possible to see how the increase in the communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
 
@@
 At 40\%, a small part of malicious nodes are yet to be detected.
-At 50\%, as is expected, the results are inconsistent as control of the network is completely divided between benign and malicious nodes; at this point, the network is completely compromised.
-The amount of malicious nodes also affects network discovery, since nodes do not trust information from malicious neighbors.
+At 50\%, as expected, the results are inconsistent the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
+The amount of malicious nodes also affects convergence of the algorithm, since nodes do not trust information from malicious neighbors.
 
@@
 
+==> AUMETEI UM POUCO AS FIGURAS, MAS ELAS PARECEM ESTAR MEIO DEFOCADAS, VE SE CONSEGUE MELHORAR UM POUCO A QUALIDADE DELAS.
+
 \begin{figure*}
 \centering
-\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
+\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
 %\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
-\hspace{1pt}
-\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
+%\hspace{1pt}
+\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
 %\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
-\hspace{1pt}
-\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/50.png}
+%\hspace{1pt}
+\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}
 \caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
@@
 \centering
-\includegraphics[width=0.3\linewidth]{Network_rA/10_1}
-\hspace{1pt}
-\includegraphics[width=0.3\linewidth]{Network_rA/10_5}
-\hspace{1pt}
-\includegraphics[width=0.3\linewidth]{Network_rA/10_10}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_1}
+%\hspace{1pt}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_5}
+%\hspace{1pt}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_10}
 \caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
@@
 \centering
-\includegraphics[width=0.3\linewidth]{Network_rA/10_30}
-\includegraphics[width=0.3\linewidth]{Network_rA/10_40}
-\includegraphics[width=0.3\linewidth]{Network_rA/10_50}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_30}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_40}
+\includegraphics[width=0.329\linewidth]{Network_rA/10_50}
 \caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
@@
  \hline
- Cost & $O(|V|* |O|)$ & n/a & n/a & n/a & n/a \\
+ Cost & $O(|V| \times |O|)$ & $O(|V|+|O|)$ & n/a & n/a & n/a \\
  \hline

### 2018-02-07 article/article.tex
added density calculation and began related work comparison

@@
 
-==> PRECISA MOSTRAR A EQUAÇÃO DE CALCULO DA DENSIDADE, SENÃO FICA DIFICIL DE ENTENDER O CALCULO.
+%==> PRECISA MOSTRAR A EQUAÇÃO DE CALCULO DA DENSIDADE, SENÃO FICA DIFICIL DE ENTENDER O CALCULO.
 
 The communication range of nodes vary from 10m to 50m, to illustrate the impact of different network densities.
-The network density is a value which abstracts the volume and frequency of connections in a vehicular network.
-It is calculated using the transmission range of the nodes, the amount of nodes, and the total area of the simulation.
-For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
+The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
+For Truman, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
+It is calculated using the transmission range of the nodes ($\sigma$), the amount of nodes($\eta$), and the total area of the simulation($\alpha$).
+
+The coverage of a single node is the circumference around it formed by the transmission radius; this is multiplied by the number of nodes in the network to estimate the maximum coverage area.
+Finally, the value is divided by the total area of the environment.
+ formula is as follows:
+
+$$ \delta = \frac{\frac{\sigma^2\pi}{2} \times \eta}{\alpha} $$
 
 Simulations shown here have densities that varying between 0.001 (10m range) and 0.04 (50m range).
-As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for the algorithm.
+As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for a satisfactory performance of the algorithm.
 
@@
 
-==> AUMETEI UM POUCO AS FIGURAS, MAS ELAS PARECEM ESTAR MEIO DEFOCADAS, VE SE CONSEGUE MELHORAR UM POUCO A QUALIDADE DELAS.
+%==> AUMETEI UM POUCO AS FIGURAS, MAS ELAS PARECEM ESTAR MEIO DEFOCADAS, VE SE CONSEGUE MELHORAR UM POUCO A QUALIDADE DELAS.
 
@@
 
-==> FALTA DETALHAR ELES.
+Several models have been proposed to solve the problem of trust in vehicular networks. 
+The analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties of a trust management model for VANETs.
+In the first part of this section, these properties are described with an assessment of whether or not Truman satisfies their conditions.
+Then, some of the most relevant models are described, considering how well they satisfy the desired properties and how they compare with Truman.
 
-Several models have been proposed to solve the problem of trust in vehicular networks. 
-In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
-None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
-Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
-There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
+The first property is \textit{decentralized trust establishment}, which means nodes must be able to form their own trust values about other nodes.
+Nodes may or not use information from other trustworthy nodes to build trust values.
+Truman satisfies this as it is built from the ground up for decentralized systems.
 
-For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-Furthermore, the algorithm is executed by a global observer which has information about the complete network.
-It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.
+The second is \textit{coping with sparsity}, meaning the model still functions when there are few nodes populating the network.
+The experiments using low density values demonstrate that Truman works in reasonably sparse networks.
+Due to its decentralized nature, it can also work on isolated chunks of the network.
+
+Then there are \textit{event/task and location/time dynamics}, which is how the model reacts to different situations depending on what, where and when events happen.
+Although this has not been used in the simulations in this paper, Truman can easily be extended to receive time and location dynamics as long as nodes store the geolocation and timestamp data when it acquires new information.
+
+The fourth desired property is \textit{scalability}, meaning the model can work on very large networks at high speeds.
+Due to the low complexity of the algorithms used in the model, Truman can be highly scalable, as it does not put substantial pressure on the vehicles' on-board units.
+It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
+
+The fifth is an \textit{integrated confidence measure}, a value that allows nodes to estimate just how useful the output of the algorithm is.
+Since nodes using Truman store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
+The closer it is to 1, the higher the chance that it is accurate.
+
+The sixth is \textit{system level security}, which requires authentication of nodes participating in the network.
+This has not been used in simulations of Truman, but can be included as a separate security model during the transmission of messages.
+
+The seventh is \textit{sensitivity to privacy concerns}, which avoids eavesdropping and stalking by malicious nodes.
+Truman has not been designed with this in mind, but it doesn't inhibit privacy protection, however it does require that nodes cannot be completely anonymous --- there must be some way of identifying the same node over time.
+
+Finally, there is \texit{robustness}, which is the model's resistance to attacks.
+Truman satisfies this as malicious nodes are usually identified quickly and accurately, making it difficult for them to perform attacks.
+
+%None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
+%Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
+%There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
+%
+%For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
+%The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
+%Furthermore, the algorithm is executed by a global observer which has information about the complete network.
+%It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
+%The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.
 
@@
 \centering
-\begin{tabular}{|p{1.5cm}||p{1.5cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
+\begin{tabular}{|p{1.5cm}||p{0.8cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
  \hline
- \textbf{Task} & Truman & \cite{vernize2013dissertation} & \cite{golle2004detecting} & \cite{patwardhan2006data} & \cite{minhas2010towards} \\
+ \textbf{Task} & Truman & \cite{vernize2013dissertation} & \cite{golle2004detecting} & \cite{minhas2010towards} & \cite{patwardhan2006data} & [00] & [00] \\
  \hline
- Decentralized & - & \checkmark & - & - & \\
+ Decentralized 	& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \\
  \hline
- Sparsity & - & \checkmark & - & - & \\
+ Sparsity 		& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \\
  \hline
- Dynamics & - & \checkmark & - & - & \\
+ Dynamics 		& \checkmark & - 		  & -		   & \checkmark & \checkmark & & \\
  \hline
- Scalability & - & \checkmark & - & - & \\
+ Scalability 	& \checkmark & \checkmark & -		   & \checkmark & \checkmark & & \\
  \hline
- Confidence & - & \checkmark & - & - & \\
+ Confidence 	& \checkmark & \checkmark & -		   & \checkmark & \checkmark & & \\
  \hline
- Security & - & \checkmark & - & - & \\
+ Security 		& \checkmark & - 		  & \checkmark & \checkmark & \checkmark & & \\
  \hline
- Privacy & - & \checkmark & - & - & \\
+ Privacy 		& - 		 & - 		  & \checkmark & \checkmark & - 		 & & \\
  \hline
- Robustness & - & \checkmark & - & - & \\
+ Robustness 	& \checkmark & - 		  & \checkmark & -			& - 		 & & \\
  \hline
- Cost & $O(|V| \times |O|)$ & $O(|V|+|O|)$ & n/a & n/a & n/a \\
- \hline
+% Cost & O(|V|* |E|) & O(|V|+|E|) & n/a & n/a & n/a \\
+% \hline
 \end{tabular}

### 2018-02-09 article/article.tex
completed section IV, layout changes to figures and tables

@@
 \usepackage{cleveref}
+\usepackage{subfig}
 
@@
 
+\begin{table}[h!]
+\caption{Simulation parameters}
+\label{table:parameters}
+\centering
+\begin{tabular}{|p{3cm}||p{3cm}|}
+ \hline
+ \textbf{Parameter}	& \textbf{Value} \\
+ \hline
+ \hline
+ Duration 			& 86400 seconds \\
+ \hline
+ Work day length 	& 28800 seconds \\
+ \hline
+ Std. dev. departure time & 7200 seconds \\
+ \hline
+ Node velocity 		& 7~10m/s \\
+ \hline
+ Simulation area	& approximately 14km\textsuperscript{2} \\
+ \hline
+ Number of nodes 	& 150 (WDMM) \\
+  					& + 10 (random) \\
+ \hline
+% Cost & O(|V|* |E|) & O(|V|+|E|) & n/a & n/a & n/a \\
+% \hline
+\end{tabular}
+\end{table}
+
 Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}.
-The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
-Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
-%The number of nodes and their communication range vary for different simulations.
-There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
+Different parameters are shown in \autoref{table:parameters}.
+%The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
+%Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
+%%The number of nodes and their communication range vary for different simulations.
+%There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
 Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
 The parameters regarding offices, meeting spots and shopping were kept intact.
+A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
 
@@
 
-\begin{figure*}
+\begin{figure*}[!t]
 \centering
-\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
-%\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
-%\hspace{1pt}
-\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
-%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
-%\hspace{1pt}
-\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}
-\caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/10.png}%
+\label{subfig:random10}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}%
+\label{subfig:random30}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}%
+\label{subfig:random50}}
+\caption{Simulation with random nodes, 10\% malicious and range of (a) 10m, (b) 30m and (c) 50m.}
 \label{fig:random103050}
 \end{figure*}
+
+%\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
+%\caption{a}
+%%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
+%%\hspace{1pt}
+%\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}
+%\caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
+%\label{fig:random103050}
+%\end{figure*}
 
@@
 
-\begin{figure*}
+\begin{figure*}[!t]
 \centering
-\includegraphics[width=0.329\linewidth]{Network_rA/10_1}
-%\hspace{1pt}
-\includegraphics[width=0.329\linewidth]{Network_rA/10_5}
-%\hspace{1pt}
-\includegraphics[width=0.329\linewidth]{Network_rA/10_10}
-\caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_1}%
+\label{subfig:random1}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_5}%
+\label{subfig:random2}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_10}%
+\label{subfig:random3}}
+
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
+\label{subfig:random4}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
+\label{subfig:random5}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
+\label{subfig:random6}}
+
+\caption{Simulation with random nodes, range of 10m and (a) 1\%, (b) 5\%, (c) 10\%, (d) 30\%, (e) 40\% or (f) 50\% malicious.}
 \label{fig:random0}
 \end{figure*}
+
+%\begin{figure*}
+%\centering
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_1}
+%%\hspace{1pt}
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_5}
+%%\hspace{1pt}
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_10}
+%\caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
+%\label{fig:random0}
+%\end{figure*}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.329\linewidth]{Network_rA/10_30}
-\includegraphics[width=0.329\linewidth]{Network_rA/10_40}
-\includegraphics[width=0.329\linewidth]{Network_rA/10_50}
-\caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
-\label{fig:random1}
-\end{figure*}
+%\begin{figure*}[!t]
+%\centering
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
+%\label{subfig:random4}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
+%\label{subfig:random5}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
+%\label{subfig:random6}}
+%\caption{Simulation with random nodes, range of 10m and (a) 30\%, (b) 40\% or (c) 50\% malicious.}
+%\label{fig:random1}
+%\end{figure*}
+
+%\begin{figure*}
+%\centering
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_30}
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_40}
+%\includegraphics[width=0.329\linewidth]{Network_rA/10_50}
+%\caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
+%\label{fig:random1}
+%\end{figure*}
 
@@
 \centering
-\includegraphics[width=0.7\linewidth]{Network_rA7/10_10}
+\includegraphics[width=\linewidth]{Network_rA7/10_10}
 \caption{Simulation with random nodes, range of 10m and 10\% malicious during 7 days.} \label{fig:random7}
@@
 
-\section{Previous work}
+\section{Related work}
 \label{section:previouswork}
@@
 In the first part of this section, these properties are described with an assessment of whether or not Truman satisfies their conditions.
-Then, some of the most relevant models are described, considering how well they satisfy the desired properties and how they compare with Truman.
+Then, some of the most relevant models are described, considering how well they satisfy the desired properties and \autoref{table:properties} shows how they compare with Truman.
+
+\subsection{Desired properties}
+\label{section:properties}
 
@@
 
-Finally, there is \texit{robustness}, which is the model's resistance to attacks.
+Finally, there is \textit{robustness}, which is the model's resistance to attacks.
 Truman satisfies this as malicious nodes are usually identified quickly and accurately, making it difficult for them to perform attacks.
+
+\subsection{Comparison with other trust models}
+\label{section:comparison}
+
+For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
+The model is proposed for complex networks in general, but is designed only for static networks and the algorithm relies on a global observer which has information about the complete network.
+It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
+The usage of strongly connected components and coloring serve as basis for Truman, which is expanded to work on distributed and dynamic networks such as vehicular networks.
+
+The model proposed in \cite{minhas2010towards} uses several criteria to judge whether or not a received message is trustworthy.
+First, nodes are classified by their roles, used for vehicles which should be automatically trustworthy (i.e. police cars).
+Nodes also store their experience each time an event message is received (if one neighboring node reported an event which did not turn out to be true, its trust value is reduced).
+Additionally, messages have higher reliability when their senders are closer in time and space to the reported event.
+When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
+%The model considers both role-based trust and experience-based trust; although the work proposed here does not use role-based trust, the authors provide a useful method of calculating and updating an experience-based trust value, which might be used or adapted.
+However, the model relies only on direct interaction between pairs of nodes, so no form of indirect trust (that is, trust values received from other nodes) is considered.
+
+In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
+By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
+When a node sends a message, one node in the cluster (the leader) must aggregate the other nodes' opinions on that message.
+Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; additionally, nodes that receive the message only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
+However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
+Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy.
+
+The ART model proposed in \cite{li2016art} is similar to Truman in the sense that there are two main steps: data gathering and malicious node detection.
+It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
+Then, it uses a Cosine-based metric to compare two nodes' trust vectors (a series of opinions regarding other nodes).
+However, these steps require several intensive calculations, which greatly increases the complexity of the algorithm.
+The authors present no details on how the model deals with sparsity, dynamics, scalability, security and privacy.
+
+The authors of \cite{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
+This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations when cellular communication is slow or unreliable.
+It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
+
+Finally, it is worth noting that, aside from MaNI, none of the mentioned related work presents complexity calculations for their algorithm.
+Considering the scale of the problem, Truman's cost of $O(|V|\times |E|)$ is very low without sacrificing quality.
+The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
 
@@
 %
-%For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-%The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-%Furthermore, the algorithm is executed by a global observer which has information about the complete network.
-%It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-%The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.
 
@@
 \caption{Properties of Truman and related work}
-\label{table:schedule}
+\label{table:properties}
 \centering
@@
  \hline
- \textbf{Task} & Truman & \cite{vernize2013dissertation} & \cite{golle2004detecting} & \cite{minhas2010towards} & \cite{patwardhan2006data} & [00] & [00] \\
+ \textbf{Property} & Truman & \cite{vernize2013dissertation} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
  \hline
- Decentralized 	& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \\
+ Decentralized 	& \checkmark & -		  & \checkmark & \checkmark & \checkmark & - \\
  \hline
- Sparsity 		& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \\
+ Sparsity 		& \checkmark & -		  & \checkmark & \checkmark & -			 & \checkmark \\
  \hline
- Dynamics 		& \checkmark & - 		  & -		   & \checkmark & \checkmark & & \\
+ Dynamics 		& \checkmark & - 		  & \checkmark & \checkmark & -			 & - \\
  \hline
- Scalability 	& \checkmark & \checkmark & -		   & \checkmark & \checkmark & & \\
+ Scalability 	& \checkmark & \checkmark & \checkmark & \checkmark & -			 & \checkmark \\
  \hline
- Confidence 	& \checkmark & \checkmark & -		   & \checkmark & \checkmark & & \\
+ Confidence 	& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - \\
  \hline
- Security 		& \checkmark & - 		  & \checkmark & \checkmark & \checkmark & & \\
+ Security 		& \checkmark & - 		  & \checkmark & \checkmark & - 		 & \checkmark \\
  \hline
- Privacy 		& - 		 & - 		  & \checkmark & \checkmark & - 		 & & \\
+ Privacy 		& - 		 & - 		  & \checkmark & - 			& -			 & \checkmark \\
  \hline
- Robustness 	& \checkmark & - 		  & \checkmark & -			& - 		 & & \\
+ Robustness 	& \checkmark & -		  & -		   & - 			& \checkmark & - \\
+ \hline
+ Efficiency		& \checkmark & \checkmark & -		   & -			& -			 & - \\
  \hline
@@
 As nodes travel across the network and collect more data from neighbor, they are able to form an abstraction of the network which can be used to detect malicious nodes.
-By placing nodes into strongly connected component with Tarjan's algorithm, a network containing a large amount of node can be simplified into a much smaller one.
+By placing nodes into strongly connected components, a network containing a large amount of node can be simplified into a much smaller one.
 And, using a simple graph coloring algorithm, most malicious nodes stand out by having colors different than the majority of nodes.
+This allows for a low complexity approach to malicious node identification in a dynamic network.
 
 The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a small amount of false positives.
-Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.
+As the network changes in shape, nodes acquire more information and are able to make even more accurate classifications of malicious nodes around them.
+%Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.
 

### 2018-02-11 article/article.tex
shortened sections on Tarjan and coloring

@@
 
-The algorithm works by performing a depth-first search, adding nodes to a stack as they are visited.
-If two nodes are present on the stack, then there is a path from the first node to the second one (in the order they were added to the stack).
-Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.
-
-In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
-If node $v$ has not yet been visited, the algorithm is called for  $v$.
-The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
-After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
-Therefore, nodes must be popped from the stack until $u$ is found.
-Each node popped, including $u$, is a member of a strongly connected component.
+%The algorithm works by performing a depth-first search, adding nodes to a stack as they are visited.
+%If two nodes are present on the stack, then there is a path from the first node to the second one (in the order they were added to the stack).
+%Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.
+%
+%In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
+%If node $v$ has not yet been visited, the algorithm is called for  $v$.
+%The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
+%After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
+%Therefore, nodes must be popped from the stack until $u$ is found.
+%Each node popped, including $u$, is a member of a strongly connected component.
 
@@
 The number of components is, at most, $|v|$: in a worst-case scenario, each node is placed into its own component.
+The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
 %; however, this would not be the case for most useful input graphs.   ---> NÃO ENTENDI <---
 
-Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
-The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
+%Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
 
@@
 
-\begin{algorithm}
-\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
-\begin{algorithmic}[1]
-
-\Function{Tarjan}{vertex $u$}
-
-\State $index[u] = count$
-\State $lowlink[u] = count$
-\State $count \gets count + 1$
-\State \textbf{push} $u$ \textbf{to} $stack$
-
-\For{$v$ \textbf{in} neighbors of $u$ }
-	\If{weight of $u \rightarrow v$ $<$ $h$}
-		\State \textbf{continue}
-	\EndIf
-
-	\If{$index[v] = -1$}
-		\Comment{\emph{v has not been visited yet}}
-		\State \textbf{Tarjan}($v$)
-	\EndIf
-
-	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
-\EndFor
-
-\If{$lowlink[u] = index[u]$}
-	\Repeat
-		\Comment{\emph{unstack nodes until u is found}}
-		\State \textbf{pop} $w$ \textbf{from} $stack$
-		\State \textbf{add} $w$ \textbf{to} $component$
-	\Until {$w = u$}
-\EndIf
-
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
+%\begin{algorithmic}[1]
+%
+%\Function{Tarjan}{vertex $u$}
+%
+%\State $index[u] = count$
+%\State $lowlink[u] = count$
+%\State $count \gets count + 1$
+%\State \textbf{push} $u$ \textbf{to} $stack$
+%
+%\For{$v$ \textbf{in} neighbors of $u$ }
+%	\If{weight of $u \rightarrow v$ $<$ $h$}
+%		\State \textbf{continue}
+%	\EndIf	
+%		
+%	\If{$index[v] = -1$}
+%		\Comment{\emph{v has not been visited yet}}
+%		\State \textbf{Tarjan}($v$)
+%	\EndIf
+%	
+%	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
+%\EndFor
+%
+%\If{$lowlink[u] = index[u]$}
+%	\Repeat
+%		\Comment{\emph{unstack nodes until u is found}}
+%		\State \textbf{pop} $w$ \textbf{from} $stack$
+%		\State \textbf{add} $w$ \textbf{to} $component$
+%	\Until {$w = u$}
+%\EndIf
+%
+%\EndFunction
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 
-Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph} \cite{vernize2013dissertation}.
+%Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph}.
+%
+%A limitation of this algorithm is that the edges must be sorted according to node indexes beforehand.
+%It does not matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
+%This is demonstrated in \cite{vernize2013dissertation}.
 
-A limitation of this algorithm is that the edges must be sorted according to node indexes beforehand.
-It does not matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
-This is demonstrated in \cite{vernize2013dissertation}.
-
-\begin{algorithm}
-\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
-\begin{algorithmic}[1]
-
-\Function{Coloring}{graph $G$}
-
-\State \textbf{color} all nodes of $G$ \textbf{with} 0
-\State $d \rightarrow 0$
-\For{$e=(u,v)$ \textbf{in} edges of $G$}
-	\If{$u$ and $v$ have the same color}
-		\If{$color[v] = d$}
-			\State $d \rightarrow d+1$
-		\EndIf
-		\State $color[v] \rightarrow d$
-	\EndIf
-\EndFor
-
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
+%\begin{algorithmic}[1]
+%
+%\Function{Coloring}{graph $G$}
+%
+%\State \textbf{color} all nodes of $G$ \textbf{with} 0
+%\State $d \rightarrow 0$
+%\For{$e=(u,v)$ \textbf{in} edges of $G$}
+%	\If{$u$ and $v$ have the same color}
+%		\If{$color[v] = d$}
+%			\State $d \rightarrow d+1$
+%		\EndIf
+%		\State $color[v] \rightarrow d$
+%	\EndIf
+%\EndFor
+%
+%
+%\EndFunction
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 %==> PRIMEIRA FRASE ESTA ESTRANHA <==
-In this study, the coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
+The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
 After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
@@
 
-The communication range of nodes vary from 10m to 50m, to illustrate the impact of different network densities.
+The communication range of nodes varies from 10m to 50m, to illustrate the impact of different network densities.
 The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
 For Truman, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
-It is calculated using the transmission range of the nodes ($\sigma$), the amount of nodes($\eta$), and the total area of the simulation($\alpha$).
+It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$).
 
@@
 Finally, the value is divided by the total area of the environment.
- formula is as follows:
+The network density formula is as follows:
 
-$$ \delta = \frac{\frac{\sigma^2\pi}{2} \times \eta}{\alpha} $$
+$$ \delta = \frac{\frac{\rho^2\pi}{2} \times \eta}{\alpha} $$
 
-Simulations shown here have densities that varying between 0.001 (10m range) and 0.04 (50m range).
-As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for a satisfactory performance of the algorithm.
+Simulations shown here have densities that varying between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
+As a comparison, the density of São Paulo was calculated as 2.24 with $\rho = 10$m, a value much higher than what is necessary for a satisfactory performance of the algorithm.
 
@@
 To improve readability, all figures in this section follow the same format:
-
-\begin{itemize}
-\item $X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
-\item $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
-\item Blue line represents the percentage of nodes detected out of the complete network;
-\item Magenta is the percentage of malicious nodes in the network (ground truth);
-\item Green represents the nodes correctly identified as malicious (true positives);
-\item Cyan represents the undetected malicious nodes (false negatives); 
-\item Red represents the benign nodes incorrectly identified as malicious (false positives).
-\end{itemize}
+$X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
+$Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
+blue line represents the percentage of nodes detected out of the complete network;
+magenta is the percentage of malicious nodes in the network (ground truth);
+green represents the nodes correctly identified as malicious (true positives);
+cyan represents the undetected malicious nodes (false negatives); 
+red represents the benign nodes incorrectly identified as malicious (false positives).
 
@@
 
-\autoref{fig:random0} and \autoref{fig:random1} show the variation of results for different amounts of malicious nodes in the network.
+\autoref{fig:random0} shows the variation of results for different amounts of malicious nodes in the network.
 By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
@@
 Several models have been proposed to solve the problem of trust in vehicular networks. 
-The analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties of a trust management model for VANETs.
+This analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
 In the first part of this section, these properties are described with an assessment of whether or not Truman satisfies their conditions.
@@
 
-\subsection{Desired properties}
+\subsection{Properties}
 \label{section:properties}
 
-The first property is \textit{decentralized trust establishment}, which means nodes must be able to form their own trust values about other nodes.
+\textit{Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes.
 Nodes may or not use information from other trustworthy nodes to build trust values.
@@
 
-The second is \textit{coping with sparsity}, meaning the model still functions when there are few nodes populating the network.
+\textit{Coping with sparsity}: the model still functions when there are few nodes populating the network.
 The experiments using low density values demonstrate that Truman works in reasonably sparse networks.
@@
 
-Then there are \textit{event/task and location/time dynamics}, which is how the model reacts to different situations depending on what, where and when events happen.
-Although this has not been used in the simulations in this paper, Truman can easily be extended to receive time and location dynamics as long as nodes store the geolocation and timestamp data when it acquires new information.
+\textit{Event/task and location/time dynamics}: how the model reacts to different situations depending on what, where and when events happen.
+Although this has not been used in the simulations in this paper, Truman can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
 
-The fourth desired property is \textit{scalability}, meaning the model can work on very large networks at high speeds.
-Due to the low complexity of the algorithms used in the model, Truman can be highly scalable, as it does not put substantial pressure on the vehicles' on-board units.
+\textit{Scalability}: the model can work on very large networks at high speeds.
+Due to the low complexity of the algorithms used in the model, Truman can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
 It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
 
-The fifth is an \textit{integrated confidence measure}, a value that allows nodes to estimate just how useful the output of the algorithm is.
+\textit{Integrated confidence measure}: a value that allows nodes to estimate just how useful the output of the algorithm is.
 Since nodes using Truman store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
@@
 
-The sixth is \textit{system level security}, which requires authentication of nodes participating in the network.
-This has not been used in simulations of Truman, but can be included as a separate security model during the transmission of messages.
+\textit{System level security}: requires authentication of nodes participating in the network.
+This has not been considered in evaluations of Truman, however it can be included as a separate security model during the transmission of messages.
 
-The seventh is \textit{sensitivity to privacy concerns}, which avoids eavesdropping and stalking by malicious nodes.
-Truman has not been designed with this in mind, but it doesn't inhibit privacy protection, however it does require that nodes cannot be completely anonymous --- there must be some way of identifying the same node over time.
+\textit{Sensitivity to privacy concerns}: avoids eavesdropping and stalking by malicious nodes.
+Truman has not been designed with this in mind, but it doesn't inhibit privacy protection, however it does require that nodes cannot be completely anonymous.
 
-Finally, there is \textit{robustness}, which is the model's resistance to attacks.
+\textit{Robustness}: the model's resistance to attacks.
 Truman satisfies this as malicious nodes are usually identified quickly and accurately, making it difficult for them to perform attacks.
+Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
+Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
+\textbf{Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
+( - CHEGUEI A ESTA CONCLUSÃO INTUITIVAMENTE, MAS NÃO MOSTRAMOS ISSO NOS EXPERIMENTOS.
+PRECISA?
+- ) }
+
+( - EU ADICIONEI CUSTO/EFICIÊNCIA COMO PROPRIEDADE NA TABELA; DEVO MENCIONAR AQUI? - )
 
@@
 For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but is designed only for static networks and the algorithm relies on a global observer which has information about the complete network.
+The model is proposed for complex networks in general, but it is designed only for static networks and the algorithm relies on a global observer which has information about the complete network.
 It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-The usage of strongly connected components and coloring serve as basis for Truman, which is expanded to work on distributed and dynamic networks such as vehicular networks.
+The usage of strongly connected components and coloring serves as basis for Truman, which is expanded to work on distributed and dynamic networks such as vehicular networks.
 
@@
 
-Finally, it is worth noting that, aside from MaNI, none of the mentioned related work presents complexity calculations for their algorithm.
-Considering the scale of the problem, Truman's cost of $O(|V|\times |E|)$ is very low without sacrificing quality.
+Finally, it is worth noting that, aside from \cite{vernize2015malicious}, none of the mentioned related work presents complexity calculations for their algorithm.
+Considering the scale of the problem, Truman's cost of $O(|V|\times |E|)$ is very low without sacrificing completeness and correctness.
 The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
@@
  \hline
- \textbf{Property} & Truman & \cite{vernize2013dissertation} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
+ \textbf{Property} & Truman & \cite{vernize2015malicious} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
  \hline

### 2018-02-20 article/article.tex
final version sent for review

@@
 
-\author{\IEEEauthorblockN{Renan Greca}
-\IEEEauthorblockA{Departamento de Informática\\
-Universidade Federal do Paraná\\
-Curitiba, Brazil\\
-Email: rdmgreca@inf.ufpr.br}
-\and
-\IEEEauthorblockN{Luiz Carlos P. Albini}
-\IEEEauthorblockA{Departamento de Informática\\
-Universidade Federal do Paraná\\
-Curitiba, Brazil\\
-Email: albini@inf.ufpr.br}}
+%\author{\IEEEauthorblockN{Renan Greca}
+%\IEEEauthorblockA{Departamento de Informática\\
+%Universidade Federal do Paraná\\
+%Curitiba, Brazil\\
+%Email: rdmgreca@inf.ufpr.br}
+%\and
+%\IEEEauthorblockN{Luiz Carlos P. Albini}
+%\IEEEauthorblockA{Departamento de Informática\\
+%Universidade Federal do Paraná\\
+%Curitiba, Brazil\\
+%Email: albini@inf.ufpr.br}}
+
+\author{\IEEEauthorblockN{Renan Greca and Luiz Carlos Pessoa Albini}
+        \IEEEauthorblockA{Department of Informatics -- Federal University of Paraná (UFPR) -- Curitiba, Brazil\\
+                          Email: rdmgreca@inf.ufpr.br, albini@inf.ufpr.br}
+}
 
@@
 A decentralized ad-hoc solution, which does not rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
-However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
+However, as it is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
 In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
-To generate these trust values, nodes use information acquired from past interactions; nodes which frequently share false or irrelevant data will have lower trust values than the ones which appear to be reliable.
+To generate these trust values, nodes use information acquired from past interactions.
+Nodes which frequently share false or irrelevant data must have lower trust values than the ones which appear to be reliable.
 This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
-The results prove to be accurate and efficient, thanks to the low complexity of the algorithms constituting the trust model.
+The results prove to be accurate and efficient, thanks to the low complexity of the algorithm constituting the trust model.
 \end{abstract}
+
+%\begin{IEEEkeywords}
+%Vehicular Network; Trust Management; V2V Communication;
+%\end{IEEEkeywords}
 
@@
 These features will allow vehicles to quickly share data with other nearby devices and can be useful tools to reduce traffic and the risk of accidents.
-Over 1 million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
+Over one million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
 %Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
 %In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
-By quickly sharing data with neighboring vehicles without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
+By quickly sharing data with neighboring vehicles without the need of an Internet connection, smart vehicles can alert drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
 
-The communication stardard for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
+The communication standard for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
 It describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
@@
 
-As is expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
+As expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
 Some issues that could be exploited in such network include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming or stalking \cite{isaac2010security}.
@@
 One way is taking advantage of the concept of trust between network members.
-By having nodes remember previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
-Trust solutions for VANETs are generally classified into data-oriented trust, which emphasizes the contents of a message, and entity-based trust, which emphasizes the sender of a message.
+By having nodes remembering previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
+Trust solutions for VANETs are generally classified into data-oriented, emphasizing the message contents, or entity-based, emphasizing message senders.
 
 This work proposes a new and improved entity-based trust model to compute the trust between any pair of nodes in a vehicular network.
-Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes might be malicious.
-Since the network is dynamic, nodes acquire more knowledge as time passes and the algorithm's results become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between frequently connected nodes.
+Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes are likely malicious.
+Since the network is dynamic, nodes acquire more knowledge as time passes and the results of the algorithm become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between frequently connected nodes.
 Simulations demonstrate that nodes are able to correctly identify more than 95\% of nodes in the network (malicious or not).
-Obviously, the algorithm correctness depends upon the velocity of the nodes, the frequency of the information exchange between them and the running time, i.e.
-the longevity, of the algorithm.
+The algorithm correctness depends upon the velocity of the nodes, the frequency of the information exchanged between them and the running time, i.e. the longevity, of the algorithm.
 %After running a series of simulations, the results were very promising; after one simulated day,
@@
 The remainder of this paper is organized as follows.
-Section \ref{section:previouswork} shows some examples of existing trust models for vehicular networks and the trust model for static networks that served as basis for this one;
-\autoref{section:algorithm} explains the algorithms used to develop the trust model;
-\autoref{section:results} shows the simulations which validate the usefulness of the trust model;
-and \autoref{section:conclusion} presents closing thoughts on the project.
+Section \ref{section:algorithm} details the proposed trust model;
+Section \ref{section:results} contains the simulation results which demonstrate the usefulness of the model;
+Section \ref{section:previouswork} presents the previously published trust models for vehicular networks and compares them with TruMan;
+and Section \ref{section:conclusion} has the conclusion and future work directions.
 
-\section{Understanding the algorithm}
+\section{TruMan}
 \label{section:algorithm}
@@
 
-The objective of the Truman trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
+The objective of the TruMan trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
 The algorithm that dictates the trust model runs continuously, with iterations happening in preset intervals.
 In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
-This information is maintained in a directed graph $T=(V,E)$, in which $V$ is the set of known vertices and $E$ is the set of trust relationships between pairs $u,v \in V$.
+This information is maintained in a directed graph $T=(V,E)$, in which $V$ is the set of known vertices and $E$ is the set of known trust relationships between pairs $u,v \in V$.
 Graph $T$ is known as the \textit{trust graph}.
@@
 
-After collecting information from other nodes, Truman performs two steps: ($i$) it divides the network graph into strongly connected components using Tarjan's strongly connected components algorithm; ($ii$) it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.
+After collecting information from other nodes, TruMan performs two steps: ($i$) it divides the network graph into strongly connected components using Tarjan's algorithm; ($ii$) it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.
 
@@
 \label{section:tarjan}
-The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of Truman's efficiency.
+The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of TruMan's efficiency.
 This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
-Given a directed graph $T = (V,E)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
+Given a directed graph $T = (V,E)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there is a path from $u$ to $v$ and a path from $v$ to $u$.
 For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
@@
 
-The number of components is, at most, $|v|$: in a worst-case scenario, each node is placed into its own component.
+The number of components is, at most, $|V|$: in a worst-case scenario, each node is placed into its own component.
 The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
@@
 
-With the results of Tarjan's algorithm, an undirected component graph $C = (V',E')$ is formed.
-Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $e' \in E'$ are edges from $T$ between nodes that do not belong in the same component.
+From the output of Tarjan's algorithm, an undirected component graph $C = (V',E')$ is formed.
+Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $e' \in E'$ are edges from $T$ between nodes that do not belong in the same component (e.g.
+if $u$ and $v$ are separated into $u'$ and $v'$, $e = (u,v)$ becomes $e'=(u',v')$.
 
@@
 %This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
-It has been mathematically proven that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
+It has been mathematically proven that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the chromatic number of the graph) is an NP-hard problem \cite{sanchez1989determining}.
 In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
@@
 % ---------------------------------------------------------------------------------------------------
-\subsection{The Truman algorithm}
+\subsection{The TruMan algorithm}
 \label{section:trustmanagement}
@@
 
-Truman is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm used here.
+TruMan is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm.
 However, MaNI was developed for static networks such as social networks, and is executed by an external supervising agent (i.e. outside of the network), making it unsuitable for a vehicular network.
@@
 
-In order to work with dynamic networks, the Truman algorithm runs iterations at predetermined intervals.
+In order to work with dynamic networks, the TruMan algorithm runs iterations at predetermined intervals.
 %As nodes acquire more information, their model of the network will change.
@@
 Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
-Every node $u$ stores a model of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows about between members of $V$.
-Since each node has its own model and the model changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
+Every node $u$ stores a representation of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes node $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows between members of $V$.
+Since each node has its own network representation and it changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
 
@@
 
-At first, the node is collecting and organizing information.
+At first, the node collects and organizes information.
 A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
-There are several ways to perform such a test, but a study between these methods is out of the scope of this paper.
-Every time a neighboring node $v$ is tested as benign, the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
-After this, a new $T^u_i$ is formed, which is then used for the following steps.
+There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
+After this, a new $T^u_i$ is formed, which is used for the remaining steps.
 
@@
 After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
-Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on their components' classifications.
+Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.
 
@@
 The complexity of the graph merge algorithm is $O(|E|)$ for each neighbor a node has in an iteration; this number is at most $|V|$.
-The total complexity of Truman is therefore $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
-Since $|O'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
+The total complexity of TruMan is, therefore, $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
+Since $|E'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
 
@@
 \begin{enumerate}
-	\item Node $u$ checks who are its neighbors (nodes within its communication range).
-		  Newly discovered nodes and newly formed edges are added to $T^u_i$.
+	\item Node $u$ checks which are its neighbors (nodes within its communication range).
+		  New discovered nodes and new formed edges are added to $T^u_i$.
 		  Edges are created with weight 0.5.
-	\item Node $u$ tests all of its neighbors to discover which ones can be directly trusted or not.
+	\item Node $u$ tests all its neighbors to discover which ones can be directly trusted or not.
 		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
@@
 
-In order to test the Truman trust model, simulations were made using an implementation of the algorithm in Python.
-To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides strong similarity with vehicle movement in real life.
+In order to test the TruMan trust model, simulations were made using an implementation of the algorithm in Python.
+To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
 Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
@@
 \centering
-\begin{tabular}{|p{3cm}||p{3cm}|}
+\begin{tabular}{|p{3cm}||p{3.5cm}|}
  \hline
@@
  \hline
- Number of nodes 	& 150 (WDMM) \\
-  					& + 10 (random) \\
+ Number of nodes 	& 150 (WDMM) + 10 (random) \\
  \hline
@@
 The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
-For Truman, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
+For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
 It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$).
 
-The coverage of a single node is the circumference around it formed by the transmission radius; this is multiplied by the number of nodes in the network to estimate the maximum coverage area.
+The coverage of a single node is the circumference around it formed by the transmission radius.
+This is divided by two to compensate for overlapping circumferences, then multiplied by the number of nodes in the network to estimate the maximum coverage area.
 Finally, the value is divided by the total area of the environment.
@@
 
-Simulations shown here have densities that varying between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
-As a comparison, the density of São Paulo was calculated as 2.24 with $\rho = 10$m, a value much higher than what is necessary for a satisfactory performance of the algorithm.
+Simulations shown here have densities between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
+As a comparison, the density of the city of São Paulo (Brazil) was calculated as 2.24 with $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+
+\begin{figure*}[!t]
+\centering
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/10.png}%
+\label{subfig:random10}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}%
+\label{subfig:random30}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}%
+\label{subfig:random50}}
+\caption{Simulation with 10\% malicious nodes and $\rho$ = (a) 10m, (b) 30m or (c) 50m.}
+\label{fig:random103050}
+\end{figure*}
+
+\begin{figure*}[!t]
+\centering
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_1}%
+\label{subfig:random1}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_5}%
+\label{subfig:random2}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_10}%
+\label{subfig:random3}}
+
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
+\label{subfig:random4}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
+\label{subfig:random5}}
+\hfil
+\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
+\label{subfig:random6}}
+
+\caption{Simulation with $\rho$ = 10m and (a) 1\%, (b) 5\%, (c) 10\%, (d) 30\%, (e) 40\% or (f) 50\% malicious.}
+\label{fig:random0}
+\end{figure*}
 
@@
 \autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
-It is possible to see how the increase in the communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
+It is possible to see how the increase in the range allows the algorithm to converge sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
 
@@
 At 40\%, a small part of malicious nodes are yet to be detected.
-At 50\%, as expected, the results are inconsistent the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
-The amount of malicious nodes also affects convergence of the algorithm, since nodes do not trust information from malicious neighbors.
+At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
+The amount of malicious nodes also affects the convergence of the algorithm, since nodes do not trust information from malicious neighbors.
 
@@
 
-\begin{figure*}[!t]
-\centering
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/10.png}%
-\label{subfig:random10}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}%
-\label{subfig:random30}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}%
-\label{subfig:random50}}
-\caption{Simulation with random nodes, 10\% malicious and range of (a) 10m, (b) 30m and (c) 50m.}
-\label{fig:random103050}
-\end{figure*}
-
 %\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
@@
 % ----
-
-\begin{figure*}[!t]
-\centering
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_1}%
-\label{subfig:random1}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_5}%
-\label{subfig:random2}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_10}%
-\label{subfig:random3}}
-
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
-\label{subfig:random4}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
-\label{subfig:random5}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
-\label{subfig:random6}}
-
-\caption{Simulation with random nodes, range of 10m and (a) 1\%, (b) 5\%, (c) 10\%, (d) 30\%, (e) 40\% or (f) 50\% malicious.}
-\label{fig:random0}
-\end{figure*}
 
@@
 % ----
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{Network_rA7/10_10}
-\caption{Simulation with random nodes, range of 10m and 10\% malicious during 7 days.} \label{fig:random7}
-\end{figure}
 
@@
 This analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
-In the first part of this section, these properties are described with an assessment of whether or not Truman satisfies their conditions.
-Then, some of the most relevant models are described, considering how well they satisfy the desired properties and \autoref{table:properties} shows how they compare with Truman.
+In the first part of this section, these properties are described with an assessment of whether or not TruMan satisfies their conditions.
+Then, some of the most relevant models are described, considering how well they satisfy the desired properties.
+\autoref{table:properties} shows how they compare with TruMan.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{Network_rA7/10_10}
+\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+\end{figure}
 
@@
 \textit{Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes.
-Nodes may or not use information from other trustworthy nodes to build trust values.
-Truman satisfies this as it is built from the ground up for decentralized systems.
+Nodes may or may not use information from other trustworthy nodes to build trust values.
+TruMan satisfies this as it is built from the ground up for decentralized systems.
 
 \textit{Coping with sparsity}: the model still functions when there are few nodes populating the network.
-The experiments using low density values demonstrate that Truman works in reasonably sparse networks.
+The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
 Due to its decentralized nature, it can also work on isolated chunks of the network.
@@
 \textit{Event/task and location/time dynamics}: how the model reacts to different situations depending on what, where and when events happen.
-Although this has not been used in the simulations in this paper, Truman can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
+Although this has not been used in the simulations in this paper, TruMan can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
 
 \textit{Scalability}: the model can work on very large networks at high speeds.
-Due to the low complexity of the algorithms used in the model, Truman can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
+Due to the low complexity of the algorithms used in the model, TruMan can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
 It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
 
-\textit{Integrated confidence measure}: a value that allows nodes to estimate just how useful the output of the algorithm is.
-Since nodes using Truman store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
+\textit{Integrated confidence measure}: allows nodes to estimate how useful the output of the algorithm is.
+Since nodes using TruMan store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
 The closer it is to 1, the higher the chance that it is accurate.
@@
 \textit{System level security}: requires authentication of nodes participating in the network.
-This has not been considered in evaluations of Truman, however it can be included as a separate security model during the transmission of messages.
+This has not been considered in evaluations of TruMan.
+However, it can be included as a separate security model during the transmission of messages.
 
 \textit{Sensitivity to privacy concerns}: avoids eavesdropping and stalking by malicious nodes.
-Truman has not been designed with this in mind, but it doesn't inhibit privacy protection, however it does require that nodes cannot be completely anonymous.
+TruMan has not been designed with this in mind, but it does not inhibit privacy protection.
+However, it does require that nodes cannot be completely anonymous.
 
 \textit{Robustness}: the model's resistance to attacks.
-Truman satisfies this as malicious nodes are usually identified quickly and accurately, making it difficult for them to perform attacks.
+TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
 Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
 Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
-\textbf{Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
-( - CHEGUEI A ESTA CONCLUSÃO INTUITIVAMENTE, MAS NÃO MOSTRAMOS ISSO NOS EXPERIMENTOS.
-PRECISA?
-- ) }
-
-( - EU ADICIONEI CUSTO/EFICIÊNCIA COMO PROPRIEDADE NA TABELA; DEVO MENCIONAR AQUI? - )
+Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
 
@@
 It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-The usage of strongly connected components and coloring serves as basis for Truman, which is expanded to work on distributed and dynamic networks such as vehicular networks.
+The usage of strongly connected components and coloring serves as a basis for TruMan, which is expanded to work on distributed and dynamic networks such as vehicular networks.
 
@@
 Additionally, messages have higher reliability when their senders are closer in time and space to the reported event.
-When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
+When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes mean a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
 %The model considers both role-based trust and experience-based trust; although the work proposed here does not use role-based trust, the authors provide a useful method of calculating and updating an experience-based trust value, which might be used or adapted.
-However, the model relies only on direct interaction between pairs of nodes, so no form of indirect trust (that is, trust values received from other nodes) is considered.
+However, the model relies only on direct interaction between pairs of nodes, so no form of indirect trust is considered.
 
 In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
-By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
-When a node sends a message, one node in the cluster (the leader) must aggregate the other nodes' opinions on that message.
-Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; additionally, nodes that receive the message only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
+By separating nodes into clusters with their geographical neighbors, it is possible to distribute the evaluation of messages using previously formed opinions.
+When a node sends a message, the cluster-leader must aggregate the other nodes' opinions on that message.
+Messages are only forwarded to other clusters if the aggregate opinion is above a certain threshold.
+Additionally, nodes that receive the message only act if the overall trust on it is above another threshold.
 However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
-Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy.
+Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster become untrustworthy.
 
-The ART model proposed in \cite{li2016art} is similar to Truman in the sense that there are two main steps: data gathering and malicious node detection.
+The ART model proposed in \cite{li2016art} works in two main steps: data gathering and malicious node detection.
 It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
 Then, it uses a Cosine-based metric to compare two nodes' trust vectors (a series of opinions regarding other nodes).
-However, these steps require several intensive calculations, which greatly increases the complexity of the algorithm.
-The authors present no details on how the model deals with sparsity, dynamics, scalability, security and privacy.
+However, these steps require several intensive calculations, which greatly increase the complexity of the algorithm.
+The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
 
 The authors of \cite{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
-This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations when cellular communication is slow or unreliable.
+This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations in which mobile communication is slow or unreliable.
 It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
@@
 Finally, it is worth noting that, aside from \cite{vernize2015malicious}, none of the mentioned related work presents complexity calculations for their algorithm.
-Considering the scale of the problem, Truman's cost of $O(|V|\times |E|)$ is very low without sacrificing completeness and correctness.
+Considering the scale of the problem, TruMan's cost of $O(|V|\times |E|)$ is very low without sacrificing completeness and correctness.
 The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 \caption{Properties of Truman and related work}
@@
 \centering
-\begin{tabular}{|p{1.5cm}||p{0.8cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
+\begin{tabular}{|p{1.5cm}||p{1.1cm}|p{1.1cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
  \hline
@@
  \hline
- Security 		& \checkmark & - 		  & \checkmark & \checkmark & - 		 & \checkmark \\
+ Security 		& -			 & - 		  & \checkmark & \checkmark & - 		 & \checkmark \\
  \hline
@@
  \hline
-% Cost & O(|V|* |E|) & O(|V|+|E|) & n/a & n/a & n/a \\
+ Cost & $|V|*|E|$ & $|V|+|E|$ & n/a & n/a & n/a & n/a\\
+ \hline
 % \hline
@@
 
-While malicious nodes can diminish the usefulness of vehicular networks, the concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
+The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
 In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
-
-As nodes travel across the network and collect more data from neighbor, they are able to form an abstraction of the network which can be used to detect malicious nodes.
+As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
 By placing nodes into strongly connected components, a network containing a large amount of node can be simplified into a much smaller one.
-And, using a simple graph coloring algorithm, most malicious nodes stand out by having colors different than the majority of nodes.
+Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
 This allows for a low complexity approach to malicious node identification in a dynamic network.
 
-The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a small amount of false positives.
+The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a very tiny amount of false positives.
 As the network changes in shape, nodes acquire more information and are able to make even more accurate classifications of malicious nodes around them.
+Future work includes the extension of the proposed model to use V2I communications.
 %Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.

### 2018-02-20 dissertation/3-vanets/chapter3.tex
article and dissertation updates
final version sent for review

--- 2017-06-01 qualification/3-vanets/chapter3.tex
@@
 %[details of security and safety issues in VANETs that can be diminished by trust]
+\section{Desired properties for VANET trust models}
+\label{section:properties}
+
+The analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
+In this section, these properties are described with an assessment of whether or not TruMan satisfies their conditions.
+\autoref{table:properties} shows how well TruMan and previous work satisfy the properties.
+
+\textit{1. Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes.
+Nodes may or may not use information from other trustworthy nodes to build trust values.
+TruMan satisfies this as it is built from the ground up for decentralized systems.
+
+\textit{2. Coping with sparsity}: the model still functions when there are few nodes populating the network.
+The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
+Due to its decentralized nature, it can also work on isolated chunks of the network.
+
+\textit{3. Event/task and location/time dynamics}: how the model reacts to different situations depending on what, where and when events happen.
+Although this has not been used in the simulations in this paper, TruMan can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
+
+\textit{4. Scalability}: the model can work on very large networks at high speeds.
+Due to the low complexity of the algorithms used in the model, TruMan can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
+It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
+
+\textit{5. Integrated confidence measure}: allows nodes to estimate how useful the output of the algorithm is.
+Since nodes using TruMan store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
+The closer it is to 1, the higher the chance that it is accurate.
+
+\textit{6. System level security}: requires authentication of nodes participating in the network.
+This has not been considered in evaluations of TruMan.
+However, it can be included as a separate security model during the transmission of messages.
+
+\textit{7. Sensitivity to privacy concerns}: avoids eavesdropping and stalking by malicious nodes.
+TruMan has not been designed with this in mind, but it does not inhibit privacy protection.
+However, it does require that nodes cannot be completely anonymous.
+
+\textit{8. Robustness}: the model's resistance to attacks.
+TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
+Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
+Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
+Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified. 
+%Furthermore, the analysis also takes into account the complexity of the algorithm necessary to implement the trust model, because a high priority of TruMan is the
+
+\begin{table}[t!]
+\caption{Properties of Truman and related work}
+\label{table:properties}
+\centering
+\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
+ \hline
+ \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ TruMan & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{dotzer2005vars} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{minhas2010towards} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{chen2010trust} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{li2016art} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \cite{chen2017cloud} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+\end{tabular}
+\end{table}
+
+Finally, it is worth noting that, considering the scale of the problem, TruMan's cost is very low without sacrificing completeness and correctness.
+The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
 
@@
 That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
-This process adds credibility to a message through validation by nodes in a non-centralized fashion.
+This process adds credibility to a message through validation by nodes in a decentralized fashion.
 It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
@@
 
+The ART model proposed in \cite{li2016art} works in two main steps: data gathering and malicious node detection.
+It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
+Then, it uses a Cosine-based metric to compare two nodes' trust vectors (a series of opinions regarding other nodes).
+However, these steps require several intensive calculations, which greatly increase the complexity of the algorithm.
+The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
+
+The authors of \cite{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
+This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations in which mobile communication is slow or unreliable.
+It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
+
+%\begin{table}[t!]
+%\caption{Properties of Truman and related work}
+%\label{table:properties}
+%\centering
+%\begin{tabular}{|p{3cm}||p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
+% \hline
+% \textbf{Property} & TruMan & \cite{vernize2015malicious} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
+% \hline
+% Decentralized 	& \checkmark & -		  & \checkmark & \checkmark & \checkmark & - \\
+% \hline
+% Sparsity 		& \checkmark & -		  & \checkmark & \checkmark & -			 & \checkmark \\
+% \hline
+% Dynamics 		& \checkmark & - 		  & \checkmark & \checkmark & -			 & - \\
+% \hline
+% Scalability 	& \checkmark & \checkmark & \checkmark & \checkmark & -			 & \checkmark \\
+% \hline
+% Confidence 	& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - \\
+% \hline
+% Security 		& -			 & - 		  & \checkmark & \checkmark & - 		 & \checkmark \\
+% \hline
+% Privacy 		& - 		 & - 		  & \checkmark & - 			& -			 & \checkmark \\
+% \hline
+% Robustness 	& \checkmark & -		  & -		   & - 			& \checkmark & - \\
+% \hline
+% Efficiency		& \checkmark & \checkmark & -		   & -			& -			 & - \\
+% \hline
+% Cost & $|V|*|E|$ & $|V|+|E|$ & n/a & n/a & n/a & n/a\\
+% \hline
+%% \hline
+%\end{tabular}
+%\end{table}
+
 %\begin{table}[h!]

### 2018-02-20 dissertation/4-project/chapter4.tex
final version sent for review

--- 2017-12-01 dissertation/4-project/chapter4.tex
@@
-\chapter{Project}
-\label{chap:project}
+\chapter{TruMan}
+\label{chap:truman}
 
@@
 \label{section:tarjan}
-An important aspect of \cite{vernize2015malicious} is the use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth}.
-This allows a large graph to be abstracted into a smaller graph, which therefore reduces the input for further algorithms.
-Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any $u, v \in O$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
+The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of TruMan's efficiency.
+This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
+Given a directed graph $T = (V,E)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there is a path from $u$ to $v$ and a path from $v$ to $u$.
+For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
 Every node of the input graph $T$ must belong to a component.
 
-The algorithm works by performing a depth-first search, adding nodes to $stack$ as they are visited.
-If two nodes are present on $stack$, then there is a path from the first node to the second one (in the order they were added to the stack).
+The algorithm works by performing a depth-first search, adding nodes to a stack as they are visited.
+If two nodes are present on the stack, then there is a path from the first node to the second one (in the order they were added to the stack).
+
 Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.
+In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessed from every call of the function.
+$index$ and $lowlink$ are arrays indexed by unique node identifiers, $count$ is an integer and $stack$ is a last-in-first-out data structure.
 
-In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessed from every call of the function.
-$index$ and $lowlink$ are arrays indexed by node IDs, $count$ is an integer and $stack$ is a last-in-first-out data structure.
-
-In the call that visits a node $u$, the algorithm must loop through each node $v$ for which $u \rightarrow v$ exists.
+In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
 If node $v$ has not yet been visited, the algorithm is called for $v$.
@@
 After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
-Therefore, nodes must be popped from the $stack$ until $u$ is found.
+Therefore, nodes must be popped from the stack until $u$ is found.
 Each node popped, including $u$, is a member of a strongly connected component.
 
-The number of components is, at most, $|v|$.
-In a worst-case scenario, each node is put into its own component; however, this would not be the case for most useful input graphs.
+The number of components is, at most, $|V|$: in a worst-case scenario, each node is placed into its own component.
+The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
 
-Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth} \cite{vernize2013dissertation}.
-The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
+Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
 
@@
 \label{section:coloring}
-The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
-Graph coloring is one of the heuristics used in \cite{vernize2013dissertation} to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
-Since it was the heuristic that presented the best results, it has been chosen as the heuristic for this current study as well.
+%The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
+Graph coloring is one of the possible heuristics used by TruMan to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
+Out of the tested heuristics, it presents the best results, so it has been chosen as the heuristic for the trust model.
 
@@
 
-In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
+In \cite{mittal2011graph}, the authors present an efficient approach to graph coloring using the minimum possible amount of colors.
 Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
-The complexity of the algorithm is $O(|E|)$ for an undirected graph $G = (V,E)$.
-As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
 For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
 
-Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph} \cite{vernize2013dissertation}.
+The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
+As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
+
+Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph}.
 
@@
 
-\section{SNAP library}
-\label{section:snap}
-In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
-The Stanford Network Analysis Platform (SNAP) library \cite{snap} was chosen primarily because it is memory efficient.
-The simulations require multiple graphs that share the same set of nodes (because each node in the network has its own knowledge of the surrounding network), and the SNAP library uses pointers to nodes and edges, saving memory by not having to duplicate the entire data structure.
-It is written in C++, but, for these simulations, the Snap.py Python library was used.
-
-The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
-
-\section{Trust management algorithm}
+\section{The TruMan algorithm}
 \label{section:algorithm}
 
-The MaNI algorithm works well and is efficient for static graphs.
-However, some important modifications had to be made to accommodate dynamic graphs.
-The main changes made are described below.
+TruMan is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm.
+However, MaNI was developed for static networks such as social networks, and is executed by an external supervising agent (i.e. outside of the network), making it unsuitable for a vehicular network.
 
-\subsection{Graph knowledge and graph building}
-First of all, there is an important change to the graph knowledge used to run the algorithm.
-The complete trust graph is a parameter to the MaNI algorithm, which executes with complete knowledge of the graph.
+In order to work with dynamic networks, the TruMan algorithm runs iterations at predetermined intervals.
+%As nodes acquire more information, their model of the network will change.
+Furthermore, the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
+Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
+Every node $u$ stores a representation of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes node $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows between members of $V$.
+Since each node has its own network representation and it changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
 
-Given the complete topology graph $G = (V, E)$, each node $u \in V$ has its own vision of the network and $u$ has to build a graph, $G_u$, to represent its surrounding network.
-In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
-For every neighbor $v \in V$, $G_u$ must me merged with $G_v$.
-In sufficient iterations, $G_u$ forms a complete or nearly complete vision of the network.
+At first, the node collects and organizes information.
+A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
+There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
+After this, a new $T^u_i$ is formed, which is used for the remaining steps.
 
-Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
-Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
-When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
-If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.
+After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}.
+For each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
+In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
+Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, E'^u_i)$.
+
+The creation of $C^u_i$ simplifies the remaining computation.
+Since each vertex $v' \in V'^u_i$ is a component of $T^u_i$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as the same.
+They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
+After the formation of $C^u_i$, a heuristic is used to classify the nodes as benign or malicious.
+
+The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
+After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
+Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.
+
+The complexity of the whole algorithm can be calculated by adding the most costly operations involved.
+As discussed above, Tarjan's algorithm has a complexity of $O(|V|+|E|)$ (for the trust graph $T$), while the graph coloring algorithm has a complexity of $O(|E'|)$ (for the component graph $C$).
+The most costly part of the algorithm is the graph merge operation that happens between trusted nodes.
+The complexity of the graph merge algorithm is $O(|E|)$ for each neighbor a node has in an iteration; this number is at most $|V|$.
+The total complexity of TruMan is, therefore, $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
+Since $|E'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
+
+In summary, every node $u$ runs the following steps in each iteration to detect malicious nodes in the network:
+
+\begin{enumerate}
+	\item Node $u$ checks which are its neighbors (nodes within its communication range).
+		  New discovered nodes and new formed edges are added to $T^u_i$.
+		  Edges are created with weight 0.5.
+	\item Node $u$ tests all its neighbors to discover which ones can be directly trusted or not.
+		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
+	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$.
+	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
+	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are classified as benign or malicious.
+\end{enumerate}
+
+%The MaNI algorithm works well and is efficient for static graphs.
+%However, some important modifications had to be made to accommodate dynamic graphs.
+%The main changes made are described below.
+%
+%\subsection{Graph knowledge and graph building}
+%First of all, there is an important change to the graph knowledge used to run the algorithm.
+%The complete trust graph is a parameter to the MaNI algorithm, which executes with complete knowledge of the graph.
+%
+%Given the complete topology graph $G = (V, E)$, each node $u \in V$ has its own vision of the network and $u$ has to build a graph, $G_u$, to represent its surrounding network.
+%In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
+%For every neighbor $v \in V$, $G_u$ must me merged with $G_v$.
+%In sufficient iterations, $G_u$ forms a complete or nearly complete vision of the network.
+%
+%Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
+%Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
+%When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
+%If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.
+%
+%[Illustrate graph building here]
+%
+%Furthermore, this implies that the malicious node detection algorithm runs using incomplete graph knowledge.
+%This is expected and generally unavoidable in vehicular networks, which have high mobility and can be extremely large.
+%
+%
+%
+%-\\
+%----------\\
+%This paragraph is probably pertinent to delimiting the size of graphs:\\
+%The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
+%Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
+%That is, data collected by the node's own sensors have the highest quality, which diminishes as information starts to come from neighbors, neighbors of neighbors and so forth.
+%They adopt a broadcast model in which RSUs send data about its surroundings and OBUs capture the information.
+%The trust model proposed here might use a similar broadcast model, although adapted so all nodes are vehicles (OBUs).\\
+%----------\\
 

### 2018-02-20 dissertation/5-simulations/chapter5.tex
final version sent for review

@@
 
-The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
+%The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
 
@@
 
+In order to test the TruMan trust model, simulations were made using an implementation of the algorithm in Python.
+To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
+Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
+
+\begin{table}[h!]
+\caption{Simulation parameters}
+\label{table:parameters}
+\centering
+\begin{tabular}{|p{5cm}||p{5cm}|}
+ \hline
+ \textbf{Parameter}	& \textbf{Value} \\
+ \hline
+ \hline
+ Duration 			& 86400 seconds \\
+ \hline
+ Work day length 	& 28800 seconds \\
+ \hline
+ Std. dev. departure time & 7200 seconds \\
+ \hline
+ Node velocity 		& 7~10m/s \\
+ \hline
+ Simulation area	& approximately 14km\textsuperscript{2} \\
+ \hline
+ Number of nodes 	& 150 (WDMM) + 10 (random) \\
+ \hline
+% Cost & O(|V|* |E|) & O(|V|+|E|) & n/a & n/a & n/a \\
+% \hline
+\end{tabular}
+\end{table}
+
+Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}.
+Different parameters are shown in \autoref{table:parameters}.
+The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
+Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
+%The number of nodes and their communication range vary for different simulations.
+There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
+Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
+The parameters regarding offices, meeting spots and shopping were kept intact.
+A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
+
+\subsection{Network Density}
+
+%==> PRECISA MOSTRAR A EQUAÇÃO DE CALCULO DA DENSIDADE, SENÃO FICA DIFICIL DE ENTENDER O CALCULO.
+
+The communication range of nodes varies from 10m to 50m, to illustrate the impact of different network densities.
+The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
+For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
+It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$).
+
+The coverage of a single node is the circumference around it formed by the transmission radius.
+This is divided by two to compensate for overlapping circumferences, then multiplied by the number of nodes in the network to estimate the maximum coverage area.
+Finally, the value is divided by the total area of the environment.
+The network density formula is as follows:
+
+$$ \delta = \frac{\frac{\rho^2\pi}{2} \times \eta}{\alpha} $$
+
+Simulations shown here have densities between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
+As a comparison, the density of the city of São Paulo (Brazil) was calculated as 2.24 with $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+
 \subsection{Validation}
@@
 \label{section:results}
+
+To improve readability, all figures in this section follow the same format:
+$X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
+$Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
+blue line represents the percentage of nodes detected out of the complete network;
+magenta is the percentage of malicious nodes in the network (ground truth);
+green represents the nodes correctly identified as malicious (true positives);
+cyan represents the undetected malicious nodes (false negatives); 
+red represents the benign nodes incorrectly identified as malicious (false positives).
+
+\autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
+It is possible to see how the increase in the range allows the algorithm to converge sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
+
+\autoref{fig:random0} shows the variation of results for different amounts of malicious nodes in the network.
+By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
+At 40\%, a small part of malicious nodes are yet to be detected.
+At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
+The amount of malicious nodes also affects the convergence of the algorithm, since nodes do not trust information from malicious neighbors.
+
+\autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
+Most malicious nodes are identified by the end of the first day; in the following iterations, the algorithm finishes building the network model and sorts out remaining false negative or false positive results.
+After iteration 20000, the results are completely consistent.
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10.png}
+\caption{$\rho$ = 10m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30.png}
+\caption{$\rho$ = 30m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50.png}
+\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+\label{fig:random103050}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
+\caption{1\% malicious.}
+\end{subfigure}
+\hspace*{0.1cm} % separation between the subfigures
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
+\caption{5\% malicious.}
+\end{subfigure}
+
+\vspace{1cm}
+
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
+\caption{10\% malicious.}
+\end{subfigure}
+\hspace*{0.1cm} % separation between the subfigures
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
+\caption{30\% malicious.}
+\end{subfigure}
+
+\vspace{1cm}
+
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
+\caption{40\% malicious.}
+\end{subfigure}
+\hspace*{0.1cm} % separation between the subfigures
+\begin{subfigure}{0.45\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
+\caption{50\% malicious.}
+\end{subfigure}
+
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes.}\label{fig:tarjan}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{images/plots/Network_rA7/10_10}
+\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+\end{figure}
+
+%\begin{figure}
+%\centering
+%
+%\begin{subfigure}{0.4\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
+%\caption{1\% malicious.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\hspace*{1cm}
+%
+%\begin{subfigure}{0.4\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
+%\caption{5\% malicious.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\vspace{1cm}
+%
+%\begin{subfigure}{0.5\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10}
+%\caption{10\% malicious.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\hspace*{\fill}
+%
+%\begin{subfigure}{0.5\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10}
+%\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\vspace{1cm}
+%
+%\begin{subfigure}{0.5\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10}
+%\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\hspace*{\fill}
+%
+%\begin{subfigure}{0.5\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10}
+%\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+%\end{subfigure}
+%
+%\caption{Simulation with $\rho$ = 10m varying malicious.}
+%\label{fig:random0}
+%\end{figure}
+
+%\begin{figure*}[!t]
+%\centering
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_1}%
+%\label{subfig:random1}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_5}%
+%\label{subfig:random2}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_10}%
+%\label{subfig:random3}}
+%
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_30}%
+%\label{subfig:random4}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_40}%
+%\label{subfig:random5}}
+%\hfil
+%\subfloat[]{\includegraphics[width=0.329\linewidth]{images/plots/Network_rA/10_50}%
+%\label{subfig:random6}}
+%
+%\caption{Simulation with $\rho$ = 10m and (a) 1\%, (b) 5\%, (c) 10\%, (d) 30\%, (e) 40\% or (f) 50\% malicious.}
+%\label{fig:random0}
+%\end{figure*}

### 2018-02-20 dissertation/6-conclusao/chapter6.tex
final version sent for review

--- 2018-02-20 article/article.tex
@@
-\documentclass[conference]{IEEEtran}
-
-\usepackage[utf8]{inputenc}	% arquivos LaTeX em Unicode (UTF8)
-\RequirePackage[plainpages,pdfpagelabels]{hyperref}	% PDF com links, metadados
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  \usepackage[pdftex]{graphicx}
-  \graphicspath{{../pdf/}{../jpeg/}{./plots/}}
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-\fi
-
-\usepackage{cleveref}
-\usepackage{subfig}
-
-\usepackage{tikz}
-\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
-\def\scalecheck{\resizebox{\widthof{\checkmark}*\ratio{\widthof{x}}{\widthof{\normalsize x}}}{!}{\checkmark}}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\floatname{algorithm}{Algorithm}
-\renewcommand{\algorithmiccomment}[1]{~~~// #1}
-
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-
-\title{TruMan: Trust Management for Vehicular Networks}
-
-%\author{\IEEEauthorblockN{Renan Greca}
-%\IEEEauthorblockA{Departamento de Informática\\
-%Universidade Federal do Paraná\\
-%Curitiba, Brazil\\
-%Email: rdmgreca@inf.ufpr.br}
-%\and
-%\IEEEauthorblockN{Luiz Carlos P. Albini}
-%\IEEEauthorblockA{Departamento de Informática\\
-%Universidade Federal do Paraná\\
-%Curitiba, Brazil\\
-%Email: albini@inf.ufpr.br}}
-
-\author{\IEEEauthorblockN{Renan Greca and Luiz Carlos Pessoa Albini}
-        \IEEEauthorblockA{Department of Informatics -- Federal University of Paraná (UFPR) -- Curitiba, Brazil\\
-                          Email: rdmgreca@inf.ufpr.br, albini@inf.ufpr.br}
-}
-
-\maketitle
-
-\begin{abstract}
-By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular ad-hoc network (VANET), in which cars share data amongst themselves in order to cooperate and make roads safer and more efficient.
-A decentralized ad-hoc solution, which does not rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
-However, as it is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
-In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
-To generate these trust values, nodes use information acquired from past interactions.
-Nodes which frequently share false or irrelevant data must have lower trust values than the ones which appear to be reliable.
-This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
-The results prove to be accurate and efficient, thanks to the low complexity of the algorithm constituting the trust model.
-\end{abstract}
-
-%\begin{IEEEkeywords}
-%Vehicular Network; Trust Management; V2V Communication;
-%\end{IEEEkeywords}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-\label{section:introduction}
-
-Within the next few years, a substantial share of new vehicles will come equipped with networking features \cite{connectedcar2016}.
-These features will allow vehicles to quickly share data with other nearby devices and can be useful tools to reduce traffic and the risk of accidents.
-Over one million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
-%Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
-%In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
-By quickly sharing data with neighboring vehicles without the need of an Internet connection, smart vehicles can alert drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
-
-The communication standard for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
-It describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
-Communication between two OBUs is called vehicle-to-vehicle (V2V) communication, while communication between an OBU and an RSU is called vehicle-to-infrastructure (V2I) communication.
-This study focuses only on V2V scenarios, and therefore, any references to Vehicular ad-hoc networks (VANETs) and their nodes refer exclusively to vehicles with on-board units.
-
-As expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
-Some issues that could be exploited in such network include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming or stalking \cite{isaac2010security}.
-
-Each of these problems require specific solutions, although there are ways of making the network safer in general.
-One way is taking advantage of the concept of trust between network members.
-By having nodes remembering previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
-Trust solutions for VANETs are generally classified into data-oriented, emphasizing the message contents, or entity-based, emphasizing message senders.
-
-This work proposes a new and improved entity-based trust model to compute the trust between any pair of nodes in a vehicular network.
-Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes are likely malicious.
-Since the network is dynamic, nodes acquire more knowledge as time passes and the results of the algorithm become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between frequently connected nodes.
-Simulations demonstrate that nodes are able to correctly identify more than 95\% of nodes in the network (malicious or not).
-The algorithm correctness depends upon the velocity of the nodes, the frequency of the information exchanged between them and the running time, i.e. the longevity, of the algorithm.
-%After running a series of simulations, the results were very promising; after one simulated day,
-
-%==> O QUE SE PRETENDE ATINGIR? O QUE TEM DE DIFERENTE? QUAIS RESULTADOS FORAM OBTIDOS? FALTA UM PARAGRAFO AQUI FALANDO DISSO.
-
-The remainder of this paper is organized as follows.
-Section \ref{section:algorithm} details the proposed trust model;
-Section \ref{section:results} contains the simulation results which demonstrate the usefulness of the model;
-Section \ref{section:previouswork} presents the previously published trust models for vehicular networks and compares them with TruMan;
-and Section \ref{section:conclusion} has the conclusion and future work directions.
-
-\section{TruMan}
-\label{section:algorithm}
-
-%==> MUDAR O TITULO, PRACISA DIZER ALGO DO TIPO UNDERNEATH ALGORITHM
-
-The objective of the TruMan trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
-The algorithm that dictates the trust model runs continuously, with iterations happening in preset intervals.
-In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
-This information is maintained in a directed graph $T=(V,E)$, in which $V$ is the set of known vertices and $E$ is the set of known trust relationships between pairs $u,v \in V$.
-Graph $T$ is known as the \textit{trust graph}.
-
-Each edge $(u, v) \in E$ stores a value between 0 and 1 which represents the degree of trust $u$ has for $v$.
-This value is called the \textit{opinion} of $u$ about $v$.
-At the start of the execution, edges have value $0.5$; this value increases when nodes have positive interactions and decreases otherwise.
-A threshold $0 < h < 1$ is used to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
-%By default, the threshold is $0.5$.
-
-After collecting information from other nodes, TruMan performs two steps: ($i$) it divides the network graph into strongly connected components using Tarjan's algorithm; ($ii$) it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.
-
-%In order to do this, a node must maintain a graph $T=(V,O)$ that represents its knowledge of the network, in which $V$ is the set of known nodes and $O$ is the set of opinions nodes have about others.
-%Opinions are directed edges, each one with a weight between 0 and 1 that represents the degree of trust from one node towards another.
-%A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
-
-The detailed descriptions of both algorithms are below, followed by the complete process of each iteration of the Truman trust model.
-
-%==> VOCE AINDA NAO FALOU DOS PESOS DAS ARESTAS E COMO ELES FORAM CONSEGUIDOS
-%==> VOCE NAO FALOU DOS GRAFOS
-
-% ---------------------------------------------------------------------------------------------------
-\subsection{Tarjan's strongly connected components algorithm}
-\label{section:tarjan}
-The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of TruMan's efficiency.
-This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
-Given a directed graph $T = (V,E)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there is a path from $u$ to $v$ and a path from $v$ to $u$.
-For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
-Every node of the input graph $T$ must belong to a component.
-
-%-----ACHO QUE ESTA DEFINIÇAO ESTA COM MUITOS DETALHES DE IMPLEMENTACAO ---
-%-----PODERIA DEIXAR ELA MAIS ALTO NIVEL, EM FORMA DE CONJUNTOS POR EXEMPLO ---
-
-%Three data structures are used by the algorithm:
-%$index$ and $lowlink$ are indexed by node IDs (predefined unique identifiers) and $stack$ is a last-in-first-out stack.
-%The $count$ variable is incremented at each visited node.
-%In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessible from every call of the function.
-%$index$ and $lowlink$ are arrays indexed by node IDs (predefined unique identifiers), $count$ is an integer and $stack$ is a last-in-first-out data structure.
-
-%The algorithm works by performing a depth-first search, adding nodes to a stack as they are visited.
-%If two nodes are present on the stack, then there is a path from the first node to the second one (in the order they were added to the stack).
-%Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.
-%
-%In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
-%If node $v$ has not yet been visited, the algorithm is called for  $v$.
-%The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
-%After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
-%Therefore, nodes must be popped from the stack until $u$ is found.
-%Each node popped, including $u$, is a member of a strongly connected component.
-
-%---------------
-
-The number of components is, at most, $|V|$: in a worst-case scenario, each node is placed into its own component.
-The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
-%; however, this would not be the case for most useful input graphs.   ---> NÃO ENTENDI <---
-
-%Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
-
-From the output of Tarjan's algorithm, an undirected component graph $C = (V',E')$ is formed.
-Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $e' \in E'$ are edges from $T$ between nodes that do not belong in the same component (e.g.
-if $u$ and $v$ are separated into $u'$ and $v'$, $e = (u,v)$ becomes $e'=(u',v')$.
-
-%\begin{algorithm}
-%\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
-%\begin{algorithmic}[1]
-%
-%\Function{Tarjan}{vertex $u$}
-%
-%\State $index[u] = count$
-%\State $lowlink[u] = count$
-%\State $count \gets count + 1$
-%\State \textbf{push} $u$ \textbf{to} $stack$
-%
-%\For{$v$ \textbf{in} neighbors of $u$ }
-%	\If{weight of $u \rightarrow v$ $<$ $h$}
-%		\State \textbf{continue}
-%	\EndIf	
-%		
-%	\If{$index[v] = -1$}
-%		\Comment{\emph{v has not been visited yet}}
-%		\State \textbf{Tarjan}($v$)
-%	\EndIf
-%	
-%	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
-%\EndFor
-%
-%\If{$lowlink[u] = index[u]$}
-%	\Repeat
-%		\Comment{\emph{unstack nodes until u is found}}
-%		\State \textbf{pop} $w$ \textbf{from} $stack$
-%		\State \textbf{add} $w$ \textbf{to} $component$
-%	\Until {$w = u$}
-%\EndIf
-%
-%\EndFunction
-%\end{algorithmic}
-%\end{algorithm}
-
-% ---------------------------------------------------------------------------------------------------
-\subsection{Graph coloring with minimum colors}
-\label{section:coloring}
-The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring.
-Graph coloring is one of the possible heuristics used to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
-Out of the tested heuristics, it presents the best results, so it has been chosen as the heuristic for the trust model.
-
-%The process of graph coloring consists of labeling each node with a color so that no two neighboring nodes share the same color.
-%This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
-It has been mathematically proven that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the chromatic number of the graph) is an NP-hard problem \cite{sanchez1989determining}.
-In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
-Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
-The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
-As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V'|^2)$ \cite{brelaz1979new}.
-%For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
-
-%Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph}.
-%
-%A limitation of this algorithm is that the edges must be sorted according to node indexes beforehand.
-%It does not matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
-%This is demonstrated in \cite{vernize2013dissertation}.
-
-%\begin{algorithm}
-%\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
-%\begin{algorithmic}[1]
-%
-%\Function{Coloring}{graph $G$}
-%
-%\State \textbf{color} all nodes of $G$ \textbf{with} 0
-%\State $d \rightarrow 0$
-%\For{$e=(u,v)$ \textbf{in} edges of $G$}
-%	\If{$u$ and $v$ have the same color}
-%		\If{$color[v] = d$}
-%			\State $d \rightarrow d+1$
-%		\EndIf
-%		\State $color[v] \rightarrow d$
-%	\EndIf
-%\EndFor
-%
-%
-%\EndFunction
-%\end{algorithmic}
-%\end{algorithm}
-
-% ---------------------------------------------------------------------------------------------------
-\subsection{The TruMan algorithm}
-\label{section:trustmanagement}
-%\subsection{Malicious Node Identification Algorithm}
-%\label{section:mani}
-
-%The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
-%The authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-%The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-%Furthermore, the algorithm is executed by a global observer which has information about the complete network.
-
-%==> ALGUMAS PARTES ESTÃO MUITO REPETITIVAS COM A SEÇAO DE CIMA. PRECISA REVER PARA NÃO FICAR MUITO REPETITIVO. <==
-
-TruMan is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm.
-However, MaNI was developed for static networks such as social networks, and is executed by an external supervising agent (i.e. outside of the network), making it unsuitable for a vehicular network.
-%The work describes here takes advantage of the efficiency of the two graph theory algorithms into a model that is viable for highly dynamic networks.
-
-%In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
-%The algorithm runs continuously, running new iterations at predetermined intervals.
-%Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
-%A snapshot $S_i$ represents the complete topology of the network at the iteration $i$.
-%%; $V_i$ is the set of nodes participating in the network at that time and $E_i$ is the set of edges connecting pairs of nodes within communication range of each other.
-%As the network is dynamic, it is expected that each $S_i$ is different from $S_{i-1}$, but also that the changes follow a determined mobility model.
-
-In order to work with dynamic networks, the TruMan algorithm runs iterations at predetermined intervals.
-%As nodes acquire more information, their model of the network will change.
-Furthermore, the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
-Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
-Every node $u$ stores a representation of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes node $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows between members of $V$.
-Since each node has its own network representation and it changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
-
-%==> VOCE NAO FALOU DOS PESOS DAS ARESTAS AINDA
-
-%==> ACHO QUE PRECISA FALAR AQUI DA PRINCIPAL DIFERENCA PARA O MaNI, QUE AQUI ESTAMOS FALANDO DE REDE MOVEL E ALGORITMO DISTRIBUIDO
-
-At first, the node collects and organizes information.
-A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
-There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
-Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
-After this, a new $T^u_i$ is formed, which is used for the remaining steps.
-
-%Each edge $w \rightarrow v \in E^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
-%The closer the value is to 1, the more $w$ trusts $v$.
-%These values are not mutual, so the value of $w \rightarrow v$ can be different from the value of $v \rightarrow w$.
-
-After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}.
-For each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
-In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
-Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, E'^u_i)$.
-
-The creation of $C^u_i$ simplifies the remaining computation.
-Since each vertex $v' \in V'^u_i$ is a component of $T^u_i$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as the same.
-They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
-After the formation of $C^u_i$, a heuristic is used to classify the nodes as benign or malicious.
-
-%==> PRIMEIRA FRASE ESTA ESTRANHA <==
-The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
-After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
-Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.
-
-%In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
-%Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
-%Therefore, for the purposes of this research, only the coloring heuristic is considered.
-
-The complexity of the whole algorithm can be calculated by adding the most costly operations involved.
-As discussed above, Tarjan's algorithm has a complexity of $O(|V|+|E|)$ (for the trust graph $T$), while the graph coloring algorithm has a complexity of $O(|E'|)$ (for the component graph $C$).
-The most costly part of the algorithm is the graph merge operation that happens between trusted nodes.
-The complexity of the graph merge algorithm is $O(|E|)$ for each neighbor a node has in an iteration; this number is at most $|V|$.
-The total complexity of TruMan is, therefore, $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
-Since $|E'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
-
-In summary, every node $u$ runs the following steps in each iteration to detect malicious nodes in the network:
-
-\begin{enumerate}
-	\item Node $u$ checks which are its neighbors (nodes within its communication range).
-		  New discovered nodes and new formed edges are added to $T^u_i$.
-		  Edges are created with weight 0.5.
-	\item Node $u$ tests all its neighbors to discover which ones can be directly trusted or not.
-		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
-	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$.
-	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
-	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are classified as benign or malicious. %according to the same rules as the MaNI algorithm.
-\end{enumerate}
-
-%Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
-%In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
-%Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
-%
-%The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
-%By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
-
-% ---------------------------------------------------------------------------------------------------
-
-%The MaNI algorithm works well and is efficient for static graphs.
-%However, some important modifications had to be made to accommodate dynamic graphs in a decentralized fashion.
-%The main changes made are described below.
-
-%First of all,
-
-%Another important change is on the trust values themselves.
-%MaNI uses binary trust (either 0 or 1).
-%This has been replaced by a floating-point number between 0 and 1; a predefined threshold defines which values are deemed trustworthy or not.
-%This was not the case in the first versions of the trust model; it was added in order to avoid abrupt changes to opinions.
-%Since the network is different on each iteration of the algorithm, there were instances in which a change to the topology caused a radical change to the output of the graph coloring algorithm.
-%Using a trust value that gradually changes over time, output variations between two iterations became small and gradual.
-%
-%At the start of an execution, a node knows only about itself.
-%From that point, iterations of the algorithm will run in a set interval of $l$ seconds.
-%In each iteration (labeled $t$ after the timestamp it occurs on) and for every node $u$, the following steps are executed:
-
-%To mention:
-%- Graph building;
-%- Change from binary trust to a weighted model;
-%- Evolution of trust over time.
-
-%First of all, there is an important change to the graph knowledge used to run the algorithm.
-%The complete trust graph is a parameter to the MaNI algorithm, which is executed by a supervisor with global knowledge of the graph.
-%
-%In this case, given a complete trust graph $T = (V, O)$, each node $u \in V$ has its own vision of the network and has to build a graph, called $T_u$, to represent its surrounding network.
-%In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
-%For every neighbor $v \in V$, $T_u$ must me merged with $T_v$.
-%An edge $u \rightarrow v$ exists if $u$ is aware of $v$.
-%Given sufficient iterations, $T_u$ forms a complete or nearly complete vision of the network.
-
-%Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
-%Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
-%When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
-%If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.
-
-%Furthermore, this implies that the malicious node detection algorithm runs using incomplete graph knowledge.
-%This is expected and generally unavoidable in vehicular networks, which have high mobility and can be extremely large.
-
-\section{Results}
-\label{section:results}
-
-In order to test the TruMan trust model, simulations were made using an implementation of the algorithm in Python.
-To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
-Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
-
-\begin{table}[h!]
-\caption{Simulation parameters}
-\label{table:parameters}
-\centering
-\begin{tabular}{|p{3cm}||p{3.5cm}|}
- \hline
- \textbf{Parameter}	& \textbf{Value} \\
- \hline
- \hline
- Duration 			& 86400 seconds \\
- \hline
- Work day length 	& 28800 seconds \\
- \hline
- Std. dev. departure time & 7200 seconds \\
- \hline
- Node velocity 		& 7~10m/s \\
- \hline
- Simulation area	& approximately 14km\textsuperscript{2} \\
- \hline
- Number of nodes 	& 150 (WDMM) + 10 (random) \\
- \hline
-% Cost & O(|V|* |E|) & O(|V|+|E|) & n/a & n/a & n/a \\
-% \hline
-\end{tabular}
-\end{table}
-
-Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}.
-Different parameters are shown in \autoref{table:parameters}.
-%The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
-%Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
-%%The number of nodes and their communication range vary for different simulations.
-%There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
-Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
-The parameters regarding offices, meeting spots and shopping were kept intact.
-A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
-
-\subsection{Network Density}
-
-%==> PRECISA MOSTRAR A EQUAÇÃO DE CALCULO DA DENSIDADE, SENÃO FICA DIFICIL DE ENTENDER O CALCULO.
-
-The communication range of nodes varies from 10m to 50m, to illustrate the impact of different network densities.
-The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
-For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
-It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$).
-
-The coverage of a single node is the circumference around it formed by the transmission radius.
-This is divided by two to compensate for overlapping circumferences, then multiplied by the number of nodes in the network to estimate the maximum coverage area.
-Finally, the value is divided by the total area of the environment.
-The network density formula is as follows:
-
-$$ \delta = \frac{\frac{\rho^2\pi}{2} \times \eta}{\alpha} $$
-
-Simulations shown here have densities between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
-As a comparison, the density of the city of São Paulo (Brazil) was calculated as 2.24 with $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
-
-\begin{figure*}[!t]
-\centering
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/10.png}%
-\label{subfig:random10}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}%
-\label{subfig:random30}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}%
-\label{subfig:random50}}
-\caption{Simulation with 10\% malicious nodes and $\rho$ = (a) 10m, (b) 30m or (c) 50m.}
-\label{fig:random103050}
-\end{figure*}
-
-\begin{figure*}[!t]
-\centering
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_1}%
-\label{subfig:random1}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_5}%
-\label{subfig:random2}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_10}%
-\label{subfig:random3}}
-
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
-\label{subfig:random4}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
-\label{subfig:random5}}
-\hfil
-\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
-\label{subfig:random6}}
-
-\caption{Simulation with $\rho$ = 10m and (a) 1\%, (b) 5\%, (c) 10\%, (d) 30\%, (e) 40\% or (f) 50\% malicious.}
-\label{fig:random0}
-\end{figure*}
-
-\subsection{Simulations}
-\label{subsection:simulations}
-To improve readability, all figures in this section follow the same format:
-$X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
-$Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
-blue line represents the percentage of nodes detected out of the complete network;
-magenta is the percentage of malicious nodes in the network (ground truth);
-green represents the nodes correctly identified as malicious (true positives);
-cyan represents the undetected malicious nodes (false negatives); 
-red represents the benign nodes incorrectly identified as malicious (false positives).
-
-\autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
-It is possible to see how the increase in the range allows the algorithm to converge sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
-
-\autoref{fig:random0} shows the variation of results for different amounts of malicious nodes in the network.
-By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
-At 40\%, a small part of malicious nodes are yet to be detected.
-At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
-The amount of malicious nodes also affects the convergence of the algorithm, since nodes do not trust information from malicious neighbors.
-
-\autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
-Most malicious nodes are identified by the end of the first day; in the following iterations, the algorithm finishes building the network model and sorts out remaining false negative or false positive results.
-After iteration 20000, the results are completely consistent.
-
-%==> ACHO QUE PODIA USAR O MULTIFIGURE PARA AGRUPAR UM POUCO AS FIGURAS. COLOCA 3 LADO A LADO. ASSIM SOBRE ESPAÇO NO FINAL PARA PODERMOS COMPARAR COM TRABALHOS RELACIONADOS.
-
-% ----
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\lin0.5ewidth]{Network_A_10.0/new_plots/10.png}
-%%\includegraphics[width=0.45\linewidth]{Network_A_10.0/new_plots/30.png}
-%\caption{Simulation without random nodes and range of 10m.} \label{fig:nonrandom1}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/30.png}
-%\caption{Simulation without random nodes and range of 30m.} \label{fig:nonrandom2}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/50.png}
-%\caption{Simulation without random nodes and range of 50m.} \label{fig:nonrandom3}
-%\end{figure}
-
-% ----
-
-%==> AUMETEI UM POUCO AS FIGURAS, MAS ELAS PARECEM ESTAR MEIO DEFOCADAS, VE SE CONSEGUE MELHORAR UM POUCO A QUALIDADE DELAS.
-
-%\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/30.png}
-%\caption{a}
-%%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
-%%\hspace{1pt}
-%\includegraphics[width=0.329\linewidth]{Network_rA_10.0/new_plots/50.png}
-%\caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
-%\label{fig:random103050}
-%\end{figure*}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/30.png}
-%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/50.png}
-%\caption{Simulation with random nodes and range of 50m.} \label{fig:random50}
-%\end{figure}
-
-% ----
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_1}
-%%\hspace{1pt}
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_5}
-%%\hspace{1pt}
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_10}
-%\caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
-%\label{fig:random0}
-%\end{figure*}
-
-%\begin{figure}
-%\centering0.5
-%\includegraphics[width=0.5\linewidth]{Network_rA/10_5}
-%\caption{Simulation with random nodes, range of 10m and 5\% malicious.} \label{fig:random1}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA/10_10}
-%\caption{Simulation with random nodes, range of 10m and 10\% malicious.} \label{fig:random2}
-%\end{figure}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA/10_20}
-%\caption{Simulation with random nodes, range of 10m and 20\% malicious.} \label{fig:random3}
-%\end{figure}
-
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_30}%
-%\label{subfig:random4}}
-%\hfil
-%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_40}%
-%\label{subfig:random5}}
-%\hfil
-%\subfloat[]{\includegraphics[width=0.329\linewidth]{Network_rA/10_50}%
-%\label{subfig:random6}}
-%\caption{Simulation with random nodes, range of 10m and (a) 30\%, (b) 40\% or (c) 50\% malicious.}
-%\label{fig:random1}
-%\end{figure*}
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_30}
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_40}
-%\includegraphics[width=0.329\linewidth]{Network_rA/10_50}
-%\caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
-%\label{fig:random1}
-%\end{figure*}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA/10_40}
-%\caption{Simulation with random nodes, range of 10m and 40\% malicious.} \label{fig:random5}
-%\end{figure}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.5\linewidth]{Network_rA/10_50}
-%\caption{Simulation with random nodes, range of 10m and 50\% malicious.} \label{fig:random6}
-%\end{figure}
-
-% ----
-
-%\begin{subfigure}{0.20\textwidth}
-%\includegraphics[width=\linewidth]{images/coloring/4.png}
-%\caption{Edge $(2,3)$ is checked and node $3$ gets the current value of $d$.} \label{fig:coloring5}
-%\end{subfigure}
-%\hspace*{2cm} % separation between the subfigures
-%\begin{subfigure}{0.20\textwidth}
-%\includegraphics[width=\linewidth]{images/coloring/5.png}
-%\caption{Edge $(2,4)$ is checked and node $4$ gets the current value of $d$.} \label{fig:coloring6}
-%\end{subfigure}
-%\hspace*{2cm} % separation between the subfigures
-%\begin{subfigure}{0.20\textwidth}
-%\includegraphics[width=\linewidth]{images/coloring/6.png}
-%\caption{Edge $(3,4)$ is checked and node $4$ gets a new color. $d=4$ (yellow).} \label{fig:coloring6}
-%\end{subfigure}
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-\section{Related work}
-\label{section:previouswork}
-
-Several models have been proposed to solve the problem of trust in vehicular networks. 
-This analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
-In the first part of this section, these properties are described with an assessment of whether or not TruMan satisfies their conditions.
-Then, some of the most relevant models are described, considering how well they satisfy the desired properties.
-\autoref{table:properties} shows how they compare with TruMan.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{Network_rA7/10_10}
-\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
-\end{figure}
-
-\subsection{Properties}
-\label{section:properties}
-
-\textit{Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes.
-Nodes may or may not use information from other trustworthy nodes to build trust values.
-TruMan satisfies this as it is built from the ground up for decentralized systems.
-
-\textit{Coping with sparsity}: the model still functions when there are few nodes populating the network.
-The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
-Due to its decentralized nature, it can also work on isolated chunks of the network.
-
-\textit{Event/task and location/time dynamics}: how the model reacts to different situations depending on what, where and when events happen.
-Although this has not been used in the simulations in this paper, TruMan can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
-
-\textit{Scalability}: the model can work on very large networks at high speeds.
-Due to the low complexity of the algorithms used in the model, TruMan can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
-It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
-
-\textit{Integrated confidence measure}: allows nodes to estimate how useful the output of the algorithm is.
-Since nodes using TruMan store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
-The closer it is to 1, the higher the chance that it is accurate.
-
-\textit{System level security}: requires authentication of nodes participating in the network.
-This has not been considered in evaluations of TruMan.
-However, it can be included as a separate security model during the transmission of messages.
-
-\textit{Sensitivity to privacy concerns}: avoids eavesdropping and stalking by malicious nodes.
-TruMan has not been designed with this in mind, but it does not inhibit privacy protection.
-However, it does require that nodes cannot be completely anonymous.
-
-\textit{Robustness}: the model's resistance to attacks.
-TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
-Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
-Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
-Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
-
-\subsection{Comparison with other trust models}
-\label{section:comparison}
-
-For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but it is designed only for static networks and the algorithm relies on a global observer which has information about the complete network.
-It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
-The usage of strongly connected components and coloring serves as a basis for TruMan, which is expanded to work on distributed and dynamic networks such as vehicular networks.
-
-The model proposed in \cite{minhas2010towards} uses several criteria to judge whether or not a received message is trustworthy.
-First, nodes are classified by their roles, used for vehicles which should be automatically trustworthy (i.e. police cars).
-Nodes also store their experience each time an event message is received (if one neighboring node reported an event which did not turn out to be true, its trust value is reduced).
-Additionally, messages have higher reliability when their senders are closer in time and space to the reported event.
-When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes mean a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
-%The model considers both role-based trust and experience-based trust; although the work proposed here does not use role-based trust, the authors provide a useful method of calculating and updating an experience-based trust value, which might be used or adapted.
-However, the model relies only on direct interaction between pairs of nodes, so no form of indirect trust is considered.
-
-In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
-By separating nodes into clusters with their geographical neighbors, it is possible to distribute the evaluation of messages using previously formed opinions.
-When a node sends a message, the cluster-leader must aggregate the other nodes' opinions on that message.
-Messages are only forwarded to other clusters if the aggregate opinion is above a certain threshold.
-Additionally, nodes that receive the message only act if the overall trust on it is above another threshold.
-However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
-Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster become untrustworthy.
-
-The ART model proposed in \cite{li2016art} works in two main steps: data gathering and malicious node detection.
-It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
-Then, it uses a Cosine-based metric to compare two nodes' trust vectors (a series of opinions regarding other nodes).
-However, these steps require several intensive calculations, which greatly increase the complexity of the algorithm.
-The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
-
-The authors of \cite{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
-This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations in which mobile communication is slow or unreliable.
-It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
-
-Finally, it is worth noting that, aside from \cite{vernize2015malicious}, none of the mentioned related work presents complexity calculations for their algorithm.
-Considering the scale of the problem, TruMan's cost of $O(|V|\times |E|)$ is very low without sacrificing completeness and correctness.
-The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
-
-%None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
-%Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
-%There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
-%
-
-\begin{table}[t!]
-\caption{Properties of Truman and related work}
-\label{table:properties}
-\centering
-\begin{tabular}{|p{1.5cm}||p{1.1cm}|p{1.1cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
- \hline
- \textbf{Property} & Truman & \cite{vernize2015malicious} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
- \hline
- Decentralized 	& \checkmark & -		  & \checkmark & \checkmark & \checkmark & - \\
- \hline
- Sparsity 		& \checkmark & -		  & \checkmark & \checkmark & -			 & \checkmark \\
- \hline
- Dynamics 		& \checkmark & - 		  & \checkmark & \checkmark & -			 & - \\
- \hline
- Scalability 	& \checkmark & \checkmark & \checkmark & \checkmark & -			 & \checkmark \\
- \hline
- Confidence 	& \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - \\
- \hline
- Security 		& -			 & - 		  & \checkmark & \checkmark & - 		 & \checkmark \\
- \hline
- Privacy 		& - 		 & - 		  & \checkmark & - 			& -			 & \checkmark \\
- \hline
- Robustness 	& \checkmark & -		  & -		   & - 			& \checkmark & - \\
- \hline
- Efficiency		& \checkmark & \checkmark & -		   & -			& -			 & - \\
- \hline
- Cost & $|V|*|E|$ & $|V|+|E|$ & n/a & n/a & n/a & n/a\\
- \hline
-% \hline
-\end{tabular}
-\end{table}
-
-\section{Conclusion}
-\label{section:conclusion}
+\chapter{Conclusion}
+\label{chap:conclusion}
 
@@
 Future work includes the extension of the proposed model to use V2I communications.
-%Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-%
-%
-%The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{article}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-%
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%
-%\end{thebibliography}
-
-% that's all folks
-\end{document}

### 2018-02-28 dissertation/1-intro/chapter1.tex
dissertation v 2017-11-30
several updates to text; possible problem with character encoding

--- 2017-06-01 qualification/1-intro/chapter1.tex
@@
 \section{Document organization}
-The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
+The remainder of this dissertation is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
 Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
-Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.
+\Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
+Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
+Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.
+
+%Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2018-02-28 dissertation/2-complexnetworks/chapter2.tex
several updates to text; possible problem with character encoding

@@
 Hence, there is the correlation between the trust graph and the topology graph of the network.
-Since MANETs are dynamic, the graph that represents one network's topology is frequently changing and, with that, the opportunities to create and update trust relationships also changes.
+Since MANETs are dynamic, the graph that represents a network's topology is frequently changing and, with that, the opportunities to create and update trust relationships also changes.
 %For two nodes to have a trust relationship at all, the topology must changed in a way so that they are neighbors to each other, or close enough that asking for a third node's

### 2018-02-28 dissertation/3-vanets/chapter3.tex
several updates to text; possible problem with character encoding

@@
 Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) which participate in it.
-If one node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
+If a node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
 Once the concept of VANETs was established, researchers have been attempting to predict ways in which malicious users might use the network to their advantage.
-Examples include triggering false alarms about inexistent accidents, lying about the average speed in one road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms.
+Examples include triggering false alarms about inexistent accidents, lying about the average speed in a road to make it less desirable for others, and falsifying geolocation data to exploit location-based routing algorithms.
 Therefore, the concept of trust must be established in the vehicular network context, allowing for nodes to judge the validity of information transmitted by others and share those conclusions with other nodes.
@@
  \hline
- TruMan & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
- \hline
- \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
+% \hline
+% \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
  \hline
@@
  \hline
- \cite{minhas2010towards} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \cite{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
  \hline
- \cite{chen2010trust} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \cite{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
  \hline
@@
  \hline
- \cite{li2016art} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \cite{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
  \hline
- \cite{chen2017cloud} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \cite{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
  \hline
@@
 It combines aspect of data- and entity- based trust, since nodes share their opinion of the data as well as their opinion of the sender. 
-One interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
+An interesting observation is setting higher trust values for certain vehicles based on their familiarity with the region (vehicles that reside in a given city may have more experience with certain types of events than newcomers).
 However, opinion piggybacking has its own share of problems.

### 2018-02-28 dissertation/4-project/chapter4.tex
several updates to text; possible problem with character encoding

@@
 
-This work proposes that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
+This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
 In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
@@
 By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
-This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as input for a malicious node detection algorithm.
+This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as in conjunction with other algorithms in order to classify nodes as correct or malicious.
 
@@
 
-In this chapter, the details for the above process are presented.
+In this chapter, the reasoning behind TruMan and details of how it works are presented.
 First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
-Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
-Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
-Finally, a schedule for the next steps of the project is shown.
+Then, two algorithms are introduced: Tarjan's strongly connected components algorithm \cite{tarjan1972depth} and an efficient algorithm for graph coloring \cite{mittal2011graph}.
+Next, the Malicious Node Identification Algorithm (MaNI) \cite{vernize2013dissertation} is explained, for it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
+Finally, TruMan itself is explained, showing how it combines features from existing algorithms and adapts them to a dynamic environment, enabled by the social properties found in vehicular networks.
+%Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.
+%Next, a previously existing trust model for complex networks is described, including its advantages and reasons why it could be well-fitted for a vehicular environment, followed by a list of changes that must be made to the model in order for it to handle dynamic networks.
+%Finally, a schedule for the next steps of the project is shown.
 
@@
 \label{section:goals}
+Building from the foundation set by MaNI \cite{vernize2013dissertation}, TruMan strives to enable efficient trust management in highly dynamic networks such as VANETs.
+In addition to efficiency, it is desirable that the trust model is both simple to understand and to implement, making it appealing for real-world usage.
+
+Furthermore, the desired properties of a VANET trust model described in \cite{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.
 
@@
 %In any case, this approach is not covered in this work.
-
-\section{Malicious Node Identification Algorithm}
-
-The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
-This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
-The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
-Furthermore, the algorithm is executed by a global observer which has information about the complete network.
-
-The input graph $T = (V,O)$ is a static, connected, and directed graph containing all trust relationships in the network.
-Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
-The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.
-
-The process for identifying malicious nodes within $T$ is as follows:
-
-First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
-In each of these components, all nodes are connected by edges of value $1$.
-In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
-Each of these components becomes a node of a component graph $C = (V', O')$.
-
-The creation of the graph $C$ simplifies the remaining computation.
-Since each node of $C$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
-They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
-After the formation of $C$, one or more heuristics can be used to classify the nodes as benign or malicious.
-
-The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
-Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
-After running either algorithms with graph $C$, the color whose nodes in $C$ represent the most nodes in $T$ is classified as correct, and all others are classified as malicious.
-Once this information in $C$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
-
-In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
-Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
-Therefore, for the purposes of this research, only the coloring heuristic is considered.
-
-Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
-In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
-Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
-
-The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
-By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
-
-%\begin{algorithm}
-%\caption{MaNI algorithm}\label{algorithm:mani}
-%\begin{algorithmic}[1]
-%
-%\Function{MaNI}{graph $T$}
-%
-%\State $C \gets$ \textbf{Tarjan}$(T[0])$
-%\State $colors \gets$ \textbf{Coloring}$(C)$
-%\State \textbf{color} nodes of $T$ \textbf{according to} $colors$
-%\State \textbf{count} 
-%
-%\EndFunction
-%\end{algorithmic}
-%\end{algorithm}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.25\textwidth}
-\includegraphics[width=\linewidth]{images/mani/0-trust.png}
-\caption{Initial graph $T$.} \label{fig:mani0}
-\end{subfigure}
-\hspace*{2cm} % separation between the subfigures
-\begin{subfigure}{0.25\textwidth}
-\includegraphics[width=\linewidth]{images/mani/1-tarjan.png}
-\caption{Tarjan's algorithm is used to identify the strongly connected components.} \label{fig:mani1}
-\end{subfigure}
-
-\vspace{1cm} % separation between the subfigures
-
-\begin{subfigure}{0.3\textwidth}
-\includegraphics[width=\linewidth]{images/mani/2-components.png}
-\caption{The graph $C$ is formed from the components of $T$.} \label{fig:mani2}
-\end{subfigure}
-\hspace*{2cm} % separation between the subfigures
-\begin{subfigure}{0.3\textwidth}
-\includegraphics[width=\linewidth]{images/mani/3-colored-components.png}
-\caption{The coloring algorithm is used to label the nodes of $C$.} \label{fig:mani3}
-\end{subfigure}
-
-\vspace{1cm} % separation between the subfigures
-
-\begin{subfigure}{0.25\textwidth}
-\includegraphics[width=\linewidth]{images/mani/4-result.png}
-\caption{The colors from $C$ are used to label nodes in $T$ as correct or malicious.} \label{fig:mani4}
-\end{subfigure}
-
-\caption{Example of an execution of the MaNI algorithm.}\label{fig:mani}
-\end{figure}
-
-\autoref{fig:mani} illustrates the execution of the MaNI algorithm.
-With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
-\autoref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
-The strongly connected components form a graph $C$ according to edges present in $T$.
-In \autoref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
-The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \autoref{fig:mani3}.
-Finally, each node in $T$ is colored according to which color its component received in $C$.
-The color with the most nodes is deemed benign, while the others are considered malicious.
-
-%$G = (V,E)$: undirected topology graph, in which $V$ are the vertices and $E$ are the edges.
-%
-%$T = (V,O)$: directed trust graph, in which $O$ are the nodes' opinions about each other.
-%
-%$G_t,T_t$: snapshots of $G$ and $T$ in a timestamp $t$.
-%
-%$G_i,T_i$: local knowledge of $G$ and $T$ of a node $i$.
-%
-%$C = (V',O')$: component graph of $T$.
 
@@
 \For{$v$ \textbf{in} neighbors of $u$ }
+
+	\If{weight of $u\rightarrow v < h$}
+		\State \textbf{continue}
+	\EndIf
 
@@
 
+\section{Malicious Node Identification Algorithm}
+
+The basis of TruMan is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, which suggests the use of strongly connected components and graph coloring for malicious node detection.
+This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
+The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
+Furthermore, the algorithm is executed by a global observer which has information about the complete network.
+
+The input graph $T = (V,E)$ is a static, connected, and directed graph containing all trust relationships in the network.
+Such relationships are binary, so there are no varying degrees of trust: either one node trusts another completely (edge value is $1$), or it distrusts the other completely (edge value is $0$).
+The relationships are also directed, meaning that if the value of $A\rightarrow B$ is $1$, $B\rightarrow A$ is not necessarily $1$.
+
+The process for identifying malicious nodes within $T$ is as follows:
+
+First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
+In each of these components, all nodes are connected by edges of value $1$.
+In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
+Each of these components becomes a node of a component graph $C = (V', E')$.
+
+The creation of the graph $C$ simplifies the remaining computation.
+Since each node of $C$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
+They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
+After the formation of $C$, one or more heuristics can be used to classify the nodes as benign or malicious.
+
+The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
+Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
+After running either algorithms with graph $C$, the color whose nodes in $C$ represent the most nodes in $T$ is classified as correct, and all others are classified as malicious.
+Once this information in $C$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
+
+In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
+Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
+Therefore, for the purposes of this research, only the coloring heuristic is considered.
+
+Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
+In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
+Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
+
+The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
+By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
+
+%\begin{algorithm}
+%\caption{MaNI algorithm}\label{algorithm:mani}
+%\begin{algorithmic}[1]
+%
+%\Function{MaNI}{graph $T$}
+%
+%\State $C \gets$ \textbf{Tarjan}$(T[0])$
+%\State $colors \gets$ \textbf{Coloring}$(C)$
+%\State \textbf{color} nodes of $T$ \textbf{according to} $colors$
+%\State \textbf{count} 
+%
+%\EndFunction
+%\end{algorithmic}
+%\end{algorithm}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.25\textwidth}
+\includegraphics[width=\linewidth]{images/mani/0-trust.png}
+\caption{Initial graph $T$.} \label{fig:mani0}
+\end{subfigure}
+\hspace*{2cm} % separation between the subfigures
+\begin{subfigure}{0.25\textwidth}
+\includegraphics[width=\linewidth]{images/mani/1-tarjan.png}
+\caption{Tarjan's algorithm is used to identify the strongly connected components.} \label{fig:mani1}
+\end{subfigure}
+
+\vspace{1cm} % separation between the subfigures
+
+\begin{subfigure}{0.3\textwidth}
+\includegraphics[width=\linewidth]{images/mani/2-components.png}
+\caption{The graph $C$ is formed from the components of $T$.} \label{fig:mani2}
+\end{subfigure}
+\hspace*{2cm} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+\includegraphics[width=\linewidth]{images/mani/3-colored-components.png}
+\caption{The coloring algorithm is used to label the nodes of $C$.} \label{fig:mani3}
+\end{subfigure}
+
+\vspace{1cm} % separation between the subfigures
+
+\begin{subfigure}{0.25\textwidth}
+\includegraphics[width=\linewidth]{images/mani/4-result.png}
+\caption{The colors from $C$ are used to label nodes in $T$ as correct or malicious.} \label{fig:mani4}
+\end{subfigure}
+
+\caption{Example of an execution of the MaNI algorithm.}\label{fig:mani}
+\end{figure}
+
+\autoref{fig:mani} illustrates the execution of the MaNI algorithm.
+With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
+\autoref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
+The strongly connected components form a graph $C$ according to edges present in $T$.
+In \autoref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
+The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \autoref{fig:mani3}.
+Finally, each node in $T$ is colored according to which color its component received in $C$.
+The color with the most nodes is deemed benign, while the others are considered malicious.
+
+%$G = (V,E)$: undirected topology graph, in which $V$ are the vertices and $E$ are the edges.
+%
+%$T = (V,O)$: directed trust graph, in which $O$ are the nodes' opinions about each other.
+%
+%$G_t,T_t$: snapshots of $G$ and $T$ in a timestamp $t$.
+%
+%$G_i,T_i$: local knowledge of $G$ and $T$ of a node $i$.
+%
+%$C = (V',O')$: component graph of $T$.
+
 \section{The TruMan algorithm}

### 2018-02-28 dissertation/5-simulations/chapter5.tex
several updates to text; possible problem with character encoding

@@
 
-\section{The ONE Simulator}
-\label{section:theone}
-
 \section{Working Day Movement Model}
@@
 
-Most VANET trust models consider the Random Waypoint mobility model, i.e.
+Most VANET trust models use the Random Waypoint mobility model for simulations, i.e.
 each node has an origin point, chooses a random location, gets to that location, then chooses another random location and goes there, and so forth.
@@
 To make use of the properties described in \autoref{section:socialvanets}, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
-Therefore, the Working Day Movement Model \cite{ekman2008working} (WDM) is useful.
+Therefore, the Working Day Movement Model \cite{ekman2008working} (WDMM) is useful.
 The model, developed for use in Delay-Tolerant Network (DTN) simulations, includes many of the features that are necessary to simulate the daily movement of a vehicular network.
 
-As the name implies, the Working Day Model abstracts people's movement from their homes to their offices and back.
+As the name implies, the Working Day Movement Model abstracts people's movement from their homes to their offices and back.
 Each node has a home and a workplace and they need to travel back and forth between those locations on a daily basis.
 Occasionally, nodes can also go to other locations for leisure.
-As mentioned above, many drivers have routes they travel on daily, so the Working Day model is a more accurate representation, although it represents the mobility of humans instead of vehicle and requires some adjustments, which are listed later in this section.
+As mentioned above, many drivers have routes they travel on daily, so the Working Day Movement Model is a reasonably accurate representation of daily movement in a city.
+
+\subsection{Original model}
+The Working Day Movement Model was developed for Delay-Tolerant Networks in which nodes are devices (such as smartphones) carried by people.
+Therefore, the Working Day model represents not only people's movements inside their cars, but also within their offices, walking on foot, or riding a bus.
+In this subsection, the word node refers to a mobile device instead of a vehicle.
 
@@
 The main mobility model defines nodes and gives them their destinations.
-Within it, five other models are used:
+Within it, five submodels are used:
 \begin{enumerate}
@@
 Nodes can go to other locations within the office (such as meeting rooms) and such movement is modeled.
-Nodes with share the same office are coworkers.
+Nodes which share the same office are coworkers.
 \item 
 The \textbf{evening activity submodel} is responsible for mobility outside the nodes' standard routine. 
-They can meet at certain locations (such as restaurants) and spend a few hours gathered with friends.
+They can meet at certain locations (such as restaurants) and spend a few hours with friends.
 \item
@@
 It includes another tier of submodels, responsible for modeling three different types of transportation: walking, driving, and riding a bus.
-Nodes which own a car always use it, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
+Nodes that own cars always use them, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
 The walking and driving submodels represent similar types of movement, although at different speeds, while the bus submodel follows cyclical routes and can take or deliver passengers at bus stops.
@@
 The \textbf{map} represents the city in which the simulation runs.
-Its streets constrain the movement of nodes and all homes and offices must be within the map boundaries.
+Its streets constrain the movement of nodes and all homes, offices and meeting spots must be within the map boundaries.
 The map can be divided into districts, which increases what the authors define as \textit{locality}.
 In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes rarely leave the district.
-Nodes which reside and work in different districts serve to connect the network with their commutes.
+Nodes which reside and work in different districts allow information to spread across different parts of the network.
 \end{enumerate}
 
-By thinking of these submodels for vehicles instead of people, it can become apparent how the frequency and length of encounters between nodes are similar in both instances.
+\subsection{Adaptation for a vehicular simulation}
+By thinking of these submodels for vehicles instead of people, it becomes apparent that the frequency and length of encounters between nodes are similar in both instances.
 If two vehicles belong to family members or neighbors, they likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
 Cars can also meet each other frequently if their drivers go out with friends.
-In the vehicular case, there is the added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, although the drivers are likely complete strangers.
+In the vehicular case, there is the added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, even if the drivers are complete strangers.
 
-In the original article, nodes are devices (such as smartphones) being carried by humans.
-Therefore, the Working Day model represents not only people's movements inside their cars, but also within their offices, walking on foot, or riding a bus.
-To adapt it to a VANET environment, changes need to be made to their definition of node, since they now represent vehicles instead of people. Some of those changes are as follows:
+To adapt the Working Day Movement Model to a VANET environment, a few changes had to be made so the nodes represent vehicles instead of people (or the devices they carry). 
+Instead of altering the model itself, all of these changes were implemented as parameters for the model.
+The changes are as follows:
 \begin{enumerate}
@@
 In both, a node can move a small amount once after reaching the office or home, to simulate parking.
-This can be done using the Working Day model's parameters.
+This was done by setting the \texttt{officeSize} parameter to 1, so vehicles do not move around while their drivers are at work.
 \item
-The walking submodel needs to be disabled, since all nodes are either cars or buses.
+The walking submodel needs to be disabled, since all nodes are cars.
+By setting the \texttt{ownCarProb} parameter to 1, mobility is always done by car.
 \item
-The bus submodel needs to be changed so that each bus is one node in the network, which follows a predefined route with bus stops.
-In the original model, each bus could carry several nodes, but this is no longer necessary.
+While the bus submodel could be used for a vehicular simulation, this was not used for TruMan's simulations.
+%The bus submodel needs to be changed so that each bus is one node in the network, which follows a predefined route with bus stops.
+%In the original model, each bus could carry several nodes, but this is no longer necessary.
 \end{enumerate}
-Other necessary changes might become apparent during the development of the study.
 
-One important topic raised in the Working Day movement model article is the use of two metrics: \textit{inter-contact times} and \textit{contact duration}.
+One important topic raised in the Working Day Movement Model article is the use of two metrics: \textit{inter-contact times} and \textit{contact duration}.
 The choice of this movement model for tests is more strongly related to inter-contact times, i.e. how much time it takes for two nodes to meet again.
 On the other hand, the contact duration is how long each meeting lasts.
-For the reasons explained earlier in this section, relatively short inter-contact times is important for the proposed trust model.
-Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although taking it into consideration might add excessive complexity to the model.
+For the reasons explained earlier in this section, relatively short inter-contact times is important for TruMan's functionality.
+Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although, for the purposes of the simulation, this was not considered.
+
+\section{The ONE Simulator}
+\label{section:theone}
+The Opportunistic Network Environment simulator \cite{keranen2009one} \cite{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.
+It was chosen for the simulations of TruMan primarily because it already includes integration with the Working Day Movement Model.
+
+In order to use data from the ONE simulator as input for TruMan, a new report module had to be created for it.
+The \texttt{AdjacencySnapshotReport} module creates a report consisting of all adjacencies in the network every $x$ number of simulated seconds.
+That is, at a given timestamp $t$, all pairs of nodes that are within communication range of each other are added to the report.
+This report is then used as input to TruMan, which uses the adjacencies to build the topology graph for each iteration of the algorithm.
+The \texttt{AdjacencySnapshotReport} module has been submitted to the ONE repository as a pull request \cite{adjacencypull}.
 
@@
 To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
-Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
+
+Snapshots of the network were taken every 10 simulated seconds using the \texttt{AdjacencySnapshotReport} module for the ONE simulator.
+However, a few experiments showed that it was not necessary to run iterations of TruMan that frequently; therefore, iterations run at an interval of 100 simulated seconds and only use one tenth of the snapshots saved.
 
@@
  \hline
- Node velocity 		& 7~10m/s \\
+ Node velocity 		& 7 to 10m/s \\
  \hline
@@
  Number of nodes 	& 150 (WDMM) + 10 (random) \\
+ \hline
+ Office size 		& 1 \\
  \hline
@@
 
-Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}.
-Different parameters are shown in \autoref{table:parameters}.
-The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
+Most of the parameters for the ONE simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}; the most important parameters are shown in \autoref{table:parameters}.
+The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds (8 hours) and a standard deviation of departure time of 7200 seconds (2 hours).
 Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
 %The number of nodes and their communication range vary for different simulations.
-There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
+There is a total of 160 nodes, 150 of which follow the Working Day Movement Model, and 10 that follow the random waypoint mobility model to simulate vehicles that do not follow daily patterns.
 Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
-The parameters regarding offices, meeting spots and shopping were kept intact.
+Aside from the office size parameter, the parameters regarding offices, meeting spots and shopping were kept intact.
 A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
+The transmission range of nodes varies from simulation to simulation, for reasons explained in \autoref{section:density}.
 
 \subsection{Network Density}
+\label{section:density}
 
@@
 The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
-For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{subsection:simulations}).
-It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$).
+For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{chap:simulations}).
+It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$, in m\textsuperscript{2}).
 
@@
 
-Simulations shown here have densities between 0.001 ($\rho = 10$m) and 0.04 ($\rho = 50$m).
-As a comparison, the density of the city of São Paulo (Brazil) was calculated as 2.24 with $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+For example, in a simulation with $\rho = 30$m, the calculation is as follows:
+
+$$ \delta = \frac{\frac{30^2\pi}{2} \times 160}{14.689.750} = 0.0154$$
+
+In \autoref{table:simdensities}, a few densities for different values of $\rho$, $\eta$ and $\alpha$ are shown. 
+Simulations for TruMan have densities between 0.0017 ($\rho = 10$m) and 0.0428 ($\rho = 50$m).
+
+\begin{table}[h!]
+\caption{Simulation densities}
+\label{table:simdensities}
+\centering
+\begin{tabular}{|p{1.5cm}|p{1.5cm}|p{2cm}|p{2cm}|}
+ \hline
+ $\rho$ & $\eta$ & $\alpha$ & $\delta$ \\
+ \hline
+ \hline
+ 10 m & 160 & 14.689.750 & 0.0017 \\
+ \hline
+ 30 m & 160 & 14.689.750 & 0.0154 \\
+ \hline
+ 50 m & 160 & 14.689.750 & 0.0428 \\
+ \hline
+ 100 m & 160 & 14.689.750 & 0.1604 \\
+ \hline
+ 150 m & 160 & 14.689.750 & 0.3609 \\
+ \hline
+ 200 m & 160 & 14.689.750 & 0.6416 \\
+ \hline
+\end{tabular}
+\end{table}
+
+As a comparison, the city of São Paulo (Brazil), which has a fleet of over 8 million vehicles in an area of 1521.11 km\textsuperscript{2}, has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+\autoref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
+Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
+
+\begin{table}[h!]
+\caption{Simulation densities}
+\label{table:citydensities}
+\centering
+\begin{tabular}{|p{3cm}||p{4cm}|p{4cm}|p{2cm}|}
+ \hline
+ \textbf{City} & \textbf{$\eta$} & $\alpha$ & $\delta$ \\
+ \hline
+ \hline
+ São Paulo (BR) & 8.603.239 & 1.521,11 & 0.8884 \\
+ \hline
+ 30 m & 160 & 14.689.750 & 0.0154 \\
+ \hline
+ 50 m & 160 & 14.689.750 & 0.0428 \\
+ \hline
+ 100 m & 160 & 14.689.750 & 0.1604 \\
+ \hline
+ 150 m & 160 & 14.689.750 & 0.3609 \\
+ \hline
+ 200 m & 160 & 14.689.750 & 0.6416 \\
+ \hline
+\end{tabular}
+\end{table}
 
@@
 green represents the nodes correctly identified as malicious (true positives);
-cyan represents the undetected malicious nodes (false negatives); 
+%cyan represents the undetected malicious nodes (false negatives); 
 red represents the benign nodes incorrectly identified as malicious (false positives).
+The lines represent average values between all benign nodes on the network, while the colored regions represent the standard deviation present in the data.
+
+It is expected that the results are very inconsistent at the beginning of the simulations, since nodes are still building their models of the network and have relatively little information to use.
 
 \autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
-It is possible to see how the increase in the range allows the algorithm to converge sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
+It is possible to see how the increase in the range allows the algorithm to produce better results.
+At $\rho = 10$m and $\delta = 0.0017$, a large amount of time is spent with only about half of malicious nodes being accurately identified.
+Results with $\rho = 30$m and $\delta = 0.0153$ are better, but still less than ideal; during this simulation, there were more false positive detections happening, although it was still a small amount.
+At $\rho = 50$m and $\delta = 0.0427$, results are solid before the 2000th iteration of the algorithm, since at that point over 90\% of the network has been acknowledged by most nodes.
+The amount of false positive detections is extremely low, and, by the end of the simulation, all benign nodes have identified 100\% of malicious nodes.
 
-\autoref{fig:random0} shows the variation of results for different amounts of malicious nodes in the network.
-By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
-At 40\%, a small part of malicious nodes are yet to be detected.
-At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is completely compromised.
-The amount of malicious nodes also affects the convergence of the algorithm, since nodes do not trust information from malicious neighbors.
+% taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
+
+\autoref{fig:randommalicious} shows the variation of results for different amounts of malicious nodes in the network.
+By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network, although there is still a small amount of false positive detections.
+At 40\%, a small part of malicious nodes are yet to be detected and the standard deviation is larger overall.
+At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is considered completely compromised.
+True positive and false positive detections are often inverted, because whenever there are more malicious nodes than benign one in a certain node's network model, the algorithm will classify the malicious ones as correct.
+The amount of malicious nodes also affects how quickly nodes are able to acquire information about the network, since nodes do not trust information from malicious neighbors.
+
+All simulations were performed with trust threshold $h=0.5$, except for the ones in \autoref{fig:randomthresholds}, performed to demonstrate the impact of different threshold values.
+The plots illustrate that there is not a significance change in results depending on the different thresholds.
+The results are slightly better at $h=0.7$, however not in a significant way.
+It still takes over 8000 iterations to detect all malicious nodes and there are still some false positive detections.
 
 \autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
-Most malicious nodes are identified by the end of the first day; in the following iterations, the algorithm finishes building the network model and sorts out remaining false negative or false positive results.
-After iteration 20000, the results are completely consistent.
+Most malicious nodes are identified by the end of the first day, before iteration 9000.
+However, not all nodes have finished building their model of the network and therefore there are still a number of false positive detections occurring.
+Around iteration 20000, all nodes finish building their models and the false positive detection rate drops to an insignificant amount.
 
@@
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10.png}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
 \caption{$\rho$ = 10m.} \label{fig:tarjan0}
@@
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30.png}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
 \caption{$\rho$ = 30m.} \label{fig:tarjan0}
@@
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50.png}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
 \caption{$\rho$ = 50m.} \label{fig:tarjan0}
@@
 
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes.}\label{fig:tarjan}
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes.}
+\label{fig:randommalicious}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{images/plots/Network_rA7/10_10}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
+\caption{$h = 0.3$.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
+\caption{$h = 0.5$.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
+\caption{$h = 0.7$.} \label{fig:tarjan0}
+\end{subfigure}
+
+\caption{Simulation with 10\% malicious nodes, $\rho = 30$m and varying values of $h$.}
+\label{fig:randomthresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{images/plots/Network_rA7/10_10}
 \caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}

### 2018-02-28 dissertation/5-simulations/chapter5.tex
finished city density table

@@
 In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
-The Stanford Network Analysis Platform (SNAP) library \cite{snap} was chosen primarily because it is memory efficient.
+The Stanford Network Analysis Platform (SNAP) library \citep{snap} was chosen primarily because it is memory efficient.
 The simulations require multiple graphs that share the same set of nodes (because each node in the network has its own knowledge of the surrounding network), and the SNAP library uses pointers to nodes and edges, saving memory by not having to duplicate the entire data structure.
@@
 
-%The datasets available from SNAP's website were used for the simulations in \cite{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
+%The datasets available from SNAP's website were used for the simulations in \citep{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
 
@@
 To make use of the properties described in \autoref{section:socialvanets}, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
-Therefore, the Working Day Movement Model \cite{ekman2008working} (WDMM) is useful.
+Therefore, the Working Day Movement Model \citep{ekman2008working} (WDMM) is useful.
 The model, developed for use in Delay-Tolerant Network (DTN) simulations, includes many of the features that are necessary to simulate the daily movement of a vehicular network.
@@
 \label{section:theone}
-The Opportunistic Network Environment simulator \cite{keranen2009one} \cite{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.
+The Opportunistic Network Environment simulator \citep{keranen2009one} \citep{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.
 It was chosen for the simulations of TruMan primarily because it already includes integration with the Working Day Movement Model.
@@
 This report is then used as input to TruMan, which uses the adjacencies to build the topology graph for each iteration of the algorithm.
-The \texttt{AdjacencySnapshotReport} module has been submitted to the ONE repository as a pull request \cite{adjacencypull}.
+The \texttt{AdjacencySnapshotReport} module has been submitted to the ONE repository as a pull request \citep{adjacencypull}.
 
@@
 In order to test the TruMan trust model, simulations were made using an implementation of the algorithm in Python.
-To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
+To generate the input graphs with node mobility, the ONE simulator \citep{keranen2009one} was used in conjunction with the Working Day Movement Model \citep{ekman2008working}, which provides a strong similarity with vehicle movement in real life.
 
@@
  \hline
- Node velocity 		& 7 to 10m/s \\
+ Node velocity 		& 7 to 10 m/s \\
  \hline
- Simulation area	& approximately 14km\textsuperscript{2} \\
+ Simulation area	& 14,689,750 m\textsuperscript{2} \\
  \hline
@@
 
-Most of the parameters for the ONE simulator were taken from the article detailing the Working Day Movement Model \cite{ekman2008working}; the most important parameters are shown in \autoref{table:parameters}.
+Most of the parameters for the ONE simulator were taken from the article detailing the Working Day Movement Model \citep{ekman2008working}; the most important parameters are shown in \autoref{table:parameters}.
 The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds (8 hours) and a standard deviation of departure time of 7200 seconds (2 hours).
-Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
+Nodes move between 7 and 10 m/s in an area of approximately 14.7 km\textsuperscript{2} based on a section of the map of Helsinki.
 %The number of nodes and their communication range vary for different simulations.
@@
 \centering
-\begin{tabular}{|p{1.5cm}|p{1.5cm}|p{2cm}|p{2cm}|}
+\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
  \hline
- $\rho$ & $\eta$ & $\alpha$ & $\delta$ \\
+ \textbf{Range ($\rho$)} & \textbf{Nodes ($\eta$)} & \textbf{Area ($\alpha$)} & \textbf{Density($\delta$)} \\
  \hline
  \hline
- 10 m & 160 & 14.689.750 & 0.0017 \\
+ 10 m & 160 & 14,689,750 & 0.0017 \\
  \hline
- 30 m & 160 & 14.689.750 & 0.0154 \\
+ 30 m & 160 & 14,689,750 & 0.0154 \\
  \hline
- 50 m & 160 & 14.689.750 & 0.0428 \\
+ 50 m & 160 & 14,689,750 & 0.0428 \\
  \hline
- 100 m & 160 & 14.689.750 & 0.1604 \\
+ 100 m & 160 & 14,689,750 & 0.1604 \\
  \hline
- 150 m & 160 & 14.689.750 & 0.3609 \\
+ 150 m & 160 & 14,689,750 & 0.3609 \\
  \hline
- 200 m & 160 & 14.689.750 & 0.6416 \\
+ 200 m & 160 & 14,689,750 & 0.6416 \\
  \hline
@@
 
-As a comparison, the city of São Paulo (Brazil), which has a fleet of over 8 million vehicles in an area of 1521.11 km\textsuperscript{2}, has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+As a comparison, the city of São Paulo (Brazil), which has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
 \autoref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
-Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
+All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan aras that surround these cities.
+%Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
+
+\defcitealias{saopauloarea}{IBGE, 2016}
+\defcitealias{newyorkfleet}{NY DMV, 2016}
+\defcitealias{losangelesfleet}{CA DMV, 2016}
+\defcitealias{usarea}{US Census, 2017}
+\defcitealias{tokyoarea}{Tokyo MG, 2017}
+\defcitealias{londonfleet}{UK DfT, 2016}
+\defcitealias{londonarea}{GLA, 2011}
 
 \begin{table}[h!]
-\caption{Simulation densities}
+\caption{Calculated densities of major cities}
 \label{table:citydensities}
@@
  \hline
- \textbf{City} & \textbf{$\eta$} & $\alpha$ & $\delta$ \\
+ \textbf{City (country)} & \textbf{Nodes ($\eta$)} & \textbf{Area ($\alpha$)} & \textbf{Density ($\delta$)} \\
  \hline
  \hline
- São Paulo (BR) & 8.603.239 & 1.521,11 & 0.8884 \\
+ São Paulo (BR) 	& 8,603,239 					& 1,521,110,000  				& 0.8884 	\\
+ 				 	& \citep{saopaulofleet}			& \citepalias{saopauloarea}		& 			\\
  \hline
- 30 m & 160 & 14.689.750 & 0.0154 \\
+ New York (US) 		& 2,162,349  					& 777,934,030  					& 0.4366 	\\
+ 					& \citepalias{newyorkfleet}		& \citepalias{usarea}			&			\\
  \hline
- 50 m & 160 & 14.689.750 & 0.0428 \\
+ Los Angeles (US) 	& 8,050,850  					& 1,213,820,883  				& 1.041 	\\
+ 					& \citepalias{losangelesfleet}	& \citepalias{usarea}			&			\\
  \hline
- 100 m & 160 & 14.689.750 & 0.1604 \\
+ Tokyo (JP) 		& 3,159,455  					& 2,191,000,000  				& 0.2265 	\\
+ 					& \citep{tokyofleet}			& \citepalias{tokyoarea}		&			\\
  \hline
- 150 m & 160 & 14.689.750 & 0.3609 \\
- \hline
- 200 m & 160 & 14.689.750 & 0.6416 \\
+ London (UK)		& 3,091,393						& 1,572,100,000					& 0.3089 	\\
+ 					& \citepalias{londonfleet}		& \citepalias{londonarea}		&			\\
  \hline
@@
 To improve readability, all figures in this section follow the same format:
-$X$ axis shows the results of sequential iterations, ranging from 0 to 8639;
-$Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
-blue line represents the percentage of nodes detected out of the complete network;
-magenta is the percentage of malicious nodes in the network (ground truth);
-green represents the nodes correctly identified as malicious (true positives);
-%cyan represents the undetected malicious nodes (false negatives); 
-red represents the benign nodes incorrectly identified as malicious (false positives).
-The lines represent average values between all benign nodes on the network, while the colored regions represent the standard deviation present in the data.
+\begin{itemize}
+	\item The $X$ axis shows the results of sequential iterations, ranging from 0 to 8639 in most cases;
+	\item The $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100;
+	\item The blue line represents the percentage of nodes detected out of the complete network;
+	\item The magenta line is the percentage of malicious nodes in the network (ground truth);
+	\item The green line represents the nodes correctly identified as malicious (true positives);
+	\item When present, the cyan line represents the undetected malicious nodes (false negatives); 
+	\item The red line represents the benign nodes incorrectly identified as malicious (false positives);
+	\item The lines represent average values between all benign nodes on the network, while the colored regions represent the standard deviation present in the data.
+\end{itemize}
 

### 2018-03-07 dissertation/1-intro/chapter1.tex
2018-03-06

@@
 As computers grow in power and shrink in size, more aspects of everyday life can be enhanced by adding processing units to common devices.
-While many of these applications focus on conveniences, such as home automation \cite{mccole2016} (the collection of connected and smart devices is dubbed the Internet of Things or IoT \cite{morgan2014}), the integration of computers with other objects and devices can also be important to save time and save lives \cite{rti2014}.
-One way of achieving this is by adding computers and wireless transmitters to vehicles — such as cars, buses, and trains — so they can share data which may increase traffic efficiency or reduce the chance of accidents \cite{saini2015close}.
+While many of these applications focus on conveniences, such as home automation \citep{mccole2016} (the collection of connected and smart devices is dubbed the Internet of Things or IoT \citep{morgan2014}), the integration of computers with other objects and devices can also be important to save time and save lives \citep{rti2014}.
+One way of achieving this is by adding computers and wireless transmitters to vehicles — such as cars, buses, and trains — so they can share data which may increase traffic efficiency or reduce the chance of accidents \citep{saini2015close}.
 
-In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \cite{whotraffic}.
-While this number has greatly reduced over the past decades \cite{johnson2010traffic} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \cite{whofactsheet}, so further actions are necessary.
-Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \cite{inrixtraffic}.
+In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \citep{whotraffic}.
+While this number has greatly reduced over the past decades \citep{johnson2010traffic} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \citep{whofactsheet}, so further actions are necessary.
+Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \citep{inrixtraffic}.
 
 Smart vehicles and vehicular networks are ways that technology can aid both of the aforementioned problems.
-Through the use of sensors and wireless communications, these vehicles are able to avoid accidents by alerting distracted drivers \cite{lee2004collision}, or by knowing in advance another vehicle's position and speed \cite{hafner2011automated}.
-By communicating, they can also collaborate to offer driving and route suggestions, therefore reducing the possibility of traffic jams \cite{knorr2012reducing}.
+Through the use of sensors and wireless communications, these vehicles are able to avoid accidents by alerting distracted drivers \citep{lee2004collision}, or by knowing in advance another vehicle's position and speed \citep{hafner2011automated}.
+By communicating, they can also collaborate to offer driving and route suggestions, therefore reducing the possibility of traffic jams \citep{knorr2012reducing}.
 %These features are possible with the development of a vehicular ad-hoc network (VANET), in which vehicles can quickly share data amongst themselves without the need of a server or an Internet connection.
 
-When dealing with safety or traffic-efficiency applications, it is crucial that network communications occur with low latency (approximately 100 milliseconds \cite{camp2005vehicle}).
-Current cellular technology, such as LTE, could be used to connect vehicles to the Internet, but the delay added by the transmission would make safety applications unfeasible or unreliable \cite{mangel2010comparison}.
+When dealing with safety or traffic-efficiency applications, it is crucial that network communications occur with low latency (approximately 100 milliseconds \citep{camp2005vehicle}).
+Current cellular technology, such as LTE, could be used to connect vehicles to the Internet, but the delay added by the transmission would make safety applications unfeasible or unreliable \citep{mangel2010comparison}.
 Cellular connections also have other problems: the connection would require an active subscription with a carrier; the connection depends on available infrastructure; the wireless frequency would be shared with phones and other mobile devices, increasing the possibility of interference and congestion; server-side issues could impact the vehicles' communications.
@@
 For these reasons, ad-hoc solutions are preferred over centralized ones.
-An ad-hoc network is one that has no reliance on pre-existing infrastructure (such as routers or access points) \cite{wu2004ad}.
+An ad-hoc network is one that has no reliance on pre-existing infrastructure (such as routers or access points) \citep{wu2004ad}.
 Instead, each node is able to communicate directly with others and a routing protocol allows for messages to be forwarded until they reach their destinations.
 Every time a node wants to send a message and the recipient is not a direct neighbor, it must choose which nearby node is the most likely one to get the message to its destination.
-Routing techniques can use either the network's topology or geographical coordinates \cite{saini2015close} to choose which node should be the next hop.
+Routing techniques can use either the network's topology or geographical coordinates \citep{saini2015close} to choose which node should be the next hop.
 
 These issues — the additional safety and efficiency as well as the low-latency communications — can be tackled through the use of a vehicular ad-hoc network (VANET), in which vehicles share data amongst themselves without relying on external devices, an Internet connection or server availability.
-Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \cite{amoozadeh2015platoon}.
-In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \cite{li2007routing}.
+Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \citep{amoozadeh2015platoon}.
+In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \citep{li2007routing}.
 That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.
@@
 
-As is the case with most new technologies, VANETs are expected to be a notable target of attacks for a diversity of reasons \cite{isaac2010security}.
-A local malicious user might alter the data his or her vehicle broadcasts in order to manipulate traffic conditions, while remote attackers could invade vehicles' computers and obtain partial control of the network  \cite{garip2015congestion}.
+As is the case with most new technologies, VANETs are expected to be a notable target of attacks for a diversity of reasons \citep{isaac2010security}.
+A local malicious user might alter the data his or her vehicle broadcasts in order to manipulate traffic conditions, while remote attackers could invade vehicles' computers and obtain partial control of the network  \citep{garip2015congestion}.
 These attacks can vary from time-consuming annoyances to life-threatening, so it is important that real-world implementations of vehicular networks are prepared to handle them.
@@
 Furthermore, once nodes form their own opinions about others, they can propagate pre-existing trust values when necessary.
-For example, if two nodes are not direct neighbors and do not have any pre-existing trust information about each other, they can ask intermediary nodes for their opinions on the other node \cite{wang2009trust}.
+For example, if two nodes are not direct neighbors and do not have any pre-existing trust information about each other, they can ask intermediary nodes for their opinions on the other node \citep{wang2009trust}.
 The management of trust values (i.e.
-how one node acquires and updates trust values) and the use of these values to derive further information (such as designating nodes as malicious or not) is called \textit{trust management} \cite{ma2011survey}.
+how one node acquires and updates trust values) and the use of these values to derive further information (such as designating nodes as malicious or not) is called \textit{trust management} \citep{ma2011survey}.
 %The combination of one node's trust of another with the opinions of other nearby nodes form what is called \textit{trust management}.
@@
 While the detection of incorrect nodes and/or messages is an important aspect of security and safety in vehicular networks, it does not address all of the problems.
-Trust solutions are not viable without a solid identity verification scheme, for instance, since nodes would not be able to form trust opinions without being sure of the others' identities (these schemes often use a Public Key Infrastructure \cite{wasef2010complementing}).
+Trust solutions are not viable without a solid identity verification scheme, for instance, since nodes would not be able to form trust opinions without being sure of the others' identities (these schemes often use a Public Key Infrastructure \citep{wasef2010complementing}).
 They also do not address issues such as driver and passenger privacy when using the facilities of a VANET.
@@
 %While this work emphasizes on the characteristics of VANETs, they are only a subset of the much broader field of complex networks.
-%The study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}.
+%The study of complex networks dates back decades, including analysis of biological or social networks \citep{wang2003complex} \citep{newmannetworks}.
 %Trust issues are present in many of these types of networks, although each one requires a specific approach that addresses the peculiarities involved.
@@
 %Such networks are only a subset of a much larger field of study: that of complex networks.
-%The study of complex networks dates back decades, including analysis of biological or social networks \cite{wang2003complex} \cite{newmannetworks}.
+%The study of complex networks dates back decades, including analysis of biological or social networks \citep{wang2003complex} \citep{newmannetworks}.
 %Technological networks, and VANETs by extension, are only a subset of that much larger field of study.
@@
 This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.
-The solution is based on a previously existing algorithm \cite{vernize2015malicious}, which was limited to centralized networks.
+The solution is based on a previously existing algorithm \citep{vernize2015malicious}, which was limited to centralized networks.
 In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.

### 2018-03-07 dissertation/2-complexnetworks/chapter2.tex
2018-03-06

@@
 
-Complex networks can describe many systems which are observed in nature and society through a collection of \textit{vertices} (or \textit{nodes}) and \textit{edges} \cite{newmannetworks}.
+Complex networks can describe many systems which are observed in nature and society through a collection of \textit{vertices} (or \textit{nodes}) and \textit{edges} \citep{newmannetworks}.
 They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).
@@
 
-Complex networks are generally divided into four categories \cite{newmannetworks}:
+Complex networks are generally divided into four categories \citep{newmannetworks}:
 
@@
 Trust is a concept studied in fields such as psychology and economics, with specific definitions.
-In complex networks, under the perspective of computer science, it is a measure of one entity's confidence that another will behave properly and provide valid and/or meaningful data \cite{sherchan2013survey}.
+In complex networks, under the perspective of computer science, it is a measure of one entity's confidence that another will behave properly and provide valid and/or meaningful data \citep{sherchan2013survey}.
 What follows is the basis of how a network can be modelled using a graph and how a trust model can be applied to it.
@@
 There are two types of social networks: real-world ones formed by relationships between people, and online ones that attempt to abstract the former into a digital environment.
-Examples of the first one are all around, present in any family, workplace, school or group of friends \cite{newmannetworks}.
+Examples of the first one are all around, present in any family, workplace, school or group of friends \citep{newmannetworks}.
 Online social networks started by connecting people who already knew each other and giving them an additional form of interaction (analogous to what telephones and email did before), but, today, it is not unusual for people to form relationships with others whom they have only met online.
@@
 For instance, upon reading an online review for a certain product, a user will be more likely to accept the review's conclusion if it was written by a close friend than if it were written by a stranger.
-Social trust is a way of estimating how much a certain recommendation will lead to a positive outcome \cite{golbeck2006inferring}.
+Social trust is a way of estimating how much a certain recommendation will lead to a positive outcome \citep{golbeck2006inferring}.
 
@@
 information shared by a close friend of a person might be considered almost as trustworthy as some collected by the person him or herself.
-Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust \cite{newmannetworks}.
-Expanding on that property, there is the concept friends of friends \cite{boissevain1974friends}.
+Therefore, it is possible to model social trust relationships as a graph, in which nodes represent people and edges represent a certain degree of trust \citep{newmannetworks}.
+Expanding on that property, there is the concept friends of friends \citep{boissevain1974friends}.
 If, for example, nodes $A$ and $B$ have mutual trust and are considered friends, then it is reasonable to assume that some of $A$'s trust for $B$ carries over to other nodes that enjoy mutual trust with $B$.
@@
 
-Another important aspect of trust in ad-hoc networks is that information is, generally, uncertain and incomplete \cite{baras2005cooperation}.
+Another important aspect of trust in ad-hoc networks is that information is, generally, uncertain and incomplete \citep{baras2005cooperation}.
 That is, since nodes form their own model and opinions of the surrounding network, it is unlikely that this data will be certain and accurate with reality.
@@
 There are few examples of MANETs implemented for consumer devices.
-Two examples are networks created to quickly share data between devices using Wi-Fi or Bluetooth \cite{krochmal2014discovery}, or ones that allow for multiplayer gaming sessions amongst multiple nearby devices \cite{sasaki2011wireless}.
+Two examples are networks created to quickly share data between devices using Wi-Fi or Bluetooth \citep{krochmal2014discovery}, or ones that allow for multiplayer gaming sessions amongst multiple nearby devices \citep{sasaki2011wireless}.
 In both cases, there is no need for a complicated system-level trust model, since those activities involve active participation from the users wielding the device (that is, the user chooses whether or not to communicate with other devices); rather, the trust relationship occurs socially amongst the users themselves.

### 2018-03-07 dissertation/3-vanets/chapter3.tex
2018-03-06

@@
 
-Today, most premium vehicles come equipped with hardware that allow for connectivity features; it is expected that, by 2022, many standard vehicles will also come with such features built-in, accounting for a substantial share of the automotive industry's revenue \cite{connectedcar2016}.
-Although these features can be useful tools to aid drivers, reducing traffic and risk of accidents, they are merely a gateway to the long-term goal of truly autonomous vehicles, which might become a reality within the next decade; many automakers and technology companies have laid out their plans for the upcoming years \cite{tesla2016}.
+Today, most premium vehicles come equipped with hardware that allow for connectivity features; it is expected that, by 2022, many standard vehicles will also come with such features built-in, accounting for a substantial share of the automotive industry's revenue \citep{connectedcar2016}.
+Although these features can be useful tools to aid drivers, reducing traffic and risk of accidents, they are merely a gateway to the long-term goal of truly autonomous vehicles, which might become a reality within the next decade; many automakers and technology companies have laid out their plans for the upcoming years \citep{tesla2016}.
 However, the proper functioning and utility of both connected and fully autonomous vehicles rely on technologies, protocols and applications that allow for the fast communication between vehicle's on-board computers.
@@
 In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
-By quickly sharing data with neighboring vehicles, without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.
+By quickly sharing data with neighboring vehicles, without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \citep{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \citep{amoozadeh2015platoon}.
 
@@
 %This approach requires appropriate routing protocols that consider the dynamic topology of VANETs.
-%While traditional MANETs often use protocols which rely on the network topology, the ones proposed for VANETs generally use geographic position instead \cite{saini2015close}; these protocols use GPS coordinates in order to send messages to the available neighbor physically closest to the destination node.
+%While traditional MANETs often use protocols which rely on the network topology, the ones proposed for VANETs generally use geographic position instead \citep{saini2015close}; these protocols use GPS coordinates in order to send messages to the available neighbor physically closest to the destination node.
 
@@
     \includegraphics[width=0.5\textwidth]{images/vanet.png}
-    \caption{Basic elements of a VANET: OBUs and RSUs. \cite{saini2015close}}
+    \caption{Basic elements of a VANET: OBUs and RSUs. \citep{saini2015close}}
     \label{fig:vanet}
@@
 
-Several current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
+Several current efforts to make VANETs viable in cities are centered around the IEEE 802.11p standard, also called Wireless Access in Vehicular Environments (WAVE) \citep{jiang2008ieee}.
 Among other aspects of the wireless technology, the WAVE standard describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
@@
 Road-side units are placed in static locations near roads; they may also have wired interfaces with other RSUs and the Internet, so it is possible to use them as anchor points for Internet access for passing vehicles.
-When referring to the communication between two OBUs, the term vehicle-to-vehicle (V2V) communication is used \cite{yang2004vehicle}; when both an OBU and an RSU are involved, it is called vehicle-to-infrastructure (V2I) communication \cite{chou2009feasibility}.
+When referring to the communication between two OBUs, the term vehicle-to-vehicle (V2V) communication is used \citep{yang2004vehicle}; when both an OBU and an RSU are involved, it is called vehicle-to-infrastructure (V2I) communication \citep{chou2009feasibility}.
 Although this nomenclature is important to understand other studies on the subject of VANETs, this study does not consider RSUs and focuses only on vehicles themselves as nodes, so references to VANETs and vehicular networks are exclusively tied to V2V communications.
@@
 This metric is not as useful in vehicular networks, since the high mobility causes the topology to change frequently.
-Instead, most VANET routing protocols use geographical coordinates to forward packets \cite{saini2015close}, that is, the physical distance between two nodes is used as the primary metric.
+Instead, most VANET routing protocols use geographical coordinates to forward packets \citep{saini2015close}, that is, the physical distance between two nodes is used as the primary metric.
 The implication is that, even if a packet requires more hops to reach its destination, it will always be traveling the generally correct direction.
@@
 			If a vehicle is broadcasting incorrect data (perhaps unknowingly) because of a hardware or software fault, it can be a serious hinderance to efficiency and safety applications.
-			It might behaving appropriately according to protocol, but the data it sends is not reliable \cite{isaac2010security}.
+			It might behaving appropriately according to protocol, but the data it sends is not reliable \citep{isaac2010security}.
 	\item	Vehicles might be deliberately broadcasting false data.
-			In this case, there might be a specific purpose (by either the vehicle's driver or a remote attacker), like altering traffic or even cause an accident \cite{golle2004detecting}.
+			In this case, there might be a specific purpose (by either the vehicle's driver or a remote attacker), like altering traffic or even cause an accident \citep{golle2004detecting}.
 	\item   Attackers with control of several vehicles can propagate junk data in an attempt to flood the network, causing a distributed denial of service (DDoS) attack.
-			Alternatively, the data propagated might have some reasoning behind it, like lying about road conditions in order to divert traffic \cite{garip2015congestion}.
+			Alternatively, the data propagated might have some reasoning behind it, like lying about road conditions in order to divert traffic \citep{garip2015congestion}.
 	\item	Instead of sending data, some vehicles might try to eavesdrop on others' communications.
@@
 			If the intermediary node is malicious, it may attempt to extract data contained in messages or refuse to forward them.
-			Related to this, there is the Sybil attack \cite{isaac2010security}, in which a node lies about its position in order to seemingly alter the physical topology of the network and be chosen as a hop \cite{leinmuller2005influence}.
-	\item	Malicious vehicles may use signal jammers or other devices in order to affect other vehicles' sensors and communications \cite{isaac2010security}.
+			Related to this, there is the Sybil attack \citep{isaac2010security}, in which a node lies about its position in order to seemingly alter the physical topology of the network and be chosen as a hop \citep{leinmuller2005influence}.
+	\item	Malicious vehicles may use signal jammers or other devices in order to affect other vehicles' sensors and communications \citep{isaac2010security}.
 			That can cause other vehicles to broadcast incorrect data, therefore obfuscating the origin of the attack.
-	\item	A malicious user or remote attacker can monitor messages shared across the network in an attempt to stalk one specific vehicle \cite{isaac2010security}.
+	\item	A malicious user or remote attacker can monitor messages shared across the network in an attempt to stalk one specific vehicle \citep{isaac2010security}.
 
@@
 %Related to trust is the issue of identity in vehicular networks.
-%Solutions to this issue can be important elements of security and privacy, and there are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}.
-%In fact, at least one study proposes that these identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+%Solutions to this issue can be important elements of security and privacy, and there are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \citep{wasef2010complementing}.
+%In fact, at least one study proposes that these identities must be refreshed with a certain frequency to maintain the privacy of each node \citep{golle2004detecting}.
 %However, refreshing identities can make maintaining a trust relationship between two nodes difficult or even impossible, since there is no way to know which nodes changed identities.
@@
 % While certification of identity is an issue in vehicular networks, and , this work will not address it.
-%There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \cite{wasef2010complementing}
-%\cite{kumar2015intelligent}.
-%One study proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \cite{golle2004detecting}.
+%There are studies focusing on node identification in VANETs (that is, making sure a node is who it claims to be), in particular through the use of a Public Key Infrastructure (PKI) \citep{wasef2010complementing}
+%\citep{kumar2015intelligent}.
+%One study proposes that such identities must be refreshed with a certain frequency to maintain the privacy of each node \citep{golle2004detecting}.
 %Although it is important that nodes are not able to lie about identity and are able to maintain their privacy, nodes having valid identities that are persistent over time, so that nodes can identify each other over a long period, is a prerequisite for the model proposed here.
@@
 %However, 
-VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \cite{yousefi2006vehicular}.
+VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \citep{yousefi2006vehicular}.
 Some of these properties include:
@@
 There is an important distinction between a malicious node and a faulty one; both of them may be sharing false data, but for different reasons and with different consequences.
-For example, a malicious node may lie about its location in order to make routing protocols use it \cite{leinmuller2005influence}, in order to try to store or alter messages, while a faulty GPS module may cause an accident because its position data was incorrect.
+For example, a malicious node may lie about its location in order to make routing protocols use it \citep{leinmuller2005influence}, in order to try to store or alter messages, while a faulty GPS module may cause an accident because its position data was incorrect.
 However, that distinction can be hard to make, because a close inspection is necessary to determine whether the incorrect data is erratic or deliberate.
@@
 In general, trust management solutions for VANETs use \textit{data-oriented trust}, \textit{entity-based trust}, or a combination of the two.
-The solutions that use data-oriented trust (or \textit{data-centric trust}) \cite{raya2008data} focus on validating messages instead of entities.
+The solutions that use data-oriented trust (or \textit{data-centric trust}) \citep{raya2008data} focus on validating messages instead of entities.
 This is important when vehicles share messages about a specific event, such as a collision, which must be quickly validated by neighbors and distributed to other nodes within a relevant area.
@@
 %\textbf{/*********}
-%In \cite{gerlach2007trust}, the author shows a general framework of how security and trust can be organized in a vehicular network.
+%In \citep{gerlach2007trust}, the author shows a general framework of how security and trust can be organized in a vehicular network.
 %It proposes three layers: a service plane, which contains the core functionalities of the network, such as communication technologies, routing algorithms and the security measures directly related to them (such as cryptography); a middleware plane which handles trust and privacy issues; and the application layer, which contains the applications that will use the other layers' data to make decisions.
@@
 
-The analysis of related work is based on \cite{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
+The analysis of related work is based on \citep{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
 In this section, these properties are described with an assessment of whether or not TruMan satisfies their conditions.
@@
  \hline
+ \hline
  TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
 % \hline
-% \cite{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+% \citep{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
  \hline
- \cite{dotzer2005vars} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \citep{dotzer2005vars} & \checkmark & - & \checkmark & - & - & - & \checkmark & -\\
  \hline
- \cite{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
+ \citep{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
  \hline
- \cite{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
+ \citep{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
  \hline
- \cite{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \citep{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
  \hline
- \cite{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \citep{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
  \hline
- \cite{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
+ \citep{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
  \hline
- \cite{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
+ \citep{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
  \hline
@@
 None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
-Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
-There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.
+Many trust management solutions for VANETs have been proposed over the years, such as \citep{patwardhan2006data}, \citep{gerlach2007trust}, \citep{raya2008data}, \citep{huang2010situation}, \citep{ding2013novel}, \citep{haddadou2013trust}, \citep{liu2016lsot}, \citep{kerrache2016detection}.
+There are also some review and/or survey articles on the subject of VANET trust models, such as \citep{zhang2011survey}, \citep{ma2011survey}, \citep{zhang2012trust}, \citep{mejri2014survey}, \citep{soleymani2015trust} \citep{sengar2016survey}, and \citep{dwivedi2016review}.
 
-\cite{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
+\citep{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
 The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents.
@@
 
-The model proposed in \cite{minhas2010towards} uses several criteria to judge whether or not a received message is trustworthy.
+The model proposed in \citep{minhas2010towards} uses several criteria to judge whether or not a received message is trustworthy.
 First, nodes are classified by their roles and previous experience with them.
@@
 
-In \cite{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
+In \citep{chen2010trust}, the authors propose to evaluate messages utilizing a cluster-based trust model.
 By separating nodes into clusters with their geographical neighbors, it is possible to efficiently distribute the evaluation of messages using previously formed opinions.
@@
 
-The trust model in \cite{park2011long} takes advantage of daily commutes.
+The trust model in \citep{park2011long} takes advantage of daily commutes.
 In this article, the focus is on the early stages of VANETs, in which a very small percentage of vehicles are equipped with OBUs.
@@
 
-The authors of \cite{huang2014social} take special note of two characteristics from social networks that can also be found in many VANET trust models: \textit{information cascading} and \textit{oversampling}.
+The authors of \citep{huang2014social} take special note of two characteristics from social networks that can also be found in many VANET trust models: \textit{information cascading} and \textit{oversampling}.
 That is, information reported by a number of original nodes (i.e. the ones that witnessed an event) may be diluted as nodes that forward it append their own opinions on the matter.
@@
 
-The ART model proposed in \cite{li2016art} works in two main steps: data gathering and malicious node detection.
+The ART model proposed in \citep{li2016art} works in two main steps: data gathering and malicious node detection.
 It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
@@
 
-The authors of \cite{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
+The authors of \citep{chen2017cloud} propose a cloud-based solution for a trust model, which requires an Internet-based global trust manager.
 This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations in which mobile communication is slow or unreliable.
@@
 % \hline
-% \textbf{Property} & TruMan & \cite{vernize2015malicious} & \cite{minhas2010towards} & \cite{chen2010trust} & \cite{li2016art} & \cite{chen2017cloud} \\
+% \textbf{Property} & TruMan & \citep{vernize2015malicious} & \citep{minhas2010towards} & \citep{chen2010trust} & \citep{li2016art} & \citep{chen2017cloud} \\
 % \hline

### 2018-03-07 dissertation/4-project/chapter4.tex
2018-03-06

@@
 First, it is shown how vehicles can form long-term relationships and trust one another in a similar way to social networks.
-Then, two algorithms are introduced: Tarjan's strongly connected components algorithm \cite{tarjan1972depth} and an efficient algorithm for graph coloring \cite{mittal2011graph}.
-Next, the Malicious Node Identification Algorithm (MaNI) \cite{vernize2013dissertation} is explained, for it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
+Then, two algorithms are introduced: Tarjan's strongly connected components algorithm \citep{tarjan1972depth} and an efficient algorithm for graph coloring \citep{mittal2011graph}.
+Next, the Malicious Node Identification Algorithm (MaNI) \citep{vernize2013dissertation} is explained, for it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
 Finally, TruMan itself is explained, showing how it combines features from existing algorithms and adapts them to a dynamic environment, enabled by the social properties found in vehicular networks.
@@
 \label{section:goals}
-Building from the foundation set by MaNI \cite{vernize2013dissertation}, TruMan strives to enable efficient trust management in highly dynamic networks such as VANETs.
+Building from the foundation set by MaNI \citep{vernize2013dissertation}, TruMan strives to enable efficient trust management in highly dynamic networks such as VANETs.
 In addition to efficiency, it is desirable that the trust model is both simple to understand and to implement, making it appealing for real-world usage.
 
-Furthermore, the desired properties of a VANET trust model described in \cite{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.
+Furthermore, the desired properties of a VANET trust model described in \citep{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.
 
@@
 \label{section:socialvanets}
-Some proposed trust models for vehicular networks, such as \cite{huang2014social}, state that the likelihood of two nodes meeting each other twice is too low to be relevant.
+Some proposed trust models for vehicular networks, such as \citep{huang2014social}, state that the likelihood of two nodes meeting each other twice is too low to be relevant.
 However, it stands to reason that, throughout the course of several days, many drivers take similar routes at similar times of day (e.g.
@@
 
-In \cite{da2013effective}, \cite{cunha2014vehicular}, and \cite{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
+In \citep{da2013effective}, \citep{cunha2014vehicular}, and \citep{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
 By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
@@
 \label{section:tarjan}
-The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of TruMan's efficiency.
+The use of Tarjan's strongly connected components algorithm \citep{tarjan1972depth} is an important aspect of TruMan's efficiency.
 This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further steps.
@@
 
-Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
+Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.
 
@@
 \label{section:coloring}
-%The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
+%The algorithm proposed in \citep{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
 Graph coloring is one of the possible heuristics used by TruMan to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
@@
 The process of graph coloring consists of giving each node a label (represented by a color) so that no two neighboring nodes share the same color.
-This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
-It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
+This problem has been studied in Computer Science since, at least, 1972 \citep{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \citep{kempe1879geographical}.
+It has been proven mathematically that any planar graph can be colored with at most four colors \citep{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \citep{sanchez1989determining}.
 
-In \cite{mittal2011graph}, the authors present an efficient approach to graph coloring using the minimum possible amount of colors.
+In \citep{mittal2011graph}, the authors present an efficient approach to graph coloring using the minimum possible amount of colors.
 Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
@@
 The complexity of the algorithm is $O(|E'|)$ for a graph $C = (V',E')$.
-As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
+As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \citep{brelaz1979new}.
 
-Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph}.
+Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.
 
@@
 It doesn't matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
-This is demonstrated in \cite{vernize2013dissertation}.
+This is demonstrated in \citep{vernize2013dissertation}.
 
@@
 
-The basis of TruMan is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, which suggests the use of strongly connected components and graph coloring for malicious node detection.
+The basis of TruMan is the Malicious Node Identification Algorithm (MaNI) proposed in \citep{vernize2015malicious}, which suggests the use of strongly connected components and graph coloring for malicious node detection.
 This article presents a malicious node identification scheme based on strongly connected components and graph coloring.
@@
 
-First, $T$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
+First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
 In each of these components, all nodes are connected by edges of value $1$.
@@
 
-The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \cite{brelaz1979new} or the algorithm proposed in \cite{mittal2011graph}.
+The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \citep{brelaz1979new} or the algorithm proposed in \citep{mittal2011graph}.
 Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
@@
 
-In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
+In the experiments shown in \citep{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
 Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
@@
 
-TruMan is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm.
+TruMan is based on the MaNI algorithm \citep{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm.
 However, MaNI was developed for static networks such as social networks, and is executed by an external supervising agent (i.e. outside of the network), making it unsuitable for a vehicular network.
@@
 
-After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}.
+After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}.
 For each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
@@
 
-The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}, although other heuristics may be considered.
+The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \citep{mittal2011graph}, although other heuristics may be considered.
 After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
@@
 %This paragraph is probably pertinent to delimiting the size of graphs:\\
-%The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
+%The authors of \citep{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
 %Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
@@
 %
-%The authors of \cite{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
+%The authors of \citep{patwardhan2006data} provide important observations regarding the collection of data from local conditions.
 %Most crucially, they mention how the quality of information is higher when the origin of such information is close to the destination node.
@@
 %It might still be useful to simulate using real-world movement traces.
-%An example of such dataset is the Zurich Trace \cite{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
+%An example of such dataset is the Zurich Trace \citep{zurichtrace}, which describes the vehicular traces in the city of Zurich, Switzerland.
 %The chosen dataset doesn't necessarily need to be created for VANETs, since with vehicular traces it is possible to assume all or a percentage of the traced vehicles are smart and can be nodes of a hypothetical VANET.

### 2018-03-26 dissertation/1-intro/chapter1.tex
updated chapters 1 and 2

@@
 
-This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.
-The solution is based on a previously existing algorithm \citep{vernize2015malicious}, which was limited to centralized networks.
-In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
-These features are found by tracing analogies to social networks, i.e.
+The study of networks is tied to graph theory, since graphs are a useful way to generate models of networks, and therefore many concepts of the two fields overlap.
+Mathematical, computational and statistical concepts developed for graphs can be translated to be useful for a variety of different types of networks.
+%, like, for example, finding articulation vertices in a graph to locate bandwidth bottlenecks in a computer network.
+Therefore, this document utilizes notation from graph theory in order to represent various kinds of networks.
+
+This work introduces TruMan, a trust management solution for vehicular networks. 
+TruMan combines solutions to two classic graph theory problems, strongly connected components and graph coloring, in order to provide an efficient approach to identifying malicious nodes in a dynamic and decentralized network.
+This is based on a previously existing algorithm, called MaNI \citep{vernize2015malicious}, which is limited to centralized networks.
+
+In order to function in a dynamic and decentralized environment, TruMan takes advantage of features that allow the development of trust relationships between members of the network.
+Since there is no unifying agent supervising the network, trust relationships are built over time, as nodes move around the network and meet other nodes.
+
+These features are discovered by tracing analogies with social networks.
+There might be groups of nodes which meet each other with a predictable frequency and at predictable intervals, for example vehicles which belong to family members, neighbors or work colleagues.
+By considering such features, TruMan allows the formation of strong trust relationships,  which in turn facilitate the discovery of malicious members of the network.
+
+%This work proposes a method for identifying malicious nodes in a vehicular network, although it may also be viable for other dynamic and decentralized networks.
+
+%In order to adapt the algorithm to a dynamic and decentralized environment, it is necessary to locate network features that allow for trust relationships to be developed amongst the network members over time, since there is no centralizing agent to store and process the whole network's trust values.
+%These features are found by tracing analogies to social networks, i.e.
 identifying how and why two nodes (in the case of VANETs, vehicles) would meet more than once and at a somewhat predictable interval.
@@
 
-\section{Document organization}
-The remainder of this dissertation is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
+%\section{Document organization}
+The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
 Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.

### 2018-03-26 dissertation/2-complexnetworks/chapter2.tex
updated chapters 1 and 2

@@
 They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).
-The study of networks is tied to graph theory, since graphs are a useful way to generate models of networks, and therefore many concepts of the two fields overlap.
-Mathematical, computational and statistical concepts developed for graphs can be translated to be useful for a variety of different types of networks, like, for example, finding articulation vertices in a graph to locate bandwidth bottlenecks in a computer network.
 
@@
 		Online social networks can also be classified as information networks, since their nodes are actually information about people rather than the people themselves.
-		Trust in information networks can be observed in some instances, like peer-to-peer networks [cite], although its usages are not relevant for this work.
+		Trust in information networks can be observed in some instances, like peer-to-peer networks, although its usages are not relevant for this work.
 	\item \textbf{Biological Networks} are the networks found in nature.
 		Their nodes can be chemicals, cells, animals, groups of lifeforms, and more.
-		An example is the brain, which contains a neural network formed by neurons, cells which enable information processing; connections in the network represent signals that are sent from one neuron to another.
+		An example is the brain, which contains a neural network formed by neurons; connections in the network represent signals that are sent from one neuron to another.
 		Another instance of biological networks are food chains, categorized as ecological webs.
@@
 
-Consider an undirected graph $G = (V,E)$, which models one complex network of any kind.
+Consider an undirected graph $G = (V, E_G)$, which models one complex network of any kind.
 The vertices are the members of the network (computers, humans, etc.) and each edge represents a pair of vertices' ability to exchange data freely.
 This graph represents the network's \textit{topology}, that is, the basic structure of the network.
-Then, there is also a directed graph $T = (V, O)$, called a \textit{trust graph}.
-$T$ contains the same nodes as $G$, although its edges represent the degree of trust (or opinion) each node has towards another.
+Then, there is also a directed graph $T = (V, E)$, called a \textit{trust graph}.
+$T$ contains the same vertices as $G$, although its edges represent the degree of trust (or opinion) each node has towards another.
 
-There are two main ways to describe the edges in $O$: they can be binary, either existing when there is trust or not otherwise, or they can hold a specific trust value within a certain range.
-This means the shape of $T$ is not necessarily similar to that of $G$.
-For instance, two people can have contact with each other but not maintain a trust relationship, thus altering the layout of the trust graph compared to the network topology.
+There are two main ways to describe the edges in $E$: they can be binary, either existing when there is trust or not otherwise, or they can hold a specific trust value within a certain range.
+In some cases, an edge $a\rightarrow b$ with value 0 indicates lack of trust, meaning $a$ has no reason to trust $b$.
+In others, it indicates \textit{negative trust}, meaning $a$ has reason to believe $b$ is malicious.
 
-In the following two sections of this chapter, trust in the contexts of social and technological networks is further explained.
+Since $G$ and $T$ represent different types of information, the shape of $T$ is not necessarily similar to that of $G$, even though they share the same vertices.
+For instance, two people can have contact with each other (therefore sharing an edge in $G$) but not maintain a trust relationship (therefore not sharing an edge in $T$), thus altering the layout of the trust graph compared to the network topology.
+
+In the following two sections of this chapter, trust is further explained in the contexts of social and technological networks.
 Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
-Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which is expanded upon in \autoref{chap:vanets}.
+Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \autoref{chap:vanets}.
 
@@
 Expanding on that property, there is the concept friends of friends \citep{boissevain1974friends}.
-If, for example, nodes $A$ and $B$ have mutual trust and are considered friends, then it is reasonable to assume that some of $A$'s trust for $B$ carries over to other nodes that enjoy mutual trust with $B$.
+If, for example, nodes $a$ and $b$ have mutual trust and are considered friends, then it is reasonable to assume that some of $a$'s trust for $b$ carries over to other nodes that enjoy mutual trust with $b$.
 In other words, a friend of a friend can be considered more trustworthy than the average stranger.
 This property is similar, although not identical, to transitivity, since trust is diminished for each extra step an origin node needs to reach a destination, and there is also the possibility that one node distrusts another even if they share a mutual friend.
-Naturally, social trust is not commutative ($A$ trusting $B$ does not imply that $B$ trusts $A$).
+Naturally, social trust is not commutative ($a$ trusting $b$ does not imply that $b$ trusts $a$).
 

### 2018-03-28 dissertation/3-vanets/chapter3.tex
corrections to chapter 3

@@
 The analysis of related work is based on \citep{zhang2011survey}, which proposes eight desired properties for a trust management model for VANETs.
-In this section, these properties are described with an assessment of whether or not TruMan satisfies their conditions.
-\autoref{table:properties} shows how well TruMan and previous work satisfy the properties.
+In this section, these properties are briefly described.
+% with an assessment of whether or not TruMan satisfies their conditions.
+%\autoref{table:properties} shows how well TruMan and previous work satisfy the properties.
 
-\textit{1. Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes.
-Nodes may or may not use information from other trustworthy nodes to build trust values.
-TruMan satisfies this as it is built from the ground up for decentralized systems.
+\textit{1. Decentralized trust establishment}: nodes must be able to form their own trust values about other nodes without the aid of an Internet connection or centralizing agents.
+Nodes may or may not use information from other trustworthy nodes to build trust values (in other words, trust might be transitive).
+%TruMan satisfies this as it is built from the ground up for decentralized systems.
 
 \textit{2. Coping with sparsity}: the model still functions when there are few nodes populating the network.
-The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
-Due to its decentralized nature, it can also work on isolated chunks of the network.
+Due to the dynamic nature of vehicular networks, it is possible that nodes will find themselves with few other nodes in range.
+In such scenarios, a trust model should be able to establish trust even if there are few neighbors with whom to share data.
+%In some cases, this can be achieved by using information that is immediately available; for example, if a vehicle broadcasting information happens to be a certified ambulance, its role is enough to guarantee trust.
+%In other cases, models must consider solutions similar to delay-tolerant networks: storing information as it is received and sharing it with other trustworthy nodes when possible.
+%The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
+%Due to its decentralized nature, it can also work on isolated chunks of the network.
 
-\textit{3. Event/task and location/time dynamics}: how the model reacts to different situations depending on what, where and when events happen.
-Although this has not been used in the simulations in this paper, TruMan can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
+\textit{3. Event/task and location/time dynamics}: the model reacts to different situations depending on what, where and when events happen.
+The event or task dynamics involve managing different situations in different ways.
+Messages can carry different types of alerts, and not all of them need to be addressed with the same urgency.
+A message about a nearby crash, for example, requires a much quicker reaction than one about an upcoming change in weather; malicious nodes that broadcast false information about critical events are especially important to detect.
+Similarly, in order to satisfy location and time dynamics, nodes might behave differently according to where and when certain messages are received.
+To do this, messages about events must contain timestamp and geolocation data attached; nodes close to the event in space and in time could be considered more trustworthy. 
+Furthermore, by attaching timestamp data to messages, it is possible to age information, allowing the model to consider only data that is recent enough to be relevant.
+%Although this has not been used in the simulations in this paper, TruMan can easily be extended to consider time and location as long as nodes store geolocation and timestamp data.
 
 \textit{4. Scalability}: the model can work on very large networks at high speeds.
-Due to the low complexity of the algorithms used in the model, TruMan can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
-It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
+This is very important in vehicular networks, since, at certain times or locations, there might be a very large number of vehicles very close to one another. 
+%In order to attain scalability, a model may consider limiting the volume or frequency of trust-related messages a node can receive at a given moment.
+In the case of a model that allows transitive trust, a high volume of nearby vehicles can be advantageous because it allows nodes to share a lot of recent data with each other.
+
+%Due to the low complexity of the algorithms used in the model, TruMan can be highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
+%It has also been demonstrated that iterations of the algorithm do not need to run extremely frequently in order to detect malicious nodes with high accuracy.
 
 \textit{5. Integrated confidence measure}: allows nodes to estimate how useful the output of the algorithm is.
-Since nodes using TruMan store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
-The closer it is to 1, the higher the chance that it is accurate.
+Along with the information of whether or not a node $a$ trusts $b$, there should also be information regarding \textit{how sure} $a$ is of its trust in $b$.
+Generally, a higher confidence measure is the result of more and/or better evidence.
+%Since nodes using TruMan store trust values as a number between 0 and 1, this value can be used as a confidence measure of the opinion.
+%The closer it is to 1, the higher the chance that it is accurate.
 
 \textit{6. System level security}: requires authentication of nodes participating in the network.
-This has not been considered in evaluations of TruMan.
-However, it can be included as a separate security model during the transmission of messages.
+There should be an infrastructure in place in order to avoid identity falsification from potentially malicious nodes as well as verifying which node is the sender of a given message.
+
+%A public-private key solution can be used to  allows nodes to verify that messages it received were truly sent by who claimed to send it.
+%This has not been considered in evaluations of TruMan.
+%However, it can be included as a separate security model during the transmission of messages.
 
 \textit{7. Sensitivity to privacy concerns}: avoids eavesdropping and stalking by malicious nodes.
-TruMan has not been designed with this in mind, but it does not inhibit privacy protection.
-However, it does require that nodes cannot be completely anonymous.
+A message should only be received by the nodes it was meant for, avoiding eavedropping of its contents.
+Additionally, it should not be possible to track the activity of a node based on the messages it sends.
+
+%TruMan has not been designed with this in mind, but it does not inhibit privacy protection.
+%However, it does require that nodes cannot be completely anonymous.
 
 \textit{8. Robustness}: the model's resistance to attacks.
-TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
-Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
-Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
-Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified. 
+There are already some studied attacks for vehicular networks, such as the Sybil Attack, Newcomer Attack and Betrayal Attack.
+Models must show that they function in the event of such attacks.
+%TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
+%Experiments show that, when fewer than 50\% of nodes in the network are malicious, Truman performs as expected.
+%Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
+%Furthermore, since nodes take into consideration experiences from other trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified. 
 %Furthermore, the analysis also takes into account the complexity of the algorithm necessary to implement the trust model, because a high priority of TruMan is the
 
-\begin{table}[t!]
-\caption{Properties of Truman and related work}
-\label{table:properties}
-\centering
-\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
- \hline
- \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
- \hline
- \hline
- TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
-% \hline
-% \citep{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
- \hline
- \citep{dotzer2005vars} & \checkmark & - & \checkmark & - & - & - & \checkmark & -\\
- \hline
- \citep{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
- \hline
- \citep{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
- \hline
- \citep{park2011long} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
- \hline
- \citep{huang2014social} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
- \hline
- \citep{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
- \hline
- \citep{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
- \hline
-\end{tabular}
-\end{table}
-
-Finally, it is worth noting that, considering the scale of the problem, TruMan's cost is very low without sacrificing completeness and correctness.
-The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
+%Finally, it is worth noting that, considering the scale of the problem, TruMan's cost is very low without sacrificing completeness and correctness.
+%The model still satisfies or permits most of the desired properties of a trust model, making it viable for real-world use.
 
@@
 Several models have been proposed to solve the problem of trust in vehicular networks. 
-In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
+In this section, the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
 None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
@@
 \citep{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
-The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents.
-That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
+The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents and the message's sender.
+In other words, when a node $a$ receives a message about a certain event from $b$, it calculates a new opinion considering  it rebroadcasts the message to other nearby nodes, but with its own opinions about the event and about node $b$ attached.
+%That opinion is formed using a combination of the forwarding node's own observations of the event, its opinion of the origin node and previous opinions appended to the message.
 This process adds credibility to a message through validation by nodes in a decentralized fashion.
@@
 First, it allows forwarding nodes to access (at least some of) the contents of a message so it can form an opinion on it, diminishing privacy; a malicious forwarding node could even attempt to alter those contents.
-Second, using previously appended opinions from other nodes to form a new opinion causes the first nodes to forward the message to have a substantially greater impact over the final opinion than the later ones.
+Second, since each new opinion appended to the message considers the previously appended opinions, the first nodes to forward the message to have a substantially greater impact over the final opinion than the later ones.
 Finally, there is an issue with scalability, since appending new information to a message on each hop may add a significant overhead to the transmission.
@@
 To make trust viable in such a scenario, the authors rely on RSUs to store reputation information from passing vehicles.
-Each vehicle must have an ``Agent RSU'', which is be tasked with sharing that vehicle's trust data to other vehicles and RSUs as well as keeping the data updated when the vehicle is near it once again.
-To make this viable, the properties of daily commutes are used: it is assumed the vehicle would be near that RSU with reasonable frequency because it is located within the driver's home-to-work route.
+Each vehicle must have an ``Agent RSU'', which is in charge of storing and sharing that vehicle's trust data to other passing vehicles and connected RSUs.
+It must also keep the data updated when the vehicle approaches it again.
+To make this viable, the properties of daily commutes are used: it is assumed that the vehicle is near its Agent RSU with reasonable frequency because it is located within the driver's home-to-work route.
 The main problem with this model is that it relies on the presence of frequent RSUs, which might not always be viable.
@@
 The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
+%extend
 
@@
 It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
+%extend
+
+%conclude
+\autoref{table:properties} shows how well each studied trust model satisfies the eight properties described in \autoref{section:properties}.
+
+\begin{table}[t!]
+\caption{Properties of the related work}
+\label{table:properties}
+\centering
+\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
+ \hline
+ \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \hline
+% TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
+% \hline
+% \citep{vernize2015malicious} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+% \hline
+ \citep{dotzer2005vars} & \checkmark & - & \checkmark & - & - & - & \checkmark & -\\
+ \hline
+ \citep{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
+ \hline
+ \citep{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
+ \hline
+ \citep{park2011long} & - & - & - & - & - & \checkmark & \checkmark & \checkmark \\
+ \hline
+ \citep{huang2014social} & \checkmark & - & \checkmark & \checkmark & \checkmark & - & - & -\\
+ \hline
+ \citep{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
+ \hline
+ \citep{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
+ \hline
+\end{tabular}
+\end{table}
 

### 2018-03-28 dissertation/4-project/chapter4.tex
corrections to chapter 3

@@
 
-This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
+%This work shows that it is possible to adapt a trust management scheme originally designed for static networks to a dynamic environment, such as a VANET.
+This work introduces an efficient solution to trust management in dynamic networks such as VANETs.
 In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
@@
 By combining a node's own opinions about familiar nodes and trust information received from its neighbors, it is possible to create a model of the surrounding network.
-This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used as in conjunction with other algorithms in order to classify nodes as correct or malicious.
+This model includes a trust graph, showing the trust relationships between pair of nodes, which can then be used in conjunction with other algorithms in order to classify nodes as correct or malicious.
 
@@
 Then, two algorithms are introduced: Tarjan's strongly connected components algorithm \citep{tarjan1972depth} and an efficient algorithm for graph coloring \citep{mittal2011graph}.
-Next, the Malicious Node Identification Algorithm (MaNI) \citep{vernize2013dissertation} is explained, for it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
-Finally, TruMan itself is explained, showing how it combines features from existing algorithms and adapts them to a dynamic environment, enabled by the social properties found in vehicular networks.
+Next, the Malicious Node Identification Algorithm (MaNI) \citep{vernize2013dissertation} is explained, because it is the work that suggests the usage of strongly connected components and graph coloring for malicious node detection.
+Finally, TruMan itself is detailed, showing how it combines features from existing algorithms and adapts them to a dynamic environment, enabled by the social properties found in vehicular networks.
 %Then, the Working Day Movement Model is introduced, along with the reasons why it is best fitted for this work.

### 2018-03-29 dissertation/4-project/chapter4.tex
Corrections to chapter 4

@@
 In \citep{da2013effective}, \citep{cunha2014vehicular}, and \citep{cunha2014possible}, the authors attempt to find features usually attributed to social networks in vehicular networks.
-By using a data set from Zurich, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
+By using a data set from the city of Zurich, Switzerland, they show that some metrics, such as clustering coefficient and number of encounters, have peaks during the rush hours.
 They note that, during rush hours, the diameter of the graph decreases to around 6 hops; additionally, the frequency of total encounters between pairs of nodes in the network increases during those hours.
@@
 The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
-
 Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.
@@
 %The algorithm proposed in \citep{mittal2011graph} is an efficient approach to graph coloring, a classic graph theory problem.
-Graph coloring is one of the possible heuristics used by TruMan to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
-Out of the tested heuristics, it presents the best results, so it has been chosen as the heuristic for the trust model.
+Graph coloring is one of the possible heuristics suggested by MaNI to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
+Out of the tested heuristics, it presents the best results, so it has been chosen as the heuristic for TruMan.
 
@@
 Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
-For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
+For the purposes of trust management, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.
 
@@
 As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \citep{brelaz1979new}.
-
 Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.
@@
 
-Figure \autoref{fig:coloring} illustrates the execution of the graph coloring algorithm.
-It is notable how few iterations the algorithm takes to fully color the graph.
-However, the fact that the result does not use the minimum amount of colors.
-By coloring node 4 as cyan and node 3 as magenta, the sample graph could have been colored with only three colors.
-However, as described above, this is not a substantial problem for the purposes of this study.
+\autoref{fig:coloring} illustrates the execution of the graph coloring algorithm.
+It is notable how the algorithm takes few iterations to fully color the graph.
+However, it is also possible to observe that the result does not use the minimum amount of colors.
+By coloring node 4 as cyan and node 3 as magenta, the sample graph could have been colored with only three colors instead of four.
+As described above, this is not a problem for the usage of the algorithm in TruMan.
 
@@
 
-The article describes the coloring heuristic, which uses a graph coloring algorithm, either DSATUR \citep{brelaz1979new} or the algorithm proposed in \citep{mittal2011graph}.
-Graph coloring is a classic problem of graph theory, in which the objective is to assign each node in a graph a color so that no two neighboring nodes share the same color.
-After running either algorithms with graph $C$, the color whose nodes in $C$ represent the most nodes in $T$ is classified as correct, and all others are classified as malicious.
-Once this information in $C$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
+In the experiments performed by the authors of MaNI, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
+The coloring heuristic uses a graph coloring algorithm, such as DSATUR \citep{brelaz1979new} or the algorithm detailed in \autoref{section:coloring}.
+Other heuristics were experimented with, but were either less effective in detecting malicious nodes, provided too many false positives, or were not efficient enough.
 
-In the experiments shown in \citep{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
-Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
-Therefore, for the purposes of this research, only the coloring heuristic is considered.
+After running a graph coloring algorithm with graph $C$ as input, the color whose nodes in $C$ represent the most nodes in $T$ is classified as correct, and all others are classified as malicious.
+Once this information from $C$ is brought back to graph $T$, it is trivial to label the nodes in $T$ as either benign or malicious based on their components' classifications.
 
@@
 The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
-By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.
+By doing this a small number of times, nearly all malicious nodes are detected by it even when randomly changing edge weights.
 
@@
 Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
-Every node $u$ stores a representation of the network in the form of a static, connected and directed trust graph $T = (V, E)$, in which $V$ is the set of nodes node $u$ is aware of and $E$ is the set of trust relationships (opinions) $u$ knows between members of $V$.
+Every node $u$ stores a representation of the network in the form of a static, connected and directed trust graph $T^u = (V^u, E^u)$, in which $V^u$ is the set of nodes node $u$ is aware of and $E^u$ is the set of trust relationships (opinions) $u$ knows of between members of $V^u$.
 Since each node has its own network representation and it changes over time, there is a $T^u_i = (V^u_i, E^u_i)$ for every node $u$ and iteration $i$.
@@
 There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
-Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
-After this, a new $T^u_i$ is formed, which is used for the remaining steps.
+Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases.
+Additionally, node $u$ performs an union between its trust graph and $v$'s trust graph, forming a new graph $T^u_i = T^u_{i-1} \bigcup T^v_{i-1}$, which is used for the remaining steps.
 
-After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}.
-For each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
-In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
+After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, although the implementation of the algorithm slightly differs from the one used in MaNI.
+Since MaNI uses binary trust, Tarjan's algorithm only checks whether edges have value 0 or 1; in TruMan, edges can have a range of values between 0 and 1.
+Therefore, a threshold $h$ is defined so Tarjan's algorithm can consider only edges represent a significant trust relationship when forming strongly connected components.
+So, for each node in a component, there is a path formed by edges of weight higher than the threshold $h$ to each other node in the same component.
 Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, E'^u_i)$.
 
-The creation of $C^u_i$ simplifies the remaining computation.
 Since each vertex $v' \in V'^u_i$ is a component of $T^u_i$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as the same.
@@
 Once this information from $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on the classifications of their components.
-
-The complexity of the whole algorithm can be calculated by adding the most costly operations involved.
-As discussed above, Tarjan's algorithm has a complexity of $O(|V|+|E|)$ (for the trust graph $T$), while the graph coloring algorithm has a complexity of $O(|E'|)$ (for the component graph $C$).
-The most costly part of the algorithm is the graph merge operation that happens between trusted nodes.
-The complexity of the graph merge algorithm is $O(|E|)$ for each neighbor a node has in an iteration; this number is at most $|V|$.
-The total complexity of TruMan is, therefore, $O(|V|+|E|)+O(|E'|)+O(|V|\times |E|)$.
-Since $|E'| \leq |E|$ and $|V|+|E| \leq |V|\times |E|$, the complexity can be simplified to $O(|V| \times |E|)$.
 
@@
 \end{enumerate}
+
+\subsection{Complexity}
+The complexity of TruMan must calculated for each iteration executed and each node in the network.
+It can be estimated by adding the complexity of most costly operations involved, which are: ($i$) Tarjan's algorithm, ($ii$) the graph coloring algorithm, and ($iii$) the graph union process.
+Tarjan's algorithm and the graph coloring algorithm are executed once per iteration on each node of the network.
+The graph union, however, is executed once every time a neighboring node shares information.
+Each neighbor shares information once per iteration, so the complexity of the graph union procedure is multiplied by $n^u_i$, the number of neighbors node $u$ has during iteration $i$.
+
+Therefore, the general complexity equation for each node and each iteration is as follows:
+
+$$ TruMan = Tarjan + Coloring + (Union\times(n^u_i)) $$
+
+As discussed above, Tarjan's algorithm has a complexity of $O(|V|+|E|)$ and is executed on graph $T$.
+Meanwhile, the graph coloring algorithm has a complexity of $O(|E'|)$ and is executed on graph $C$.
+The complexity of the graph merge algorithm is $O(|E|)$ and, in a worst-case scenario, $n^u_i$ is at most $|V|$ (every node in the network is a neighbor), resulting in a total complexity of $O(|V|\times |E|)$ for a whole iteration.
+The total complexity of each iteration of TruMan is, therefore:
+
+$$ O(|V|+|E|) + O(|E'|) + O(|V|\times |E|)$$
+
+However, $|E'| \leq |E|$ is always true, because the graph $C$ is a reduction of graph $T$.
+Furthermore, $|V|+|E| \leq |V|\times |E|$ is also true, except for the irrelevant scenarios of $|V| \leq 1$ or $|E| \leq 1$.
+
+Therefore, the complexity can be simplified to: $O(|V| \times |E|)$.
 

### 2018-03-29 dissertation/5-simulations/chapter5.tex
2018-03-06
started updating chapter 5

@@
-\chapter{Simulations}
+\chapter{Evaluation}
 \label{chap:simulations}
+
+In order to validate TruMan's functionality and efficiency, several simulations were executed, attempting to replicate real-world scenarios.
+This chapter includes all information relevant to these simulations, including the tools used, the chosen movement model, the parameters and methodology, and the results.
 
@@
 \label{section:snap}
-In order to run the simulations necessary to validate the project, a robust library was required to handle graph data structures.
+The simulations necessary to validate the project require a robust library to handle graph data structures.
 The Stanford Network Analysis Platform (SNAP) library \citep{snap} was chosen primarily because it is memory efficient.
@@
 %The datasets available from SNAP's website were used for the simulations in \citep{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
+
+\section{The ONE Simulator}
+\label{section:theone}
+The Opportunistic Network Environment simulator \citep{keranen2009one} \citep{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.
+It was chosen for the simulations of TruMan primarily because it already includes integration with the Working Day Movement Model.
+
+The ONE Simulator comes with a usable map of the city of Helsinki, Finland, so the city was chosen as the map for the simulations of TruMan.
+
+In order to use data from the ONE simulator as input for TruMan, a new report module had to be created for it.
+The \texttt{AdjacencySnapshotReport} module creates a report consisting of all adjacencies in the network every $x$ number of simulated seconds.
+That is, at a given timestamp $t$, all pairs of nodes that are within communication range of each other are added to the report.
+This report is then used as input to TruMan, which uses the adjacencies to build the topology graph for each iteration of the algorithm.
+The \texttt{AdjacencySnapshotReport} module has been submitted to the ONE repository as a pull request \citep{adjacencypull}.
 
@@
 \subsection{Original model}
-The Working Day Movement Model was developed for Delay-Tolerant Networks in which nodes are devices (such as smartphones) carried by people.
+The Working Day Movement Model was developed for Delay-Tolerant Networks in which network members are devices (such as smartphones) carried by people.
 Therefore, the Working Day model represents not only people's movements inside their cars, but also within their offices, walking on foot, or riding a bus.
-In this subsection, the word node refers to a mobile device instead of a vehicle.
+% In this subsection, the word node refers to a mobile device instead of a vehicle.
 
 The model proposed by the authors makes use of several other models for specific tasks.
-The main mobility model defines nodes and gives them their destinations.
+The main mobility model places devices in the network and sets their destinations.
 Within it, five submodels are used:
@@
 \item 
-The \textbf{home activity submodel} describes what nodes do at night, within their homes.
+The \textbf{home activity submodel} describes what devices do at night, within their owners' homes.
 No movement is modeled.
-Nodes can be relatives or neighbors, and therefore share the same home.
+Devices can belong to relatives or neighbors, and therefore share the same home location.
 \item 
-The \textbf{office activity submodel} describes the nodes' routines within their offices.
-Nodes can go to other locations within the office (such as meeting rooms) and such movement is modeled.
-Nodes which share the same office are coworkers.
+The \textbf{office activity submodel} describes the devices' movement routines within their owners' offices.
+Devices can move to other locations within the office (such as meeting rooms) and such movement is modeled.
+Devices whose owners are coworkers share the same office space.
 \item 
-The \textbf{evening activity submodel} is responsible for mobility outside the nodes' standard routine. 
-They can meet at certain locations (such as restaurants) and spend a few hours with friends.
+The \textbf{evening activity submodel} is responsible for mobility outside the devices' owners' standard routines. 
+Devices can be carried by people who meet at certain locations (such as restaurants) and spend a few hours with friends.
 \item
-The \textbf{transport submodel} shows how nodes move around the city.
+The \textbf{transport submodel} shows how devices move around the city.
 It includes another tier of submodels, responsible for modeling three different types of transportation: walking, driving, and riding a bus.
-Nodes that own cars always use them, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
+People who own cars always use them, while the others can decide to walk or ride a bus depending on the distance between the origin and destination and the available bus stops.
 The walking and driving submodels represent similar types of movement, although at different speeds, while the bus submodel follows cyclical routes and can take or deliver passengers at bus stops.
@@
 The \textbf{map} represents the city in which the simulation runs.
-Its streets constrain the movement of nodes and all homes, offices and meeting spots must be within the map boundaries.
+Its streets constrain the movement of devices, and all homes, offices and meeting spots must be within the map boundaries.
 The map can be divided into districts, which increases what the authors define as \textit{locality}.
-In the simulation parameters, the number of nodes which reside and work within the same district can be chosen, which means those nodes rarely leave the district.
-Nodes which reside and work in different districts allow information to spread across different parts of the network.
+It is possible to configure how many people work in the same district where they reside; devices carried by these people rarely leave their district.
+%In the simulation parameters, the number of devices which reside and work within the same district can be chosen, which means those devices rarely leave the district.
+People who reside and work in different districts allow information to spread across different parts of the network by carrying their device with them.
 \end{enumerate}
@@
 \subsection{Adaptation for a vehicular simulation}
-By thinking of these submodels for vehicles instead of people, it becomes apparent that the frequency and length of encounters between nodes are similar in both instances.
+By thinking of these submodels for vehicles instead of people, it becomes apparent that the frequency and length of encounters between members of the network are similar in both instances.
 If two vehicles belong to family members or neighbors, they likely spend most of the night within communication distance, while coworkers' cars spend the office hours close by.
-Cars can also meet each other frequently if their drivers go out with friends.
-In the vehicular case, there is the added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, even if the drivers are complete strangers.
+Cars can also meet each other frequently if their drivers are friends who go out together after work.
+In the vehicular case, there is an added layer of encounters: cars can communicate frequently with buses and other cars that take the same route daily, even if their drivers are complete strangers.
 
-To adapt the Working Day Movement Model to a VANET environment, a few changes had to be made so the nodes represent vehicles instead of people (or the devices they carry). 
-Instead of altering the model itself, all of these changes were implemented as parameters for the model.
+To adapt the Working Day Movement Model to a VANET environment, a few changes had to be made so the network members are vehicles instead of people (or the devices they carry). 
+Rather than altering the model itself, all of these changes were implemented as parameters for the model.
 The changes are as follows:
@@
 \item
-While the bus submodel could be used for a vehicular simulation, this was not used for TruMan's simulations.
+While the bus submodel could be used for a vehicular simulation, this was not used in this evaluation.
 %The bus submodel needs to be changed so that each bus is one node in the network, which follows a predefined route with bus stops.
@@
 
-One important topic raised in the Working Day Movement Model article is the use of two metrics: \textit{inter-contact times} and \textit{contact duration}.
-The choice of this movement model for tests is more strongly related to inter-contact times, i.e. how much time it takes for two nodes to meet again.
-On the other hand, the contact duration is how long each meeting lasts.
-For the reasons explained earlier in this section, relatively short inter-contact times is important for TruMan's functionality.
-Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although, for the purposes of the simulation, this was not considered.
+One important topic raised in the Working Day Movement Model article is the use of two metrics for a movement model: \textit{inter-contact times} and \textit{contact duration}.
+Inter-contact time is the average time it takes for two nodes to meet repeatedly in the network.
+For example, two vehicles who belong to neighbors might have an inter-contact time of about 12 hours, since that is how long they are apart before connecting again.
+Meanwhile, the contact duration is the time nodes spend connected when they do meet.
+In the case of the two vehicles owned by neighbors, the contact duration might also be about 12 hours, while their owners are at home and leave the vehicles close to each other.
 
-\section{The ONE Simulator}
-\label{section:theone}
-The Opportunistic Network Environment simulator \citep{keranen2009one} \citep{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.
-It was chosen for the simulations of TruMan primarily because it already includes integration with the Working Day Movement Model.
-
-In order to use data from the ONE simulator as input for TruMan, a new report module had to be created for it.
-The \texttt{AdjacencySnapshotReport} module creates a report consisting of all adjacencies in the network every $x$ number of simulated seconds.
-That is, at a given timestamp $t$, all pairs of nodes that are within communication range of each other are added to the report.
-This report is then used as input to TruMan, which uses the adjacencies to build the topology graph for each iteration of the algorithm.
-The \texttt{AdjacencySnapshotReport} module has been submitted to the ONE repository as a pull request \citep{adjacencypull}.
+The choice of the Working Day Movement Model for evaluations is more strongly related to inter-contact times.
+For reasons explained in \autoref{chapter:truman}, relatively short inter-contact times is important for TruMan's functionality.
+Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although, for this evaluation, it was not considered.
 
@@
 
+\defcitealias{helsinkifleet}{Helsinki, 2011}
+\defcitealias{helsinkiarea}{NLSF, 2018}
 \defcitealias{saopauloarea}{IBGE, 2016}
@@
  \hline
+ \hline
+ Helsinki (FI)		& 250,000						& 214,250,000 					& 0.1833	\\
+ 					& \citepalias{helsinkifleet}	& \citepalias{helsinkiarea}		&			\\
  \hline
@@
 
-\subsection{Validation}
-
-\section{Restrictions}
-\label{section:restrictions}
+%\subsection{Validation}
+%
+%\section{Restrictions}
+%\label{section:restrictions}
 

### 2018-04-02 dissertation/3-vanets/chapter3.tex
updated chapters 3, 4, 5

@@
 \section{Special properties of VANETs}
+\label{section:properties}
 %As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
@@
 
-The ART model proposed in \citep{li2016art} works in two main steps: data gathering and malicious node detection.
-It uses the Dempster-Shafter theory of evidence to merge data coming from other nodes.
-Then, it uses a Cosine-based metric to compare two nodes' trust vectors (a series of opinions regarding other nodes).
-However, these steps require several intensive calculations, which greatly increase the complexity of the algorithm.
-The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
+The Attack-Resistant Trust Management Scheme (ART) from \citep{li2016art} proposes to resist three types of attacks: 
+simple attacks, in which malicious nodes do not cooperate with the network;
+bad-mouth attacks, in which malicious nodes perform simple attacks but also share false information about benign nodes;
+and zig-zag attacks, in which nodes vary their behaviour from benign to malicious in order to be harder to detect.
+It works in two main steps: data gathering and malicious node detection.
+For data gathering, it uses the Dempster-Shafter theory of evidence, which establishes \textit{belief} and \textit{plausibility} values, both real numbers ranging from 0 to 1.
+The former refers to the amount of evidence indicating the truthfulness of a hypothesis (for example, a node sending false data corroborates the hypothesis of it being malicious), while the latter is 1 minus the amount of evidence that supports the possibility of the hypothesis being false.
+These evidences are acquired through observations by a node and through data received from other nodes.
+The resulting probability is the basic trust value, which is stored in a trust vector (a series of trust values regarding other nodes).
+The malicious node detection step uses a Cosine-based metric to compare two nodes' trust vectors.
+When two nodes share similar opinions about other nodes, they will consider each other trustworthy.
+The downside of this model is that each step is mathematically costly, requiring several intensive calculations in order to achieve its goals.
+This likely increases the complexity of the algorithm, which is not detailed by the authors.
+Because of this, it is uncertain how the model would scale to large networks, while it might also underperform in small networks in which there is little evidence to collect.
+%The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
 %extend

### 2018-04-02 dissertation/4-project/chapter4.tex
updated chapters 3, 4, 5

@@
 
+In a network in which malicious nodes are a minority (under 50\%), it is expected that the benign nodes will form components with large numbers of nodes, because these benign nodes will share their networks with each other and it is easy to form trust paths between pairs of benign nodes.
+Malicious nodes, on the other hand, do not send their own networks of false information to benign nodes, and might not always trust other malicious nodes, causing them to become isolated in small strongly connected components (in some cases, these contain a single malicious node).
+The result is that most benign nodes become members of a small number of large components; when the component graph is colored, these components are likely to receive the same color, because two benign components are never adjacent.
+Because of this, the coloring heuristic works as a classification method.
+%Classifying the color that represents the most nodes as correct works because
+
 In summary, every node $u$ runs the following steps in each iteration to detect malicious nodes in the network:
@@
 		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
-	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$.
+	\item If a neighbor $v$ is trustworthy, $u$ performs a union with $T^u_{i-1}$ and $T^v_{i-1}$, establishing $T^u_i$.
 	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.

### 2018-04-02 dissertation/5-simulations/chapter5.tex
updated chapters 3, 4, 5

@@
 
-As a comparison, the city of São Paulo (Brazil), which has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+As a comparison, the city of São Paulo (Brazil) has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, and thus has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
 \autoref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
-All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan aras that surround these cities.
+All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan areas that surround these cities.
 %Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
@@
 \end{table}
+
+Simulations of TruMan were performed with densities as low as 0.0017, a value much lower than even the real world density of Helsinki, which is a city with relatively few vehicles for its size.
+It can be expected that the model will perform even better in real-world scenarios in cities with even higher densities, which is common, as exemplified by \autoref{table:citydensities}.
 
@@
 %\end{figure*}
+
+\section{Satisfaction of desired properties}
+\label{section:satisfaction}
+
+Eight desired properties for VANET trust models were presented in \autoref{section:properties}.
+Here, TruMan's ability to satisfy each property is evaluated, ending with a final comparison between TruMan and previously proposed trust models for VANETs.

### 2018-04-02 dissertation/4-project/chapter4.tex
Added conclusion to Chap3 and added section 4.6.1

@@
 After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, although the implementation of the algorithm slightly differs from the one used in MaNI.
-Since MaNI uses binary trust, Tarjan's algorithm only checks whether edges have value 0 or 1; in TruMan, edges can have a range of values between 0 and 1.
+Since MaNI uses binary trust, Tarjan's algorithm only checks whether edges have value 0 or 1; in TruMan, each edge stores a trust value $t \in [0, 1]$.%can have a range of values between 0 and 1.
 Therefore, a threshold $h$ is defined so Tarjan's algorithm can consider only edges represent a significant trust relationship when forming strongly connected components.
@@
 \end{enumerate}
+
+\subsection{Information aging}
+
+In order to make TruMan resistant to attacks, it is necessary to age the information nodes store about the network, so old information does not affect the identification of malicious nodes.
+
+To do this, each edge stores a timestamp value $s$ in addition to the trust value $t$.
+When two nodes $u$ and $v$ interact with each other, the edge $u\rightarrow v$ in $T_u$ stores the timestamp $s$, which is set according to node $v$'s internal clock.
+
+Then, when node $u$ interacts with another node $w$ in the future, $u$'s opinion of $v$ comes with the timestamp attached, so $w$ has the information to know how much time has passed since the interaction between $u$ and $v$ happened.
+Once $w$ performs the graph merge procedure with $T_u$ and $T_w$, but $T_w$ already has an edge $u\rightarrow v$ stored, it checks the two timestamps and only updates the edge if the incoming information is more recent than what is already stored.
+Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
+%This was implemented as an additional feature of TruMan, and was not used on every test.
+
+For example, when $u$ interacts with $v$ on and establishes that $v$ is benign, it creates the edge $u \rightarrow v$ with $t \geq 0.5$ and $s = s_0$, where $s_0$ is the timestamp of the moment in which the interaction occurred according to $v$.
+Then, when $w$ interacts with $u$ and establishes that $u$ is benign, it receives network information from $u$, including the edge $u\rightarrow v$.
+Since $w$ didn't have that edge in its graph before, it is added maintaining the original timestamp $s_0$.
+Later, when $w$ interacts with $u$ again or with another trustworthy node that recently interacted with $u$, it receives information about the edge $u\rightarrow v$ again.
+If this new information includes the timestamp $s_1$ which is more recent than $s_0$, $w$ updates its own graph so that the edge $u\rightarrow v$ stores the new timestamp and the new trust value.
+
+This is necessary because, in the event that a node converts from benign to malicious, old information might still indicate that it is benign, as shown in ------figure------.
+
+Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
+This is done by setting a maximum age for edges.
+Once a edge is older than that maximum value, it is discarded.
+A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).
 

### 2018-04-02 dissertation/3-vanets/chapter3.tex
Added conclusion to Chap3 and added section 4.6.1
Small addition to chapter 3 and added algorithms to chapter 4.6

@@
 \autoref{table:properties} shows how well each studied trust model satisfies the eight properties described in \autoref{section:properties}.
+The numbers used in the table are the same ones from \autoref{section:properties}.
+This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.
 

### 2018-04-02 dissertation/4-project/chapter4.tex
Small addition to chapter 3 and added algorithms to chapter 4.6

@@
 \State \textbf{color} all nodes of $G$ \textbf{with} 0
-\State $d \rightarrow 0$
+\State $d \gets 0$
 \For{$e=(u,v)$ \textbf{in} edges of $G$}
@@
 		\If{$color[v] = d$}
-			\State $d \rightarrow d+1$
+			\State $d \gets d+1$
 		\EndIf
-		\State $color[v] \rightarrow d$
+		\State $color[v] \gets d$
 	\EndIf
@@
 Additionally, node $u$ performs an union between its trust graph and $v$'s trust graph, forming a new graph $T^u_i = T^u_{i-1} \bigcup T^v_{i-1}$, which is used for the remaining steps.
+Algorithm  \autoref{algorithm:interaction} shows the basic interaction between two nodes, while algorithm \autoref{algorithm:union} details the steps of the graph union.
+
+\begin{algorithm}
+\caption{Interaction between two nodes}\label{algorithm:interaction}
+\begin{algorithmic}[1]
+
+\Function{Interaction}{node $u$, node $v$}
+
+\If{$v\notin T^u$}
+	\State \textbf{add} $v$ \textbf{to} $T^u$
+	\State \textbf{add} $u\rightarrow v$ \textbf{to} $T^u$
+	\State $T^u(u\rightarrow v).trustvalue = 0.5$
+	\State $T^u(u\rightarrow v).timestamp = $ \textbf{now}
+\EndIf
+
+\State $u$ \textbf{tests} $v$
+
+\If{$v$ is benign}
+	\State $T^u(u\rightarrow v).trustvalue$ \textbf{increases}
+	\State \textbf{Union}($T^u$, $T^v$)
+\Else
+	\State $T^u(u\rightarrow v).trustvalue$ \textbf{decreases}
+\EndIf
+
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Graph union}\label{algorithm:union}
+\begin{algorithmic}[1]
+
+\Function{Union}{graph $T^u$, graph $T^v$}
+
+\For{$a\rightarrow b$ \textbf{in} edges of $T^v$ }
+
+	\If{$a\notin T^u$}
+		\State \textbf{add} $a$ \textbf{to} $T^u$
+	\EndIf
+
+	\If{$b\notin T^u$}
+		\State \textbf{add} $b$ \textbf{to} $T^u$
+	\EndIf
+
+	\If{$a\rightarrow b \notin T^u$}
+		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
+	\EndIf
+\EndFor
+
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 
@@
 Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
-%This was implemented as an additional feature of TruMan, and was not used on every test.
+Algorithm \autoref{algorithm:union_timestamp} shows how the union function was changed to accommodate timestamps and information aging.
+
+\begin{algorithm}
+\caption{Graph union with timestamps}\label{algorithm:union_timestamp}
+\begin{algorithmic}[1]
+
+\Function{Union}{graph $T^u$, graph $T^v$}
+
+\For{$a\rightarrow b$ \textbf{in} edges of $T^v$ }
+
+	\If{$a\notin T^u$}
+		\State \textbf{add} $a$ \textbf{to} $T^u$
+	\EndIf
+
+	\If{$b\notin T^u$}
+		\State \textbf{add} $b$ \textbf{to} $T^u$
+	\EndIf
+
+	\If{$a\rightarrow b \notin T^u$}
+		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
+	\Else
+		\State $s_0 \gets T^u(a\rightarrow b).timestamp$
+		\State $s_1 \gets T^v(a\rightarrow b).timestamp$
+		\If{$s_1$ \textbf{is more recent than} $s_0$}
+			\State $T^u(a\rightarrow b).trustvalue \gets T^v(a\rightarrow b).trustvalue$
+			\State $T^u(a\rightarrow b).timestamp \gets T^v(a\rightarrow b).timestamp$
+		\EndIf	
+	\EndIf
+\EndFor
+
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
 

### 2018-04-05 dissertation/2-complexnetworks/chapter2.tex
Extended chapter 5 and small improvements elsewhere

@@
 
+\autoref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
+The topology graph changes as one node moves across the network, with edges being added and removed as nodes move into and out of each other's ranges.
+The trust graph, however, is constantly being built, maintaining past relationships even if the nodes are no longer in communication range.
+
 Naturally, VANETs are an instance of MANETs and therefore share some of the same features.
@@
 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.4\textwidth}
 \includegraphics[width=\linewidth]{images/manet_topology_1.png}
-\caption{Initial topology and trust graph} \label{fig:2_2a}
+\caption{Initial topology graph} \label{fig:2_2a}
 \end{subfigure}
 
-\vspace*{\fill} % separation between the subfigures
+\vspace*{0.5cm} % separation between the subfigures
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/manet_topology_2.png}
-\caption{Final topology graph} \label{fig:2_2b}
+\begin{subfigure}{0.08\textwidth}
+\includegraphics[width=\linewidth]{images/manet_trust_1.png}
+\caption{} \label{fig:2_3b}
 \end{subfigure}
 
-\vspace*{\fill} % separation between the subfigures
+\vspace*{0.5cm} % separation between the subfigures
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/manet_trust.png}
-\caption{Final trust graph} \label{fig:2_3b}
+\begin{subfigure}{0.4\textwidth}
+\includegraphics[width=\linewidth]{images/manet_topology_2.png}
+\caption{Final topology graph} \label{fig:2_2c}
+\end{subfigure}
+
+\vspace*{0.5cm} % separation between the subfigures
+
+\begin{subfigure}{0.4\textwidth}
+\includegraphics[width=\linewidth]{images/manet_trust_2.png}
+\caption{Final trust graph} \label{fig:2_3d}
 \end{subfigure}

### 2018-04-05 dissertation/4-project/chapter4.tex
Extended chapter 5 and small improvements elsewhere

@@
 
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.30\textwidth}
+\includegraphics[width=\linewidth]{images/rogue/0.png}
+\caption{All nodes are benign and belong to the same strongly connected component.} \label{fig:rogue0}
+\end{subfigure}
+\hspace*{1cm} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+\includegraphics[width=\linewidth]{images/rogue/1.png}
+\caption{Node $0$ becomes malicious, but still belongs to the component.} \label{fig:rogue1}
+\end{subfigure}
+
+\vspace{1cm} % separation between the subfigures
+
+\begin{subfigure}{0.3\textwidth}
+\includegraphics[width=\linewidth]{images/rogue/2.png}
+\caption{Node $2$ updates its opinion of node $0$, identifying it as malicious.} \label{fig:rogue2}
+\end{subfigure}
+\hspace*{1cm} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+\includegraphics[width=\linewidth]{images/rogue/3.png}
+\caption{$3$'s opinion becomes too old and is discarded.} \label{fig:rogue3}
+\end{subfigure}
+
+\caption{Example of what happens when a node becomes malicious.}\label{fig:rogue}
+\end{figure}
+
 For example, when $u$ interacts with $v$ on and establishes that $v$ is benign, it creates the edge $u \rightarrow v$ with $t \geq 0.5$ and $s = s_0$, where $s_0$ is the timestamp of the moment in which the interaction occurred according to $v$.
@@
 
-This is necessary because, in the event that a node converts from benign to malicious, old information might still indicate that it is benign, as shown in ------figure------.
+Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
+This is done by setting a maximum age $m$ for edges; once a edge is older than $m$, it is discarded.
+A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).
 
-Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
-This is done by setting a maximum age for edges.
-Once a edge is older than that maximum value, it is discarded.
-A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).
+Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \autoref{fig:rogue}.
+Immediately after node $0$ becomes malicious, it is still part of a benign strongly connected component, because there are paths from every other node to $0$, and from $0$ to every other node.
+Some time later, node $2$ interacts with $0$ again and identifies it as malicious; other nodes that interact with $2$ will receive that updated information.
+However, the outdated edge $3\rightarrow 0$ keeps $0$ in the component, since other nodes trust $3$ and $3$ still hasn't updated its own opinion of $0$.
+Once that edge is old enough, it is discarded and node $0$ is correctly identified as malicious by all other nodes.
 

### 2018-04-05 dissertation/5-simulations/chapter5.tex
Extended chapter 5 and small improvements elsewhere

@@
 
-\autoref{fig:randommalicious} shows the variation of results for different amounts of malicious nodes in the network.
+\autoref{fig:randommalicious1} and \autoref{fig:randommalicious2} show the variation of results for different amounts of malicious nodes in the network.
 By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network, although there is still a small amount of false positive detections.
@@
 Around iteration 20000, all nodes finish building their models and the false positive detection rate drops to an insignificant amount.
+
+\autoref{fig:extramaliciousaging1} and \autoref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
+These were the only simulations to consider information aging.
+Here, several different values for the maximum age, $m$, were tested.
+For the simulation, information timestamps are set as the iteration in which they were added to the graph ($s \gets i$), and the age is calculated as the difference between the current iteration and the timestamp ($age \gets i - s$).
+With $m=10$, information is discarded too quickly, resulting in highly imprecise results.
+Starting with $m=250$, results are reasonable, and the additional malicious node is detected approximately 250 iterations after it converts.
+With higher values such as $m=1000$, detection takes longer, but not much else changes. 
+When the value is very high, like in the case of $m=5000$, the extra malicious node is not detected by the end of one simulated day.
+
+\section{Satisfaction of desired properties}
+\label{section:satisfaction}
+
+Eight desired properties for VANET trust models were presented in \autoref{section:properties}.
+Here, TruMan's ability to satisfy each property is evaluated, ending with a final comparison between TruMan and previously proposed trust models for VANETs.
+
+\textit{1. Decentralized trust establishment}:
+TruMan is built from the ground up for decentralized systems.
+Nodes form their own abstractions of the network and the model does not rely on a central observer.
+
+\textit{2. Coping with sparsity}:
+The experiments using low density values demonstrate that TruMan works in reasonably sparse networks.
+Since nodes carry information from previous interactions, the model can also work on temporarily isolated parts of the network.
+
+\textit{3. Event/task and location/time dynamics}:
+Aside from the time-related information used for information aging, dynamics were not used in the simulations presented here.
+However, TruMan can be extended to consider event, task, location and time dynamics when managing trust.
+
+\textit{4. Scalability}: 
+Due to the low complexity of the algorithms used in the model, TruMan is highly scalable, as it does not incur substantial pressure on the vehicles' on-board units.
+It has also been demonstrated that iterations of the algorithm do not need to run very frequently in order to detect malicious nodes with high accuracy --- once every ten seconds is more than enough.
+
+\textit{5. Integrated confidence measure}:
+Since nodes using TruMan store trust values as a number between 0 and 1 that increase or decrease the more nodes interact (i.e.
+the more evidence they gather), this value can be used as a confidence measure of the opinion.
+Setting the threshold $h$ to a value closer to 1 makes it unlikely that untrustworthy nodes will be incorrectly labeled as correct.
+
+\textit{6. System level security}:
+A public-private key solution can be used to verify message integrity.
+This has not been considered in evaluations of TruMan.
+However, it can be included as a separate security model during the transmission of messages.
+
+\textit{7. Sensitivity to privacy concerns}:
+TruMan has not been designed with this in mind, as it requires that nodes maintain a constant identity the whole time.
+However, this does not inhibit other types of privacy protection, which can be implemented in addition to TruMan.
+
+\textit{8. Robustness}: the model's resistance to attacks.
+TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
+Experiments show that, when fewer than 50\% of nodes in the network are malicious, TruMan performs as expected.
+Collusion attacks must be performed by more than half of the entire network, in which case the network is considered compromised.
+Furthermore, since nodes take into consideration experiences from several trustworthy nodes, a malicious node that occasionally behaves correctly can still be identified.
+Experiments with information aging show that it is possible to detect nodes that suddenly become malicious, although further experimentation with known attacks is still necessary.
+%Furthermore, the analysis also takes into account the complexity of the algorithm necessary to implement the trust model, because a high priority of TruMan is the
+
+Finally, it is worth noting that, considering the scale of the problem, TruMan's cost is very low without sacrificing completeness and correctness.
+The model satisfies or permits most of the desired properties of a trust model with low computational complexity, making it viable for real-world use.
 
@@
 
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.6\textwidth}
 \includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
@@
 \end{subfigure}
-\hspace*{0.1cm} % separation between the subfigures
-\begin{subfigure}{0.45\textwidth}
+
+\begin{subfigure}{0.6\textwidth}
 \includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
@@
 
-\vspace{1cm}
-
-\begin{subfigure}{0.45\textwidth}
+\begin{subfigure}{0.6\textwidth}
 \includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
@@
 \end{subfigure}
-\hspace*{0.1cm} % separation between the subfigures
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
-\caption{30\% malicious.}
-\end{subfigure}
 
-\vspace{1cm}
-
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
-\caption{40\% malicious.}
-\end{subfigure}
-\hspace*{0.1cm} % separation between the subfigures
-\begin{subfigure}{0.45\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
-\caption{50\% malicious.}
-\end{subfigure}
-
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes.}
-\label{fig:randommalicious}
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1/2).}
+\label{fig:randommalicious1}
 \end{figure}
@@
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
-\caption{$h = 0.3$.} \label{fig:tarjan0}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
+\caption{30\% malicious.}
 \end{subfigure}
 
-\begin{subfigure}{0.5\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
-\caption{$h = 0.5$.} \label{fig:tarjan0}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
+\caption{40\% malicious.}
 \end{subfigure}
 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
+\caption{50\% malicious.}
+\end{subfigure}
+
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (2/2).}
+\label{fig:randommalicious2}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
+\caption{$h = 0.3$.} \label{fig:threshold03}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
+\caption{$h = 0.5$.} \label{fig:threshold05}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
 \includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
-\caption{$h = 0.7$.} \label{fig:tarjan0}
+\caption{$h = 0.7$.} \label{fig:threshold07}
 \end{subfigure}
@@
 \caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
+\caption{$m = 10$.} \label{fig:extra10}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
+\caption{$m = 250$.} \label{fig:extra250}
+\end{subfigure}
+
+\caption{Simulation with information aging, with different maximum age values (1/2).}
+\label{fig:extramaliciousaging1}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
+\caption{$m = 1000$.} \label{fig:extra1000}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
+\caption{$m = 5000$.} \label{fig:extra5000}
+\end{subfigure}
+
+\caption{Simulation with information aging, with different maximum age values (2/2).}
+\label{fig:extramaliciousaging2}
 \end{figure}
@@
 %\end{figure*}
-
-\section{Satisfaction of desired properties}
-\label{section:satisfaction}
-
-Eight desired properties for VANET trust models were presented in \autoref{section:properties}.
-Here, TruMan's ability to satisfy each property is evaluated, ending with a final comparison between TruMan and previously proposed trust models for VANETs.

### 2018-04-05 dissertation/6-conclusao/chapter6.tex
Extended chapter 5 and small improvements elsewhere

@@
 The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
-In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
+In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previously proposed algorithms in order to generate fast and accurate results.
+
 As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
-By placing nodes into strongly connected components, a network containing a large amount of node can be simplified into a much smaller one.
+By placing nodes into strongly connected components, a network containing a large amount of nodes can be simplified into a much smaller one.
 Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.

### 2018-04-06 dissertation/0-preambulo/abstract.tex
updated abstracts

--- 2017-06-01 qualification/0-preambulo/abstract.tex
@@
 To generate these trust values, nodes use information acquired from past interactions; nodes which frequently share false or irrelevant data will have lower trust values than the ones which appear to be reliable.
-This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
+This work introduces TruMan, a trust management model for vehicular networks in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
 This movement model allows the comparison of VANETs to traditional social networks, because it can be observed that pairs of vehicles are likely to meet more than once in several scenarios: for example, they can belong to neighbors or work colleagues, or simply take similar routes every day.
@@
 The resulting trust value can also be used to aid other nodes which might not have a developed relationship with each other.
-The proposed algorithm is based on a previously existing one, which was developed for centralized networks and focused on static ad-hoc models; the previous algorithm will be adapted to serve a decentralized and dynamic network, which is the case of VANETs.
-Using existing trust values, a directed graph is modeled in which edges represent the trust relationship between pairs of nodes.
-Then, graph components are formed in which no pair of nodes within a single component has a negative trust relationship.
+TruMan is based on a previously existing algorithm, which was developed for centralized networks and focused on static ad-hoc models; its concepts were adapted to serve a decentralized and dynamic network, which is the case of VANETs.
+Using trust values formed by node interactions, a trust graph is modeled; its edges represent trust relationships between pairs of nodes.
+Then, strongly connected components are formed so that each node in each component trusts other nodes in the same component directly or indirectly.
 A graph coloring algorithm is used on the resulting components graph and, using the coloring results, it is possible to infer which nodes are considered malicious by the consensus of the network.
-The complete algorithm is expected to be fast, so it can be executed frequently, and will allow nodes to maintain a model of the surrounding networks indicating which neighboring nodes can be trusted or not.
+TruMan is fast, so it incurs low pressure on on-board computers, and is able to satisfy most desired properties for vehicular trust management models.
 

### 2018-04-10 dissertation/3-vanets/chapter3.tex
Final draft of the dissertation

@@
 This has the advantage of simplifying properties such as handling sparsity and scalability, but also makes the system slower in general, especially in situations in which mobile communication is slow or unreliable.
-It also makes the system prone to attacks, since the whole system collapses if the global trust manager is attacked.
+In general, it goes against several established concepts for vehicular networks, such as being completely decentralized and ad-hoc.
+It also makes the system somewhat, since the whole system collapses if the global trust manager is attacked, or could leave the network members unaided in the event of a server or connection failure.
 %extend

### 2018-04-10 dissertation/6-conclusao/chapter6.tex
Final draft of the dissertation

--- 2018-02-20 dissertation/6-conclusao/chapter6.tex
@@
 
-The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
-In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
-As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
-By placing nodes into strongly connected components, a network containing a large amount of node can be simplified into a much smaller one.
-Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
-This allows for a low complexity approach to malicious node identification in a dynamic network.
+In the coming years, vehicular networks or VANETs will an important part of safety and security in transportation, optimizing traffic and reducing the number of accidents.
+However, they are an appealing target for entities with malicious intents, creating the need of robust solutions to maintain the integrity of such networks.
 
-The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a very tiny amount of false positives.
+The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information among members of a network as much as possible.
+In this paper, a new trust model for vehicular networks called TruMan was introduced, which combines the efficiency of previously proposed algorithms in order to generate fast and accurate results.
+The solution works in a decentralized fashion and is built for the dynamic environment of vehicular networks, although it could also be adapted to other types of networks.
+
+%As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
+%By placing nodes into strongly connected components, a network containing a large amount of nodes can be simplified into a much smaller one.
+%Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
+%This allows for a low complexity approach to malicious node identification in a dynamic network.
+
+TruMan was evaluated using mobility data gathered from the ONE simulator using the Working Day Movement Model, which approximates node mobility to that of real-world vehicles.
+Several simulations were performed, changing certain parameters to understand how the model performs in different scenarios.
+The experiments show that vehicles within a network can form a sufficient abstraction of the network in around one day, and with that information they are able to detect nearly every malicious node in the network, with a very small amount of false positives.
 As the network changes in shape, nodes acquire more information and are able to make even more accurate classifications of malicious nodes around them.
-Future work includes the extension of the proposed model to use V2I communications.
+With the implementation of information aging, TruMan is also able to detect nodes that start benign and become malicious during the simulation.
+
+A next step for future work would be performing experiments simulating more types of known attacks against vehicular networks, then adapting the model to resist such attacks.
+Furthermore, TruMan could be extended to consider more types of information in its algorithm, using, for example, V2I communications with road-side units, vehicular roles, or location and time parameters.
+Finally, the concepts of the algorithm could be applied to different types of networks.

### 2018-04-11 dissertation/4-project/chapter4.tex
prepare for final draft

@@
 Because of this, the coloring heuristic works as a classification method.
+A high scale collusion attack could affect this heuristic, as malicious nodes would form large components and distinguishing these components from the benign ones would be a challenge.
 %Classifying the color that represents the most nodes as correct works because
@@
 Then, when node $u$ interacts with another node $w$ in the future, $u$'s opinion of $v$ comes with the timestamp attached, so $w$ has the information to know how much time has passed since the interaction between $u$ and $v$ happened.
-Once $w$ performs the graph merge procedure with $T_u$ and $T_w$, but $T_w$ already has an edge $u\rightarrow v$ stored, it checks the two timestamps and only updates the edge if the incoming information is more recent than what is already stored.
+Once $w$ performs the graph merge procedure with $T_u$ and $T_w$, but $T_w$ already has an edge $u\rightarrow v$ stored, it checks the two timestamps and only updates the edge if the incoming information is more recent than the node already had stored.
 Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
@@
 \subsection{Complexity}
-The complexity of TruMan must calculated for each iteration executed and each node in the network.
+The complexity of TruMan must be calculated for each iteration executed and each node in the network.
 It can be estimated by adding the complexity of most costly operations involved, which are: ($i$) Tarjan's algorithm, ($ii$) the graph coloring algorithm, and ($iii$) the graph union process.

### 2018-04-11 dissertation/5-simulations/chapter5.tex
prepare for final draft

@@
 The choice of the Working Day Movement Model for evaluations is more strongly related to inter-contact times.
-For reasons explained in \autoref{chapter:truman}, relatively short inter-contact times is important for TruMan's functionality.
+For reasons explained in \autoref{chap:truman}, relatively short inter-contact times is important for TruMan's functionality.
 Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although, for this evaluation, it was not considered.
@@
 
-$$ \delta = \frac{\frac{30^2\pi}{2} \times 160}{14.689.750} = 0.0154$$
+$$ \delta = \frac{\frac{30^2\pi}{2} \times 160}{14,689,750} = 0.0154$$
 
@@
 
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
+\caption{$\rho$ = 10m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+\caption{$\rho$ = 30m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+\end{subfigure}
+
+\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+\label{fig:random103050}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
+\caption{1\% malicious.}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
+\caption{5\% malicious.}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
+\caption{10\% malicious.}
+\end{subfigure}
+
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1\%, 5\% and 10\%).}
+\label{fig:randommalicious1}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
+\caption{30\% malicious.}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
+\caption{40\% malicious.}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
+\caption{50\% malicious.}
+\end{subfigure}
+
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (30\%, 40\% and 50\%).}
+\label{fig:randommalicious2}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
+\caption{$h = 0.3$.} \label{fig:threshold03}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
+\caption{$h = 0.5$.} \label{fig:threshold05}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
+\caption{$h = 0.7$.} \label{fig:threshold07}
+\end{subfigure}
+
+\caption{Simulation with 10\% malicious nodes, $\rho = 30$m and varying values of $h$.}
+\label{fig:randomthresholds}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{images/plots/Network_rA7/10_10}
+\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
+\caption{$m = 10$.} \label{fig:extra10}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
+\caption{$m = 250$.} \label{fig:extra250}
+\end{subfigure}
+
+\caption{Simulation with information aging, with different maximum age values ($m=10, 250$).}
+\label{fig:extramaliciousaging1}
+\end{figure}
+
+\begin{figure}
+\centering
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
+\caption{$m = 1000$.} \label{fig:extra1000}
+\end{subfigure}
+
+\begin{subfigure}{0.6\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
+\caption{$m = 5000$.} \label{fig:extra5000}
+\end{subfigure}
+
+\caption{Simulation with information aging, with different maximum age values ($m=1000, 5000$).}
+\label{fig:extramaliciousaging2}
+\end{figure}
+
 \section{Satisfaction of desired properties}
@@
 The model satisfies or permits most of the desired properties of a trust model with low computational complexity, making it viable for real-world use.
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
-\caption{$\rho$ = 10m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
-\caption{$\rho$ = 30m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
-\caption{$\rho$ = 50m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
-\label{fig:random103050}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
-\caption{1\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
-\caption{5\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
-\caption{10\% malicious.}
-\end{subfigure}
-
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1/2).}
-\label{fig:randommalicious1}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
-\caption{30\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
-\caption{40\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
-\caption{50\% malicious.}
-\end{subfigure}
-
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (2/2).}
-\label{fig:randommalicious2}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
-\caption{$h = 0.3$.} \label{fig:threshold03}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
-\caption{$h = 0.5$.} \label{fig:threshold05}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
-\caption{$h = 0.7$.} \label{fig:threshold07}
-\end{subfigure}
-
-\caption{Simulation with 10\% malicious nodes, $\rho = 30$m and varying values of $h$.}
-\label{fig:randomthresholds}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{images/plots/Network_rA7/10_10}
-\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
-\caption{$m = 10$.} \label{fig:extra10}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
-\caption{$m = 250$.} \label{fig:extra250}
-\end{subfigure}
-
-\caption{Simulation with information aging, with different maximum age values (1/2).}
-\label{fig:extramaliciousaging1}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
-\caption{$m = 1000$.} \label{fig:extra1000}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
-\caption{$m = 5000$.} \label{fig:extra5000}
-\end{subfigure}
-
-\caption{Simulation with information aging, with different maximum age values (2/2).}
-\label{fig:extramaliciousaging2}
-\end{figure}
 

### 2018-04-14 dissertation/5-simulations/chapter5.tex
dissertation local revision

--- 2018-04-05 dissertation/5-simulations/chapter5.tex
@@
 Eight desired properties for VANET trust models were presented in \autoref{section:properties}.
-Here, TruMan's ability to satisfy each property is evaluated, ending with a final comparison between TruMan and previously proposed trust models for VANETs.
+Here, TruMan's ability to satisfy each property is evaluated.
+\autoref{table:properties_truman} replicates the one from \autoref{section:properties} with the addition of TruMan, providing a full comparison between itself and the related work.
 
@@
 \textit{3. Event/task and location/time dynamics}:
-Aside from the time-related information used for information aging, dynamics were not used in the simulations presented here.
-However, TruMan can be extended to consider event, task, location and time dynamics when managing trust.
+TruMan can be extended to consider event, task, location and time dynamics when managing trust.
+However, aside from the time-related information used for information aging, dynamics were not used in the simulations presented here.
 
@@
 the more evidence they gather), this value can be used as a confidence measure of the opinion.
-Setting the threshold $h$ to a value closer to 1 makes it unlikely that untrustworthy nodes will be incorrectly labeled as correct.
+Setting the threshold $h$ to a value closer to 1 makes it unlikely that malicious nodes will be incorrectly labeled as benign.
 
 \textit{6. System level security}:
-A public-private key solution can be used to verify message integrity.
-This has not been considered in evaluations of TruMan.
-However, it can be included as a separate security model during the transmission of messages.
+A public-private key solution can be used to verify message integrity, although it is not a requirement for the functioning of TruMan.
+Any cryptographic solution can be integrated as a separate security model during the transmission of messages without affecting TruMan's efficiency and efficacy.
 
@@
 
-\textit{8. Robustness}: the model's resistance to attacks.
-TruMan satisfies this property. Malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
+\textit{8. Robustness}:
+TruMan satisfies this property, because malicious nodes are quickly and accurately identified, making it difficult for them to perform attacks.
 Experiments show that, when fewer than 50\% of nodes in the network are malicious, TruMan performs as expected.
@@
 The model satisfies or permits most of the desired properties of a trust model with low computational complexity, making it viable for real-world use.
+
+\begin{table}[t!]
+\caption{Properties of the related work}
+\label{table:properties_truman}
+\centering
+\begin{tabular}{|p{4cm}||p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|p{0.5cm}|}
+ \hline
+ \textbf{Property} & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8\\
+ \hline
+ \hline
+ TruMan & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & - & \checkmark \\
+ \hline
+ \citep{dotzer2005vars} & \checkmark & - & \checkmark & - & - & - & \checkmark & -\\
+ \hline
+ \citep{minhas2010towards} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & -\\
+ \hline
+ \citep{chen2010trust} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & - & -\\
+ \hline
+ \citep{park2011long} & - & - & - & - & - & \checkmark & \checkmark & \checkmark \\
+ \hline
+ \citep{huang2014social} & \checkmark & - & \checkmark & \checkmark & \checkmark & - & - & -\\
+ \hline
+ \citep{li2016art} & \checkmark & - & - & - & \checkmark & - & - & \checkmark\\
+ \hline
+ \citep{chen2017cloud} & - & \checkmark & - & \checkmark & - & \checkmark & \checkmark & -\\
+ \hline
+\end{tabular}
+\end{table}
 

### 2018-04-16 article/article.tex
update from last week

@@
 Nodes which frequently share false or irrelevant data must have lower trust values than the ones which appear to be reliable.
-This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
-The results prove to be accurate and efficient, thanks to the low complexity of the algorithm constituting the trust model.
+This work proposes TruMan, a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
+The results prove to be accurate, detecting nearly all malicious nodes with very few false positives when they constitute up to 50\% of the network.
+The model is also very efficient thanks to the low complexity of the algorithm constituting the trust model.
 \end{abstract}
@@
 
-Each of these problems require specific solutions, although there are ways of making the network safer in general.
+Each of these problems requires specific solutions, although there are ways of making the network safer in general.
 One way is taking advantage of the concept of trust between network members.
@@
 A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
-There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+Testing the correctness of neighboring nodes is a problem in and of itself, which is beyond the scope of this paper, but studies on this topic can be found on \cite{golle2004detecting}, \cite{li2016defective}, \cite{kerrache2016detection}.
+%There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+
 Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
@@
 Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
+Malicious nodes misbehave by randomizing the values of their opinions of neighboring nodes.
 
@@
 
-Finally, it is worth noting that, aside from \cite{vernize2015malicious}, none of the mentioned related work presents complexity calculations for their algorithm.
+Finally, it is worth noting that, aside from \cite{vernize2015malicious}, none of the mentioned related work presents complexity calculations for its algorithm.
 Considering the scale of the problem, TruMan's cost of $O(|V|\times |E|)$ is very low without sacrificing completeness and correctness.
@@
 In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
+Nearly all malicious nodes are detected when they constitute up to 50\%, with very few false positives polluting the results, without incurring substantial cost to the vehicles' on-board units.
+%The results prove to be accurate, detecting nearly all malicious nodes with very few false positives when they constitute up to 50\% of the network.
+%The model is also very efficient thanks to the low complexity of the algorithm constituting the trust model.
+
 As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.

### 2018-04-17 dissertation/0-preambulo/abstract.tex

@@
 A decentralized ad-hoc solution, which doesn't rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
-However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
+However, as is the case with most new technologies, VANETs might be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
 In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
@@
 This work introduces TruMan, a trust management model for vehicular networks in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
-This movement model allows the comparison of VANETs to traditional social networks, because it can be observed that pairs of vehicles are likely to meet more than once in several scenarios: for example, they can belong to neighbors or work colleagues, or simply take similar routes every day.
+This movement model allows the comparison of VANETs to traditional social networks, as it can be observed that pairs of vehicles are likely to meet more than once in several scenarios: for example, they can belong to neighbors or work colleagues, or simply take similar routes every day.
 Through these repeated encounters, a trust relationship can be developed between a pair of nodes.

### 2018-04-17 dissertation/0-preambulo/resumoextendido.tex

\begin{resumoextendido}

	\section*{Introdução}
	Este é o resumo extendido em português.

	\section*{Redes Complexas}

	\section*{Redes Veiculares}

	\section*{TruMan}

	\section*{Simulações}

	\section*{Conclusão}
\end{resumoextendido}

### 2018-04-17 dissertation/5-simulations/chapter5.tex

@@
 
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1/2).}
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1\%, 5\% and 10\%).}
 \label{fig:randommalicious1}
@@
 
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (2/2).}
+\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (30\%, 40\% and 50\%).}
 \label{fig:randommalicious2}
@@
 \centering
-\includegraphics[width=\textwidth]{images/plots/Network_rA7/10_10}
+\includegraphics[width=0.6\textwidth]{images/plots/Network_rA7/10_10}
 \caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
@@
 
-\caption{Simulation with information aging, with different maximum age values (1/2).}
+\caption{Simulation with information aging, with different maximum age values ($m=10, 250$).}
 \label{fig:extramaliciousaging1}
@@
 
-\caption{Simulation with information aging, with different maximum age values (2/2).}
+\caption{Simulation with information aging, with different maximum age values ($m=1000, 5000$).}
 \label{fig:extramaliciousaging2}

### 2018-04-18 article/article.tex

--- 2018-02-20 article/article.tex
@@
 This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
-The results prove to be accurate and efficient, thanks to the low complexity of the algorithm constituting the trust model.
+The results prove to be accurate and efficient, detecting nearly all malicious nodes when they constitute up to 50\% of the network, with very few false positive detections.
 \end{abstract}
@@
 A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
-There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+Testing the correctness of neighboring nodes is a problem in and of itself, which is beyond the scope of this paper, but studies on this topic can be found on \cite{golle2004detecting}, \cite{kerrache2016detection}, \cite{li2016defective}.
+%There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+
 Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases and the trust graph $T^v_{i-1}$ is merged into $u$'s trust graph.
@@
 Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
+Malicious nodes in the simulation misbehave by randomizing their opinions of neighbors (i.e. the trust value of outgoing edges to other nodes).
+%The behavior of malicious nodes is to randomize their opinions regarding their neighbors.
 
@@
 The concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
-In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
+In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results
+
 As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.

### 2018-04-25 article/article.tex

--- 2018-04-16 article/article.tex
@@
 Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.
-Malicious nodes misbehave by randomizing the values of their opinions of neighboring nodes.
+Malicious nodes in the simulation misbehave by randomizing their opinions of neighbors (i.e. the trust value of outgoing edges to other nodes).
+%The behavior of malicious nodes is to randomize their opinions regarding their neighbors.
 
@@
 In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.
-Nearly all malicious nodes are detected when they constitute up to 50\%, with very few false positives polluting the results, without incurring substantial cost to the vehicles' on-board units.
+Nearly all malicious nodes are detected when they constitute up to 50\% of the network, with very few false positives polluting the results, without incurring substantial cost to the vehicles' on-board units.
 %The results prove to be accurate, detecting nearly all malicious nodes with very few false positives when they constitute up to 50\% of the network.

### 2018-04-25 dissertation/4-project/chapter4.tex
dissertation local revision

@@
 A prerequisite of this step is a test that correctly classifies a neighboring node as benign or malicious.
-There are several ways to perform such a test, but a study between these methods is beyond the scope of this paper.
+Testing the correctness of neighboring nodes is a problem in and of itself, which is beyond the scope of this project, but studies on this topic can be found on \citep{golle2004detecting}, \citep{kerrache2016detection}, \citep{li2016defective}.
+
 Every time a neighboring node $v$ is tested as benign, the value of $u \rightarrow v$ increases.
@@
 Furthermore, after each iteration, nodes go through the edges they have stored and discard the ones that are too old.
-This is done by setting a maximum age $m$ for edges; once a edge is older than $m$, it is discarded.
+This is done by setting a maximum age $m$ for edges; once an edge is older than $m$, it is discarded.
 A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).

### 2018-04-25 dissertation/5-simulations/chapter5.tex

@@
 However, a few experiments showed that it was not necessary to run iterations of TruMan that frequently; therefore, iterations run at an interval of 100 simulated seconds and only use one tenth of the snapshots saved.
+
+Malicious nodes in the simulation misbehave by randomizing their opinions of neighbors.
+This means that an edge from a malicious node $a$ to another node $b$ is not reliable; its trust value can be anything regardless of the behavior of $b$.
+In collusion attacks, malicious nodes intentionally trust other malicious nodes, but such situations were not considered for the evaluation of TruMan.
 

### 2018-06-18 dissertation/1-intro/chapter1.tex
Updated chapters 1 and 2

@@
 These features are discovered by tracing analogies with social networks.
-There might be groups of nodes which meet each other with a predictable frequency and at predictable intervals, for example vehicles which belong to family members, neighbors or work colleagues.
+For example, certain groups of nodes might come into communication range of each other with predictable frequency, such as vehicles which belong to family members, neighbors or work colleagues as their owners perform their daily activities.
 By considering such features, TruMan allows the formation of strong trust relationships,  which in turn facilitate the discovery of malicious members of the network.

### 2018-06-18 dissertation/2-complexnetworks/chapter2.tex
Updated chapters 1 and 2

@@
-\chapter{Complex Networks}
+\chapter{Trust in Complex Networks}
 \label{chap:complexnetworks}
@@
 They can be comprised of palpable components (such as computers and cables), somewhat abstract entities (such as the World Wide Web's collection of webpages and URLs), or both (like the people and relationships that form a social network).
-
 Complex networks are generally divided into four categories \citep{newmannetworks}:
@@
 Trust is a concept studied in fields such as psychology and economics, with specific definitions.
-In complex networks, under the perspective of computer science, it is a measure of one entity's confidence that another will behave properly and provide valid and/or meaningful data \citep{sherchan2013survey}.
+In complex networks, under the perspective of computer science, trust is a measure of how much confidence one member of a network has that another member of the same network will behave properly and provide valid and/or meaningful data \citep{sherchan2013survey}.
 What follows is the basis of how a network can be modelled using a graph and how a trust model can be applied to it.
@@
 Another important aspect of trust in ad-hoc networks is that information is, generally, uncertain and incomplete \citep{baras2005cooperation}.
-That is, since nodes form their own model and opinions of the surrounding network, it is unlikely that this data will be certain and accurate with reality.
+That is, since nodes form their own model and opinions of the surrounding network, it is unlikely that this data will always be certain and accurate with reality.
 For this reason, it is also possible to use data gathered by neighboring nodes to complement the model.
-This data itself is subject to the trust evaluation of the neighbors, but it is crucial to better approximate the trust values of other nodes.
+Data received from neighbors is also subject to the trust evaluation of whoever requests it, but it is crucial to have better knowledge about other members of the network.
 Incompleteness is an inherit trace of MANETs, since it is entirely possible for nodes to be too distant to communicate, and only occasionally come into contact.

### 2018-06-22 dissertation/1-intro/chapter1.tex
Updated preface and chapter 1

@@
 In 2013, an estimated 1.25 million people lost their lives due to traffic accidents globally \citep{whotraffic}.
-While this number has greatly reduced over the past decades \citep{johnson2010traffic} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \citep{whofactsheet}, so further actions are necessary.
-Furthermore, as the car population increases, congestions consume ever more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \citep{inrixtraffic}.
+While this number has greatly reduced over the past decades \citep{johnson2010traffic} \citep{iihsfatality} thanks to better safety features (seat belts, air bags, ABS, etc.) and stronger laws (drunk driving, motorcycle helmets, speed limits, etc.), it may still rise as a major cause of death in the years to come \citep{whofactsheet}, so further actions are necessary.
+Furthermore, as the car population increases, congestions consume even more time of the daily commuter, peaking at over 100 hours per year for the residents of Los Angeles, CA \citep{inrixtraffic}.
+Moscow, New York, Bogotá, São Paulo, London and Paris are also among the ten most congested cities in the world \citep{inrixtraffic}.
 
@@
 %These features are possible with the development of a vehicular ad-hoc network (VANET), in which vehicles can quickly share data amongst themselves without the need of a server or an Internet connection.
+
+Today, certain vehicle manufacturers already include the on-board technology required to enable vehicular communications in the real world \citep{toyota2015}.
+However, this technology only becomes truly useful when there are other vehicles or infrastructure with whom to communicate, so, at the moment, the benefits are notable but limited \citep{cadillac2017}.
 
@@
 Neighboring vehicles can share their position and velocity data at high frequencies, allowing, for example, for autonomous vehicles to plan a platooning approach to traffic \citep{amoozadeh2015platoon}.
-In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \citep{li2007routing}.
+In the case of a collision or other event, nearby nodes can broadcast alerts, which other nodes pick up and forward \citep{li2007routing}, as illustrated by \Cref{fig:collision}.
 That way, an alert can travel long distances in little time, allowing approaching vehicles to safely slow down or pick alternative routes.
@@
 The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
-Then, \autoref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
+Then, \Cref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
 \Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
-Next, \autoref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
-Finally \autoref{chap:conclusion} presents the final thoughts on the project and concludes this document.
+Next, \Cref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
+Finally \Cref{chap:conclusion} presents the final thoughts on the project and concludes this document.
 
-%Finally, \autoref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.
+%Finally, \Cref{chap:proposal} shows the fundamentals of this proposed work, including the similarities found between VANETs and social networks, a realistic movement model that may be used for simulations, the previously existing study on malicious node identification for complex networks, and the important aspects that must be adapted to the dynamic vehicular environment.

### 2018-06-22 dissertation/2-complexnetworks/chapter2.tex
Updated preface and chapter 1

@@
-\chapter{Trust in Complex Networks}
+\chapter{Complex Networks and Trust}
 \label{chap:complexnetworks}
@@
 		Although not of widespread use, MANETs can provide a way to create a network without pre-existing infrastructure, as long as each device is equipped with the proper hardware and software.
-		Trust issues in technological networks and MANETs are detailed in \autoref{section:trusttechnological}.
-		VANETs, which are special types of MANETs, are introduced in \autoref{chap:vanets}, along with several details regarding trust in those types of networks.
+		Trust issues in technological networks and MANETs are detailed in \Cref{section:trusttechnological}.
+		VANETs, which are special types of MANETs, are introduced in \Cref{chap:vanets}, along with several details regarding trust in those types of networks.
 	\item \textbf{Social Networks} are formed of relationships between people, or groups of people.
@@
 		For the purposes of this work, the most relevant type of relationship is that of trust.
-		The details surrounding trust relationships in social networks are shown in \autoref{section:trustsocial}.
+		The details surrounding trust relationships in social networks are shown in \Cref{section:trustsocial}.
 	\item \textbf{Information Networks} are the ones in which nodes are pieces of data or information and the edges are the connections between those pieces.
@@
 Both types of network can be fitted into the model above, but contain distinct features that demand closer examination.
-Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \autoref{chap:vanets}.
+Furthermore, features of both are relevant when analyzing network structure and trust graphs for vehicular networks, which are expanded upon in \Cref{chap:vanets}.
 
@@
 While positive trust is often tied to the social topology, it is not always the case: one example is two work colleagues who may have a professional relationship, but wouldn't trust each other on other matters; another is the trust people place in authority figures without necessarily having met.
-\autoref{fig:social_network} shows an example of a small social network containing a family and an office.
+\Cref{fig:social_network} shows an example of a small social network containing a family and an office.
 
-In \autoref{section:socialvanets}, the argument is made that VANETs can be considered social networks in several occasions and how this can be used to develop trust in vehicular networks.
+In \Cref{section:socialvanets}, the argument is made that VANETs can be considered social networks in several occasions and how this can be used to develop trust in vehicular networks.
 
@@
 
-% In ad-hoc networks, such as vehicular networks, which are detailed in \autoref{chap:vanets}, trust can be handled in very different ways.
+% In ad-hoc networks, such as vehicular networks, which are detailed in \Cref{chap:vanets}, trust can be handled in very different ways.
 While the centralized trust solutions above serve their purpose on the security and privacy of Internet users, they would be too slow to be viable in a dynamic ad-hoc network, which cannot rely on a back-end infrastructure to distribute those lists.
@@
 
-\autoref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
+\Cref{fig:manet} is an example highlighting the difference between the topology graph and the trust graph.
 The topology graph changes as one node moves across the network, with edges being added and removed as nodes move into and out of each other's ranges.

### 2018-06-22 dissertation/3-vanets/chapter3.tex
Updated chapters 1 and 2
Updated preface and chapter 1

@@
-\chapter{Vehicular Ad-hoc Networks}
+\chapter{Trust in Vehicular Ad-hoc Networks}
 \label{chap:vanets}
@@
 
-%The advantages of an ad-hoc approach, as opposed to using cellular infrastructure or other server-based solutions, are explained in \autoref{chap:introduction}.
+%The advantages of an ad-hoc approach, as opposed to using cellular infrastructure or other server-based solutions, are explained in \Cref{chap:introduction}.
 %For messages to travel to nodes out of the sender's wireless range, they must use other nodes as \textit{hops}.
@@
 Each of these possible attacks requires a unique approach, though there are some broader ways to help the security and safety of VANET users.
-Trust, as is described in \autoref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
+Trust, as is described in \Cref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
 It does not, however, avoid all possible attacks, such as a signal jammer or stalking.
@@
 % with an assessment of whether or not TruMan satisfies their conditions.
-%\autoref{table:properties} shows how well TruMan and previous work satisfy the properties.
+%\Cref{table:properties} shows how well TruMan and previous work satisfy the properties.
 
@@
 
-%Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \autoref{chap:proposal}.
+%Although the model proposed here will not take role-based trust into consideration and does not emphasize event-related messages, the author's approach to experience-based trust resembles what will be described in \Cref{chap:proposal}.
 
@@
 %conclude
-\autoref{table:properties} shows how well each studied trust model satisfies the eight properties described in \autoref{section:properties}.
-The numbers used in the table are the same ones from \autoref{section:properties}.
+\Cref{table:properties} shows how well each studied trust model satisfies the eight properties described in \Cref{section:properties}.
+The numbers used in the table are the same ones from \Cref{section:properties}.
 This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.

### 2018-06-22 dissertation/4-project/chapter4.tex
updates to keynote and minor updates to dissertation
Updated preface and chapter 1

@@
 In order to make this possible, it is necessary to identify features in VANETs that show that nodes can share a long-term relationship, as is the case for social networks.
-%The arguments for such features are presented in \autoref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \autoref{section:workingday}.
+%The arguments for such features are presented in \Cref{section:socialvanets}, while the movement model that enables the simulation of the features is explained in \Cref{section:workingday}.
 Through these long-term relationships, it then becomes feasible for nodes to store trust data and share it with other nodes.
@@
 
-Furthermore, the desired properties of a VANET trust model described in \citep{zhang2011survey} (explained in detail in \autoref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.
+Furthermore, the desired properties of a VANET trust model described in \citep{zhang2011survey} (explained in detail in \Cref{section:properties}) were considered, so TruMan attempts to fulfill or enable as many of those properties as possible.
 
@@
 The complexity of the algorithm is $O(|V|+|E|)$ for a graph $T = (V,E)$.
-Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.
+Algorithm \Cref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \citep{tarjan1972depth}.
 
@@
 
-\autoref{fig:tarjan} illustrates the execution of Tarjan's algorithm.
+\Cref{fig:tarjan} illustrates the execution of Tarjan's algorithm.
 The algorithm starts from node 0, with $index[0] = 0$ and $lowlink[0] = 0$.
@@
 As a comparison, the DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \citep{brelaz1979new}.
-Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.
+Algorithm \Cref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \citep{mittal2011graph}.
 
@@
 
-\autoref{fig:coloring} illustrates the execution of the graph coloring algorithm.
+\Cref{fig:coloring} illustrates the execution of the graph coloring algorithm.
 It is notable how the algorithm takes few iterations to fully color the graph.
@@
 
-First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \autoref{section:tarjan}.
+First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \Cref{section:tarjan}.
 In each of these components, all nodes are connected by edges of value $1$.
@@
 In the experiments performed by the authors of MaNI, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
-The coloring heuristic uses a graph coloring algorithm, such as DSATUR \citep{brelaz1979new} or the algorithm detailed in \autoref{section:coloring}.
+The coloring heuristic uses a graph coloring algorithm, such as DSATUR \citep{brelaz1979new} or the algorithm detailed in \Cref{section:coloring}.
 Other heuristics were experimented with, but were either less effective in detecting malicious nodes, provided too many false positives, or were not efficient enough.
@@
 
-\autoref{fig:mani} illustrates the execution of the MaNI algorithm.
+\Cref{fig:mani} illustrates the execution of the MaNI algorithm.
 With the starting graph $T$, whose edges represent the trust relationships between nodes, Tarjan's strongly connected component algorithm is executed.
-\autoref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
+\Cref{fig:mani1} shows nodes colored according to their placement in a strongly connected component.
 The strongly connected components form a graph $C$ according to edges present in $T$.
-In \autoref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
-The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \autoref{fig:mani3}.
+In \Cref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
+The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \Cref{fig:mani3}.
 Finally, each node in $T$ is colored according to which color its component received in $C$.
@@
 Additionally, node $u$ performs an union between its trust graph and $v$'s trust graph, forming a new graph $T^u_i = T^u_{i-1} \bigcup T^v_{i-1}$, which is used for the remaining steps.
-Algorithm  \autoref{algorithm:interaction} shows the basic interaction between two nodes, while algorithm \autoref{algorithm:union} details the steps of the graph union.
+Algorithm  \Cref{algorithm:interaction} shows the basic interaction between two nodes, while algorithm \Cref{algorithm:union} details the steps of the graph union.
 
 \begin{algorithm}
-\caption{Interaction between two nodes}\label{algorithm:interaction}
+\caption{Interaction between two nodes $u$ and $v$}\label{algorithm:interaction}
 \begin{algorithmic}[1]
-
-\Function{Interaction}{node $u$, node $v$}
 
@@
 	\State $T^u(u\rightarrow v).trustvalue$ \textbf{increases}
-	\State \textbf{Union}($T^u$, $T^v$)
+	\State $T^u \gets $ \textbf{Union}($T^u$, $T^v$)
 \Else
@@
 
-\EndFunction
 \end{algorithmic}
@@
 
+\State $T \gets T^u$
+
 \For{$a\rightarrow b$ \textbf{in} edges of $T^v$ }
 
-	\If{$a\notin T^u$}
-		\State \textbf{add} $a$ \textbf{to} $T^u$
+	\If{$a\notin T$}
+		\State \textbf{add} $a$ \textbf{to} $T$
 	\EndIf
 
-	\If{$b\notin T^u$}
-		\State \textbf{add} $b$ \textbf{to} $T^u$
+	\If{$b\notin T$}
+		\State \textbf{add} $b$ \textbf{to} $T$
 	\EndIf
 
-	\If{$a\rightarrow b \notin T^u$}
-		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
+	\If{$a\rightarrow b \notin T$}
+		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T$
 	\EndIf
+
 \EndFor
+
+\State \textbf{return} $T$
 
@@
 
-The coloring heuristic is used to classify nodes, which uses the algorithm described in \autoref{section:coloring} \citep{mittal2011graph}, although other heuristics may be considered.
+The coloring heuristic is used to classify nodes, which uses the algorithm described in \Cref{section:coloring} \citep{mittal2011graph}, although other heuristics may be considered.
 After running the graph coloring algorithm with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
@@
 Since the timestamp is always set to the destination node of the edge, it can be used for comparison regardless of which nodes handled the information. 
-Algorithm \autoref{algorithm:union_timestamp} shows how the union function was changed to accommodate timestamps and information aging.
+Algorithm \Cref{algorithm:union_timestamp} shows how the union function was changed to accommodate timestamps and information aging.
 
@@
 
+\State $T \gets T^u$
+
 \For{$a\rightarrow b$ \textbf{in} edges of $T^v$ }
 
-	\If{$a\notin T^u$}
-		\State \textbf{add} $a$ \textbf{to} $T^u$
+	\If{$a\notin T$}
+		\State \textbf{add} $a$ \textbf{to} $T$
 	\EndIf
 
-	\If{$b\notin T^u$}
-		\State \textbf{add} $b$ \textbf{to} $T^u$
+	\If{$b\notin T$}
+		\State \textbf{add} $b$ \textbf{to} $T$
 	\EndIf
 
-	\If{$a\rightarrow b \notin T^u$}
-		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T^u$
+	\If{$a\rightarrow b \notin T$}
+		\State \textbf{add} $a\rightarrow b$ \textbf{to} $T$
 	\Else
-		\State $s_0 \gets T^u(a\rightarrow b).timestamp$
-		\State $s_1 \gets T^v(a\rightarrow b).timestamp$
-		\If{$s_1$ \textbf{is more recent than} $s_0$}
-			\State $T^u(a\rightarrow b).trustvalue \gets T^v(a\rightarrow b).trustvalue$
-			\State $T^u(a\rightarrow b).timestamp \gets T^v(a\rightarrow b).timestamp$
+		\State $s_u \gets T(a\rightarrow b).timestamp$
+		\State $s_v \gets T^v(a\rightarrow b).timestamp$
+		\If{$s_u$ \textbf{is more recent than} $s_v$}
+			\State $T(a\rightarrow b).trustvalue \gets T^v(a\rightarrow b).trustvalue$
+			\State $T(a\rightarrow b).timestamp \gets T^v(a\rightarrow b).timestamp$
 		\EndIf	
@@
 \EndFor
+
+\State \textbf{return} $T$
 
@@
 
-Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \autoref{fig:rogue}.
+Information aging is necessary because, when a node converts from benign to malicious, old information might still indicate that it is benign, as shown in \Cref{fig:rogue}.
 Immediately after node $0$ becomes malicious, it is still part of a benign strongly connected component, because there are paths from every other node to $0$, and from $0$ to every other node.

### 2018-06-22 dissertation/5-simulations/chapter5.tex
Updated preface and chapter 1

@@
 
-%The datasets available from SNAP's website were used for the simulations in \citep{vernize2015malicious} and the static network simulations shown in \autoref{chap:simulations}.
+%The datasets available from SNAP's website were used for the simulations in \citep{vernize2015malicious} and the static network simulations shown in \Cref{chap:simulations}.
 
@@
 
-To make use of the properties described in \autoref{section:socialvanets}, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
+To make use of the properties described in \Cref{section:socialvanets}, it is important to choose a mobility pattern that properly represents the way vehicles move on a daily basis in the real world.
 Therefore, the Working Day Movement Model \citep{ekman2008working} (WDMM) is useful.
@@
 The choice of the Working Day Movement Model for evaluations is more strongly related to inter-contact times.
-For reasons explained in \autoref{chap:truman}, relatively short inter-contact times is important for TruMan's functionality.
+For reasons explained in \Cref{chap:truman}, relatively short inter-contact times is important for TruMan's functionality.
 Contact duration time is an important metric to measure how much data can be exchanged during each encounter, although, for this evaluation, it was not considered.
@@
 
-Most of the parameters for the ONE simulator were taken from the article detailing the Working Day Movement Model \citep{ekman2008working}; the most important parameters are shown in \autoref{table:parameters}.
+Most of the parameters for the ONE simulator were taken from the article detailing the Working Day Movement Model \citep{ekman2008working}; the most important parameters are shown in \Cref{table:parameters}.
 The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds (8 hours) and a standard deviation of departure time of 7200 seconds (2 hours).
@@
 A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
-The transmission range of nodes varies from simulation to simulation, for reasons explained in \autoref{section:density}.
+The transmission range of nodes varies from simulation to simulation, for reasons explained in \Cref{section:density}.
 
@@
 The network density ($\delta$) is a value which abstracts the volume and frequency of connections in a vehicular network by estimating how much of the environment is covered by the network.
-For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \autoref{chap:simulations}).
+For TruMan, higher densities yield better results, since nodes can construct and update their models of the network faster (this is demonstrated in \Cref{chap:simulations}).
 It is calculated using the transmission range of the nodes ($\rho$), the amount of nodes ($\eta$), and the total area of the simulation ($\alpha$, in m\textsuperscript{2}).
@@
 
-In \autoref{table:simdensities}, a few densities for different values of $\rho$, $\eta$ and $\alpha$ are shown. 
+In \Cref{table:simdensities}, a few densities for different values of $\rho$, $\eta$ and $\alpha$ are shown. 
 Simulations for TruMan have densities between 0.0017 ($\rho = 10$m) and 0.0428 ($\rho = 50$m).
@@
 As a comparison, the city of São Paulo (Brazil) has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, and thus has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
-\autoref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
+\Cref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
 All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan areas that surround these cities.
@@
 Simulations of TruMan were performed with densities as low as 0.0017, a value much lower than even the real world density of Helsinki, which is a city with relatively few vehicles for its size.
-It can be expected that the model will perform even better in real-world scenarios in cities with even higher densities, which is common, as exemplified by \autoref{table:citydensities}.
+It can be expected that the model will perform even better in real-world scenarios in cities with even higher densities, which is common, as exemplified by \Cref{table:citydensities}.
 
@@
 
-\autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
+\Cref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
 It is possible to see how the increase in the range allows the algorithm to produce better results.
@@
 
-\autoref{fig:randommalicious1} and \autoref{fig:randommalicious2} show the variation of results for different amounts of malicious nodes in the network.
+\Cref{fig:randommalicious1} and \Cref{fig:randommalicious2} show the variation of results for different amounts of malicious nodes in the network.
 By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network, although there is still a small amount of false positive detections.
@@
 
-All simulations were performed with trust threshold $h=0.5$, except for the ones in \autoref{fig:randomthresholds}, performed to demonstrate the impact of different threshold values.
+All simulations were performed with trust threshold $h=0.5$, except for the ones in \Cref{fig:randomthresholds}, performed to demonstrate the impact of different threshold values.
 The plots illustrate that there is not a significance change in results depending on the different thresholds.
@@
 
-\autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
+\Cref{fig:random7} shows the execution of the algorithm over the course of 7 days.
 Most malicious nodes are identified by the end of the first day, before iteration 9000.
@@
 
-\autoref{fig:extramaliciousaging1} and \autoref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
+\Cref{fig:extramaliciousaging1} and \Cref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
 These were the only simulations to consider information aging.
@@
 
-Eight desired properties for VANET trust models were presented in \autoref{section:properties}.
+Eight desired properties for VANET trust models were presented in \Cref{section:properties}.
 Here, TruMan's ability to satisfy each property is evaluated.
-\autoref{table:properties_truman} replicates the one from \autoref{section:properties} with the addition of TruMan, providing a full comparison between itself and the related work.
+\Cref{table:properties_truman} replicates the one from \Cref{section:properties} with the addition of TruMan, providing a full comparison between itself and the related work.
 

### 2018-06-22 dissertation/2-complexnetworks/chapter2.tex
updated chapter 2

@@
 For instance, two people can have contact with each other (therefore sharing an edge in $G$) but not maintain a trust relationship (therefore not sharing an edge in $T$), thus altering the layout of the trust graph compared to the network topology.
+
+Trust can be hard to define in the context of biological networks, as these networks are often formed by members who have no distinctive properties to categorize as trustworthy or not and, when such properties do exist, the network is probably closer to a social network than a biological one.
+The same applies to information networks, which are formed by pieces of data that do not need to maintain active relationships with each other.
+Again, an information network that requires trust is likely closer to the category of technological networks.
 

### 2018-07-04 dissertation/1-intro/chapter1.tex
Updated chapter 2

@@
 %\section{Document organization}
-The remainder of this document is organized as follows. \Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in technological and social networks.
-Then, \Cref{chap:vanets} goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
+The remainder of this document is organized as follows.
+\Cref{chap:complexnetworks} explains the broad study of complex networks and the importance of trust in social and technological networks, then goes into details regarding VANETs and the importance of trust solutions in the field, presenting previous studies made on the subject.
 \Cref{chap:truman} describes the goals of TruMan, its underlying hypothesis and the work done to achieve those goals, as well as the previously existing algorithms which form parts of TruMan.
 Next, \Cref{chap:simulations} explains the tools used to validate TruMan's functionality and presents the results of the experiments that were performed.
-Finally \Cref{chap:conclusion} presents the final thoughts on the project and concludes this document.
+Finally, \Cref{chap:conclusion} presents the final thoughts on the project and concludes this document.
 

### 2018-07-04 dissertation/2-complexnetworks/chapter2.tex
Updated chapter 2

@@
-\chapter{Complex Networks and Trust}
+\chapter{Background and Related Work}
 \label{chap:complexnetworks}
 
+In order to develop a new trust management scheme for vehicular networks, a review of the study of networks as a whole is required.
+This chapter introduces the concept of complex networks, digging into the meaning and importance of trust in the cases of social networks and technological networks.
+Then, vehicular ad-hoc networks are explained in details, with the current state of the art, unique properties compared to other networks, desired properties for a vehicular network trust model and details on the most relevant works in the subject.
+
+\section {Complex Networks}
 %Often, the term ``network'' is associated only with computer

### 2018-07-04 dissertation/3-vanets/chapter3.tex
Updated chapter 2

@@
-\chapter{Trust in Vehicular Ad-hoc Networks}
+\section{Trust in Vehicular Ad-hoc Networks}
 \label{chap:vanets}
@@
 Instead, most VANET routing protocols use geographical coordinates to forward packets \citep{saini2015close}, that is, the physical distance between two nodes is used as the primary metric.
-The implication is that, even if a packet requires more hops to reach its destination, it will always be traveling the generally correct direction.
+The implication is that, even if a packet requires more hops to reach its destination, it always travelling in the generally correct geographical direction.
 
@@
 
-\section{Special properties of VANETs}
-\label{section:properties}
-%As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
-%However, 
-VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \citep{yousefi2006vehicular}.
-Some of these properties include:
-
-\begin{enumerate}
-
-	\item 	Rapidly changing topology.
-			Since the nodes are vehicles, they move frequently and at relatively high speeds.
-			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
-%			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
-%			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
-
-	\item	Node mobility is constrained to a pre-existing grid of roads.
-			Within those roads, nodes usually travel in predictable directions according to local laws and historical data.
-			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.
-
-	\item	VANETs are prone to fragmentation, since a gap in the network topology can make two parts of it unable to communicate with each other.
-			Combined with the property above, this fragmentation can appear and disappear frequently, depending on the node density.
-
-	\item	Due to the changing topology and possible disconnection, connection with distant nodes is not reliable.
-			Therefore, the effective diameter of the network is relatively small for important applications.
-
-	\item 	Compared to devices like smartphones, vehicles have no notable power constraints.
-
-	\item 	In certain locations and/or moments, large vehicle density results in a large-scale network, since there are many nodes concentrated in a relatively small space.
-
-	\item 	The topology is susceptible to driver behavior.
-			First, this means the topology can occasionally change in unpredictable ways.
-			Second, contents of a message sent through the network can alter the driver's behavior and therefore change the topology.
-
-\end{enumerate}
-
-Some of these properties provide advantages or disadvantages when developing trust models for vehicular networks, although all of them must be considered.
-
-\section{Trust in VANETs}
-\label{section:trustvanet}
-Like in other types of networks, the proper functioning of a VANET depends on the reliability of the vehicles (nodes) which participate in it.
+Like in other types of networks, the proper functioning of a VANET depends on the reliability of its participants.
 If a node is malicious or faulty, it can spread incorrect data that may compromise the network's utility.
@@
 %[details of security and safety issues in VANETs that can be diminished by trust]
-\section{Desired properties for VANET trust models}
+
+\subsection{Special properties of VANETs}
 \label{section:properties}
+%As was described above, VANETs are a special type of MANETs, which in turn are technological networks.
+%However, 
+VANETs feature several unique properties which distinguish them, and the behavior of its members, from other types of networks \citep{yousefi2006vehicular}.
+Some of these properties include:
+
+\begin{enumerate}
+
+	\item 	Rapidly changing topology.
+			Since the nodes are vehicles, they move frequently and at relatively high speeds.
+			Each node's wireless communications also have a certain range, so the other nodes within that range (and, therefore, network neighbors) can change very quickly.
+%			To build a topology model of its surrounding network, a node must often ping its neighbor to gather information.
+%			Using velocity and location information, however, it is possible to create reasonable assumptions about future states of the topology.
+
+	\item	Node mobility is constrained to a pre-existing grid of roads.
+			Within those roads, nodes usually travel in predictable directions according to local laws and historical data.
+			The spaces in the grid, like city blocks, provide a challenge to communication both because of distance and because buildings can cause obstructions to radio transmissions.
+
+	\item	VANETs are prone to fragmentation, since a gap in the network topology can make two parts of it unable to communicate with each other.
+			Combined with the property above, this fragmentation can appear and disappear frequently, depending on the node density.
+
+	\item	Due to the changing topology and possible disconnection, connection with distant nodes is not reliable.
+			Therefore, the effective diameter of the network is relatively small for important applications.
+
+	\item 	Compared to devices like smartphones, vehicles have no notable power constraints.
+
+	\item 	In certain locations and/or moments, large vehicle density results in a large-scale network, since there are many nodes concentrated in a relatively small space.
+
+	\item 	The topology is susceptible to driver behavior.
+			First, this means the topology can occasionally change in unpredictable ways.
+			Second, contents of a message sent through the network can alter the driver's behavior and therefore change the topology.
+
+\end{enumerate}
+
+Some of these properties provide advantages or disadvantages when developing trust models for vehicular networks, although all of them must be considered.
+
+%\subsection{Trust in VANETs}
+%\label{section:trustvanet}
+
+\subsection{Desired properties for VANET trust models}
+\label{section:desired}
 
@@
 
-\section{Existing trust models for VANETs}
+\subsection{Existing trust models for VANETs}
 
@@
 \citep{dotzer2005vars} is one of the earliest examples of VANET trust models, establishing a system called VARS, based on the reputation of nodes and messages throughout the network.
-The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node appends its opinion of the message's contents and the message's sender.
+The authors use what they call \textit{opinion piggybacking}, which means that, for each hop between the origin and the destination of an event-related message, the forwarding node includes its opinion of the message's contents and the message's sender.
 In other words, when a node $a$ receives a message about a certain event from $b$, it calculates a new opinion considering  it rebroadcasts the message to other nearby nodes, but with its own opinions about the event and about node $b$ attached.
@@
 Nodes also store their experience each time an event message is received (if one neighboring node reported an event which did not turn out to be true, its trust value is reduced).
-Additionally, messages have higher reliability when their senders are closer in time and space to the reported event.
-When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority (fewer chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
+Additionally, messages are considered more reliable when their senders were direct witnesses to an event (i.e. were close to the event when it happened).
+When several messages about the same event are received, a node can either choose the $n$ most trustworthy senders, according to the priority of the event.
+For example, events that require a fast response might cause a decision to be made using fewer messages about it, which diminishes precision.
+% chosen nodes means a faster, but less precise, decision), or compute the majority opinion of the messages according to each sender's trust value.
 The model considers both role-based trust and experience-based trust; although the work proposed here does not use role-based trust, the authors provide a useful method of calculating and updating an experience-based trust value, which might be used or adapted.
@@
 Afterward, the message is only forwarded to another cluster if that aggregate opinion is above a certain threshold; furthermore, nodes that receive the message only act upon it if the overall trust on it is above another threshold, which can be different according to the nature of the message.
-However, it is unclear how the model behaves when the network is too sparse to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
-Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the cluster leader itself is malicious, all the information from that cluster becomes untrustworthy.
+However, it is unclear how the model behaves when the network is too sparse -- nodes are few and/or far apart -- to form relevant clusters, neither do the authors inform how the aforementioned thresholds are decided.
+Furthermore, maintaining clusters in a highly dynamic network is a costly job and, if the node chosen to be the leader of a cluster happens to be malicious, all the information from that cluster could be compromised.
 %The whole premise of organizing a VANET in clusters is  
@@
 To make this viable, the properties of daily commutes are used: it is assumed that the vehicle is near its Agent RSU with reasonable frequency because it is located within the driver's home-to-work route.
-The main problem with this model is that it relies on the presence of frequent RSUs, which might not always be viable.
+The main problem with this model is that it relies on the existence of a reasonably large number of RSUs, which might not always be viable due to infrastructure costs.
 It also does not make it clear what should happen when a vehicle stops using a route or does not have a daily predictable path (it does, however, handle occasions in which a vehicle chooses an alternate route or is absent for some days such as weekends and holidays).
@@
 That is, information reported by a number of original nodes (i.e. the ones that witnessed an event) may be diluted as nodes that forward it append their own opinions on the matter.
-An algorithm is proposed to diminish that effect by assigning higher weights to the opinions of origin nodes and lower weights  to others.
+Following the VARS \citep{dotzer2005vars} approach, an algorithm is proposed to diminish that effect by assigning higher weights to the opinions of nodes that were witnesses to an event and lower weights to forwarding nodes.
 However, the authors conclude that the optimal scenario is to assign no weight at all to forwarding nodes, therefore allowing each node to form an opinion based only on the original nodes' reports.
-Furthermore, the authors are quick to dismiss the validity of entity-based trust, instead opting for a pure data-oriented approach.
+Furthermore, the authors are quick to dismiss the validity of entity-based trust, instead opting for a pure data-oriented approach, considering only the contents of a message and disregarding who sent it.
 %Since their model is based only on events, it also does not consider the usefulness of trust in more mundane cases such as the frequent sharing of location and speed among vehicles.
@@
 %The authors present no details on how it deals with sparsity, dynamics, scalability, security and privacy.
-%extend
 
@@
 In general, it goes against several established concepts for vehicular networks, such as being completely decentralized and ad-hoc.
-It also makes the system somewhat, since the whole system collapses if the global trust manager is attacked, or could leave the network members unaided in the event of a server or connection failure.
-%extend
+It also makes the system somewhat unreliable, since the whole system collapses if the global trust manager is attacked, or could leave the network members unaided in the event of a server or connection failure.
 
-%conclude
-\Cref{table:properties} shows how well each studied trust model satisfies the eight properties described in \Cref{section:properties}.
-The numbers used in the table are the same ones from \Cref{section:properties}.
-This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.
+\section{Discussion}
+\label{section:background_discussion}
 
-\begin{table}[t!]
+Analyzing the previous work in the subject of trust management for vehicular networks, it is possible to observe that many interesting solutions to the problem have been attempted, but all of them show at least one major deficiency when thinking about real-world implementation.
+\Cref{table:properties} shows how well each studied trust model satisfies the eight properties described in \Cref{section:desired}, which are based on \citep{zhang2011survey}.
+The numbers used in the table are the same ones from \Cref{section:desired}.
+
+\begin{table}[hpbt!]
 \caption{Properties of the related work}
@@
 \end{table}
+
+Furthermore, vehicular ad-hoc networks are a special kind of network, but, despite the unique properties explained in \Cref{section:properties}, they still share many features with mobile ad-hoc networks and even with social networks.
+
+With this information in mind, a new trust model is introduced, taking advantage of social features in order to establish long-term trust relationships between participants of a vehicular ad-hoc network.
+This new trust model emphasizes efficiency while maintaining correctness and most of the desired properties for such a model.
+In the following chapter, the model is explained in detail.
+%This analysis shows that, although several previously proposed trust models provide reasonable solutions to the trust management problem, all of them have issues in terms of correctness, completeness or efficiency, indicating that the problem is not yet truly solved.
 

### 2018-07-04 dissertation/4-project/chapter4.tex
Updated chapter 2

@@
-\chapter{TruMan}
+\chapter{Design and Implementation of TruMan}
 \label{chap:truman}
@@
 
+\section{Discussion}
+\label{section:truman_discussion}
+
+This chapter shows the concepts and algorithms that went into the development of TruMan.
+Using efficient algorithms as a foundation, TruMan is able to efficiently detect and identify malicious nodes operating within a dynamic network.
+
+Although TruMan could be a viable solution for multiple kinds of networks, this work is focused primarily on vehicular networks and, therefore, TruMan was developed with such networks in mind.
+The following chapter explains how TruMan was evaluated in the context of vehicular networks, including results of the performed experiments.
+
 %The MaNI algorithm works well and is efficient for static graphs.

### 2018-07-04 dissertation/5-simulations/chapter5.tex
Updated chapter 2

@@
 
-\section{SNAP library}
-\label{section:snap}
+\section{Tools}
+\label{section:tools}
+
+\textbf{SNAP library}
+
 The simulations necessary to validate the project require a robust library to handle graph data structures.
@@
 %The datasets available from SNAP's website were used for the simulations in \citep{vernize2015malicious} and the static network simulations shown in \Cref{chap:simulations}.
+%\\
+\textbf{The ONE Simulator}
 
-\section{The ONE Simulator}
-\label{section:theone}
 The Opportunistic Network Environment simulator \citep{keranen2009one} \citep{onerepo} is a Delay-Tolerant Network simulator, used in this study to generate mobility patterns used as input for simulations.

### 2018-07-04 dissertation/4-project/chapter4.tex
Updated chapter 3

@@
 
-The process of graph coloring consists of giving each node a label (represented by a color) so that no two neighboring nodes share the same color.
+The process of graph coloring consists of giving each node a label so that no two neighboring nodes share the same label.
+When visualizing a graph, labels are represented by colors, although they can be any type of data.
 This problem has been studied in Computer Science since, at least, 1972 \citep{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \citep{kempe1879geographical}.
@@
 First, $T$ is separated into strongly connected components using Tarjan's algorithm \citep{tarjan1972depth}, which is described in detail in \Cref{section:tarjan}.
-In each of these components, all nodes are connected by edges of value $1$.
-In other words, within a single component, all nodes trust one another; nodes connected by edges of value $0$ are separated into different components.
+In each of these components, there are paths formed by edges of value $1$ connecting each pair of nodes. 
+% there is a path of trust ball nodes are connected by edges of value $1$.
+In other words, within a single component, all nodes trust one another directly or indirectly; nodes which do not do not meet this criteria are separated into different components.
 Each of these components becomes a node of a component graph $C = (V', E')$.
@@
 The creation of the graph $C$ simplifies the remaining computation.
-Since each node of $C$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
+Since each node of $C$ is a vertex $v' \in V'$ and each vertex $v'$ is a component of $T$ in which all nodes trust each other directly or indirectly, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as one.
 They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
@@
 The strongly connected components form a graph $C$ according to edges present in $T$.
-In \Cref{fig:mani2}, component $0$ is the one containing node $5$; component $1$ contains nodes $6$, $7$ and $8$; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
+In \Cref{fig:mani2}, component $0$ is the one containing nodes 6, 7 and 7; component $1$ contains node 5; component $2$ contains nodes $0$, $1$, $2$ and $3$; and component $3$ contains node $4$.
 The graph coloring with minimum colors algorithm is executed on $C$, producing the coloration shown in \Cref{fig:mani3}.
@@
 Malicious nodes, on the other hand, do not send their own networks of false information to benign nodes, and might not always trust other malicious nodes, causing them to become isolated in small strongly connected components (in some cases, these contain a single malicious node).
-The result is that most benign nodes become members of a small number of large components; when the component graph is colored, these components are likely to receive the same color, because two benign components are never adjacent.
+The result is that most benign nodes become members of a small number of large components; when the component graph is colored, these components are likely to receive the same color, because two benign components are almost never adjacent
+\footnote{
+	When a node is building its local representation of a network, new nodes are almost certainly identified as malicious. 
+	This happens because there is a trust edge going from an already established node to the new one, but the returning edge is not there yet. 
+	Therefore, two adjacent benign nodes are separated into different components. 
+	This causes inaccurate results in the very beginning of execution, and becomes a minor detail afterwards.
+}.
 Because of this, the coloring heuristic works as a classification method.
-A high scale collusion attack could affect this heuristic, as malicious nodes would form large components and distinguishing these components from the benign ones would be a challenge.
+A large scale collusion attack (in which malicious nodes form nearly half the network or more) could affect this heuristic, as malicious nodes would form large components and distinguishing these components from the benign ones would be a challenge.
 %Classifying the color that represents the most nodes as correct works because
@@
 This is done by setting a maximum age $m$ for edges; once an edge is older than $m$, it is discarded.
-A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increacing the likelihood of false positive detections).
+A low value for this setting allows for faster detection of converted nodes, however  also increases the likelihood of good information being thrown away (therefore increasing the likelihood of false positive detections).
 

### 2018-07-06 dissertation/5-simulations/chapter5.tex
updated results chapter with better graph positioning

@@
-\chapter{Evaluation}
+\chapter{Evaluation of TruMan}
 \label{chap:simulations}
@@
 %\\
+\vspace{5mm}
 \textbf{The ONE Simulator}
@@
 
-As a comparison, the city of São Paulo (Brazil) has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, and thus has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
-\Cref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
-All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan areas that surround these cities.
-%Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
-
 \defcitealias{helsinkifleet}{Helsinki, 2011}
@@
 
+As a comparison, the city of São Paulo (Brazil) has a fleet of over 8 million vehicles \citep{saopaulofleet} in an area of 1,521.11 km\textsuperscript{2} \citep{saopauloarea}, and thus has a density of 0.8884 at $\rho = 10$m, a much higher value than what is necessary for a satisfactory performance of the algorithm.
+\Cref{table:citydensities} shows the densities of a few major cities around the world, using $\rho = 10$m for all of them.
+All data is taken from local government sources regarding the number of licensed vehicles in each city; these numbers do not include vehicles from the larger metropolitan areas that surround these cities.
+%Areas are shown in km\textsuperscript{2} (values were converted to m\textsuperscript{2} for calculation).
+
 Simulations of TruMan were performed with densities as low as 0.0017, a value much lower than even the real world density of Helsinki, which is a city with relatively few vehicles for its size.
@@
 
+%\begin{figure}
+%\centering
+%
+%\begin{subfigure}{0.4\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
+%\caption{$\rho$ = 10m.}
+%\label{fig:random10}
+%\end{subfigure}
+%
+%\hspace*{\fill} % separation between the subfigures
+%
+%\begin{subfigure}{0.4\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+%\caption{$\rho$ = 30m.} 
+%\label{fig:random30}
+%\end{subfigure}
+
+%\begin{subfigure}{0.6\textwidth}
+%\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+%\caption{$\rho$ = 50m.} 
+%\label{fig:random50}
+%\end{subfigure}
+%
+%\caption{Simulation of TruMan with 10\% malicious nodes and varying values of $\rho$.}
+%\label{fig:random103050}
+%\end{figure}
+
+%\begin{figure}
+%\centering
+%
+%	\includegraphics[width=0.6\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
+%	\caption{$\rho$ = 10m.} 
+%	
+%%\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+%\label{fig:random10}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%
+%	\includegraphics[width=0.6\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+%	\caption{$\rho$ = 30m.} 
+%	
+%%\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+%\label{fig:random30}
+%\end{figure}
+
+%\end{subfigure}%
+
+%\hspace*{\fill}
+
+%\begin{subfigure}{0.4\textwidth}
+%	\centering
+%	\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+%	\caption{$\rho$ = 30m.} 
+%	\label{fig:tarjan0}
+%\end{subfigure}
+%
+%\begin{subfigure}{0.6\textwidth}
+%\includegraphics[width=0.5\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+%\caption{$\rho$ = 50m.} \label{fig:tarjan0}
+%\end{subfigure}
+
+%\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+%\label{fig:random103050}
+%\end{figure}
+
+\pagebreak
 \Cref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
@@
 
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
+\caption{$\rho$ = 10m.} 
+\label{fig:random10}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+\caption{$\rho$ = 30m.} 
+\label{fig:random30}
+\end{subfigure}
+
+\centering
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+\caption{$\rho$ = 50m.} 
+\label{fig:random50}
+\end{subfigure}
+
+\caption{Simulation of TruMan with 10\% malicious nodes and varying values of $\rho$.}
+\label{fig:random103050}
+\end{figure}
+
+%\begin{figure}
+%\centering
+%
+%	\includegraphics[width=0.6\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+%	\caption{$\rho$ = 50m.} 
+%	
+%%\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
+%\label{fig:random50}
+%\end{figure}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
+%\caption{$\rho$ = 10m.} \label{fig:tarjan0}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
+%\caption{$\rho$ = 30m.} \label{fig:tarjan1}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.5\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
+%\caption{$\rho$ = 50m.} \label{fig:tarjan2}
+%\end{figure}
+
 % taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.
 
+\pagebreak
 \Cref{fig:randommalicious1} and \Cref{fig:randommalicious2} show the variation of results for different amounts of malicious nodes in the network.
@@
 
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
+\caption{1\% malicious.}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
+\caption{5\% malicious.}
+\end{subfigure}
+
+\centering
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
+\caption{10\% malicious.}
+\end{subfigure}
+
+\caption{Simulation of TruMan with $\rho$ = 10m and varying percentages of malicious nodes (1\%, 5\% and 10\%).}
+\label{fig:randommalicious1}
+\end{figure}
+
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
+\caption{30\% malicious.}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
+\caption{40\% malicious.}
+\end{subfigure}
+
+\centering
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
+\caption{50\% malicious.}
+\end{subfigure}
+
+\caption{Simulation of TruMan with $\rho$ = 10m and varying percentages of malicious nodes (30\%, 40\% and 50\%).}
+\label{fig:randommalicious2}
+\end{figure}
+
+\pagebreak
 All simulations were performed with trust threshold $h=0.5$, except for the ones in \Cref{fig:randomthresholds}, performed to demonstrate the impact of different threshold values.
@@
 It still takes over 8000 iterations to detect all malicious nodes and there are still some false positive detections.
+
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
+\caption{$h = 0.3$.} \label{fig:threshold03}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
+\caption{$h = 0.5$.} \label{fig:threshold05}
+\end{subfigure}
+
+\centering
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
+\caption{$h = 0.7$.} \label{fig:threshold07}
+\end{subfigure}
+
+\caption{Simulation of TruMan with 10\% malicious nodes, $\rho = 30$m and varying values of $h$.}
+\label{fig:randomthresholds}
+\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{images/plots/Network_rA7/10_10}
+\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+\end{figure}
+
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
+\caption{$m = 10$.} \label{fig:extra10}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
+\caption{$m = 250$.} \label{fig:extra250}
+\end{subfigure}
+
+\centering
+\caption{Simulation of TruMan with information aging, with different maximum age values ($m=10, 250$).}
+\label{fig:extramaliciousaging1}
+\end{figure}
+
+\begin{figure}
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
+\caption{$m = 1000$.} \label{fig:extra1000}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
+\caption{$m = 5000$.} \label{fig:extra5000}
+\end{subfigure}
+
+\centering
+\caption{Simulation with information aging, with different maximum age values ($m=1000, 5000$).}
+\label{fig:extramaliciousaging2}
+\end{figure}
+
+\pagebreak
 \Cref{fig:extramaliciousaging1} and \Cref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
@@
 
-\begin{table}[t!]
+\begin{table}[hpbt!]
 \caption{Properties of the related work}
@@
 \end{table}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/10_10.png}
-\caption{$\rho$ = 10m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/30_10.png}
-\caption{$\rho$ = 30m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA_10.0/new_plots/50_10.png}
-\caption{$\rho$ = 50m.} \label{fig:tarjan0}
-\end{subfigure}
-
-\caption{Simulation with 10\% malicious nodes and varying values of $\rho$.}
-\label{fig:random103050}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_1.png}
-\caption{1\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_5.png}
-\caption{5\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_10.png}
-\caption{10\% malicious.}
-\end{subfigure}
-
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (1\%, 5\% and 10\%).}
-\label{fig:randommalicious1}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_30.png}
-\caption{30\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_40.png}
-\caption{40\% malicious.}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/Network_rA/10_50.png}
-\caption{50\% malicious.}
-\end{subfigure}
-
-\caption{Simulation with $\rho$ = 10m and varying percentages of malicious nodes (30\%, 40\% and 50\%).}
-\label{fig:randommalicious2}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/03_30_10}
-\caption{$h = 0.3$.} \label{fig:threshold03}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/05_30_10}
-\caption{$h = 0.5$.} \label{fig:threshold05}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/thresholds/07_30_10}
-\caption{$h = 0.7$.} \label{fig:threshold07}
-\end{subfigure}
-
-\caption{Simulation with 10\% malicious nodes, $\rho = 30$m and varying values of $h$.}
-\label{fig:randomthresholds}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{images/plots/Network_rA7/10_10}
-\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_10}
-\caption{$m = 10$.} \label{fig:extra10}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_250}
-\caption{$m = 250$.} \label{fig:extra250}
-\end{subfigure}
-
-\caption{Simulation with information aging, with different maximum age values ($m=10, 250$).}
-\label{fig:extramaliciousaging1}
-\end{figure}
-
-\begin{figure}
-\centering
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_1000}
-\caption{$m = 1000$.} \label{fig:extra1000}
-\end{subfigure}
-
-\begin{subfigure}{0.6\textwidth}
-\includegraphics[width=\linewidth]{images/plots/extra_malicious/1_5000}
-\caption{$m = 5000$.} \label{fig:extra5000}
-\end{subfigure}
-
-\caption{Simulation with information aging, with different maximum age values ($m=1000, 5000$).}
-\label{fig:extramaliciousaging2}
-\end{figure}
 

### 2018-07-06 dissertation/6-conclusao/chapter6.tex
small improvement to conclusion

@@
 
-In the coming years, vehicular networks or VANETs will an important part of safety and security in transportation, optimizing traffic and reducing the number of accidents.
+In the coming years, vehicular networks or VANETs will be an important part of safety and security in transportation, optimizing traffic and reducing the number of accidents.
 However, they are an appealing target for entities with malicious intents, creating the need of robust solutions to maintain the integrity of such networks.
@@
 
+In comparison with the related work, TruMan was able to satisfy most of the desired properties for vehicular network trust models, while not inhibiting the properties that were not desired.
+Most importantly, TruMan put emphasis on efficiency and is the first model that clearly displays the complexity of its algorithm.
+
+The work done on TruMan was published as a conference paper on the 2018 IEEE Symposium on Computers and Communications \citep{greca2018truman}.
+
 A next step for future work would be performing experiments simulating more types of known attacks against vehicular networks, then adapting the model to resist such attacks.

### 2018-07-10 dissertation/5-simulations/chapter5.tex
updated conclusion

@@
 %The number of nodes and their communication range vary for different simulations.
+
 There is a total of 160 nodes, 150 of which follow the Working Day Movement Model, and 10 that follow the random waypoint mobility model to simulate vehicles that do not follow daily patterns.
 Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
-Aside from the office size parameter, the parameters regarding offices, meeting spots and shopping were kept intact.
+This configuration was chosen in order to make the network homogenous, but buses and pedestrians could be considered as an addition to the scope.
+
+Aside from the office size parameter, which is 1 in order to inhibit in-office mobility, the parameters regarding offices, meeting spots and shopping were kept intact.
 A small part of nodes move randomly to simulate vehicles that do not follow daily patterns.
@@
 At 50\%, as expected, the results are inconsistent as the network is completely split between benign and malicious nodes; at this point, the network is considered completely compromised.
-True positive and false positive detections are often inverted, because whenever there are more malicious nodes than benign one in a certain node's network model, the algorithm will classify the malicious ones as correct.
+True positive and false positive detections are often inverted, because whenever there are more malicious nodes than benign ones in a certain node's network model, the algorithm will classify the malicious ones as correct.
 The amount of malicious nodes also affects how quickly nodes are able to acquire information about the network, since nodes do not trust information from malicious neighbors.

### 2018-07-10 dissertation/6-conclusao/chapter6.tex
updated conclusion

@@
 
-%As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
-%By placing nodes into strongly connected components, a network containing a large amount of nodes can be simplified into a much smaller one.
-%Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
-%This allows for a low complexity approach to malicious node identification in a dynamic network.
+As nodes travel across the network and collect more data from neighbors, they are able to form an abstraction of the network which can be used to detect malicious nodes.
+By placing nodes into strongly connected components, a network containing a large amount of nodes can be simplified into a much smaller one.
+Using a simple graph coloring algorithm, most malicious nodes stand out by having different colors than the majority of nodes.
+This allows for a low complexity approach to malicious node identification in a dynamic network.
 
@@
 Most importantly, TruMan put emphasis on efficiency and is the first model that clearly displays the complexity of its algorithm.
+Furthermore, TruMan begins taking advantage of social network features found in vehicular networks, although more can be done with this idea.
 
@@
 
-A next step for future work would be performing experiments simulating more types of known attacks against vehicular networks, then adapting the model to resist such attacks.
-Furthermore, TruMan could be extended to consider more types of information in its algorithm, using, for example, V2I communications with road-side units, vehicular roles, or location and time parameters.
-Finally, the concepts of the algorithm could be applied to different types of networks.
+\pagebreak
+Several paths could be considered for future work on TruMan, such as:
+\begin{itemize}
+\item TruMan could be tested in more varied scenarios, using different maps and various different amounts of nodes in the simulations. It would be even better to use real-world mobility data.
+\item TruMan takes advantage of social features found in vehicular networks, but even more could be done with this.
+For example, vehicles that belong to the same family could have strong ties and share a lot of data with each other.
+Certain vehicles, such as police cars and ambulances, could have privileged roles within the model.
+\item The model could be adapted to include public transportation vehicles (trains, buses) with predictable routes, as well as vehicle-to-infrastructure communication.
+\item Well-known vehicular network attacks, such as the ones in \citep{isaac2010security}, could be used against TruMan, attempting to break it.
+Such experiments would fully validate TruMan's robustness.
+\item TruMan should be tested in real-world networks.
+Although it was designed for vehicular networks, other types of networks with mobility, such as mobile ad-hoc networks, could be used for experiments.
+\item Finally, it is possible that TruMan might be a valuable tool for more than just vehicular network.
+A decentralized and dynamic trust management scheme could be useful for social networks, mobile ad-hoc networks, Internet-based peer-to-peer networks, and others.
+These scenarios should be tested.
+\end{itemize}
+
+%Rede complexas: métricas (dinâmica)? Grafo de Componentes para reduzir a complexidade da rede
+%
+%Limitações do One? Dtn
+%Preservação da rede (slide) mesmo sem alcance da comunicação?
+%
+%Trust: trusted vehicles (minhas 2010)
+%
+%Fog computing: cluster em bus stops ou rsup
+%
+%Contribuição: correção & eficiência, redes sociais wdmm, rotina,
+%
+%Janela de tempo para descoberta e atualização dinâmica da topologia- execução dos 3 passos seguintes (limitação do tempo - iteração?) alguma frequência ? F(v, d)
+%M ? 
+%
+%
+%Diagnóstico (ok - fora do escopo) - sem análise do teste do nó.
+%
+%Noticias boas mais rápido que más: retirada da aresta?
+%Modelo de ataques?

### 2018-07-13 dissertation/0-preambulo/resumoextendido.tex
continued Portuguese section

@@
 	\section*{Introdução}
-	Este é o resumo extendido em português.
+%	Computadores estão cada vez menores e mais poderosos, permitindo que diversos aspectos da vida possam ser melhorados ao adicionar unidades de processamento a dispositivos comuns.
+%	Muitas dessas aplicações são focadas em conveniências, a integração de objetos com computadores pode ser usado para economizar tempo e salvar vidas \citep{rti2014}.
+%	Uma forma de fazer isto é adicionando computadores e dispositivos de comunicação a veículos -- como carros, ônibus e trens -- para aumentar a eficiência e a segurança do trânsito.
+%	
+%	Em 2013, cerca de 1,25 milhão de pessoas perderam suas vidas devido a acidentes de trânsito \citep{whofactsheet}.
+%	Nos EUA, este número tem diminuído ao longo das últimas décadas, mas ainda é uma grande causa de morte ao redor do mundo \citep{whotraffic}.
+%	Além disso, com o aumento da população de veículos, congestionamento tomam muito tempo das pessoas em suas rotinas.
+%	Em cidades como Los Angeles, Moscou, Nova Iorque, Bogotá e São Paulo, pessoas passam dezenas de horas por ano no trânsito \citep{inrixtraffic}.
+%	
+%	Veículos inteligentes e redes veiculares são ferramentas tecnológicas para remediar os problemas acima.
+%	Através do uso de sensores e de comunicação sem fio, os veículos podem evitar acidentes ao alertar motoristas distraídos \citep{lee2004collision} ou ao acompanhar a velocidade e posição de outros veículos \citep{hafner2011automated}.
+%	Além disso, podem sugerir rotas para distribuir o tráfego e evitar congestionamentos.
+%	
+%	Se tratando de aplicações de segurança e eficiência no trânsito, é crucial que comunicações ocorram com baixa latência (cerca de 100 milissegundos \citep{camp2005vehicle}).
+%	Tecnologia celular contemporânea, como LTE, poderia ser usada para conectar veículos à internet, mas a latência adicionada pela transmissão tornaria aplicações de segurança lentas e pouco confiáveis \citep{mangel2010comparison}.
+%	Outros potenciais problemas da tecnologia celular para o uso veicular incluem a necessidade de infraestrutura, o custo do serviço, e o compartilhamento da rede com outros tipos de dispositivos.
+%	
+%	Portanto, uma solução \textit{ad-hoc} é preferível, sem depender de infraestrutura e permitindo a comunicação direta entre participantes da rede.
+%	
+%	
+%	Porém, como é o caso em muitas tecnologias novas, VANETs podem ser um alvo de ataques de usuários maliciosos.
+%	Um usuário malicioso local pode alterar dados para manipular o trânsito, enquanto atacantes remotos podem invadir veículos para controlar a rede \citep{garip2015congestion}.
+%	Ataques podem ser apenas inconveniências até ameaças à vida, portanto é importante que redes veiculares estejam preparadas para mitigá-los.
+%	
+%	Em redes \textit{ad-hoc}, uma forma de mitigar certos ataques é usando dados coletados previamente para filtrar mensagens que aparentam ser maliciosas, incorretas ou irrelevantes.
+%	Para tal, emprega-se o conceito de \textit{confiança}.
+%	Ao receber mensagens de outros nós, um membro da rede pode construir uma relação de confiança com os outros.
+%	Caso o valor de confiança da origem de uma mensagem seja muito baixo, essa mensagem pode ser considerada não confiável.
+%	Além disso, essas relações de confiança podem ser propagadas pela rede, permitindo que nós que ainda não formaram suas próprias opiniões possam se beneficiar da informação que já circula pela rede.
+%
 
-	\section*{Redes Complexas}
+	Dentro dos próximos anos, uma grande parte de novos veículos virão equipados com funcionalidades de comunicação.
+	Essas funcionalidades permitirão o compartilhamento de dados com outros dispositivos e podem ser ferramentas importantes para reduzir o trânsito e o risco de acidentes.
+	Acidentes de trânsito são uma das maiores causas de morte no mundo \citep{whofactsheet}, tornando necessárias soluções para melhorar a segurança nas ruas.
 
-	\section*{Redes Veiculares}
+	O compartilhamento rápido de dados entre veículos permite, por exemplo, que veículos inteligentes alertem seus motoristas sobre condições de trânsito \citep{lee2004collision} e que veículos autônomos formem pelotões \citep{amoozadeh2015platoon}.
 
-	\section*{TruMan}
+	Assim, surge a necessidade de redes veiculares \textit{ad-hoc} (\textit{vehicular ad-hoc networks}, ou VANETs), nas quais veículos são os nós ou membros da rede e compartilham informações entre si, sem depender da internet ou de infraestrutura.
+	O padrão de comunicação mais usado para redes veiculares é o IEEE 802.11p, que descreve dois tipos de membros (ou nós) para redes veiculares: unidades a bordo (\textit{on-board units} ou OBUs) e unidades de beira de estrada \textit(\textit{roadside units} ou RSUs).
+	Comunicação entre pares de OBUs é denominada \textit{vehicle-to-vehicle} (V2V), enquanto comunicação entre OBUs e RSUs é chamada de \textit{vehicle-to-infrastructure} (V2I).
+	Este estudo aborda apenas casos V2V e, portanto, refere-se apenas a veículos como membros de uma rede veicular.
 
-	\section*{Simulações}
+	Porém, como é o caso em muitas tecnologias novas, VANETs podem ser um alvo de ataques de usuários maliciosos.
+	Um usuário malicioso local pode alterar dados para manipular o trânsito, enquanto atacantes remotos podem invadir veículos para controlar a rede \citep{garip2015congestion}.
+	Ataques podem ser apenas inconveniências até ameaças à vida, portanto é importante que redes veiculares estejam preparadas para mitigá-los.
+
+	Em redes \textit{ad-hoc}, uma forma de mitigar certos ataques é usando dados coletados previamente para filtrar mensagens que aparentam ser maliciosas, incorretas ou irrelevantes.
+	Para tal, emprega-se o conceito de \textit{confiança}.
+	Ao receber mensagens de outros nós, um membro da rede pode construir uma relação de confiança com os outros.	
+	Caso o valor de confiança da origem de uma mensagem seja muito baixo, essa mensagem pode ser considerada não confiável.
+	Além disso, essas relações de confiança podem ser propagadas pela rede, permitindo que nós que ainda não formaram suas próprias opiniões possam se beneficiar da informação que já circula pela rede.
+
+	Este trabalho propõe um novo modelo de confiança, TruMan, para gerenciar relações de confiança em uma rede veicular.
+	Usando o modelo proposto, nós de uma rede veicular podem rapidamente identificar quais outros nós são dignos de confiança ou não.
+	Como redes veiculares são altamente dinâmicas, nós adquirem mais informações à medida do tempo e podem se beneficiar de propriedades sociais de VANETs para construir relações fortes com outros nós encontrados frequentemente.
+
+	O modelo TruMan é baseado em outro já existente, chamado MaNI \citep{vernize2015malicious}, que era restrito para redes estáticas.
+	Utilizando algoritmos de grafos, TruMan demonstra-se uma solução eficiente para o problema de gerenciamento de confiança em redes veiculares.
+
+	As próximas seções são as seguintes.
+	A \textbf{Revisão Bibliográfica} aborda os conceitos de redes complexas, redes sociais, redes tecnológicas e redes veiculares, além de estudos relevantes na área de confiança para redes veiculares.
+	Em \textbf{Projeto e Implementação do TruMan}, os objetivos e hipóteses do TruMan são apresentados, além das explicações dos algoritmos que compõe o modelo.
+	A seguir, \textbf{Avaliação do TruMan} mostra as ferramentas usadas para validar o TruMan e os resultados dos experimentos realizados.
+	Por fim, a \textbf{Conclusão} contém os pensamentos finais sobre o projeto.
+
+	\section*{Revisão Bibliográfica}
+
+	Uma revisão do estudo de redes como um todo foi feito para o desenvolvimento de um novo modelo de confiança para redes veiculares.
+	Esta seção introduz conceitos relacionados a redes complexas, redes sociais, redes tecnológicas e redes veiculares, incluindo o significado e importância de confiança em cada tipo de rede.
+	Por fim, trabalhos relevantes relacionados à área de confiança em redes veiculares são apresentados.
+
+	\subsection*{Redes Complexas}
+	Redes complexas podem descrever diversos sistemas observados na natureza e na sociedade com \textit{vértices} (ou \textit{nós}) e \textit{arestas} \citep{newmannetworks}.
+	Geralmente são dividas em quatro categorias:
+	\begin{enumerate}
+		\item \textbf{Redes Tecnológicas} são construídas para prover serviços. Exemplos incluem a Internet, a rede telefônica e redes de transporte.
+		\item \textbf{Redes Sociais} são compostas de pessoas ou grupos de pessoas e a relação entre elas. Podem ser relações de parentesco, amizades, relações profissionais, etc.
+		\item \textbf{Redes de Informação} são redes nas quais os nós são informações e as arestas são relações entre conjuntos de informações.
+Por exemplo, a World Wide Web é uma rede de informação que existe na Internet (que, por sua vez, é uma rede tecnológica).
+		\item \textbf{Redes Biológicas} são as redes encontradas na natureza. Podem ser compostas de animais, células, conjuntos de seres vivos, etc.
+	\end{enumerate}
+
+	\subsection*{Redes Sociais}
+	\subsection*{Redes Tecnológicas}
+	\subsection*{Redes Veiculares}
+	\subsection*{Trabalhos Relacionados}
+
+	\section*{Projeto e Implementação do TruMan}
+
+	\section*{Avaliação do TruMan}
 

### 2018-07-17 dissertation/0-preambulo/resumoextendido.tex
continuing portuguese section

@@
 
-	\subsection*{Redes Sociais}
-	\subsection*{Redes Tecnológicas}
+	Confiança pode ser uma ferramenta útil em diversos tipos de rede para proteger seus participantes.
+	Sob a perspectiva da ciência da computação, confiança pode ser definida como uma medida de quanta certeza um membro da rede tem de que outro membro se comporta de forma adequada e fornece informações válidas e/ou significativas \citep{sherchan2013survey}.
+
+	\subsection*{Confiança em Redes Sociais}
+	Redes sociais são aquelas formadas por pessoas e as relações entre elas. 
+	Em uma rede social, é simples observar a importância de confiança, pois é algo usado por pessoas todos os dias para tomar decisões.
+	De forma geral, essas relações de confiança também funcionam quando adaptadas ao meio digital, pois relacionamentos construídos no mundo real permanecem válidos.
+
+	Uma característica importante de redes sociais é a possibilidade de transmitir confiança de um relacionamento para outro, formando o conceito de ``amigos de amigos'' \citep{boissevain1974friends}.
+	Isso permite que pares de integrantes de uma rede social mantenham um laço de confiança mesmo que não se conheçam diretamente, graças a um elo de confiança em comum, permitindo rápida disseminação de informação.
+
+	Em geral, redes sociais são estáticas.
+	Mesmo que laços de confiança sejam criados ou terminados, o formato geral da rede tende a mudar pouco, graças a uma série de outros laços formados entre nós próximos ao evento.
+
+	\subsection*{Confiança em Redes Tecnológicas}
+	Em muitas redes tecnológicas, como na Internet, confiança é aplicada de maneira centralizada através de serviços que fornecem segurança a seus usuários.
+	Ou seja, no contexto da Internet, confiança geralmente é derivada de uma fonte secundária; usuários e provedores de serviço raramente mantêm suas próprias listas de confiabilidade.
+
+	Essas soluções são válidas para a Internet, mas seriam lentas demais para serem usadas em redes dinâmicas \textit{ad-hoc}.
+	Redes \textit{ad-hoc} exigem soluções decentralizadas para gerenciamento de confiança.
+	Ou seja, cada membro da rede armazena suas próprias opiniões sobre membros, originadas a partir de contatos anteriores.
+
+	Além disso, como essas informações são adquiridas aos poucos, o conhecimento que um membro da rede tem é geralmente incerto e incompleto \citep{baras2005cooperation}.
+	Raramente pode-se ter certeza de que determinada informação é completamente precisa.
+	Por isso é importante também levar em consideração as informações adquiridas por vizinhos confiáveis, assim expandido o conhecimento disponível.
+
+	Redes veiculares \textit{ad-hoc} são um caso especial de rede tecnológica, mas, devido a diversas peculiaridades em topologia e mobilidade, também apresentam peculiaridades quando se trata de confiança.
+
 	\subsection*{Redes Veiculares}

### 2018-07-17 dissertation/0-preambulo/resumoextendido.tex
continuing portuguese section

@@
 	Assim, surge a necessidade de redes veiculares \textit{ad-hoc} (\textit{vehicular ad-hoc networks}, ou VANETs), nas quais veículos são os nós ou membros da rede e compartilham informações entre si, sem depender da internet ou de infraestrutura.
-	O padrão de comunicação mais usado para redes veiculares é o IEEE 802.11p, que descreve dois tipos de membros (ou nós) para redes veiculares: unidades a bordo (\textit{on-board units} ou OBUs) e unidades de beira de estrada \textit(\textit{roadside units} ou RSUs).
-	Comunicação entre pares de OBUs é denominada \textit{vehicle-to-vehicle} (V2V), enquanto comunicação entre OBUs e RSUs é chamada de \textit{vehicle-to-infrastructure} (V2I).
-	Este estudo aborda apenas casos V2V e, portanto, refere-se apenas a veículos como membros de uma rede veicular.
-
 	Porém, como é o caso em muitas tecnologias novas, VANETs podem ser um alvo de ataques de usuários maliciosos.
@@
 
-	\subsection*{Redes Veiculares}
+	\subsection*{Confiança em Redes Veiculares}
+	Hoje, muitos veículos já vêm equipados com o hardware necessário para processamento e comunicação veicular.
+	É esperado que, até 2022, a maioria dos veículos comuns também venham com tais funcionalidades \citep{connectedcar2016}.
+	Redes veiculares \textit{ad-hoc} (VANETs) são uma aplicação muito estudada quando se trata de veículos inteligentes ou autônomos.
+	Nelas, todos os nós são relacionados ao trânsito, como veículos ou unidades posicionadas em infraestrutura ao lado das ruas.
+
+	O padrão de comunicação mais usado para redes veiculares é o IEEE 802.11p, que descreve dois tipos de membros (ou nós) para redes veiculares: unidades a bordo (\textit{on-board units} ou OBUs) e unidades de beira de estrada \textit(\textit{roadside units} ou RSUs).
+	Comunicação entre pares de OBUs é denominada \textit{vehicle-to-vehicle} (V2V), enquanto comunicação entre OBUs e RSUs é chamada de \textit{vehicle-to-infrastructure} (V2I).
+	Este estudo aborda apenas casos V2V e, portanto, refere-se apenas a veículos como membros de uma rede veicular.
+
+	Como é esperado para novas tecnologias, redes veiculares podem se tornar um alvo relevante para usuários maliciosos e atacantes.
+	Alguns exemplos de potenciais problemas são: 
+	módulos e sensores, como GPS e velocímetro, defeituosos, inibindo aplicações de segurança ou eficiência \citep{isaac2010security}; 
+	veículos intencionalmente transmitindo dados falsos \citep{golle2004detecting}; 
+	atacantes remotos controlando múltiplos veículos para congestionar a rede \citep{garip2015congestion}; 
+	invasão de privacidade ao tentar decifrar e ler mensagens alheias \citep{isaac2010security};
+	disrupção de sinal para impedir a comunicação de outros veículos \citep{isaac2010security}.
+
+	Como em outros tipos de rede, VANETs dependem de membros que se comportam de maneira correta e previsível e informações incorretas comprometem a utilidade da rede.
+	Há uma distinção importante a ser feita entre nós maliciosos e defeituosos, porém, em termos de confiabilidade, é possível tratá-los da mesma forma, pois, afinal, a principal característica de ambos é a transmissão de informação incorreta.
+
+	Em geral, soluções de gerenciamento de confiança são divididos em dois tipos: os que usam \textit{confiança orientada a entidade}, nos quais confiança é relacionada a membros da rede e leva-se em consideração quem transmitiu certa mensagem, ou \textit{confiança orientada a dados}, nos quais o conteúdo da mensagem é mais importante do que quem a transmitiu.
+	Existem também algumas soluções que combinam ambos métodos.
+
+	Algumas das propriedades únicas de redes veiculares, que afetam soluções de confiança para elas, são:
+	topologia que muda constantemente e rapidamente;
+	mobilidade de nós restringidas às ruas disponíveis;
+	fragmentação, quando duas ou mais partes da rede estão distantes demais para se comunicar;
+	comunicação pouco confiável com nós distantes;
+	nenhuma restrição notável de energia, quando comparadas a redes de dispositivos móveis;
+	densidade potencialmente muito alta;
+	topologia suscetível a comportamentos erráticos de motoristas.
+
+	Em \citep{zhang2011survey}, oito propriedades desejáveis para modelos de confiança para redes veiculares são apresentadas:
+	construção de confiança descentralizada;
+	lidar bem com baixas densidades;
+	dinâmicas relacionadas a local, tempo, eventos e tarefas;
+	escalabilidade;
+	medida de certeza integrada;
+	segurança a nível de sistema;
+	sensibilidade a privacidade;
+	robustez.
+
 	\subsection*{Trabalhos Relacionados}
+
+	Muitas soluções para confiança em redes veiculares foram propostas ao longo dos anos, como \citep{patwardhan2006data}, \citep{gerlach2007trust}, \citep{raya2008data}, \citep{huang2010situation}, \citep{ding2013novel}, \citep{haddadou2013trust}, \citep{liu2016lsot}, \citep{kerrache2016detection}.
+	Além disso, alguns trabalhos oferecem revisões sobre propostas já apresentadas, como \citep{zhang2011survey}, \citep{ma2011survey}, \citep{zhang2012trust}, \cite{mejri2014survey}, \citep{soleymani2015trust}, \citep{sengar2016survey}, \citep{dwivedi2016review}.
+	Nesta seção, alguns dos trabalhos mais relevantes são apresentados.
+
+	\citep{minhas2010towards}
+
+	\citep{chen2010trust}
+
+	\citep{li2016art}
+
+	\citep{chen2017cloud}
 

### 2018-07-17 dissertation/3-vanets/chapter3.tex
continuing portuguese section

@@
 Each of these possible attacks requires a unique approach, though there are some broader ways to help the security and safety of VANET users.
-Trust, as is described in \Cref{section:trustvanet}, can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
+Trust can be an important feature in vehicular networks, especially when attempting to filter out malicious or incorrect messages.
 It does not, however, avoid all possible attacks, such as a signal jammer or stalking.

### 2018-07-18 dissertation/0-preambulo/resumoextendido.tex
continuing portuguese section

@@
 
-	\citep{minhas2010towards}
+	No modelo proposto em \citep{minhas2010towards} usa diversos critérios para julgar se uma mensagem é confiável ou não.
+	Ele utiliza uma combinação de confiança baseada em função (por exemplo, viaturas policiais são automaticamente mais confiáveis) e confiança baseada em experiência (baseada em interações anteriores).
+	Além disso, uma mensagem é considerada mais confiáveis quando sua origem estava próximo do evento sendo relatado por ela.
+	Quando múltiplas mensagens sobre o mesmo evento são recebidas, um nó pode optar por considerar as que foram enviadas por nós mais confiáveis, ou ponderar um consenso baseado em diversas opiniões alheias.
+	Porém, este modelo depende apenas de interações diretas, e confiança não é propagada pela rede.
 
-	\citep{chen2010trust}
+	Em \citep{chen2010trust}, os autores propõem avaliar mensagens com um método que utiliza grupos.
+	Nós são separados em grupos e, cada vez que um deseja enviar uma mensagem, os outros membros do grupo oferecem suas opiniões sobre o emissor.
+	Finalmente, um dos nós, designado como líder do grupo, coleta as opiniões e decide se a mensagem é válida de acordo com o consenso.
+	Porém, é incerto como o modelo funcionaria em redes esparsas, manter grupos em uma rede altamente dinâmica pode ser uma tarefa de alto custo e um grupo todo pode ser comprometido se o líder não for confiável.
 
-	\citep{li2016art}
+	O modelo ART \citep{li2016art} busca um modelo robusto e resistente a ataques.
+	Ele tem dois passos principais: coleta de dados e detecção de nós maliciosos.
+	Utiliza a teoria de evidências Dempster-Shafter para agregar dados vindos de outros nós.
+	Então, usa uma métrica baseada em cosseno para comparar vetores de confiança de dois nós (cada vetor é uma sequência de opiniões que um nó tem sobre outros).
+	Nós com vetores de confiança próximos confiam uns nos outros.
+	O problema dessa solução é a dependência em cálculos custosos que podem atrapalhar o desempenho em situações que exigem baixa latência.
 
-	\citep{chen2017cloud}
+	Os autores de \citep{chen2017cloud} propõem uma solução de confiança baseada em nuvem, que exige um gerenciamento de confiança via internet.
+	A vantagem disso é simplificar diversas dificuldades de redes veiculares, como redes esparsas e altamente dinâmicas.
+	Contudo, o modelo é problemático em regiões com pouco ou nenhum sinal de comunicação celular e o sistema todo é suscetível a instabilidades no serviço.
+
+	Por fim, é importante notar que nenhum dos trabalhos acima oferece análises de custo e complexidade de seus algoritmos.
+	Portanto, manter uma baixa complexidade é um objetivo chave do modelo TruMan.
 
 	\section*{Projeto e Implementação do TruMan}
+
+	TruMan é um modelo de gerenciamento de confiança para redes veiculares, possibilitando a detecção de nós maliciosos em uma rede e a disseminação dados de confiança para outros nós.
+	TruMan busca gerenciamento de confiança eficiente em redes altamente dinâmicas, mantendo baixo custo computacional e um modelo simples de entender e implementar.
+	Esta seção apresenta os fundamentos e algoritmos por trás de TruMan, assim como detalhes de sua implementação.
+
+	\textbf{Redes Sociais e VANETs}
+
+	\subsection*{Algoritmo de Tarjan}
+
+	O algoritmo de Tarjan para componentes fortemente conexos \citep{tarjan1972depth} é uma parte importante da eficiência atingida pelo TruMan.
+	Este algoritmo permite abstrair um grafo grande em outro menor, reduzindo o custo de passos subsequentes.
+	Dado um grafo direcionado $T = (V,E)$, um componente fortemente conexo é um grupo de nós nos quais, para cada par $u, v \in V$, há um caminho de $u$ para $v$ e de $v$ para $u$.
+	Pensando em gerenciamento de confiança, podemos extender essa definição para considerar apenas caminhos formados por arestas que representam relações de confiança (ou um valor de confiança alto).
+
+	O número de componentes é, no máximo, $|V|$: no pior caso, cada nó forma seu próprio componente.
+	A complexidade do algoritmo é de $O(|V|+|E|)$.
+
+	Utilizando os resultados do algoritmo de Tarjan, um grafo não direcionado $C = (V', E')$ é formado.
+	Cada $v' \in V'$ representa um componente de $T$, enquanto as arestas $e' \in E'$ representam arestas entre nós de $T$ que não pertencem ao mesmo componente.
+
+	\subsection*{Algoritmo de coloração de grafos}
+	Coloração de grafos é um problema clássico da teoria dos grafos, no qual deve-se atribuir a cada nó uma cor de forma que não existam dois nós vizinhos com a mesma cor.
+	Coloração de grafos é a heurística usada para a detecção de nós maliciosos no TruMan.
+	Descobrir o menor número de cores possível para colorir um grafo arbitrário é um problema NP-difícil \citep{sanchez1989determining}.
+
+	Os autores de \citep{mittal2011graph} apresentam uma solução eficiente de colorir grafos, apesar de não provar que o algoritmo sempre usa o menor número possível de cores.
+	Porém, o algoritmo é eficiente e seus resultados são corretos, tornando-se extremamente útil para o TruMan.
+	O custo do algoritmo é de apenas $O(|E|)$ -- em comparação, o algoritmo DSATUR tem complexidade de $O(|V|^2)$.
+
+	\subsection*{O algoritmo TruMan}
+
+	TruMan é baseado no algoritmo MaNI \citep{vernize2013dissertation}, que sugeriu o uso de componentes fortemente conexos e de coloração de grafos para a detecção de nós maliciosos em uma rede.
+	Porém, o MaNI foi desenvolvido para redes estáticas e é executado por um agente externo à rede, tornando-se inapropriado para redes veiculares.
+	Para funcionar em redes dinâmicas, TruMan roda iterativamente em intervalos pré-determinados.
+	Além disso, o algoritmo roda de forma descentralizada, com uma instância rodando em cada membro da rede.
+
+	Cada nó $u$ armazena um grafo direcionado de confiança $T = (V,E)$ que é uma abstração da rede real e começa apenas com $V = {u}$.
+	Cada nó em $V$ representa um membro da rede e cada aresta em $E$ representa uma relação de confiança entre dois nós.
+	Como cada nó armazena sua própria representação da rede e essa representação evolui com o tempo, há um $T_i^u = (V_i^u, E_i^u)$ para cada nó $u$ e iteração $i$.
+
+	No começo de cada iteração, nós coletam informações sobre seus vizinhos.
+	Um pré-requisito deste passo é a existência de um teste que classifica um nó adjacente como benigno ou malicioso.
+	Tal teste é um problema grande por si próprio, e sai do escopo deste trabalho.
+	Estudos sobre isso podem ser encontrados em \citep{golle2004detecting}, \citep{li2016defective}, \citep{kerrache2016detection}.
+
+	Cada vez que um nó vizinho $v$ é identificado como benigno, o valor de confiança armazenado em $u \rightarrow v$ aumenta, e o grafo $T_i-1^v$ é unido com o grafo armazenado por $u$.
+	Após coletar informações de todos os seus vizinhos naquele instante, um novo grafo $T_i^u$ é formado, que é utilizado para os próximos passos.
+
+	Em seguida, $T_i^u$ é separado em componentes fortemente conexos usando o algoritmo de Tarjan.
+	Cada par de nós em um componente é conectado por um caminho de confiança.
+	Ou seja, todos os nós de um mesmo componente confiam uns nos outros direta ou indiretamente.
+	Portanto, em termos de confiança, nós dentro de um mesmo componente podem ser considerados como um só: se um deles é confiável, pode-se assumir que todos são.
+	Os componentes tornam-se nós de um grafo $C_i^u = (V'_i^u, E'_i^u)$
 

### 2018-07-19 dissertation/0-preambulo/resumoextendido.tex
continuing portuguese section

@@
 
-	Em seguida, $T_i^u$ é separado em componentes fortemente conexos usando o algoritmo de Tarjan.
-	Cada par de nós em um componente é conectado por um caminho de confiança.
+	Em seguida, $T_i^u$ é separado em componentes fortemente conexos usando o algoritmo de Tarjan \citep{tarjan1972depth}, de forma que cada par de nós em um componente é conectado por um caminho de confiança.
 	Ou seja, todos os nós de um mesmo componente confiam uns nos outros direta ou indiretamente.
 	Portanto, em termos de confiança, nós dentro de um mesmo componente podem ser considerados como um só: se um deles é confiável, pode-se assumir que todos são.
-	Os componentes tornam-se nós de um grafo $C_i^u = (V'_i^u, E'_i^u)$
+	Os componentes tornam-se nós do grafo $C^u_i = (V'^u_i, E'^u_i)$.
+
+	O algoritmo de coloração de grafos \citep{mittal2011graph} é usado como heurística para classificar nós como benignos ou não.
+	Após a execução do algoritmo, a cor cujos nós em $C_i^u$ representam a maior quantidade de nós em $T_i^u$ é classificada como correta, e as outras cores são classificadas como incorretas.
+	Isto funciona porque, em uma rede na qual a maior parte dos nós são benignos, estes tendem a formar poucos componentes grandes, enquanto os nós maliciosos pertencem a componentes pequenos.
+	Assume-se que a maior parte dos nós seja sempre benigna -- caso contrário, a rede como um todo está comprometida e perde completamente sua função.
+
+	A complexidade do algoritmo pode ser calculada ao somar as operações mais custosas.
+	Para cada iteração $i$ e nó $u$, e sendo $n$ o número de vizinhos de $u$, o cálculo é o seguinte:
+
+	$$ TruMan = Tarjan + Colora\text{\textit{\c{c}}}\tilde ao + (Uni\tilde ao\times(n^u_i)) $$
+
+	Como discutido acima, o algoritmo de Tarjan tem complexidade de $O(|V|+|E|)$ para o grafo de confiança $T$.
+	Já o algoritmo de coloração tem complexidade de $O(|E'|)$ para o grafo de componentes $C$.
+	A parte mais custosa do algoritmo é a união de grafos que acontece após a comunicação entre nós confiáveis.
+	A complexidade desse processo é de $O(|E|)$ para vizinho que um nó tem em uma determinada iteração; o número de vizinhos é, no máximo, $|V|$.
+	A complexidade total do Truman é, portanto:
+
+	$$ O(|V|+|E|) + O(|E'|) + O(|V|\times |E|)$$
+
+	Porém, $|E'| \leq |E|$ é sempre verdade, porque o grafo $C$ é uma redução do grafo $T$.
+	Além disso, $|V|+|E| \leq |V|\times |E|$ também é verdade, com a exceção do cenário irrelevante no qual $|V| \leq 1$ ou $|E| \leq 1$.
+	Portanto, a complexidade do algoritmo TruMan pode ser simplificada como:
+
+	$$O(|V| \times |E|)$$
 
@@
 
+	Para testar o TruMan, uma implementação do algoritmo foi feita usando Python.
+	Grafos com mobilidade de nós foram gerados no simulador ONE \citep{keranen2009one}, usando o \textit{Working Day Movement Model} \citep{ekman2008working} para providenciar mobilidade próxima ao do mundo real, e um mapa de parte da cidade de Helsinki, Finlândia.
+	Imagens da topologia da rede foram salvas a cada 10 segundos simulados, e essas imagens foram usadas como entrada para o algoritmo TruMan.
+	O comportamento de nós maliciosos é aleatorizar suas opiniões sobre seus vizinhos.
+
+	A maioria dos parâmetros da simulação no ONE foram tiradas do artigo do Working Day Movement Model \citep{ekman2008working}.
+	Alguns parâmetros diferentes foram usados, mostrados na Tabela \ref{table:parameters}.
+	Todos os nós da simulação são carros; para o propósito deste trabalho, nenhum outro tipo de veículo foi considerado.
+	Uma parte pequena dos nós movimenta-se aleatoriamente, para simular veículos que não seguem padrões de movimento diários.
+
+	O raio de transmissão dos nós varia de 10m a 50m, ilustrando as diferenças entre diferentes densidades de rede.
+	A densidade de rede ($\delta$) é um valor que abstrai o volume e a frequência de conexões em redes veiculares, estimando a cobertura da rede pelo ambiente.
+	Para o TruMan, densidades mais altas trazem melhores resultados, pois nós podem adquirir informações mais rapidamente.
+	A densidade é calculada usando o raio de transmissão ($\rho$), o número de nós ($\eta$) e a área da simulação ($\alpha$).
+
+	A cobertura de um único nó é o círculo ao redor dele formado pelo raio de transmissão.
+	Esse valor é dividido por dois para compensar círculos sobrepostos, e então multiplicado pelo número de nós na rede.
+	Por fim, esse valor é dividido pela área total do ambiente.
+	A fórmula de densidade da rede é a seguinte:
+
+	$$ \delta = \frac{\frac{\rho^2\pi}{2} \times \eta}{\alpha} $$
+
+	As densidades de algumas simulações realizadas são exibidas na Tabela \ref{table:simdensities}.
+	Já a tabela \ref{table:citydensities} mostra as densidades de rede hipotéticas de algumas cidades do mundo, usando dados geográficos reais e supondo um raio de transmissão de apenas 10m.
+	É possível ver que, mesmo com um raio de transmissão pequeno, as cidades oferecem densidades de rede maiores do que as das simulações, assumindo que uma parcela substancial de seus automóveis seja equipada com dispositivos de conexão.
+
+	\subsection*{Simulações}
+
+	Para validar o desempenho e a corretude de TruMan, diversas simulações foram executadas.
+
+	A figura \ref{fig:random103050} mostra simulações com 10\% dos nós agindo maliciosamente, com raio de comunicação entre 10m e 50m.
+	É possível ver como o aumento do raio de comunicação melhora bastante os resultados:
+	com 10m, quase 8000 iterações são necessárias para atingir um bom resultado, enquanto com 50 são apenas cerca de 1000.
+
+	As figuras \ref{fig:randommalicious1} e \ref{fig:randommalicious2} mostram simulações com quantidades diferentes de nós maliciosos e raio de comunicação de apenas 10m.
+	Com até 30\% de nós maliciosos, os resultados são bons.
+	Com 40\%, uma parte pequena de nós maliciosos não são detectados.
+	Já com 50\%, os resultados são erráticos, pois o controle da rede é dividido entre os nós benignos e maliciosos.
+
 	\section*{Conclusão}

### 2018-07-22 dissertation/0-preambulo/resumoextendido.tex
continuing portuguese section

@@
 	As densidades de algumas simulações realizadas são exibidas na Tabela \ref{table:simdensities}.
-	Já a tabela \ref{table:citydensities} mostra as densidades de rede hipotéticas de algumas cidades do mundo, usando dados geográficos reais e supondo um raio de transmissão de apenas 10m.
+	Já a Tabela \ref{table:citydensities} mostra as densidades de rede hipotéticas de algumas cidades do mundo, usando dados geográficos reais e supondo um raio de transmissão de apenas 10m.
 	É possível ver que, mesmo com um raio de transmissão pequeno, as cidades oferecem densidades de rede maiores do que as das simulações, assumindo que uma parcela substancial de seus automóveis seja equipada com dispositivos de conexão.
@@
 
-	A figura \ref{fig:random103050} mostra simulações com 10\% dos nós agindo maliciosamente, com raio de comunicação entre 10m e 50m.
+	A Figura \ref{fig:random103050} mostra simulações com 10\% dos nós agindo maliciosamente, com raio de comunicação entre 10m e 50m.
 	É possível ver como o aumento do raio de comunicação melhora bastante os resultados:
@@
 
-	As figuras \ref{fig:randommalicious1} e \ref{fig:randommalicious2} mostram simulações com quantidades diferentes de nós maliciosos e raio de comunicação de apenas 10m.
+	As Figuras \ref{fig:randommalicious1} e \ref{fig:randommalicious2} mostram simulações com quantidades diferentes de nós maliciosos e raio de comunicação de apenas 10m.
 	Com até 30\% de nós maliciosos, os resultados são bons.
@@
 
+	A maioria das simulações foram feitas com o limiar de confiança $h=0.5$, que significa que, para um nó confiar em outro, o valor de confiança deve ser acima de $0.5$.
+	A Figura {fig:randomthresholds} demonstra o impacto de mudar esse limiar.
+	Pode-se ver que o impacto não é muito significativo, porém, com $h=0.7$, os resultados são um pouco melhores.
+
+	A Figura \ref{fig:random7} mostra os resultados da execução do algoritmo durante 7 dias simulados.
+	A maioria dos nós maliciosos é detectada ao fim do primeiro dia, e a rede é completamente descoberta pouco tempo depois.
+	Com o tempo, o número de falsos positivos cai, até tornar-se um número insignificante.
+
+	Por fim, as Figuras \ref{fig:extramaliciousaging1} e \ref{fig:extramaliciousaging2} mostram como TruMan reage a um potencial ataque.
+	Nessas simulações, um nó passa a ser malicioso na metade do tempo.
+	O parâmetro $m$ determina quantas iterações o algoritmo leva para descartar arestas antigas, e, portanto, afeta a agilidade do modelo ao detectar um nó convertido.
+	Com $m$ muito baixo, informações são descartadas muito rapidamente e o número de falsos positivos aumenta drasticamente.
+
 	\section*{Conclusão}

### 2018-07-24 dissertation/0-preambulo/resumoextendido.tex
finished portuguese section

@@
 	A maioria das simulações foram feitas com o limiar de confiança $h=0.5$, que significa que, para um nó confiar em outro, o valor de confiança deve ser acima de $0.5$.
-	A Figura {fig:randomthresholds} demonstra o impacto de mudar esse limiar.
+	A Figura \ref{fig:randomthresholds} demonstra o impacto de mudar esse limiar.
 	Pode-se ver que o impacto não é muito significativo, porém, com $h=0.7$, os resultados são um pouco melhores.
@@
 	\section*{Conclusão}
+
+	Nos próximos anos, comunicação veicular será uma importante ferramenta para segurança e eficiência em transportes.
+	Porém, elas também serão um alvo de atacantes e usuários maliciosos.
+	Confiança é um conceito poderoso para evitar a disseminação de dados falsos entre membros de uma rede.
+	Neste projeto, um novo modelo de confiança para redes veiculares, chamado TruMan, foi apresentado.
+	TruMan combina algoritmos eficientes para providenciar gerenciamento de confiança rápido para redes altamente dinâmicas.
+
+	Enquanto nós (veículos) viajam pela rede, eles adquirem mais informações sobre outros membros da rede, e podem usar essas informações para detectar nós maliciosos.
+	A utilização de componentes fortemente conexos permite que o grafo da rede seja simplificado em um menor, no qual cada nó é uma abstração de diversos membros da rede.
+	Então, com um algoritmo de coloração de grafo, membros maliciosos se destacam ao ter cores diferentes da maioria dos outros nós.
+
+	TruMan foi testado usando dados gerados com o simulador ONE e o Working Day Movement Model.
+	Os experimentos mostram que veículos podem formar uma abstração suficiente da rede para detectar nós maliciosos ao redor deles.
+	Em geral, quanto mais tempo durar a simulação e quanto maior for a densidade de nós na rede, melhores são os resultados.
+
+	Em comparação aos trabalhos relacionados, TruMan satisfaz a maior parte das propriedades desejáveis para modelos de mobilidade em redes veiculares, e, ao mesmo tempo, permite que as outras propriedades sejam implementadas de outras formas.
+	O foco do TruMan é eficiência e é o primeiro modelo de confiança para redes veiculares a apresentar claramente a complexidade de seu algoritmo.
+
+	Há diversas opções para trabalhos futuros relacionados ao TruMan, como, por exemplo:
+	testes em uma variedade maior de cenários, usando dados reais de mobilidade;
+	aproveitar mais as propriedades sociais de redes veiculares;
+	a integração com veículos de transporte público e infraestrutura;
+	testes de resistência contra ataques conhecidos;
+	testes em redes móveis reais;
+	experimentação em outros tipos de rede.
+
 \end{resumoextendido}

### 2018-08-10 dissertation/0-preambulo/resumoextendido.tex
applied final revision

@@
 
-	Este trabalho propõe um novo modelo de confiança, TruMan, para gerenciar relações de confiança em uma rede veicular.
-	Usando o modelo proposto, nós de uma rede veicular podem rapidamente identificar quais outros nós são dignos de confiança ou não.
-	Como redes veiculares são altamente dinâmicas, nós adquirem mais informações à medida do tempo e podem se beneficiar de propriedades sociais de VANETs para construir relações fortes com outros nós encontrados frequentemente.
-
-	O modelo TruMan é baseado em outro já existente, chamado MaNI \citep{vernize2015malicious}, que era restrito para redes estáticas.
-	Utilizando algoritmos de grafos, TruMan demonstra-se uma solução eficiente para o problema de gerenciamento de confiança em redes veiculares.
-
-	As próximas seções são as seguintes.
-	A \textbf{Revisão Bibliográfica} aborda os conceitos de redes complexas, redes sociais, redes tecnológicas e redes veiculares, além de estudos relevantes na área de confiança para redes veiculares.
-	Em \textbf{Projeto e Implementação do TruMan}, os objetivos e hipóteses do TruMan são apresentados, além das explicações dos algoritmos que compõe o modelo.
-	A seguir, \textbf{Avaliação do TruMan} mostra as ferramentas usadas para validar o TruMan e os resultados dos experimentos realizados.
-	Por fim, a \textbf{Conclusão} contém os pensamentos finais sobre o projeto.
-
-	\section*{Revisão Bibliográfica}
-
-	Uma revisão do estudo de redes como um todo foi feito para o desenvolvimento de um novo modelo de confiança para redes veiculares.
-	Esta seção introduz conceitos relacionados a redes complexas, redes sociais, redes tecnológicas e redes veiculares, incluindo o significado e importância de confiança em cada tipo de rede.
-	Por fim, trabalhos relevantes relacionados à área de confiança em redes veiculares são apresentados.
-
-	\subsection*{Redes Complexas}
-	Redes complexas podem descrever diversos sistemas observados na natureza e na sociedade com \textit{vértices} (ou \textit{nós}) e \textit{arestas} \citep{newmannetworks}.
-	Geralmente são dividas em quatro categorias:
-	\begin{enumerate}
-		\item \textbf{Redes Tecnológicas} são construídas para prover serviços. Exemplos incluem a Internet, a rede telefônica e redes de transporte.
-		\item \textbf{Redes Sociais} são compostas de pessoas ou grupos de pessoas e a relação entre elas. Podem ser relações de parentesco, amizades, relações profissionais, etc.
-		\item \textbf{Redes de Informação} são redes nas quais os nós são informações e as arestas são relações entre conjuntos de informações.
-Por exemplo, a World Wide Web é uma rede de informação que existe na Internet (que, por sua vez, é uma rede tecnológica).
-		\item \textbf{Redes Biológicas} são as redes encontradas na natureza. Podem ser compostas de animais, células, conjuntos de seres vivos, etc.
-	\end{enumerate}
-
-	Confiança pode ser uma ferramenta útil em diversos tipos de rede para proteger seus participantes.
-	Sob a perspectiva da ciência da computação, confiança pode ser definida como uma medida de quanta certeza um membro da rede tem de que outro membro se comporta de forma adequada e fornece informações válidas e/ou significativas \citep{sherchan2013survey}.
-
-	\subsection*{Confiança em Redes Sociais}
-	Redes sociais são aquelas formadas por pessoas e as relações entre elas. 
-	Em uma rede social, é simples observar a importância de confiança, pois é algo usado por pessoas todos os dias para tomar decisões.
-	De forma geral, essas relações de confiança também funcionam quando adaptadas ao meio digital, pois relacionamentos construídos no mundo real permanecem válidos.
-
-	Uma característica importante de redes sociais é a possibilidade de transmitir confiança de um relacionamento para outro, formando o conceito de ``amigos de amigos'' \citep{boissevain1974friends}.
-	Isso permite que pares de integrantes de uma rede social mantenham um laço de confiança mesmo que não se conheçam diretamente, graças a um elo de confiança em comum, permitindo rápida disseminação de informação.
-
-	Em geral, redes sociais são estáticas.
-	Mesmo que laços de confiança sejam criados ou terminados, o formato geral da rede tende a mudar pouco, graças a uma série de outros laços formados entre nós próximos ao evento.
-
-	\subsection*{Confiança em Redes Tecnológicas}
-	Em muitas redes tecnológicas, como na Internet, confiança é aplicada de maneira centralizada através de serviços que fornecem segurança a seus usuários.
-	Ou seja, no contexto da Internet, confiança geralmente é derivada de uma fonte secundária; usuários e provedores de serviço raramente mantêm suas próprias listas de confiabilidade.
-
-	Essas soluções são válidas para a Internet, mas seriam lentas demais para serem usadas em redes dinâmicas \textit{ad-hoc}.
-	Redes \textit{ad-hoc} exigem soluções decentralizadas para gerenciamento de confiança.
-	Ou seja, cada membro da rede armazena suas próprias opiniões sobre membros, originadas a partir de contatos anteriores.
-
-	Além disso, como essas informações são adquiridas aos poucos, o conhecimento que um membro da rede tem é geralmente incerto e incompleto \citep{baras2005cooperation}.
-	Raramente pode-se ter certeza de que determinada informação é completamente precisa.
-	Por isso é importante também levar em consideração as informações adquiridas por vizinhos confiáveis, assim expandido o conhecimento disponível.
-
-	Redes veiculares \textit{ad-hoc} são um caso especial de rede tecnológica, mas, devido a diversas peculiaridades em topologia e mobilidade, também apresentam peculiaridades quando se trata de confiança.
-
-	\subsection*{Confiança em Redes Veiculares}
 	Hoje, muitos veículos já vêm equipados com o hardware necessário para processamento e comunicação veicular.
@@
 	segurança a nível de sistema;
-	sensibilidade a privacidade;
+	sensibilidade a privacidade; e
 	robustez.
 
-	\subsection*{Trabalhos Relacionados}
+	Este trabalho propõe um novo modelo de confiança, TruMan, para gerenciar relações de confiança em uma rede veicular.
+	Usando o modelo proposto, nós de uma rede veicular podem rapidamente identificar quais outros nós são dignos de confiança ou não.
+	Como redes veiculares são altamente dinâmicas, nós adquirem mais informações à medida do tempo e podem se beneficiar de propriedades sociais de VANETs para construir relações fortes com outros nós encontrados frequentemente.
+
+	O modelo TruMan é baseado em outro já existente, chamado MaNI \citep{vernize2015malicious}, que era restrito para redes estáticas.
+	Utilizando algoritmos de grafos, TruMan demonstra-se uma solução eficiente para o problema de gerenciamento de confiança em redes veiculares.
+
+	As próximas seções são as seguintes.
+	A \textbf{Revisão Bibliográfica} mostra estudos relevantes na área de confiança para redes veiculares.
+	Em \textbf{Projeto e Implementação do TruMan}, os objetivos e hipóteses do TruMan são apresentados, além das explicações dos algoritmos que compõe o modelo.
+	A seguir, \textbf{Avaliação do TruMan} mostra as ferramentas usadas para validar o TruMan e os resultados dos experimentos realizados.
+	Por fim, a \textbf{Conclusão} contém os pensamentos finais sobre o projeto.
+
+%	\section*{Revisão Bibliográfica}
+%	
+%	Uma revisão do estudo de redes como um todo foi feito para o desenvolvimento de um novo modelo de confiança para redes veiculares.
+%	Esta seção introduz conceitos relacionados a redes complexas, redes sociais, redes tecnológicas e redes veiculares, incluindo o significado e importância de confiança em cada tipo de rede.
+%	Por fim, trabalhos relevantes relacionados à área de confiança em redes veiculares são apresentados.
+%	
+%	\subsection*{Redes Complexas}
+%	Redes complexas podem descrever diversos sistemas observados na natureza e na sociedade com \textit{vértices} (ou \textit{nós}) e \textit{arestas} \citep{newmannetworks}.
+%	Geralmente são dividas em quatro categorias:
+%	\begin{enumerate}
+%		\item \textbf{Redes Tecnológicas} são construídas para prover serviços. Exemplos incluem a Internet, a rede telefônica e redes de transporte.
+%		\item \textbf{Redes Sociais} são compostas de pessoas ou grupos de pessoas e a relação entre elas. Podem ser relações de parentesco, amizades, relações profissionais, etc.
+%		\item \textbf{Redes de Informação} são redes nas quais os nós são informações e as arestas são relações entre conjuntos de informações.
+Por exemplo, a World Wide Web é uma rede de informação que existe na Internet (que, por sua vez, é uma rede tecnológica).
+%		\item \textbf{Redes Biológicas} são as redes encontradas na natureza. Podem ser compostas de animais, células, conjuntos de seres vivos, etc.
+%	\end{enumerate}
+%	
+%	Confiança pode ser uma ferramenta útil em diversos tipos de rede para proteger seus participantes.
+%	Sob a perspectiva da ciência da computação, confiança pode ser definida como uma medida de quanta certeza um membro da rede tem de que outro membro se comporta de forma adequada e fornece informações válidas e/ou significativas \citep{sherchan2013survey}.
+%	
+%	\subsection*{Confiança em Redes Sociais}
+%	Redes sociais são aquelas formadas por pessoas e as relações entre elas. 
+%	Em uma rede social, é simples observar a importância de confiança, pois é algo usado por pessoas todos os dias para tomar decisões.
+%	De forma geral, essas relações de confiança também funcionam quando adaptadas ao meio digital, pois relacionamentos construídos no mundo real permanecem válidos.
+%	
+%	Uma característica importante de redes sociais é a possibilidade de transmitir confiança de um relacionamento para outro, formando o conceito de ``amigos de amigos'' \citep{boissevain1974friends}.
+%	Isso permite que pares de integrantes de uma rede social mantenham um laço de confiança mesmo que não se conheçam diretamente, graças a um elo de confiança em comum, permitindo rápida disseminação de informação.
+%	
+%	Em geral, redes sociais são estáticas.
+%	Mesmo que laços de confiança sejam criados ou terminados, o formato geral da rede tende a mudar pouco, graças a uma série de outros laços formados entre nós próximos ao evento.
+%	
+%	\subsection*{Confiança em Redes Tecnológicas}
+%	Em muitas redes tecnológicas, como na Internet, confiança é aplicada de maneira centralizada através de serviços que fornecem segurança a seus usuários.
+%	Ou seja, no contexto da Internet, confiança geralmente é derivada de uma fonte secundária; usuários e provedores de serviço raramente mantêm suas próprias listas de confiabilidade.
+%	
+%	Essas soluções são válidas para a Internet, mas seriam lentas demais para serem usadas em redes dinâmicas \textit{ad-hoc}.
+%	Redes \textit{ad-hoc} exigem soluções decentralizadas para gerenciamento de confiança.
+%	Ou seja, cada membro da rede armazena suas próprias opiniões sobre membros, originadas a partir de contatos anteriores.
+%	
+%	Além disso, como essas informações são adquiridas aos poucos, o conhecimento que um membro da rede tem é geralmente incerto e incompleto \citep{baras2005cooperation}.
+%	Raramente pode-se ter certeza de que determinada informação é completamente precisa.
+%	Por isso é importante também levar em consideração as informações adquiridas por vizinhos confiáveis, assim expandido o conhecimento disponível.
+%	
+%	Redes veiculares \textit{ad-hoc} são um caso especial de rede tecnológica, mas, devido a diversas peculiaridades em topologia e mobilidade, também apresentam peculiaridades quando se trata de confiança.
+%	
+%	\subsection*{Confiança em Redes Veiculares}
+
+	\section*{Revisão Bibliográfica}
 
@@
 
-	\textbf{Redes Sociais e VANETs}
-
-	\subsection*{Algoritmo de Tarjan}
-
-	O algoritmo de Tarjan para componentes fortemente conexos \citep{tarjan1972depth} é uma parte importante da eficiência atingida pelo TruMan.
-	Este algoritmo permite abstrair um grafo grande em outro menor, reduzindo o custo de passos subsequentes.
-	Dado um grafo direcionado $T = (V,E)$, um componente fortemente conexo é um grupo de nós nos quais, para cada par $u, v \in V$, há um caminho de $u$ para $v$ e de $v$ para $u$.
-	Pensando em gerenciamento de confiança, podemos extender essa definição para considerar apenas caminhos formados por arestas que representam relações de confiança (ou um valor de confiança alto).
-
-	O número de componentes é, no máximo, $|V|$: no pior caso, cada nó forma seu próprio componente.
-	A complexidade do algoritmo é de $O(|V|+|E|)$.
-
-	Utilizando os resultados do algoritmo de Tarjan, um grafo não direcionado $C = (V', E')$ é formado.
-	Cada $v' \in V'$ representa um componente de $T$, enquanto as arestas $e' \in E'$ representam arestas entre nós de $T$ que não pertencem ao mesmo componente.
-
-	\subsection*{Algoritmo de coloração de grafos}
-	Coloração de grafos é um problema clássico da teoria dos grafos, no qual deve-se atribuir a cada nó uma cor de forma que não existam dois nós vizinhos com a mesma cor.
-	Coloração de grafos é a heurística usada para a detecção de nós maliciosos no TruMan.
-	Descobrir o menor número de cores possível para colorir um grafo arbitrário é um problema NP-difícil \citep{sanchez1989determining}.
-
-	Os autores de \citep{mittal2011graph} apresentam uma solução eficiente de colorir grafos, apesar de não provar que o algoritmo sempre usa o menor número possível de cores.
-	Porém, o algoritmo é eficiente e seus resultados são corretos, tornando-se extremamente útil para o TruMan.
-	O custo do algoritmo é de apenas $O(|E|)$ -- em comparação, o algoritmo DSATUR tem complexidade de $O(|V|^2)$.
-
-	\subsection*{O algoritmo TruMan}
+%	\subsection*{Algoritmo de Tarjan}
+%	
+%	O algoritmo de Tarjan para componentes fortemente conexos \citep{tarjan1972depth} é uma parte importante da eficiência atingida pelo TruMan.
+%	Este algoritmo permite abstrair um grafo grande em outro menor, reduzindo o custo de passos subsequentes.
+%	Dado um grafo direcionado $T = (V,E)$, um componente fortemente conexo é um grupo de nós nos quais, para cada par $u, v \in V$, há um caminho de $u$ para $v$ e de $v$ para $u$.
+%	Pensando em gerenciamento de confiança, podemos extender essa definição para considerar apenas caminhos formados por arestas que representam relações de confiança (ou um valor de confiança alto).
+%
+%	O número de componentes é, no máximo, $|V|$: no pior caso, cada nó forma seu próprio componente.
+%	A complexidade do algoritmo é de $O(|V|+|E|)$.
+%	
+%	Utilizando os resultados do algoritmo de Tarjan, um grafo não direcionado $C = (V', E')$ é formado.
+%	Cada $v' \in V'$ representa um componente de $T$, enquanto as arestas $e' \in E'$ representam arestas entre nós de $T$ que não pertencem ao mesmo componente.
+%	
+%	\subsection*{Algoritmo de coloração de grafos}
+%	Coloração de grafos é um problema clássico da teoria dos grafos, no qual deve-se atribuir a cada nó uma cor de forma que não existam dois nós vizinhos com a mesma cor.
+%	Coloração de grafos é a heurística usada para a detecção de nós maliciosos no TruMan.
+%	Descobrir o menor número de cores possível para colorir um grafo arbitrário é um problema NP-difícil \citep{sanchez1989determining}.
+%	
+%	Os autores de \citep{mittal2011graph} apresentam uma solução eficiente de colorir grafos, apesar de não provar que o algoritmo sempre usa o menor número possível de cores.
+%	Porém, o algoritmo é eficiente e seus resultados são corretos, tornando-se extremamente útil para o TruMan.
+%	O custo do algoritmo é de apenas $O(|E|)$ -- em comparação, o algoritmo DSATUR tem complexidade de $O(|V|^2)$.
+%	
+%	\subsection*{O algoritmo TruMan}
 
@@
 
-	\section*{Avaliação do TruMan}
+	\subsection*{Avaliação do TruMan}
 
@@
 
-	\subsection*{Simulações}
+%	\subsection*{Simulações}
 
@@
 
-	A Figura \ref{fig:random103050} mostra simulações com 10\% dos nós agindo maliciosamente, com raio de comunicação entre 10m e 50m.
-	É possível ver como o aumento do raio de comunicação melhora bastante os resultados:
-	com 10m, quase 8000 iterações são necessárias para atingir um bom resultado, enquanto com 50 são apenas cerca de 1000.
+	Nas simulações com 10\% dos nós agindo maliciosamente, com raio de comunicação entre 10m e 50m, é possível ver como o aumento do raio de comunicação melhora bastante os resultados: com 10m, quase 8000 iterações são necessárias para atingir um bom resultado, enquanto com 50 são apenas cerca de 1000.
 
-	As Figuras \ref{fig:randommalicious1} e \ref{fig:randommalicious2} mostram simulações com quantidades diferentes de nós maliciosos e raio de comunicação de apenas 10m.
-	Com até 30\% de nós maliciosos, os resultados são bons.
-	Com 40\%, uma parte pequena de nós maliciosos não são detectados.
-	Já com 50\%, os resultados são erráticos, pois o controle da rede é dividido entre os nós benignos e maliciosos.
+	Em simulações com raio de comunicação de apenas 10m e até 30\% de nós maliciosos, os resultados são bons. 
+	Porém, com 40\% de nós maliciosos, uma parte pequena desses nós não são detectados.
+	Já com 50\% de maliciosos, os resultados são erráticos, pois o controle da rede é dividido entre os nós benignos e maliciosos.
 
 	A maioria das simulações foram feitas com o limiar de confiança $h=0.5$, que significa que, para um nó confiar em outro, o valor de confiança deve ser acima de $0.5$.
-	A Figura \ref{fig:randomthresholds} demonstra o impacto de mudar esse limiar.
-	Pode-se ver que o impacto não é muito significativo, porém, com $h=0.7$, os resultados são um pouco melhores.
+	Simulações com $h=0.3$ e $h=0.7$ demonstram o impacto de mudar esse limiar.
+	É possível observar que o impacto não é muito significativo, porém, com $h=0.7$, os resultados são um pouco melhores.
 
-	A Figura \ref{fig:random7} mostra os resultados da execução do algoritmo durante 7 dias simulados.
-	A maioria dos nós maliciosos é detectada ao fim do primeiro dia, e a rede é completamente descoberta pouco tempo depois.
+	Quando o algoritmo é executado durante 7 dias simulados, a maioria dos nós maliciosos é detectada ao fim do primeiro dia, e a rede é completamente descoberta pouco tempo depois.
 	Com o tempo, o número de falsos positivos cai, até tornar-se um número insignificante.
 
-	Por fim, as Figuras \ref{fig:extramaliciousaging1} e \ref{fig:extramaliciousaging2} mostram como TruMan reage a um potencial ataque.
-	Nessas simulações, um nó passa a ser malicioso na metade do tempo.
+	Por fim, as simulações nas quais um nó passa a ser malicioso na metade do tempo mostram como TruMan reage a um possível ataque.
 	O parâmetro $m$ determina quantas iterações o algoritmo leva para descartar arestas antigas, e, portanto, afeta a agilidade do modelo ao detectar um nó convertido.

### 2018-08-10 dissertation/5-simulations/chapter5.tex
created legend image
applied final revision

@@
 
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.3\linewidth]{images/legend2.png}
+%\caption{Legend showing the meaning of each color in the following graphs.} 
+%\label{fig:legend}
+%\end{figure}
+
 It is expected that the results are very inconsistent at the beginning of the simulations, since nodes are still building their models of the network and have relatively little information to use.
@@
 \end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+%\centering
+\includegraphics[width=\linewidth]{images/legend2.png}
+%\caption{Legend} 
+%\label{fig:legend}
+\end{subfigure}
 
@@
 
-\centering
+%\centering
 \begin{subfigure}{0.5\textwidth}
@@
 \caption{10\% malicious.}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+%\centering
+\includegraphics[width=\linewidth]{images/legend2.png}
+%\caption{Legend} 
+%\label{fig:legend}
 \end{subfigure}
@@
 \caption{50\% malicious.}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+%\centering
+\includegraphics[width=\linewidth]{images/legend2.png}
+%\caption{Legend} 
+%\label{fig:legend}
 \end{subfigure}
@@
 \caption{$h = 0.7$.} \label{fig:threshold07}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+%\centering
+\includegraphics[width=\linewidth]{images/legend2.png}
+%\caption{Legend} 
+%\label{fig:legend}
 \end{subfigure}
@@
 \begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{images/plots/Network_rA7/10_10}
-\caption{7 days scenario: 10m range and 10\% malicious nodes.} \label{fig:random7}
+%\centering
+
+\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=\textwidth]{images/plots/Network_rA7/10_10}
+\end{subfigure}
+\hspace*{\fill} % separation between the subfigures
+\begin{subfigure}{0.3\textwidth}
+%\centering
+\includegraphics[width=\linewidth]{images/legend2.png}
+%\caption{Legend} 
+%\label{fig:legend}
+\end{subfigure}
+
+\caption{7 days scenario: 10m range and 10\% malicious nodes.}
+\label{fig:random7}
 \end{figure}
@@
 
-\centering
-\caption{Simulation of TruMan with information aging, with different maximum age values ($m=10, 250$).}
-\label{fig:extramaliciousaging1}
-\end{figure}
+\vspace{1cm}
 
-\begin{figure}
+%\centering
+%\caption{Simulation of TruMan with information aging, with different maximum age values ($m=10, 250$).}
+%\label{fig:extramaliciousaging1}
+%\end{figure}
+%
+%\begin{figure}
 
@@
 \pagebreak
-\Cref{fig:extramaliciousaging1} and \Cref{fig:extramaliciousaging2} show how TruMan reacts to one node converting from benign to malicious.
+\Cref{fig:extramaliciousaging2} shows how TruMan reacts to one node converting from benign to malicious.
 These were the only simulations to consider information aging.
@@
 \begin{table}[hpbt!]
-\caption{Properties of the related work}
+\caption{Satisfaction of desired properties compared to related work}
 \label{table:properties_truman}
@@
 \end{tabular}
+
+\begin{flushleft}
+	\begin{center} 
+		\textbf{Desired properties} 
+	\end{center}
+	\hspace{10.5em} 1. Decentralized trust establishment\\
+	\hspace{10.5em} 2. Coping with sparsity\\
+	\hspace{10.5em} 3. Event/task and location/time dynamics\\
+	\hspace{10.5em} 4. Scalability\\
+	\hspace{10.5em} 5. Integrated confidence measure\\
+	\hspace{10.5em} 6. System level security\\
+	\hspace{10.5em} 7. Sensitivity to privacy concerns\\
+	\hspace{10.5em} 8. Robustness
+\end{flushleft}
 \end{table}
